### 2015-02-11 manuscript.tex
Upload to github

\documentclass{pnastwo}
\usepackage{cite}
\usepackage{color}

\begin{document}

\title{Demography and selection during maize domestication}
\author{Timothy M. Beissinger\affil{1}{University of
    California,Davis}, Li Wang\affil{2}{Iowa State University}, Arun
  Durvasula\affil{1}{}, Kate Crosby\affil{1}{}, \and Jeffrey
  Ross-Ibrarra\affil{1}{}\affil{2}{Genome Center, UC
    Davis}\affil{3}{Center for population biology, UC Davis} }

\significancetext{This work is insignificant ;-)}

\maketitle

\begin{article}

\begin{abstract}
This is the abstract. It should probably be somewhere around 200 words.
\end{abstract}

\dropcap{T}his is the beginning of the article. Notice the
dropcap... that is a neat feature that PNAS likes, I think it looks
pretty neat too! This is where the introduction will live.

\section{Results}
\subsection{Patterns of variability differ between genic and
  nongenic regions of the genome}
Text goes here

\subsection{Hard sweeps do not shape maize (or teosinte) diversity}
A mutation that is immediately beneficial and positively selected leaves a classical hard
sweep signature in the genome, whereby genetic diversity surrounding
that mutation is reduced as the haplotype where the mutation first
arose increases in frequency to fixation. The prevalence of such hard
sweeps was evaluated by comparing diversity
surrounding non-synonymous and synonymous substitutions in maize and
teosinte since divergence from tripsicum, roughly \textcolor{red}{ZZZ} years before
present \textcolor{red}{(ref here)}. For both taxa, no difference in diversity
surrounding these classes of substitutions was identified \textcolor{red}{(figure
here)}. Additionally, diversity around maize substitutions 
not seen in teosinte was investigated. These sites have the
potential to correspond to recent maize sweeps, occuring after the
split from teosinte. Again, no difference was observed
between diversity around synonymous and nonsynonymous
substitutions \textcolor{red}{(Figure here)}. Together, these observations suggest that hard sweeps
are not a primary form of selection for either maize or teosinte.

\subsection{Patterns of purifying and background selection can be explained by demography}
Purifying selection refers to the situation where deleterious
mutations arising in a population are continuously selected against. When this
form of selection is operating, it can serve to reduce genetic
variability at linked neutral sites, a phenomenon referred to as
background selection \cite{charlesworth93}. Purifying and background selection lead
to lower diversity within genes and other functional sites relative
to neutral regions. We investigated purifying
selection in maize and teosinte by evaluating the average magnitude of reduced
diversity within genes and recovery away from genes in both taxa
\textcolor{red}{(figure here)}. When standardized by neutral
levels (diversity far from genes), a stronger reduction of diversity
and slower recovery was observed for teosinte than for maize, implying
that purifying selection has left a more pronounced signature in the
teosinte genome. This conflicted with our \emph{a priori} hypothesis;
we expected that strong artificial selection since domestication would
have caused enhanced purifying selection for maize. We therefore
conducted the same analysis based on singleton diversity. As a class,
singleton alleles depict the most recent patterns of evolution, but
also have the lowest effect on pairwise diversity. Therefore, unlike
pairwise diversity, patterns of singleton
diversity reflect recent patterns of evolution. When
evaluating the data in this manner, an opposite relationship was
observed. Maize singleton diversity was lower than teosinte singleton
diversity near genes, and recovered more slowly
\textcolor{red}{(figure here)}, implying that in the
recent past maize has been more influenced by purifying selection than
teosinte. Together, these findings imply that demographic history has
a strong influence on the effect of purifying selection. Historically,
teosinte has had a larger population size than maize, and only
recently has maize population size overcome that of teosinte. Since
the efficacy of purifying selection scales with population size,
these results likely reflect changes in $N_e$ more than they reflect
underlying changes in selection pressure.

\textcolor{red}{This might be a good place to stick in a paragraph
  about integrating to estimate BGS based on Jeff's work.}

\subsection{Demography of maize domestication}
Text goes here

\section{Discussion}
\subsection{Discussion 1}
Text goes here

\subsection{Discussion 2}
Text goes here

\begin{materials}
\subsection{Plant materials}
Accessions studied were selected from the Maize HapMap2
panel \cite{hapmap2} . Principal component analysis was employed to ensure that
closely related individuals were not included due to their potential
to bias results \textcolor{red}{(maybe a supplemental figure here)}. Ultimately, 23 maize inbreds derived from a diverse
assortment of landraces were selected for inclusion. Thirteen teosinte
inbred lines, all members of the subspecies Z. \emph{mays}
ssp. \emph{parviglumis}, were utilized. Sequences were mapped to the
maize B73 version 3 reference genome \cite{maizeGenome}
(ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/).

\subsection{Interpolating genetic position}
For many of the following analyses, physical position along a
chromosome was a less relevant measure of map location than was genetic
position. Therefore, physical positions were converted to genetic
positions by interpolating from the NAM genetic map (REFERENCE), which
provides a 1 cM resolution for physical to genetic conversion. Within
R \cite{R}, physical positions with corresponding genetic positions in
the NAM map were used as anchors. Physical positions in our dataset
wihout corresponding genetic position were assigned genetic positions by scaling
the anchored genetic positions according to the physical distance
between the unlabeled position and the flanking anchors.

\subsection{Estimating the site frequency spectrum}
To estimate individual  and joint site frequency spectra (SFS) for
maize and teosinte from inbred lines, each inbred individual was treated as representing
a single haplotype from its population. These were
separately computed for genic and intergenic regions, as well as for
the whole genome together. First, genic and intergenic regions were isolated using
the biomaRt package \cite{biomaRt1,biomaRt2} of R \cite{R}. Genic
regions were defined as DNA between the start and stop position of a
gene, while intergenic regions were required to be at least 5kb up- or
down-stream from a gene start/stop. With regions defined, SFS were
estimated with ANGSD \cite{angsd}. Individual population SFS were
estimated using all positions observed in at least 80\% of the
individuals in the population, and joint SFS were estimated using all
positions observed in at least 80\% of individuals in both
populations. Individuals were
assumed to be fully inbred (-doSaf 2), and subsequently allele frequencies were
divided by two to indicate haplotype frequencies. Quality filters were
employed such that reads with quality score below 30 and bases with
quality score below 20 were discarded (-minMapq 30 and -minQ 20), as were reads that didn't map
uniquely (-uniqueOnly 0). Quality scores around indels were adjusted
as in Samtools (-baq 0). Genotype likelihoods were estimated using the samtools
method (-GL 1). Major and minor alleles were inferred from the data
(-doMaf 1). Because ANGSD cannot calculate a folded joint SFS, the
maize reference genome was used for polarization and then unfolded spectra
were folded using dadi \cite{dadi}.

\subsection{Demographic inference}

\subsection{Evaluating diversity around substitutions}
To investigate diversity around substitutions, maize and teosinte pairwise diversity was first
calculated in 1,000 kb non-overlapping windows using ANGSD
\cite{angsd}. This was performed separately for both maize and
teosinte, using the same filters as employed for estimating the
SFS. Next, SNPs and genotypes among maize, teosinte, and tripsicum were called. Tripsicum bam files
were downloaded from (TRIPSICUM FILES), and then all SNPs with a
p-value less than 1e-6 were called using ANGSD. Quality filters were
as the same as before, and genotypes were only called when the
posterior probability was above 0.95. From the set of called SNPs and
genotypes, substitutions between maize and tripsicum, as well as
between teosinte and tripsicum were identified using R \cite{R} as all positions with
no more than 20\% missing data for which every maize or teosinte
allele differed from the observed tripsicum allele. At each class of
substitution, effects were estimated using the ensembl variant effects
predictor \cite{vep}.

For each diversity window with at least 100 bps observed, the distance from the window center to the
nearest synonymous and nonsynonymous (missense) substitution was
computed. Then, following the methods of \cite{hernandez11}, a loess
curve was plotted for diversity values against the distance to the
nearest synonymous or nonsynonymous substitution. A span of 0.01 was
utilized. Unlike
\cite{sattath11}, we did not fit separate loess curves in the up- and
down-stream directions, but instead fit single curves encompassing
both directions.

\subsection{Evaluating diversity around genes and conserved sequences}
Two types of diversity surrounding genes were investigated. The first
was pairwise diversity in 1kb windows, as described previously. The
second was singleton diversity in 1kb windows. Singletons represent
the rarest class of alleles that this dataset can identify, and
collectively demonstrate the most recent patterns of evolution. Minor
allele frequencies were estimated with ANGSD \cite{angsd} using the
same quality filters previously described. Then, the number of
singletons in each non-overlapping 1kb window was calculated with R
\cite{R}. BiomaRt \cite{biomaRt1, biomaRt2} was then used to identify
the center of each gene. Next, the distance from each diversity window
to the nearest gene center was computed. Teosinte diversity is
generally higher then maize diversity. Therefore, to enable comparisons between
the reduction of diversity around genes in maize and teosinte, a
neutral measure of pairwise and singleton diversity for each taxa was estimated according to
mean pairwise and singleton diversity at windows greater than 0.01 cM from the nearest
gene. Then, pairwise and singleton diversity at each window was
standardized by dividing by the corresponding neutral
measure. Separately for pairwise and singleton diversity in maize and
teosinte, cubic smoothing splines were fit to
describe diversity levels according to the distance to the nearest
gene. Significant differences were assessed by taking 100 bootstrap
samples and re-fitting the cubic smoothing spline to each. Then, the
2.5\% and 97.5\% quantiles of values along the bootstrapped splines
were identified.

\subsection{Simulations}

\end{materials}

\begin{acknowledgments}
Various thankyous will be in order.
\end{acknowledgments}

\begin{thebibliography}{10}
\bibitem{dadi}
R.~Gutenkunst, R.~Hernandez,S.~Williamson, and C.~Bustamante, {\em
  Inferring the joint demographic history of multiple populations from
  multidimensional SNP frequency data}, PLoS Genetics., 5:10 (2009), e1000695.

\bibitem{hapmap2}
J.~Chia, C.~Song, P.~Bradbury, D.~Costich, N.~de~Leon, and others,
\emph{Maize HapMap2 identifies extant variation from a genome in
  flux}, Nature Genetics., 44 (2012), 803-807.

\bibitem{maizeGenome}
P.~Schnable, D.~Ware, R.~Fulton, J.~Stein, F.~Wei, \emph{The B73 maize
genome:complexity, diversity, and dynamics}, Science., 326:5956
(2009), 1112-1115.

\bibitem{angsd}
T.~Korneliussen, A.~Albrechtsen, and R.~Nielsen, \emph{ANGSD: Analysis of
next generation sequencing data}, BMC Bioinformatics., 15:356 (2014),
10.1186/s12859-014-0356-4.

\bibitem{R}
R Core Team, \emph{R: A language and environment for statistical
  computing}, R Foundation for Statistical Computing., Vienna,
Austria.

\bibitem{biomaRt1}
S.~Durinck, P.T.~Soekknabm E.~Birney, and W.~Huber, \emph{Mapping
  identifiers for the integration of genomic datasets with the
  R/Bioconductor package biomaRt}, Nature Protocols., 4 (2009),
1184-1191.

\bibitem{biomaRt2}
S.~Durinck, Y.~Moreau, A.~Kasprzyk, S.~Davis, B.~De~Moor, and others,
\emph{BioMart and Bioconductor: a powerful link between biological
  databases and microarray data analysis}, Bioinformatics, 21 (2005),
3439-3440.

\bibitem{vep}
W.~McLaren, B.~Pritchard, D.~Rios, Y.~Chen, P.~Flicek, and
F.~Cunningham, \emph{Deriving the consequences of genomic variants
  with the Ensembl API and SNP Effect Predictor}, Bioinformatics,
26:16 (2010), 2069-2070.

\bibitem{hernandez11}
R.D.~Hernandez, J.L.~Kelley, E.~Elyashiv, S.C.~Melton, A.~Auton, and
others, \emph{Classic selective sweeps were rare in recent human
  evolution}, Science, 331:6019 (2011), 920-924.

\bibitem{sattath11}
S.~Sattath, E.~Elyashiv, O.~Kolodny, Y.~Rinott, and G.~Sella,
\emph{Pervasive adaptative protein evolution apparent in diversity
  patterns around amino acid substitutions in Drosophila simulans},
PLoS Genetics, 7:2 (2011), e1001302.

\bibitem{charlesworth93}
B.~Charlesworth, M.~T.~Morgan, and D.~Charlesworth, \emph{The Effect
  of Deleterious Mutations on Neutral Molecular Variation}, Genetics,
134 (1993), 1289-1303.

\end{thebibliography}
\end{article}
\end{document}

### 2015-04-30 manuscript.tex
updated

@@
     California,Davis}, Li Wang\affil{2}{Iowa State University}, Arun
-  Durvasula\affil{1}{}, Kate Crosby\affil{1}{}, \and Jeffrey
-  Ross-Ibrarra\affil{1}{}\affil{2}{Genome Center, UC
-    Davis}\affil{3}{Center for population biology, UC Davis} }
+  Durvasula\affil{1}{}, Kate Crosby\affil{1}{}, Matthew Hufford\affil{2}{}, \and Jeffrey
+  Ross-Ibrarra\affil{1}{}\affil{3}{Genome Center, UC
+    Davis}\affil{4}{Center for population biology, UC Davis} }
 
@@
 \subsection{Demography of maize domestication}
-Text goes here
+To explore whether differences in purifying selection between maize
+and teosinte can be explained purely by demographic processes, we
+first estimated the parameters of maize domestication using large-scale
+sequencing data involving 23 inbred maize
+landraces and 13 teosinte inbred lines included in the HapMap 2 panel
+\cite{hapmap2} \textcolor{red}{(supplemental table here)}. The maize
+lines were collected from across the Americas and the teosinte lines
+came from central Mexico. Before estimating demography, we compared
+the site frequency spectrum (SFS) in genic and non-genic regions, and
+observed substantial differences in the evolution of these classes of
+sites. For both maize and teosinte, the SFS within genes showed a
+dearth of low-frequency alleles and Tajima's D
+\textcolor{red}{(reference here)} was therefore shifted
+to more positive values \textcolor{red}{(figure here)}. This is consistent with the aforementioned purifying
+selection and indicates that genic regions are not evolving
+neutrally. Therefore, for demographic modeling we restricted analysis
+to non-genic sites.
+
+We used diffusion approximation as implemented in
+dadi \cite{dadi} to find the domestication parameters that best explain
+the joint site frequency spectrum of maize and teosinte.  The model we optimized began with an
+equilibrium ancestral population of size $N_a$
+splitting into separate maize and teosinte populations $T_B$ generations in the past. Moving forward in
+time from the split, teosinte maintained the ancestral population size, while
+maize experienced an immediate effective population size change to
+$N_b$ individuals, followed by exponential growth to size
+$N_m$. Additionally, since the split a $M_{tm}$ individuals migrated
+from teosinte into maize and $M_{mt}$ individuals migrated from maize
+to teosinte.
+
+The most likely model suggested an ancestral for $\theta$ ($4N\mu$) of
+0.014734, which is similar to previous estimates
+\textcolor{red}{(references here)}. Assuming a mutation rate of $\mu =
+3 \times 10^{-8}$ \textcolor{red}{(reference here)}, this suggests an effective
+population size of $N_a = 122,783$ individuals, a population split
+$T_B = 15,523$ generations in the past, a maize bottleneck
+effective populations size of $N_b = 6,455$ individuals (5.26\% of
+$N_a$), a modern maize effective population size of $N_m = 366,973$
+individuals, $M_{tm} = 1.35$ migrants per generation from teosinte to
+maize, and $M_{mt} = 1.72$ migrants per generation from maize to
+teosinte \textcolor{red}{(figure here)}. We note that a population
+split over 15 thousand generations before present precedes estimates
+from 
+archaeological data which suggests maize domestication began
+approximately 9,000 years before present \textcolor{red}{(reference
+  here)}. This could result from multiple generations per year, or in
+may reflect teosinte population structure that was present before
+domestication. Also, note that the genetic time of the population
+split must precede morphological changes that could be identified
+morphologically.   Additionally, because recent expansion is
+most evidenced by rare alleles, and because these data provide low
+power to detect rare alleles, we expect that the estimate of  $N_m =
+366,973$, or $\sim 3N_a$, is likely an underestimate. \textcolor{red}{(This
+  is where evaluation using Kate's SFS should go)}.
 

### 2015-05-01 manuscript.tex
updates

@@
 \usepackage{color}
+\usepackage{pgf}
 
@@
 
-\title{Demography and selection during maize domestication}
+\title{Demography and selection since maize domestication}
 \author{Timothy M. Beissinger\affil{1}{University of
@@
 present \textcolor{red}{(ref here)}. For both taxa, no difference in diversity
-surrounding these classes of substitutions was identified \textcolor{red}{(figure
-here)}. Additionally, diversity around maize substitutions 
+surrounding these classes of substitutions was identified (Figure \ref{hardSweeps}). Additionally, diversity around maize substitutions 
 not seen in teosinte was investigated. These sites have the
@@
 
+\begin{figure}[h]
+%\vspace*{.05in}
+\centering
+\includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_unNeutralized}
+\caption{Pairwise diversity surrounding synonymous and nonsynonymous
+  substitutions in maize. Observe that the reduction of diversity
+  surrounding nonsynonymous subsitutions is no more severe than the
+  reduction of diversity surrounding synonymous substitutions. This
+  equivalency suggests that hard sweeps are not a primary driver of
+  maize diversity.}
+\label{hardSweeps}
+\end{figure}
+
 \subsection{Patterns of purifying and background selection can be explained by demography}
@@
 form of selection is operating, it can serve to reduce genetic
-variability at linked neutral sites, a phenomenon referred to as
-background selection \cite{charlesworth93}. Purifying and background selection lead
+variability at linked neutral sites, a phenomenon referred called
+background selection \cite{charlesworth1993}. Purifying and background selection lead
 to lower diversity within genes and other functional sites relative
-to neutral regions. We investigated purifying
+to neutral regions \textcolor{red}{(ref here)}. We investigated purifying
 selection in maize and teosinte by evaluating the average magnitude of reduced
@@
 \textcolor{red}{This might be a good place to stick in a paragraph
-  about integrating to estimate BGS based on Jeff's work.}
+  about curve fitting B to estimate s, mu.}
 
@@
 landraces and 13 teosinte inbred lines included in the HapMap 2 panel
-\cite{hapmap2} \textcolor{red}{(supplemental table here)}. The maize
+\cite{chia2012} \textcolor{red}{(supplemental table here)}. The maize
 lines were collected from across the Americas and the teosinte lines
@@
 We used diffusion approximation as implemented in
-dadi \cite{dadi} to find the domestication parameters that best explain
+dadi \cite{gutenkunst2009} to find the domestication parameters that best explain
 the joint site frequency spectrum of maize and teosinte.  The model we optimized began with an
@@
 Accessions studied were selected from the Maize HapMap2
-panel \cite{hapmap2} . Principal component analysis was employed to ensure that
+panel \cite{chia2012} . Principal component analysis was employed to ensure that
 closely related individuals were not included due to their potential
@@
 ssp. \emph{parviglumis}, were utilized. Sequences were mapped to the
-maize B73 version 3 reference genome \cite{maizeGenome}
+maize B73 version 3 reference genome \cite{schnable2009}
 (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/).
@@
 provides a 1 cM resolution for physical to genetic conversion. Within
-R \cite{R}, physical positions with corresponding genetic positions in
+R \cite{R2014}, physical positions with corresponding genetic positions in
 the NAM map were used as anchors. Physical positions in our dataset
@@
 the whole genome together. First, genic and intergenic regions were isolated using
-the biomaRt package \cite{biomaRt1,biomaRt2} of R \cite{R}. Genic
+the biomaRt package \cite{durinck2009,durinck2005} of R \cite{R2014}. Genic
 regions were defined as DNA between the start and stop position of a
@@
 down-stream from a gene start/stop. With regions defined, SFS were
-estimated with ANGSD \cite{angsd}. Individual population SFS were
+estimated with ANGSD \cite{korneliussen2014}. Individual population SFS were
 estimated using all positions observed in at least 80\% of the
@@
 maize reference genome was used for polarization and then unfolded spectra
-were folded using dadi \cite{dadi}.
+were folded using dadi \cite{gutenkunst2009}.
 
@@
 calculated in 1,000 kb non-overlapping windows using ANGSD
-\cite{angsd}. This was performed separately for both maize and
+\cite{korneliussen2014}. This was performed separately for both maize and
 teosinte, using the same filters as employed for estimating the
@@
 genotypes, substitutions between maize and tripsicum, as well as
-between teosinte and tripsicum were identified using R \cite{R} as all positions with
+between teosinte and tripsicum were identified using R \cite{R2014} as all positions with
 no more than 20\% missing data for which every maize or teosinte
@@
 substitution, effects were estimated using the ensembl variant effects
-predictor \cite{vep}.
+predictor \cite{mclaren2010}.
 
@@
 nearest synonymous and nonsynonymous (missense) substitution was
-computed. Then, following the methods of \cite{hernandez11}, a loess
+computed. Then, following the methods of \cite{hernandez2011}, a loess
 curve was plotted for diversity values against the distance to the
@@
 utilized. Unlike
-\cite{sattath11}, we did not fit separate loess curves in the up- and
+\cite{sattath2011}, we did not fit separate loess curves in the up- and
 down-stream directions, but instead fit single curves encompassing
@@
 collectively demonstrate the most recent patterns of evolution. Minor
-allele frequencies were estimated with ANGSD \cite{angsd} using the
+allele frequencies were estimated with ANGSD \cite{korneliussen2014} using the
 same quality filters previously described. Then, the number of
 singletons in each non-overlapping 1kb window was calculated with R
-\cite{R}. BiomaRt \cite{biomaRt1, biomaRt2} was then used to identify
+\cite{R2014}. BiomaRt \cite{durinck2009, durinck2005} was then used to identify
 the center of each gene. Next, the distance from each diversity window
@@
 
-\begin{thebibliography}{10}
-\bibitem{dadi}
-R.~Gutenkunst, R.~Hernandez,S.~Williamson, and C.~Bustamante, {\em
-  Inferring the joint demographic history of multiple populations from
-  multidimensional SNP frequency data}, PLoS Genetics., 5:10 (2009), e1000695.
+% \begin{thebibliography}{10}
+% \bibitem{dadi}
+% R.~Gutenkunst, R.~Hernandez,S.~Williamson, and C.~Bustamante, {\em
+%   Inferring the joint demographic history of multiple populations from
+%   multidimensional SNP frequency data}, PLoS Genetics., 5:10 (2009), e1000695.
 
-\bibitem{hapmap2}
-J.~Chia, C.~Song, P.~Bradbury, D.~Costich, N.~de~Leon, and others,
-\emph{Maize HapMap2 identifies extant variation from a genome in
-  flux}, Nature Genetics., 44 (2012), 803-807.
+% \bibitem{hapmap2}
+% J.~Chia, C.~Song, P.~Bradbury, D.~Costich, N.~de~Leon, and others,
+% \emph{Maize HapMap2 identifies extant variation from a genome in
+%   flux}, Nature Genetics., 44 (2012), 803-807.
 
-\bibitem{maizeGenome}
-P.~Schnable, D.~Ware, R.~Fulton, J.~Stein, F.~Wei, \emph{The B73 maize
-genome:complexity, diversity, and dynamics}, Science., 326:5956
-(2009), 1112-1115.
+% \bibitem{maizeGenome}
+% P.~Schnable, D.~Ware, R.~Fulton, J.~Stein, F.~Wei, \emph{The B73 maize
+% genome:complexity, diversity, and dynamics}, Science., 326:5956
+% (2009), 1112-1115.
 
-\bibitem{angsd}
-T.~Korneliussen, A.~Albrechtsen, and R.~Nielsen, \emph{ANGSD: Analysis of
-next generation sequencing data}, BMC Bioinformatics., 15:356 (2014),
-10.1186/s12859-014-0356-4.
+% \bibitem{angsd}
+% T.~Korneliussen, A.~Albrechtsen, and R.~Nielsen, \emph{ANGSD: Analysis of
+% next generation sequencing data}, BMC Bioinformatics., 15:356 (2014),
+% 10.1186/s12859-014-0356-4.
 
-\bibitem{R}
-R Core Team, \emph{R: A language and environment for statistical
-  computing}, R Foundation for Statistical Computing., Vienna,
-Austria.
+% \bibitem{R}
+% R Core Team, \emph{R: A language and environment for statistical
+%   computing}, R Foundation for Statistical Computing., Vienna,
+% Austria.
 
-\bibitem{biomaRt1}
-S.~Durinck, P.T.~Soekknabm E.~Birney, and W.~Huber, \emph{Mapping
-  identifiers for the integration of genomic datasets with the
-  R/Bioconductor package biomaRt}, Nature Protocols., 4 (2009),
-1184-1191.
+% \bibitem{biomaRt1}
+% S.~Durinck, P.T.~Soekknabm E.~Birney, and W.~Huber, \emph{Mapping
+%   identifiers for the integration of genomic datasets with the
+%   R/Bioconductor package biomaRt}, Nature Protocols., 4 (2009),
+% 1184-1191.
 
-\bibitem{biomaRt2}
-S.~Durinck, Y.~Moreau, A.~Kasprzyk, S.~Davis, B.~De~Moor, and others,
-\emph{BioMart and Bioconductor: a powerful link between biological
-  databases and microarray data analysis}, Bioinformatics, 21 (2005),
-3439-3440.
+% \bibitem{biomaRt2}
+% S.~Durinck, Y.~Moreau, A.~Kasprzyk, S.~Davis, B.~De~Moor, and others,
+% \emph{BioMart and Bioconductor: a powerful link between biological
+%   databases and microarray data analysis}, Bioinformatics, 21 (2005),
+% 3439-3440.
 
-\bibitem{vep}
-W.~McLaren, B.~Pritchard, D.~Rios, Y.~Chen, P.~Flicek, and
-F.~Cunningham, \emph{Deriving the consequences of genomic variants
-  with the Ensembl API and SNP Effect Predictor}, Bioinformatics,
-26:16 (2010), 2069-2070.
+% \bibitem{vep}
+% W.~McLaren, B.~Pritchard, D.~Rios, Y.~Chen, P.~Flicek, and
+% F.~Cunningham, \emph{Deriving the consequences of genomic variants
+%   with the Ensembl API and SNP Effect Predictor}, Bioinformatics,
+% 26:16 (2010), 2069-2070.
 
-\bibitem{hernandez11}
-R.D.~Hernandez, J.L.~Kelley, E.~Elyashiv, S.C.~Melton, A.~Auton, and
-others, \emph{Classic selective sweeps were rare in recent human
-  evolution}, Science, 331:6019 (2011), 920-924.
+% \bibitem{hernandez11}
+% R.D.~Hernandez, J.L.~Kelley, E.~Elyashiv, S.C.~Melton, A.~Auton, and
+% others, \emph{Classic selective sweeps were rare in recent human
+%   evolution}, Science, 331:6019 (2011), 920-924.
 
-\bibitem{sattath11}
-S.~Sattath, E.~Elyashiv, O.~Kolodny, Y.~Rinott, and G.~Sella,
-\emph{Pervasive adaptative protein evolution apparent in diversity
-  patterns around amino acid substitutions in Drosophila simulans},
-PLoS Genetics, 7:2 (2011), e1001302.
+% \bibitem{sattath11}
+% S.~Sattath, E.~Elyashiv, O.~Kolodny, Y.~Rinott, and G.~Sella,
+% \emph{Pervasive adaptative protein evolution apparent in diversity
+%   patterns around amino acid substitutions in Drosophila simulans},
+% PLoS Genetics, 7:2 (2011), e1001302.
 
-\bibitem{charlesworth93}
-B.~Charlesworth, M.~T.~Morgan, and D.~Charlesworth, \emph{The Effect
-  of Deleterious Mutations on Neutral Molecular Variation}, Genetics,
-134 (1993), 1289-1303.
+% \bibitem{charlesworth93}
+% B.~Charlesworth, M.~T.~Morgan, and D.~Charlesworth, \emph{The Effect
+%   of Deleterious Mutations on Neutral Molecular Variation}, Genetics,
+% 134 (1993), 1289-1303.
 
-\end{thebibliography}
+% \end{thebibliography}
+
+\bibliography{Reference}
+\bibliographystyle{abbrvnat}
+%\nocite{*}
+
 \end{article}

### 2015-05-01 manuscript.tex
update

@@
   nongenic regions of the genome}
-Text goes here
+Previous research of the maize domestication process has relied upon
+observations drawn from genic DNA \textcolor{red}{(several references
+  here)}. Our data were generated through whole genome sequencing,
+which eliminated this constraint. Importantly, we observed substantial
+differences in patterns of diversity between genic and non-genic regions
+of the genome for both maize and teosinte. For maize, mean pairwise
+diversity ($\pi$) within genes was significantly lower than at
+positions at least 5kb away from genes (0.00668 within, 0.00691 away, p$<$2e-44). The same
+pattern of significantly greater diversity outside of genes was observed in
+teosinte, but to a larger extent. For teosinte we observed $\pi$
+within genes to be 0.0088 and at least 5kb from genes it was 0.1153
+(p$\approx$0). These obserrvations suggest that genes are not evolving
+neutrally in maize or teosinte. Instead, some form of selection must
+be reducing diversity within genes. Additionally, these observations
+suggest that demographic inference, which relies on the assumption of
+neutrally evolving DNA, will be more accurate if it is based on
+observations from non-genic genetic material.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[b]
 %\vspace*{.05in}

### 2015-05-05 manuscript.tex
updates

@@
 
-\dropcap{T}his is the beginning of the article. Notice the
-dropcap... that is a neat feature that PNAS likes, I think it looks
-pretty neat too! This is where the introduction will live.
+\dropcap{T}his is the beginning of the article. For now, I will give a
+bulleted list of the introduction:
+\begin{enumerate}
+\item Maize
+\item Plant domestication
+\item Maize domestication
+\item Importance of demography
+\item Background selection
+\end{enumerate}
 
@@
 (p$\approx$0). These obserrvations suggest that genes are not evolving
-neutrally in maize or teosinte. Instead, some form of selection must
-be reducing diversity within genes. Additionally, these observations
+neutrally in maize or teosinte. Instead, some form of selection is likely
+reducing diversity within genes. Additionally, these observations
 suggest that demographic inference, which relies on the assumption of
@@
 between diversity around synonymous and nonsynonymous
-substitutions \textcolor{red}{(Figure here)}. Together, these observations suggest that hard sweeps
+substitutions \textcolor{red}{(Supplemental figure here)}. Together, these observations suggest that hard sweeps
 are not a primary form of selection for either maize or teosinte.
@@
 selection in maize and teosinte by evaluating the average magnitude of reduced
-diversity within genes and recovery away from genes in both taxa
-\textcolor{red}{(figure here)}. When standardized by neutral
-levels (diversity far from genes), a stronger reduction of diversity
+diversity within genes and recovery away from genes in both taxa. When standardized by neutral
+levels (which we defined as diversity at least \textcolor{red}{0.01} cM distal
+from genes), a stronger reduction of diversity
 and slower recovery was observed for teosinte than for maize, implying
 that purifying selection has left a more pronounced signature in the
-teosinte genome. This conflicted with our \emph{a priori} hypothesis;
+teosinte genome (Figure \ref{purify}). This conflicted with our \emph{a priori} hypothesis;
 we expected that strong artificial selection since domestication would
@@
 diversity near genes, and recovered more slowly
-\textcolor{red}{(figure here)}, implying that in the
+(Figure \ref{purify}), implying that in the
 recent past maize has been more influenced by purifying selection than
@@
 underlying changes in selection pressure.
+
+\begin{figure*}[b]
+%\vspace*{.05in}
+\centering
+\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_manuscript.png}
+\caption{Relative level of diversity versus distance to the nearest
+  gene, in maize and teosinte. Two measures of diversity were
+  investigated. \textbf{A} displays pairwise
+  diversity, which is most influenced by intermediate frequency
+  alleles and therefore depicts more ancient evolutionary patterns,
+  and \textbf{B} depicts singleton diversity, influenced by rare
+  alleles and thus depicting evolutionary patterns in the recent past.}
+\label{purify}
+\end{figure*}
 
@@
 \bibliography{Reference}
-\bibliographystyle{abbrvnat}
+\bibliographystyle{pnas}
 %\nocite{*}

### 2015-05-05 manuscript.tex
updates

@@
 
-\subsection{Patterns of purifying and background selection can be explained by demography}
+\subsection{Patterns of purifying and background selection are
+  influenced by demographic history}
 Purifying selection refers to the situation where deleterious
@@
 maize, and $M_{mt} = 1.72$ migrants per generation from maize to
-teosinte \textcolor{red}{(figure here)}. We note that a population
+teosinte (Figure \ref{bottleneck}). We note that a population
 split over 15 thousand generations before present precedes estimates
@@
   is where evaluation using Kate's SFS should go)}.
+
+\begin{figure}[b]
+%\vspace*{.05in}
+\centering
+\includegraphics[width=.5\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
+\caption{This is what domestication looked like. WRITE CAPTION!}
+\label{bottleneck}
+\end{figure}
+
+\subsection{Simulations confirm the relationship between demographic
+  processes and background selection}
+\textcolor{red}{We explored whether or not the demographic patterns
+  that correspond to maize domestication are capable of generating the
+  interesting pattern of background selection that is observed. That
+  is, does demography alone explain the weak effect of background
+  selection on pairwise diversity in maize? We begain by estimating
+  the distribution of fitness effects and mutation rate in both maize and teosinte from
+  (formulas)... These were observed to be ZZZZZZ. Then, we simulated
+  mating according to our estimated demographic model and evaluated
+  the resulting levels of pairwise and singleton diversity. A LOT MORE
+GOES HERE, BUT WHAT IS WRITTEN NOW CAN FRAME THINGS.}
 

### 2015-05-15 manuscript.tex
updates

@@
 morphologically.   Additionally, because recent expansion is
-most evidenced by rare alleles, and because these data provide low
+most evidenced by rare alleles, and since these data provide low
 power to detect rare alleles, we expect that the estimate of  $N_m =
-366,973$, or $\sim 3N_a$, is likely an underestimate. \textcolor{red}{(This
-  is where evaluation using Kate's SFS should go)}.
+366,973$, or $\sim 3N_a$, is likely an underestimate.
+
+We utilized a complementary dataset of 4,808 maize landrace samples collected
+from across the Americas \textcolor{red}{(SeeDs reference here)} to
+obtain a less downward-biased estimate of the current maize effective
+population size. According to a singleton-based estimate of $\theta$
+\cite{fu1993}, which was chosen since rapid post-bottleneck expansion
+implies that rare alleles will most accurately reflect current maize
+demography, we obtained the estimate \textcolor{red}{$N_e =
+  1,011,686$}. Although less biased than the previously mentioned
+esitmate, we note that ascertainment bias of the
+genotyping platform used to generate these data again biases this
+estimate downward.
 
@@
 \includegraphics[width=.5\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
-\caption{This is what domestication looked like. WRITE CAPTION!}
+\caption{Parameters of domestication as estimated by dadi. Notably,
+  the maize effective population size ($N_e$) during the domestication
+  bottleneck appears to have consisted of approximately 5.26\%
+  the ancestral $N_e$, before recovering two at least 2.98
+  times as large as the ancestral $N_e$.}
 \label{bottleneck}

### 2015-05-18 manuscript.tex
minor whinings

@@
 \usepackage{pgf}
+\usepackage{hyperref}
+\newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }
 
@@
 \title{Demography and selection since maize domestication}
-\author{Timothy M. Beissinger\affil{1}{University of
-    California,Davis}, Li Wang\affil{2}{Iowa State University}, Arun
+\author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
+    California, Davis, CA, USA}, Li Wang\affil{2}{Iowa State University, Ames, IA, USA}, Arun
   Durvasula\affil{1}{}, Kate Crosby\affil{1}{}, Matthew Hufford\affil{2}{}, \and Jeffrey
-  Ross-Ibrarra\affil{1}{}\affil{3}{Genome Center, UC
-    Davis}\affil{4}{Center for population biology, UC Davis} }
+  Ross-Ibrarra\affil{1}{}\affil{3}{Genome Center and Center for population biology, University of
+    California, Davis, CA, USA} }
 
@@
 surrounding non-synonymous and synonymous substitutions in maize and
-teosinte since divergence from tripsicum, roughly \textcolor{red}{ZZZ} years before
+teosinte since divergence from \emph{Tripsacum}, roughly \textcolor{red}{ZZZ} years before
 present \textcolor{red}{(ref here)}. For both taxa, no difference in diversity
@@
 \caption{Pairwise diversity surrounding synonymous and nonsynonymous
-  substitutions in maize. Observe that the reduction of diversity
-  surrounding nonsynonymous subsitutions is no more severe than the
-  reduction of diversity surrounding synonymous substitutions. This
-  equivalency suggests that hard sweeps are not a primary driver of
-  maize diversity.}
+  substitutions in maize. \jri{needs larger legend/axis labels. x axis needs to say cM}}
 \label{hardSweeps}
@@
   and \textbf{B} depicts singleton diversity, influenced by rare
-  alleles and thus depicting evolutionary patterns in the recent past.}
+  alleles and thus depicting evolutionary patterns in the recent past.
+\jri{are these done with latest values we used for curve fitting?
+should also add the singleton vs pi comparison within species as a supplemental figure}}
 \label{purify}
@@
 \centering
-\includegraphics[width=.5\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
+\includegraphics[width=.4\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
 \caption{Parameters of domestication as estimated by dadi. Notably,
@@
 teosinte, using the same filters as employed for estimating the
-SFS. Next, SNPs and genotypes among maize, teosinte, and tripsicum were called. Tripsicum bam files
-were downloaded from (TRIPSICUM FILES), and then all SNPs with a
+SFS. Next, SNPs and genotypes among maize, teosinte, and \emph{Tripsacum} were called. \emph{Tripsacum}\emph{Tripsacum} bam files
+were downloaded from \url{}(TRIPSICUM FILES), and then all SNPs with a
 p-value less than 1e-6 were called using ANGSD. Quality filters were
@@
 genotypes, substitutions between maize and tripsicum, as well as
-between teosinte and tripsicum were identified using R \cite{R2014} as all positions with
+between teosinte and \emph{Tripsacum} were identified using R \cite{R2014} as all positions with
 no more than 20\% missing data for which every maize or teosinte
-allele differed from the observed tripsicum allele. At each class of
+allele differed from the observed \emph{Tripsacum} allele. At each class of
 substitution, effects were estimated using the ensembl variant effects
@@
 \begin{acknowledgments}
-Various thankyous will be in order.
+Various thankyous will be in order. \jri{Coop and his lab. funding from NSF Biology of Rare alleles -- need grant number.}
 \end{acknowledgments}

### 2015-05-18 manuscript.tex
JRI ideas added as comments
more ideas
updates

@@
 
-\dropcap{T}his is the beginning of the article. For now, I will give a
+\dropcap{D}omesticated plant species evolve in a unique fashion
+compared to their wild relatives \textcolor{red}{(ref here)}. This
+results from both the anthropomorphic nature of selection on
+domesticates as well as the characteristics of the domestication
+bottleneck that they usually experience \textcolor{red}{(ref here)}.
+
+\textcolor{red}{
+This is the beginning of the article. For now, I will give a
 bulleted list of the introduction:
 \begin{enumerate}
+\item Plant domestication
 \item Maize
-\item Plant domestication
 \item Maize domestication
@@
 \end{enumerate}
+}
 
@@
 
+%JRI IDEAS:
+% look at SFS of SNPs by GERP score in maize vs teosinte. shift to lower freq. should be seen at lower GERP scores in maize
+% discuss our estimates of deleterious mutation rate from B stuff.  these suggest only ~25% of mutations are deleterious, which contrasts results from a standard DFE
+% do a standard DFE! this gives us the proportion of mutations in categories of Ns. Can use Eyre-Walker software.
+% i think we do need to say something about load. add up GERP scores like Li?
+
 \subsection{Simulations}

### 2015-05-18 manuscript.tex
a bit of intro

@@
 \usepackage{pgf}
-\usepackage{hyperref}
+%\usepackage{hyperref}
 \newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }
@@
 compared to their wild relatives \textcolor{red}{(ref here)}. This
-results from both the anthropomorphic nature of selection on
-domesticates as well as the characteristics of the domestication
-bottleneck that they usually experience \textcolor{red}{(ref here)}.
+is a result of both the anthropomorphic nature of selection on
+domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
+bottleneck that they tend to have experienced
+\cite{ross2007}. However, the
+complex interplay between selective pressures and demographic limitations in
+domesticated species is not well understood. Despite the fact that,
+researchers have searched the genomes of several species for evidence
+of positive selection during domestication
+\textcolor{red}{(citations here)}, these types of studies tend to focus on
+identifying or mapping particular genes or regions that play an
+important role in phenotypic evolution. Our knowledge regarding the impacts of demography and
+selection have on whole-genome patterns of genetic variability therefore remain limited.
 
@@
 present \textcolor{red}{(ref here)}. For both taxa, no difference in diversity
-surrounding these classes of substitutions was identified (Figure \ref{hardSweeps}). Additionally, diversity around maize substitutions
-not seen in teosinte was investigated. These sites have the
-potential to correspond to recent maize sweeps, occuring after the
-split from teosinte. Again, no difference was observed
-between diversity around synonymous and nonsynonymous
-substitutions \textcolor{red}{(Supplemental figure here)}. Together, these observations suggest that hard sweeps
-are not a primary form of selection for either maize or teosinte.
+surrounding these classes of substitutions was identified (Figure
+\ref{hardSweeps}). Additionally, diversity around maize substitutions
+not seen in teosinte was investigated. These sites have the potential
+to correspond to recent maize sweeps, occuring after the split from
+teosinte. Again, no difference was observed between diversity around
+synonymous and nonsynonymous substitutions
+\textcolor{red}{(Supplemental figure here)}. Together, these
+observations suggest that hard sweeps are not a primary form of
+selection for either maize or teosinte.
 

### 2015-05-20 manuscript.tex
updates

@@
 compared to their wild relatives \textcolor{red}{(ref here)}. This
-is a result of both the anthropomorphic nature of selection on
+is a result of both the anthropomorphic nature of artificial selection on
 domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication

### 2015-05-21 manuscript.tex
updates

@@
 form of selection is operating, it can serve to reduce genetic
-variability at linked neutral sites, a phenomenon referred called
+variability at linked neutral sites, a phenomenon called
 background selection \cite{charlesworth1993}. Purifying and background selection lead
@@
 diversity within genes and recovery away from genes in both taxa. When standardized by neutral
-levels (which we defined as diversity at least \textcolor{red}{0.01} cM distal
-from genes), a stronger reduction of diversity
+levels, which were defined as mean diversity at postions at least 0.01 cM distal
+from genes, a stronger reduction of diversity
 and slower recovery was observed for teosinte than for maize, implying
@@
 teosinte genome (Figure \ref{purify}). This conflicted with our \emph{a priori} hypothesis;
-we expected that strong artificial selection since domestication would
-have caused enhanced purifying selection for maize. We therefore
-conducted the same analysis based on singleton diversity. As a class,
+we expected that strong artificial selection since domestication may
+have elevated the intensity of purifying selection for maize. We therefore
+conducted a parallel analysis based on singleton diversity. As a class,
 singleton alleles depict the most recent patterns of evolution, but
@@
 pairwise diversity, patterns of singleton
-diversity reflect recent patterns of evolution. When
-evaluating the data in this manner, an opposite relationship was
-observed. Maize singleton diversity was lower than teosinte singleton
-diversity near genes, and recovered more slowly
+diversity reflect recent patterns of evolution. Maize was downsampled
+to reflect the expected level of singletons from a sample size of 13,
+equivalent to that of teosinte, and for singletons our definition of
+neutral diversity was modified to be mean diversity at positions at
+least 0.02 cM distal from genes. When
+evaluating the data in this manner, a very different pattern
+emerged. Maize singleton diversity was just as low as teosinte singleton
+diversity near genes, but recovered more slowly
 (Figure \ref{purify}), implying that in the
-recent past maize has been more influenced by purifying selection than
+recent past maize has been at least as influenced by purifying selection as
 teosinte. Together, these findings imply that demographic history has
@@
 \centering
-\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_manuscript.png}
+\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_Downsampled_manuscript.png}
 \caption{Relative level of diversity versus distance to the nearest
@@
 teosinte, using the same filters as employed for estimating the
-SFS. Next, SNPs and genotypes among maize, teosinte, and \emph{Tripsacum} were called. \emph{Tripsacum}\emph{Tripsacum} bam files
+SFS. Next, SNPs and genotypes among maize, teosinte, and \emph{Tripsacum} were called. \emph{Tripsacum} bam files
 were downloaded from \url{}(TRIPSICUM FILES), and then all SNPs with a
@@
 
-\subsection{Evaluating diversity around genes and conserved sequences}
+\subsection{Evaluating diversity around genes}
 Two types of diversity surrounding genes were investigated. The first
@@
 singletons in each non-overlapping 1kb window was calculated with R
-\cite{R2014}. BiomaRt \cite{durinck2009, durinck2005} was then used to identify
-the center of each gene. Next, the distance from each diversity window
-to the nearest gene center was computed. Teosinte diversity is
+\cite{R2014}. For maize, we generated a parallel set of downsampled
+singleton data, for which binomial sampling within R was conducted
+based on allele frequencies at each site, to generate a singleton
+psuedo dataset of sample size 13, equivalent to our sample size for
+teosinte. BiomaRt \cite{durinck2009, durinck2005} was then used to identify
+the center of each gene. Next, the distance from each diversity window center
+to the nearest gene center was computed.
+
+Teosinte diversity is
 generally higher then maize diversity. Therefore, to enable comparisons between
 the reduction of diversity around genes in maize and teosinte, a
-neutral measure of pairwise and singleton diversity for each taxa was estimated according to
-mean pairwise and singleton diversity at windows greater than 0.01 cM from the nearest
-gene. Then, pairwise and singleton diversity at each window was
+neutral measure of pairwise and singleton diversity for each taxa was
+estimated. For pairwise diversity, this nuetral measure was defined according to
+mean pairwise diversity at windows greater than 0.01 cM from the nearest
+gene. For singletons, both maize and teosinte still showed reduced
+diversity at a distance of 1 cM, so neutral diversity in this case was
+defined as mean singleton diversity at positions 0.02 cM or greater
+from genes. Then, pairwise and singleton diversity at each window was
 standardized by dividing by the corresponding neutral
@@
 gene. Significant differences were assessed by taking 100 bootstrap
-samples and re-fitting the cubic smoothing spline to each. Then, the
+samples from each set of diversity windows and re-fitting the cubic smoothing spline to each. Then, the
 2.5\% and 97.5\% quantiles of values along the bootstrapped splines

### 2015-05-22 manuscript.tex
Singleton/pi figs updated

@@
 pairwise diversity, patterns of singleton
-diversity reflect recent patterns of evolution. Maize was downsampled
-to reflect the expected level of singletons from a sample size of 13,
-equivalent to that of teosinte, and for singletons our definition of
+diversity reflect recent patterns of evolution. Since our sample size
+for maize (n=23) was larger than teosite (n=13), maize singletons were
+analyzed directly as well as downsampled to reflect the sample size of
+teosinte. For singletons, our definition of
 neutral diversity was modified to be mean diversity at positions at
@@
 evaluating the data in this manner, a very different pattern
-emerged. Maize singleton diversity was just as low as teosinte singleton
+emerged. Maize singleton diversity was just at least as low as teosinte singleton
 diversity near genes, but recovered more slowly
@@
 recent past maize has been at least as influenced by purifying selection as
-teosinte. Together, these findings imply that demographic history has
+teosinte. Together, these findings suggest that demographic history has
 a strong influence on the effect of purifying selection. Historically,
@@
 \centering
-\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_Downsampled_manuscript.png}
+\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_Downsampled_threeLines_manuscript.png}
 \caption{Relative level of diversity versus distance to the nearest
@@
 singletons in each non-overlapping 1kb window was calculated with R
-\cite{R2014}. For maize, we generated a parallel set of downsampled
+for both maize and teosinte \cite{R2014}. For maize, we also generated a parallel set of downsampled
 singleton data, for which binomial sampling within R was conducted

### 2015-06-04 manuscript.tex
update

@@
 %\usepackage{hyperref}
+\usepackage[normalem]{ulem}
 \newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }
@@
 domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
-bottleneck that they tend to have experienced
+bottleneck(s) that they tend to have experienced
 \cite{ross2007}. However, the
-complex interplay between selective pressures and demographic limitations in
-domesticated species is not well understood. Despite the fact that,
-researchers have searched the genomes of several species for evidence
-of positive selection during domestication
-\textcolor{red}{(citations here)}, these types of studies tend to focus on
+complex interplay between selective pressures and demographic
+limitations, and the impact that this interplay has on identifying
+selection and understanding demography, is not fully understood. Although a large body of
+research that involves searching the genomes of domesticated species for evidence
+of positive selection exists \cite{hufford2012, he2011, OTHERS?}, these studies tend to focus on
 identifying or mapping particular genes or regions that play an
-important role in phenotypic evolution. Our knowledge regarding the impacts of demography and
-selection have on whole-genome patterns of genetic variability therefore remain limited.
+important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
+selection have on whole-genome patterns of genetic variability therefore remains limited.
 
-\textcolor{red}{
-This is the beginning of the article. For now, I will give a
-bulleted list of the introduction:
-\begin{enumerate}
-\item Plant domestication
-\item Maize
-\item Maize domestication
-\item Importance of demography
-\item Background selection
-\end{enumerate}
-}
+For instance, supposed neutrally-evolving DNA is often used to estimate historical demographic
+parameters of a population such as effective population size, structure, and expansion
+history \cite{luikart2003, gutenkunst2009}. However, researchers have
+called into question whether or not there are reasonable approaches
+for identifying neutral regions of the genome, since the effects of
+selection can be wide-ranging \cite{li2012, slotte2014}. For example, in
+\emph{Drosophila} it appears that the majority of the genome is
+impacted by the effects of selection through linkage
+\cite{sella2009}. This observation begs the question, how can
+demographic parameters be estimated independently of
+selective parameters? Human researchers have attempted to address this
+problem by limiting analyses to only sites far from genes
+\cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is
+difficult to be certain that sites even in gene-poor regions of the
+genome are not influenced by linked selection. Ultimately, an understanding
+of which sites have the potential to be adaptive, and how these affect
+genome-wide patterns of variability through linkage, is required for
+reasonable demographic inference.
+
+Maize represents an excellent organism to study these
+phenomena. Maize is a species of tremendous importance worldwide as
+both a staple crop \cite{SOMETHING} and as a model for
+understanding crop evolution \cite{SOMETHINGS}. Broadly speaking, archaeological and genetic studies have
+established that maize domestication is likely to have taken place in
+Central Mexico approximately 9,000 years bp
+\cite{SOMETHING,SOMETHING}. Teosinte, the most
+recent wild ancestor to maize, remains extant throughout much of the
+Americas \cite{SOMETHING}. Additionally, several large-effect
+domestication loci \cite{SOMETHINGS} and putative domestication
+regions \cite{hufford2012} have been identified. But despite all that is
+known about maize domestication, the parameters of the
+domestication process remain uncertain. Specifically, the size of the
+maize domestication bottleneck has not been estimated independently of
+the bottleneck's duration, nor are there sequence-based estimates of the effective
+population size of modern maize. Sequence information from maize and
+teosinte plants may therefore be utilized to address these questions.
+
+To that end, the objectives of our study were to 1) investigate the
+relative importance of different forms of selection on whole-genome
+variability in both maize and teosinte, as different forms of selection affect DNA
+variability remarkably differently; 2) research the impact that the
+domestication process has had on genetic variability in maize, and how
+this compares to the impact of a different demographic history in
+teosinte; and 3) precisely estimate the parameters of the
+maize domestication bottleneck. We show that our third objective, estimating the
+parameters of domestication, is not possible without first completing
+ objectives one and two, which demonstrate that as in humans \cite{gazave2014}, the
+majority of maize non-genic DNA may reasonably be treated as
+neutral. We achieve these objectives by utilizing
+whole-genome-sequence information from 23 maize and 13 teosinte lines
+sequenced as part of the Maize HapMap 2 panel \cite{chia2012}.
 

### 2015-06-05 manuscript.tex
update

@@
 
-\begin{figure}[b]
+\begin{figure*}[b]
 %\vspace*{.05in}
 \centering
-\includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_unNeutralized}
+\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_unNeutralized_June} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_unNeutralized_June}
 \caption{Pairwise diversity surrounding synonymous and nonsynonymous
-  substitutions in maize. \jri{needs larger legend/axis labels. x axis needs to say cM}}
+  substitutions in maize and teosinte. A: Maize diversity; B: Teosinte
+diversity.}
 \label{hardSweeps}
-\end{figure}
+\end{figure*}
 

### 2015-06-08 manuscript.tex
new figs

@@
 \centering
-\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_unNeutralized_June} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_unNeutralized_June}
+\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_June}
+\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_June}
 \caption{Pairwise diversity surrounding synonymous and nonsynonymous

### 2015-06-17 manuscript.tex
updated the methods--demographic inference

@@
 surrounding these classes of substitutions was identified (Figure
-\ref{hardSweeps}). Additionally, diversity around maize substitutions
+\ref{fig:hardSweeps}). Additionally, diversity around maize substitutions
 not seen in teosinte was investigated. These sites have the potential
@@
   substitutions in maize and teosinte. A: Maize diversity; B: Teosinte
-diversity.}
-\label{hardSweeps}
+diversity. \label{fig:hardSweeps}}
 \end{figure*}
@@
 that purifying selection has left a more pronounced signature in the
-teosinte genome (Figure \ref{purify}). This conflicted with our \emph{a priori} hypothesis;
+teosinte genome (Figure \ref{fig:purify}). This conflicted with our \emph{a priori} hypothesis;
 we expected that strong artificial selection since domestication may
@@
 diversity near genes, but recovered more slowly
-(Figure \ref{purify}), implying that in the
+(Figure \ref{fig:purify}), implying that in the
 recent past maize has been at least as influenced by purifying selection as
@@
   and \textbf{B} depicts singleton diversity, influenced by rare
-  alleles and thus depicting evolutionary patterns in the recent past.
-\jri{are these done with latest values we used for curve fitting?
-should also add the singleton vs pi comparison within species as a supplemental figure}}
-\label{purify}
+  alleles and thus depicting evolutionary patterns in the recent
+  past. \label{fig:purify} \jri{are these done with latest values we used for curve fitting? should also add the singleton vs pi comparison within species as a supplemental figure}}
 \end{figure*}
@@
 maize, and $M_{mt} = 1.72$ migrants per generation from maize to
-teosinte (Figure \ref{bottleneck}). We note that a population
+teosinte (Figure \ref{fig:bottleneck}). We note that a population
 split over 15 thousand generations before present precedes estimates
@@
 
-We utilized a complementary dataset of 4,808 maize landrace samples collected
+We utilized a complementary dataset of 4,021 maize landrace individuals collected
 from across the Americas \textcolor{red}{(SeeDs reference here)} to
@@
 demography, we obtained the estimate \textcolor{red}{$N_e =
-  1,011,686$}. Although less biased than the previously mentioned
+  992,713.5$}. Although less biased than the previously mentioned
 esitmate, we note that ascertainment bias of the
@@
   the ancestral $N_e$, before recovering two at least 2.98
-  times as large as the ancestral $N_e$.}
-\label{bottleneck}
+  times as large as the ancestral $N_e$. \label{fig:bottleneck} }
 \end{figure}
@@
 maize and teosinte from inbred lines, each inbred individual was treated as representing
-a single haplotype from its population. These were
+a single haplotype from its population. SFS were
 separately computed for genic and intergenic regions, as well as for
@@
 populations. Individuals were
-assumed to be fully inbred (-doSaf 2), and subsequently allele frequencies were
-divided by two to indicate haplotype frequencies. Quality filters were
+assumed to be fully inbred (-doSaf 2). Quality filters were
 employed such that reads with quality score below 30 and bases with
@@
 \subsection{Demographic inference}
+We used dadi \cite{gutenkunst2009} to estimate the parameters of maize
+domestication based on diffusion approximation of the 2-dimensional
+SFS for maize and teosinte. Our first step was to specify a model, as
+shown in (Figure \ref{fig:bottleneck}). In words, our model began with an ancestral teosinte
+population with effective population size $N_a$ individuals. The
+ancestral population split into
+two distinct populations, one of maize and the other of teosinte,
+$T_b$ generations in the past. The maize population immediately shrank
+to size $N_b$, and then grew exponentially to size $N_{maize}$
+today. From the split until today, the teosinte effective population
+size remained constant. Each year, $M_{mt}$ individuals migrated from
+maize to teosinte, and $M_{tm}$ individuals migrated from teosinte to
+maize.The free parameters of our model were $N_a, N_b, N_{maize}, T_b,
+M_{mt}, \text{and } M_{tm} $. Python code specifying this model is
+provided in \textcolor{red}{(SUPPLEMENT)}. We implimented 1,000 dadi
+optimizations of this model using the ``dadi.Inference.optimize\_log\_fmin''
+optimizer. In dadi optimizations,
+initial values for $N_b, N_{maize}, T_b,
+M_{mt}, \text{and } M_{tm}$, respectively, were randomly perturbed up
+to a factor of two from 0.02, 3, 0.04, 1e-5, and 1e-5. Lower
+bounds were set at 1e-7, 1e-7,  0, 1e-10, and
+1e-10, respecively, while upper bounds were respectively set at 2,
+200, 1, .001, and .001. $N_a$ did not have an initial value, upper, or
+lower bounds because it is estimated from equilibrium ancestral theta
+($\theta = 4N_a\mu$), which dadi determines automatically regardless
+of model. Population sizes were converted from dadi's
+units using a mutation rate estimate of 3e-8
+\textcolor{red}{(REFERENCE)}. Among the 1,000 dadi runs, parameters
+from that with the lowest log-likelihood were taken as estimates of
+the parameters of the domestication process.
+
+Recent population expansion, as is observed for maize, most influences
+the abundance of rare alleles. However, our demographic parameter
+estimation using dadi suffered from a relatively small sample size of
+23 maize individuals. Therefore, our approach likely underestimated
+current maize effective population size. To address this, we utilized
+a complimentary dataset of 4,021 maize landrace individuals collected
+from across the Americas \textcolor{red}{(SeeDs reference here)} to
+generate an independent estimate of
+$N_{maize}$. These individuals were genotyped using GBS
+\cite{elshire2011}, a protocol that suffers from a high rate of
+missing data \cite{beissinger2013}. To minimiaze biases imposed from
+missing data, we used R to isolate only those SNPs that
+were observed in at least 1,500 individuals and subsequently projected the
+SFS down to a sample of 500 individuals. Then, we utilized the singleton-based estimate of
+theta ($\theta = 4N_e\mu = \frac{S}{L}$) to estimate
+modern maize $N_e$ \cite{fu1993}, where $S$ is the total number of
+singletons and $L$ is the total length of DNA sequenced. Based on the
+GBS protocol implemented \cite{elshire2011}, $L$ is equal to the total
+number of SNPs in the dataset.
 

### 2015-07-01 manuscript.tex
significance

@@
 \usepackage[normalem]{ulem}
-\newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }
+\newcommand{\jri}[1]{\textcolor{red}{\emph{#1}} }
+\newcommand{\citex}{\textcolor{red}{\bf CITE}}
+\newcommand{\X}{\textcolor{red}{\bf X}}
 
@@
 
-\title{Demography and selection since maize domestication}
+\title{The impacts of demography on linked selection in maize}
 \author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
@@
 
-\significancetext{This work is insignificant ;-)}
+\significancetext{patterns of linked selection have been carefully evaluated in only a few species.
+how demography interacts with selection to impact genome-wide pattersn of diversity is still not well understood.
+also this work is significant (Beissinger b-test, p<0.01) }
 
@@
 \begin{abstract}
-This is the abstract. It should probably be somewhere around 200 words.
+This is the abstract. It should probably be longer than a tweet and shorter than 250 words.
 \end{abstract}

### 2015-07-01 manuscript.tex
comments

@@
 within genes to be 0.0088 and at least 5kb from genes it was 0.1153
-(p$\approx$0). These obserrvations suggest that genes are not evolving
+(p$\approx$0). \jri{that number is impossible, check again. off by maybe an order of magnitude!}
+These observations suggest that genes are not evolving
 neutrally in maize or teosinte. Instead, some form of selection is likely
@@
 surrounding these classes of substitutions was identified (Figure
-\ref{fig:hardSweeps}). Additionally, diversity around maize substitutions
-not seen in teosinte was investigated. These sites have the potential
+\ref{fig:hardSweeps}).
+\jri{weird formatting that the first figure referenced is fig.
+2.
+something about PNAS style and methods?}Additionally, diversity around maize substitutions not seen in teosinte was investigated.
+These sites have the potential
 to correspond to recent maize sweeps, occuring after the split from
@@
 observations suggest that hard sweeps are not a primary form of
-selection for either maize or teosinte.
+selection for either maize or teosinte. \jri{add Enard citation and GERP results (supp. figures)}
 
@@
 \begin{materials}
+
+\jri{either we need to provide all scripts in a single github repo with a readme, or provide links to individual scripts in the methods as we go along}
+
 \subsection{Plant materials}
 Accessions studied were selected from the Maize HapMap2
-panel \cite{chia2012} . Principal component analysis was employed to ensure that
-closely related individuals were not included due to their potential
-to bias results \textcolor{red}{(maybe a supplemental figure here)}. Ultimately, 23 maize inbreds derived from a diverse
-assortment of landraces were selected for inclusion. Thirteen teosinte
-inbred lines, all members of the subspecies Z. \emph{mays}
-ssp. \emph{parviglumis}, were utilized. Sequences were mapped to the
-maize B73 version 3 reference genome \cite{schnable2009}
-(ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/).
+panel \cite{chia2012}. \jri{aren't you using the high coverage teosinte? if so I think you need to cite Lemmon 2014 Plos Gen} 
+Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results \textcolor{red}{(maybe a supplemental figure here)}.
+\jri{yes supp.
+figure good idea}
+Ultimately, 23 maize inbreds derived from a diverse assortment of landraces were selected for inclusion. 
+Thirteen teosinte inbred lines, all members of  subspecies Z. \emph{mays}
+ssp. \emph{parviglumis}, were utilized. 
+Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} \url(ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/).
 
 \subsection{Interpolating genetic position}
-For many of the following analyses, physical position along a
-chromosome was a less relevant measure of map location than was genetic
-position. Therefore, physical positions were converted to genetic
-positions by interpolating from the NAM genetic map (REFERENCE), which
-provides a 1 cM resolution for physical to genetic conversion. Within
-R \cite{R2014}, physical positions with corresponding genetic positions in
-the NAM map were used as anchors. Physical positions in our dataset
-wihout corresponding genetic position were assigned genetic positions by scaling
-the anchored genetic positions according to the physical distance
-between the unlabeled position and the flanking anchors.
+For many of the following analyses, physical position along a chromosome was a less relevant measure of map location than was genetic position. 
+Therefore, physical positions were converted to genetic positions by interpolating from the NAM genetic map \citex \jri{Wallace 2014 no?}, which provides a 1 cM resolution for physical to genetic conversion.
+Within R \cite{R2014}, physical positions with corresponding genetic positions in the NAM map were used as anchors. 
+Physical positions in our dataset wihout corresponding genetic position were assigned genetic positions by scaling the anchored genetic positions according to the physical distance between the unlabeled position and the flanking anchors.
 
 \subsection{Estimating the site frequency spectrum}
-To estimate individual  and joint site frequency spectra (SFS) for
-maize and teosinte from inbred lines, each inbred individual was treated as representing
-a single haplotype from its population. SFS were
+To estimate individual  and joint site frequency spectra (SFS) for maize and teosinte from inbred lines, each inbred individual was treated as representing a single haplotype from its population. 
+SFS were
 separately computed for genic and intergenic regions, as well as for

### 2015-07-02 manuscript.tex
edits/comments

@@
 
-\subsection{Patterns of purifying and background selection are
+{\bf Outline}
+\begin{enumerate}
+\item Diversity differs between genes, not genes
+\item No hard sweeps
+\item soft sweeps? (putting it here means if we find a lot we may have to redo stuff below on the subset of genes w/o evidence of soft sweeps.)
+\item purifying selection differs ($\pi$)
+\item Estimating the bneck
+\item purifying selection (singletons)
+\item fit the B curve
+\item simulations
+\end{enumerate}
+
+\subsection{Patterns of purifying and background selection \jri{we will need to define background selection and be careful about ``purifying selection'' vs ``background selection'', as technically background selection does not include very weak deleterious alleles.}are
   influenced by demographic history}
-Purifying selection refers to the situation where deleterious
-mutations arising in a population are continuously selected against. When this
-form of selection is operating, it can serve to reduce genetic
-variability at linked neutral sites, a phenomenon called
-background selection \cite{charlesworth1993}. Purifying and background selection lead
-to lower diversity within genes and other functional sites relative
-to neutral regions \textcolor{red}{(ref here)}. We investigated purifying
-selection in maize and teosinte by evaluating the average magnitude of reduced
-diversity within genes and recovery away from genes in both taxa. When standardized by neutral
-levels, which were defined as mean diversity at postions at least 0.01 cM distal
-from genes, a stronger reduction of diversity
-and slower recovery was observed for teosinte than for maize, implying
-that purifying selection has left a more pronounced signature in the
-teosinte genome (Figure \ref{fig:purify}). This conflicted with our \emph{a priori} hypothesis;
-we expected that strong artificial selection since domestication may
-have elevated the intensity of purifying selection for maize. We therefore
-conducted a parallel analysis based on singleton diversity. As a class,
-singleton alleles depict the most recent patterns of evolution, but
+Purifying selection refers to the situation where deleterious mutations arising in a population are continuously selected against. 
+When this form of selection is operating, it can serve to reduce genetic variability at linked neutral sites, a phenomenon called background selection \cite{charlesworth1993}.
+Purifying and background selection lead to lower diversity within genes and other functional sites relative to neutral regions \textcolor{red}{(ref here)}.
+We investigated purifying selection in maize and teosinte by evaluating the average magnitude of reduced diversity within genes and recovery away from genes in both taxa. 
+When standardized by neutral levels, which were defined as mean diversity at postions at least 0.01 cM distal
+from genes, a stronger reduction of diversity and slower recovery was observed for teosinte than for maize, implying that purifying selection has left a more pronounced signature in the teosinte genome (Figure \ref{fig:purify}).
+This conflicted with our \emph{a priori} hypothesis;
+we expected that strong artificial selection since domestication may \jri{don't think we should say this; this wasn't my a priori prediction}
+have elevated the intensity of purifying selection for maize. 
+We therefore conducted a parallel analysis based on singleton diversity. \jri{disagree singletons should go here. see outline above.}
+As a class, singleton alleles depict the most recent patterns of evolution, but
 also have the lowest effect on pairwise diversity. Therefore, unlike
@@
   alleles and thus depicting evolutionary patterns in the recent
-  past. \label{fig:purify} \jri{are these done with latest values we used for curve fitting? should also add the singleton vs pi comparison within species as a supplemental figure}}
+  past.
+\label{fig:purify} \jri{are these done with latest values we used for curve fitting?
+should also add the singleton vs pi comparison within species as a supplemental figure.
+if $\pi$/singleton curves {\bf don't} match, will need to explain this suggests teosinte not in equilibrium.
+good justification i think for msmc model, which we then use to say ``gee whiz, not perfect equilibrium but who cares because the big interesting demography is in maize.'' }
 \end{figure*}
@@
 selection and indicates that genic regions are not evolving
-neutrally. Therefore, for demographic modeling we restricted analysis
-to non-genic sites.
+neutrally. \jri{need to cite Hufford 2012 here as we made this observation in that paper} Therefore, for demographic modeling we restricted analysis
+to non-genic sites. \jri{need to clarify 5kb from genes}
 
@@
 approximately 9,000 years before present \textcolor{red}{(reference
-  here)}. This could result from multiple generations per year, or in
+  here)}. This could result from multiple generations per year \jri{this seems unlikely, i'd probably drop multiple generations as an explanation}, or in
 may reflect teosinte population structure that was present before
@@
 Text goes here
+\jri{include back of envelope calculatio based on land area!}
 

### 2015-07-20 manuscript.tex
update

@@
 \usepackage[normalem]{ulem}
-\newcommand{\jri}[1]{\textcolor{red}{\emph{#1}} }
-\newcommand{\citex}{\textcolor{red}{\bf CITE}}
-\newcommand{\X}{\textcolor{red}{\bf X}}
+\newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }
+\newcommand{\citex}{\textcolor{blue}{\bf CITE}}
+\newcommand{\X}{\textcolor{blue}{\bf X}}
 
@@
 \dropcap{D}omesticated plant species evolve in a unique fashion
-compared to their wild relatives \textcolor{red}{(ref here)}. This
+compared to their wild counterparts \textcolor{red}{(ref here)}. This
 is a result of both the anthropomorphic nature of artificial selection on
@@
 \section{Results}
-\subsection{Patterns of variability differ between genic and
+\subsection{Patterns of diversity differ between genic and
   nongenic regions of the genome}
@@
 teosinte, but to a larger extent. For teosinte we observed $\pi$
-within genes to be 0.0088 and at least 5kb from genes it was 0.1153
-(p$\approx$0). \jri{that number is impossible, check again. off by maybe an order of magnitude!}
-These observations suggest that genes are not evolving
+within genes to be 0.0088 and at least 5kb from genes it was 0.0115
+(p$\approx$0). These observations suggest that genes are not evolving
 neutrally in maize or teosinte. Instead, some form of selection is likely
@@
 surrounding non-synonymous and synonymous substitutions in maize and
-teosinte since divergence from \emph{Tripsacum}, roughly \textcolor{red}{ZZZ} years before
-present \textcolor{red}{(ref here)}. For both taxa, no difference in diversity
+teosinte since divergence from \emph{Tripsacum}, roughly 1-1.2 million years before
+present \cite{ross2009}. For both taxa, no difference in diversity
 surrounding these classes of substitutions was identified (Figure
-\ref{fig:hardSweeps}).
-\jri{weird formatting that the first figure referenced is fig.
-2.
-something about PNAS style and methods?}Additionally, diversity around maize substitutions not seen in teosinte was investigated.
-These sites have the potential
-to correspond to recent maize sweeps, occuring after the split from
-teosinte. Again, no difference was observed between diversity around
-synonymous and nonsynonymous substitutions
-\textcolor{red}{(Supplemental figure here)}. Together, these
-observations suggest that hard sweeps are not a primary form of
-selection for either maize or teosinte. \jri{add Enard citation and GERP results (supp. figures)}
+\ref{fig:hardSweeps}). This observation suggests that hard sweeps are not a primary form of
+selection for either maize or teosinte.
+
+To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene tga1, one of the few known instances of a hard sweep corresponding to a causal nonsynonymous mutuation in maize \cite{wang2015}.
+In the case of this known sweep, we observed a reduction in diversity surrounding the causitive substitution to levels much lower than seen for synonymous substitutions \textcolor{red}{(SUPP FIG)}, confirming that this method should be informative if hard sweeps were the prevaling mode of selection in maize.
+Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize.
+When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites \textcolor{red}{(SUPP FIG)}.
+
+\subsection{Soft Sweeps?}
+In this section, I'll write about what we did to look for soft sweeps...
 
@@
 
-{\bf Outline}
-\begin{enumerate}
-\item Diversity differs between genes, not genes
-\item No hard sweeps
-\item soft sweeps? (putting it here means if we find a lot we may have to redo stuff below on the subset of genes w/o evidence of soft sweeps.)
-\item purifying selection differs ($\pi$)
-\item Estimating the bneck
-\item purifying selection (singletons)
-\item fit the B curve
-\item simulations
-\end{enumerate}
+%%%%
+%%%% This outline no longer needed in-text because results headers follow
+%%%%
+% \jri{
+% {\bf Outline}
+% \begin{enumerate}
+% \item Diversity differs between genes, not genes
+% \item No hard sweeps
+% \item soft sweeps? (putting it here means if we find a lot we may have to redo stuff below on the subset of genes w/o evidence of soft sweeps.)
+% \item purifying selection differs ($\pi$)
+% \item Estimating the bneck
+% \item purifying selection (singletons)
+% \item fit the B curve
+% \item simulations
+% \end{enumerate}
+% }
 
-\subsection{Patterns of purifying and background selection \jri{we will need to define background selection and be careful about ``purifying selection'' vs ``background selection'', as technically background selection does not include very weak deleterious alleles.}are
+\subsection{Patterns of purifying and background selection are
   influenced by demographic history}
+ \jri{we will need to define background selection and be careful about ``purifying selection'' vs ``background selection'', as technically background selection does not include very weak deleterious alleles.}
 Purifying selection refers to the situation where deleterious mutations arising in a population are continuously selected against.
@@
 if $\pi$/singleton curves {\bf don't} match, will need to explain this suggests teosinte not in equilibrium.
-good justification i think for msmc model, which we then use to say ``gee whiz, not perfect equilibrium but who cares because the big interesting demography is in maize.'' }
+good justification i think for msmc model, which we then use to say ``gee whiz, not perfect equilibrium but who cares because the big interesting demography is in maize.'' }}
 \end{figure*}
-
-\textcolor{red}{This might be a good place to stick in a paragraph
-  about curve fitting B to estimate s, mu.}
 
@@
 
+\section{purifying selection (singletons)}
+If this is separated out from the previous purifying selection section
+and placed after the demography section, it should go here.
+
 \subsection{Simulations confirm the relationship between demographic
   processes and background selection}
+\textcolor{red}{This might be a good place to stick in a paragraph
+  about curve fitting B to estimate s, mu.}
+
 \textcolor{red}{We explored whether or not the demographic patterns
@@
 Accessions studied were selected from the Maize HapMap2
-panel \cite{chia2012}. \jri{aren't you using the high coverage teosinte? if so I think you need to cite Lemmon 2014 Plos Gen} 
+panel \cite{chia2012, lemmon2014}.
 Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results \textcolor{red}{(maybe a supplemental figure here)}.
@@
 
-% \begin{thebibliography}{10}
-% \bibitem{dadi}
-% R.~Gutenkunst, R.~Hernandez,S.~Williamson, and C.~Bustamante, {\em
-%   Inferring the joint demographic history of multiple populations from
-%   multidimensional SNP frequency data}, PLoS Genetics., 5:10 (2009), e1000695.
-
-% \bibitem{hapmap2}
-% J.~Chia, C.~Song, P.~Bradbury, D.~Costich, N.~de~Leon, and others,
-% \emph{Maize HapMap2 identifies extant variation from a genome in
-%   flux}, Nature Genetics., 44 (2012), 803-807.
-
-% \bibitem{maizeGenome}
-% P.~Schnable, D.~Ware, R.~Fulton, J.~Stein, F.~Wei, \emph{The B73 maize
-% genome:complexity, diversity, and dynamics}, Science., 326:5956
-% (2009), 1112-1115.
-
-% \bibitem{angsd}
-% T.~Korneliussen, A.~Albrechtsen, and R.~Nielsen, \emph{ANGSD: Analysis of
-% next generation sequencing data}, BMC Bioinformatics., 15:356 (2014),
-% 10.1186/s12859-014-0356-4.
-
-% \bibitem{R}
-% R Core Team, \emph{R: A language and environment for statistical
-%   computing}, R Foundation for Statistical Computing., Vienna,
-% Austria.
-
-% \bibitem{biomaRt1}
-% S.~Durinck, P.T.~Soekknabm E.~Birney, and W.~Huber, \emph{Mapping
-%   identifiers for the integration of genomic datasets with the
-%   R/Bioconductor package biomaRt}, Nature Protocols., 4 (2009),
-% 1184-1191.
-
-% \bibitem{biomaRt2}
-% S.~Durinck, Y.~Moreau, A.~Kasprzyk, S.~Davis, B.~De~Moor, and others,
-% \emph{BioMart and Bioconductor: a powerful link between biological
-%   databases and microarray data analysis}, Bioinformatics, 21 (2005),
-% 3439-3440.
-
-% \bibitem{vep}
-% W.~McLaren, B.~Pritchard, D.~Rios, Y.~Chen, P.~Flicek, and
-% F.~Cunningham, \emph{Deriving the consequences of genomic variants
-%   with the Ensembl API and SNP Effect Predictor}, Bioinformatics,
-% 26:16 (2010), 2069-2070.
-
-% \bibitem{hernandez11}
-% R.D.~Hernandez, J.L.~Kelley, E.~Elyashiv, S.C.~Melton, A.~Auton, and
-% others, \emph{Classic selective sweeps were rare in recent human
-%   evolution}, Science, 331:6019 (2011), 920-924.
-
-% \bibitem{sattath11}
-% S.~Sattath, E.~Elyashiv, O.~Kolodny, Y.~Rinott, and G.~Sella,
-% \emph{Pervasive adaptative protein evolution apparent in diversity
-%   patterns around amino acid substitutions in Drosophila simulans},
-% PLoS Genetics, 7:2 (2011), e1001302.
-
-% \bibitem{charlesworth93}
-% B.~Charlesworth, M.~T.~Morgan, and D.~Charlesworth, \emph{The Effect
-%   of Deleterious Mutations on Neutral Molecular Variation}, Genetics,
-% 134 (1993), 1289-1303.
-
-% \end{thebibliography}
-
 \bibliography{Reference}

### 2015-07-22 manuscript.tex
updates

@@
 
-\subsection{Patterns of purifying and background selection are
-  influenced by demographic history}
- \jri{we will need to define background selection and be careful about ``purifying selection'' vs ``background selection'', as technically background selection does not include very weak deleterious alleles.}
+\subsection{Pairwise diversity is heavily influenced by purifying selection}
+% \jri{we will need to define background selection and be careful about ``purifying selection'' vs ``background selection'', as technically background selection does not include very weak deleterious alleles.}
 Purifying selection refers to the situation where deleterious mutations arising in a population are continuously selected against.
@@
 We investigated purifying selection in maize and teosinte by evaluating the average magnitude of reduced diversity within genes and recovery away from genes in both taxa.
-When standardized by neutral levels, which were defined as mean diversity at postions at least 0.01 cM distal
-from genes, a stronger reduction of diversity and slower recovery was observed for teosinte than for maize, implying that purifying selection has left a more pronounced signature in the teosinte genome (Figure \ref{fig:purify}).
-This conflicted with our \emph{a priori} hypothesis;
-we expected that strong artificial selection since domestication may \jri{don't think we should say this; this wasn't my a priori prediction}
-have elevated the intensity of purifying selection for maize.
-We therefore conducted a parallel analysis based on singleton diversity. \jri{disagree singletons should go here. see outline above.}
-As a class, singleton alleles depict the most recent patterns of evolution, but
-also have the lowest effect on pairwise diversity. Therefore, unlike
-pairwise diversity, patterns of singleton
-diversity reflect recent patterns of evolution. Since our sample size
-for maize (n=23) was larger than teosite (n=13), maize singletons were
-analyzed directly as well as downsampled to reflect the sample size of
-teosinte. For singletons, our definition of
-neutral diversity was modified to be mean diversity at positions at
-least 0.02 cM distal from genes. When
-evaluating the data in this manner, a very different pattern
-emerged. Maize singleton diversity was just at least as low as teosinte singleton
-diversity near genes, but recovered more slowly
-(Figure \ref{fig:purify}), implying that in the
-recent past maize has been at least as influenced by purifying selection as
-teosinte. Together, these findings suggest that demographic history has
-a strong influence on the effect of purifying selection. Historically,
-teosinte has had a larger population size than maize, and only
-recently has maize population size overcome that of teosinte. Since
-the efficacy of purifying selection scales with population size,
-these results likely reflect changes in $N_e$ more than they reflect
-underlying changes in selection pressure.
+To enable comparisons between maize and teosinte, we standardized by neutral levels within each taxa, where neutral levels were defined as mean diversity at positions far from genes at least 0.01 cM distal from genes).
+For both maize and teosinte, pairwise diversity was reduced within genes and a recovery of diversity away from genes was observed.
+Interestingly, a stronger reduction of diversity and slower recovery was observed for teosinte than for maize, implying that purifying selection has left a more pronounced signature on pairwise diversity in the teosinte genome (Figure \ref{fig:purify}).
 
@@
 To explore whether differences in purifying selection between maize
-and teosinte can be explained purely by demographic processes, we
+and teosinte can be explained by demographic processes, we
 first estimated the parameters of maize domestication using large-scale
@@
 selection and indicates that genic regions are not evolving
-neutrally. \jri{need to cite Hufford 2012 here as we made this observation in that paper} Therefore, for demographic modeling we restricted analysis
-to non-genic sites. \jri{need to clarify 5kb from genes}
+neutrally. This observation is consistent with previous research \cite{hufford2012}. Therefore, for demographic modeling we restricted analysis
+to sites at least 5kb from gene start/stop postions.
 
@@
 approximately 9,000 years before present \textcolor{red}{(reference
-  here)}. This could result from multiple generations per year \jri{this seems unlikely, i'd probably drop multiple generations as an explanation}, or in
-may reflect teosinte population structure that was present before
+  here)}. This could reflect teosinte population structure that was present before
 domestication. Also, note that the genetic time of the population
@@
   992,713.5$}. Although less biased than the previously mentioned
-esitmate, we note that ascertainment bias of the
+estimate, we note that ascertainment bias of the
 genotyping platform used to generate these data again biases this
@@
 
-\section{purifying selection (singletons)}
-If this is separated out from the previous purifying selection section
-and placed after the demography section, it should go here.
+\section{Post-domestiction expansion is apparent in patterns of singleton diversity}
+Motivated by the rapid post-domestication expansion in maize that is apparent from our demographic analysis, we conducted an analysis of genome-wide patterns of singleton diversity.
+Singleton diversity measures the abundance of alleles that are only observed in one individual among the sample.
+As a class, singleton alleles depict the most recent patterns of evolution, but
+they also have the lowest effect on pairwise diversity. Therefore, unlike
+pairwise diversity, patterns of singleton
+diversity reflect recent patterns of evolution. Since our sample size
+for maize (n=23) was larger than teosite (n=13), maize singletons were
+analyzed directly as well as downsampled to reflect the sample size of
+teosinte. For singletons, our definition of
+neutral diversity was modified to be mean diversity at positions at
+least 0.02 cM distal from genes. When
+evaluating the data in this manner, a very different pattern
+emerged compared to what we saw with respect to pairwise diversity. Maize singleton diversity was at least as reduced low as teosinte singleton
+diversity near genes, but recovered more slowly
+(Figure \ref{fig:purify}), implying that in the
+recent past maize has been more influenced by purifying selection than has
+teosinte.
 
-\subsection{Simulations confirm the relationship between demographic
-  processes and background selection}
-\textcolor{red}{This might be a good place to stick in a paragraph
-  about curve fitting B to estimate s, mu.}
+Together, these findings suggest that demographic history has
+a strong influence on the effect of purifying selection. Historically,
+teosinte has had a larger population size than maize, and only
+recently has maize population size overcome that of teosinte. Since
+the efficacy of purifying selection scales with population size,
+these results likely reflect changes in $N_e$ more than they reflect
+underlying changes in selection pressure.
 
-\textcolor{red}{We explored whether or not the demographic patterns
-  that correspond to maize domestication are capable of generating the
-  interesting pattern of background selection that is observed. That
-  is, does demography alone explain the weak effect of background
-  selection on pairwise diversity in maize? We begain by estimating
-  the distribution of fitness effects and mutation rate in both maize and teosinte from
-  (formulas)... These were observed to be ZZZZZZ. Then, we simulated
-  mating according to our estimated demographic model and evaluated
-  the resulting levels of pairwise and singleton diversity. A LOT MORE
-GOES HERE, BUT WHAT IS WRITTEN NOW CAN FRAME THINGS.}
+\subsection{Purifying seleciton results are robust to soft-sweeps}
+As a category, instances of soft-sweeps, or cases of selection on alleles that have already reached intermediate frequency before becoming advantageous, are difficult to identify \textcolor{red}{(refs here)}.
+However, this form of selection is likely to be frequent during crop domestication \textcolor{red}{(refs here)}.
+We were concerned that the disparate patterns of pairwise and singleton diversity in maize and teosinte, which seem to reflect differences in demographic history, may instead reflect differences in the location and abundance of soft-sweeps.
+Therefore, we re-analyzed patterns of pairwise and singleton diversity in maize and teosinte including only a subset of sites that are least likely to correspond to soft-sweeps.
+To achieve this, we performed a genome-wide scan for soft or hard sweeps in maize or teosinte based on the H12 statistic \cite{garud2015}.
+H12 is sensitive to both hard and soft sweeps.
+Sites nearest to genes that displayed among the top 20\% of H12 values in either maize or teosinte were removed from our dataset, and both pairwise and singleton diversity at remaining sites, as a function of the distance to the nearest gene, was calculated \textcolor{red}{(Supp Fig)}.
+This subset of the data has the potentially confounding effects of soft-sweeps removed, and again we see nearly identical patterns of pairwise and singleton diversity surrounding genes.
+This provides further evidence that the patterns of diversity we observe in maize as compared to teosinte are the result of rapid population expansion as opposed to differing selective patterns in the domesticated and wild taxa.
 

### 2015-07-28 manuscript.tex
updates

@@
 
-\significancetext{patterns of linked selection have been carefully evaluated in only a few species.
-how demography interacts with selection to impact genome-wide pattersn of diversity is still not well understood.
-also this work is significant (Beissinger b-test, p<0.01) }
+\significancetext{Patterns of linked selection, or the impact of selection on sites that neighbor the functional variant, have been carefully evaluated in only a few species.
+In this work, we demonstrate that selection against deleterious mutations leaves a pronounced signature on the maize genome, reducing diversity in and immediately around genes.
+We show how demography interacts with selection to impact genome-wide patterns of diversity, including the important observation that rapid population expansion can increase the efficiency of selection as much as a sudden population bottleneck can weaken it.
+Along the way, we develop the first estimate the domegraphic parameters of the maize domestication from whole genome sequence data.}
 
@@
 \begin{abstract}
-This is the abstract. It should probably be longer than a tweet and shorter than 250 words.
+  Unique selective and demographic forces operate on plants during domestication.
+These forces interact to generate the patterns of DNA variability that are persistent in crop species today.
+To quantify this interplay, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide.
+We utilized whole genome sequence data from 23 maize and 13 teosinte individuals to make inferences.
+We obtained a complete estimate of the population size fluctuation and other demographic parameters experienced by maize as it was domesticated from teosinte.
+Here, we show that maize experienced a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid population size expansion post-domestication.
+We observe that hard sweeps, specifically positive selection on new genic mutations, are not the primary force driving maize and teosinte apart.
+Also, we find that reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize.
+However, expansion after domestication has since increased the efficiency of purifying selection to levels superior to those seen in teosinte.
+Our results demonstrate that in domesticated species, demography and selection are constantly fluctuating through time.
 \end{abstract}
@@
 Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize.
-When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites \textcolor{red}{(SUPP FIG)}.
-
-\subsection{Soft Sweeps?}
-In this section, I'll write about what we did to look for soft sweeps...
+When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score, putatively corresponding to genes undergoing the strongest and weakest purifying selection \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites \textcolor{red}{(SUPP FIG)}.
 
@@
 \caption{Pairwise diversity surrounding synonymous and nonsynonymous
-  substitutions in maize and teosinte. A: Maize diversity; B: Teosinte
-diversity. \label{fig:hardSweeps}}
+  substitutions in maize and teosinte. {\bf A:} Maize diversity; {\bf B:} Teosinte
+diversity. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis. \label{fig:hardSweeps}}
 \end{figure*}
@@
 Purifying selection refers to the situation where deleterious mutations arising in a population are continuously selected against.
-When this form of selection is operating, it can serve to reduce genetic variability at linked neutral sites, a phenomenon called background selection \cite{charlesworth1993}.
+When this form of selection is operating, it can serve to reduce genetic variability at linked neutral sites, a phenomenon often called background selection \cite{charlesworth1993}.
 Purifying and background selection lead to lower diversity within genes and other functional sites relative to neutral regions \textcolor{red}{(ref here)}.
 We investigated purifying selection in maize and teosinte by evaluating the average magnitude of reduced diversity within genes and recovery away from genes in both taxa.
-To enable comparisons between maize and teosinte, we standardized by neutral levels within each taxa, where neutral levels were defined as mean diversity at positions far from genes at least 0.01 cM distal from genes).
+To enable comparisons between maize and teosinte, we standardized by neutral levels within each taxa, where neutral levels were defined as mean diversity at positions far from genes at least 0.01 cM distal from genes.
 For both maize and teosinte, pairwise diversity was reduced within genes and a recovery of diversity away from genes was observed.
@@
 
-\begin{figure*}[b]
+\begin{figure*}[t]
 %\vspace*{.05in}
@@
   past.
-\label{fig:purify} \jri{are these done with latest values we used for curve fitting?
-should also add the singleton vs pi comparison within species as a supplemental figure.
+Bootstrap-based 95\% confidence intervals are depicted via shading.
+Inset plots depict a smaller range on the x-axis.
+\label{fig:purify} \jri{should also add the singleton vs pi comparison within species as a supplemental figure.
 if $\pi$/singleton curves {\bf don't} match, will need to explain this suggests teosinte not in equilibrium.
@@
 population size. According to a singleton-based estimate of $\theta$
-\cite{fu1993}, which was chosen since rapid post-bottleneck expansion
+\cite{fu1993}, which was utilized since rapid post-bottleneck expansion
 implies that rare alleles will most accurately reflect current maize
@@
 \section{Post-domestiction expansion is apparent in patterns of singleton diversity}
-Motivated by the rapid post-domestication expansion in maize that is apparent from our demographic analysis, we conducted an analysis of genome-wide patterns of singleton diversity.
+Motivated by the rapid post-domestication expansion in maize is seen in our demographic analysis, we conducted an analysis of genome-wide patterns of singleton diversity.
 Singleton diversity measures the abundance of alleles that are only observed in one individual among the sample.
@@
 evaluating the data in this manner, a very different pattern
-emerged compared to what we saw with respect to pairwise diversity. Maize singleton diversity was at least as reduced low as teosinte singleton
+emerged compared to what we saw with respect to pairwise diversity. Maize singleton diversity was at least as reduced as teosinte singleton
 diversity near genes, but recovered more slowly
@@
 
+Similarly, to complement our investigation of pairwise diversity surrounding substitutions, we studied patterns of singleton diversity surrounding synonymous and nonsynonymous substitutions in maize \textcolor{red}{(Supplemental Fig)}.
+A nearly identical pattern to that shown in Figure \ref{fig:hardSweeps} was observed.
+This further demonstrates that even the most recent selection patterns in maize are not dominated by hard sweeps.
+
 \subsection{Purifying seleciton results are robust to soft-sweeps}
@@
 However, this form of selection is likely to be frequent during crop domestication \textcolor{red}{(refs here)}.
-We were concerned that the disparate patterns of pairwise and singleton diversity in maize and teosinte, which seem to reflect differences in demographic history, may instead reflect differences in the location and abundance of soft-sweeps.
+We were concerned that the disparate patterns of pairwise and singleton diversity in maize and teosinte, which seem to reflect differences in the efficiency of purifying selection due to demographic history, may instead reflect differences in the location and abundance of soft-sweeps.
 Therefore, we re-analyzed patterns of pairwise and singleton diversity in maize and teosinte including only a subset of sites that are least likely to correspond to soft-sweeps.
@@
 Sites nearest to genes that displayed among the top 20\% of H12 values in either maize or teosinte were removed from our dataset, and both pairwise and singleton diversity at remaining sites, as a function of the distance to the nearest gene, was calculated \textcolor{red}{(Supp Fig)}.
-This subset of the data has the potentially confounding effects of soft-sweeps removed, and again we see nearly identical patterns of pairwise and singleton diversity surrounding genes.
+This subset of the data in no longer subject to the potentially confounding effects of soft-sweeps, and nearly identical patterns of pairwise and singleton diversity surrounding genes are observed.
 This provides further evidence that the patterns of diversity we observe in maize as compared to teosinte are the result of rapid population expansion as opposed to differing selective patterns in the domesticated and wild taxa.
@@
 \section{Discussion}
-\subsection{Discussion 1}
-Text goes here
-\jri{include back of envelope calculatio based on land area!}
+\subsection{Little positive selection on new genic mutations}
+Our findings indicate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize.
+More specifically, we observe that positive selection on new genic mutations in maize is not common compared to other drivers of diversity.
+This finding does not exclude the possibility that hard sweeps have taken place at non-genic sites, such as in those in enhancer regions.
+One known example of positive selection on a non-genic mutation involves the \emph{tb1} locus of maize, which is one of the best characterized examples of positive selection on a ``domestication gene'' in any crop \cite{clark2006}.
+However, the maize \emph{tb1} allele was already present in teosinte before domestication \cite{studer2011}, so this even this example demonstrates supports our finding that hard sweeps are rare.
+The \emph{gt1} locus is another well-characterized case of positive selection operating on standing variation at an enhancer region \cite{wills2013}.
+Instances of selection on standing variation such as these, often called soft sweeps, may be a major contributor to maize patterns of diversity \cite{beissinger2014}, and this could be part of the explanation as to why hard sweeps appear to be so rare, despite obvious morphological differences between maize and teosinte.
 
-\subsection{Discussion 2}
-Text goes here
+Unfortunately, our ability to accurately identify soft sweeps, and particularly to distinguish them from hard sweeps, remains limited \cite{innan2004,messer2013}.
+However, we implemented a scan based on the H12 statistic, which is designed to identify both hard and soft sweeps with reasonable power \cite{garud2015}.
+Although the goal of this study was not to identify specific sweeps, it may be that the outlier sites distinguished by H12 are primarily composed of soft sweeps instead of hard.
+Similarly, previous studies scanning for evidence of positive selection in maize \cite{hufford2012} may have picked up primarily evidence of soft sweeps.
+In domesticated species, artificial selection beginning at the onset of domestication represents a drastic shift in selective pressure.
+One can easily imagine a scenario in which previously neutral or slightly deleterious variants that were segregating in the population when this shift took place suddenly became beneficial, leading to soft sweeps.
+As appealing as this explanation is, however, our data do not speak to the abundance of soft sweeps, beyond demonstrating that they are one possibility for the lack of observed hard sweeps.
+
+We should additionally note that our observations do not exclude the possibility of infrequent hard sweeps having taken place during maize evolution.
+In fact, there is one maize locus, \emph{tga1}, that has been shown to correspond to a hard sweep at an amino-acid changing mutation \cite{wang2015}.
+Surrounding \emph{tga1},  our data demonstrate a pattern consistent with a hard sweep.
+But, our data do demonstrate that instances such as this are infrequent.
+This contrasts sharply with \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, where differences in diversity surrounding synonymous and nonsynonymous substitutions are clear.
+However, it agrees with what has been seen for humans \cite{hernandez2011}.
+
+\subsection{Demography of domestication}
+Since we observed that classical hard sweeps do not explain patterns of maize diversity, it is perhaps unsurprising that we instead found evidence of demographic history contributing substantially to genetic diversity.
+Our estimate of the demographic history of maize, which is the first estimate of maize demographic history to utilize whole-genome-sequence (WGS) data, allowed us to disentangle the confounding parameters of bottleneck strength and bottleneck intensity.
+This confounding limited previous domestication estimates \cite{wright2005}.
+By utilizing the two-dimensional site frequency spectrum (SFS) between maize and teosinte for inference \cite{gutenkunst2009} we estimate that the maize effective population size during the bottleneck was approximately 5\% that of teosinte.
+Elevated drift during a bottleneck of this magnitude may have enabled moderately deleterious alleles to reach intermediate frequency in maize, while also increasing the probability that strongly deleterious alleles segregating at low frequency in teosinte were purged from the maize gene pool, as is predicted during a domestication bottleneck \cite{somebody}.
+
+After the initial bottleneck, our analysis suggests that maize $N_e$ expanded to reach levels much greater than that of teosinte.
+Our WGS-based estimate shows that the $N_e$ of landrace maize is at least 3X that of ancestral teosinte, while our singleton-based estimate from the the SeeDs project data \cite{somebody} implies an $N_e$ of close to 1 million individuals, or $\sim$8X that of ancestral teosinte.
+From the perspective of a new mutation, both of these estimates are biased downward, since they represent past reductions in diversity due to the aforementioned bottleneck.
+A back-of-the-envelope calculation of modern $N_e$ can be made by assuming there are 47.9 million ha of landrace maize in production \cite{cimmyt1999}, 42,000 plants per ha for open pollinated maize varieties \cite{van2010}, and conservatively that no more than one in 1,000 plants contribute gametes to the next generation.
+This calculation implies a modern $N_e \geq  2$ billion individuals, and incorporates no individuals from hybrid breeding programs.
+If for some reason this back-of-the-envelope calculation is an overestimate, even by several orders of magnitude, it is clear that the post-domesticaiton expansion in maize $N_e$ is enormous.
+Patterns of singleton diversity in maize, as discussed in more detail in the next section, demonstrate that this recent expansion has a notable impact on the maize genome.
+
+\subsection{Demography strongly impacts efficiency of purifying selection}
+Explain that singleton patterns are flipped compared to pi.
+Put this in the context of large pop recent, small pop past, so efficient selection recent, inefficient selection past.
+Explain the importance that this has on old deleterious alleles (potential for many due to drift during bottleneck) and recent deleterious alleles (efficiently purged from population).
 
@@
 
-\subsection{Simulations}
-
 \end{materials}

### 2015-07-29 manuscript.tex
updates... "complete", depending on your def...

@@
 \subsection{Demography strongly impacts efficiency of purifying selection}
-Explain that singleton patterns are flipped compared to pi.
-Put this in the context of large pop recent, small pop past, so efficient selection recent, inefficient selection past.
-Explain the importance that this has on old deleterious alleles (potential for many due to drift during bottleneck) and recent deleterious alleles (efficiently purged from population).
+The observation that maize pairwise diversity is less impacted by the distance to the nearest gene than is teosinte pairwise diversity is reasonable from a long-term evolutionary standpoint.
+Theory demonstrates that purifying selection is more efficient in a large population than in a small one \cite{somebody}, and this observation most likely reflects that prediction.
+More specifically, if teosinte $N_e$ remained relatively constant while maize bottlenecked and recovered exponentially, this provides that the average $N_e$ of maize over the previous several thousand generations is much smaller than that of teosinte, regardless of how much maize has ballooned in the recent past.
+Therefore, our observation shows that purifying selection in maize has not purged deleterious alleles, or the neutral alleles they are linked to, as effectively in maize as in teosinte.
+
+The reversal of this trend when we analyze only singleton diversity instead of pairwise diversity, however, stands out as a notable observation.
+Every mutation begins as a singleton (an allele present in only one individual), and therefore singletons are, on average, the youngest class of alleles that we can observe.
+Therefore unlike pairwise patterns of diversity, which are most heavily influenced by intermediate frequency alleles based on the definition of $\pi$ \cite{nei1979}, singleton diversity is most influenced by recent patterns of evolution.
+Hence, because our demographic estimation indicates dramatic expansion of maize $N_e$ in the recent past, we expect for purifying selection to presently operate more efficiently in maize than in teosinte.
+This observation is very clearly demonstrated by the fact that singleton diversity in maize is more impacted by the distance to the nearest gene than it is in teosinte, as was shown in Figure \ref{fig:purify}.
+
+A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more deleterious alleles segregating at intermediate frequency than does teosinte.
+This could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{something}; if deleterious alleles tend to be recessive and are particularly frequent, they will have increased opportunities to display their phenotypes in inbreds.
+These results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously selected.
+The sample we studied contained only open-pollinated varieties of both maize and teosinte, so we cannot explicitly test this prediction.
+
+Ultimately, we have shown that purifying selection in maize has operated very differently than purifying selection in teosinte.
+This observation along with the potential for soft sweeps appear to explain the phenotypic divergence between maize and teosinte much more completely than do hard sweeps on coding variation.
+Importantly, our estimation of the parameters of the maize domestication bottleneck contribute to the understanding of how the deomography of crop domestication can impact crop diversity.
+The bottleneck-effects from a sudden collapse in population size have been well studied and are known to impact crops for thousands of generations.
+Our observations demonstrate that the rapid expansion experienced by many crops after domestication can also have a profound influence on patterns of diversity, and the effects of this expansion should be accounted for as important contributors to long-term evolution.
 

### 2015-07-30 manuscript.tex
even closer

@@
 
-\title{The impacts of demography on linked selection in maize}
+\title{Demographic history and patterns of selection in maize during and since domestication}
 \author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
@@
 We utilized whole genome sequence data from 23 maize and 13 teosinte individuals to make inferences.
-We obtained a complete estimate of the population size fluctuation and other demographic parameters experienced by maize as it was domesticated from teosinte.
+We obtained a complete estimate of the population size fluctuations and other demographic parameters experienced by maize as it was domesticated from teosinte.
 Here, we show that maize experienced a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid population size expansion post-domestication.
-We observe that hard sweeps, specifically positive selection on new genic mutations, are not the primary force driving maize and teosinte apart.
-Also, we find that reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize.
+We observe that hard sweeps, specifically positive selection on new genic mutations, are not the primary force driving maize evolution.
+Also, we find that a reduced population size during domestication decreased the efficiency of purifying selection at purging deleterious alleles from maize.
 However, expansion after domestication has since increased the efficiency of purifying selection to levels superior to those seen in teosinte.
@@
 \dropcap{D}omesticated plant species evolve in a unique fashion
-compared to their wild counterparts \textcolor{red}{(ref here)}. This
+compared to their wild counterparts \cite{doebley2006}. This
 is a result of both the anthropomorphic nature of artificial selection on
@@
 research that involves searching the genomes of domesticated species for evidence
-of positive selection exists \cite{hufford2012, he2011, OTHERS?}, these studies tend to focus on
+of positive selection exists \cite{hufford2012, he2011, vigouroux2002, chapman2008}, these studies tend to focus on
 identifying or mapping particular genes or regions that play an
@@
 impacted by the effects of selection through linkage
-\cite{sella2009}. This observation begs the question, how can
+\cite{sella2009}. A natural next question, therefore, is how can
 demographic parameters be estimated independently of
@@
 phenomena. Maize is a species of tremendous importance worldwide as
-both a staple crop \cite{SOMETHING} and as a model for
-understanding crop evolution \cite{SOMETHINGS}. Broadly speaking, archaeological and genetic studies have
+both a staple crop \cite{shiferaw2011} and as a model for
+understanding plant evolution \cite{strable2009}. Broadly speaking, archaeological and genetic studies have
 established that maize domestication is likely to have taken place in
-Central Mexico approximately 9,000 years bp
-\cite{SOMETHING,SOMETHING}. Teosinte, the most
+ Mexico approximately 9,000 years bp
+\cite{smith1995,matsuoka2002}. Teosinte, the most
 recent wild ancestor to maize, remains extant throughout much of the
-Americas \cite{SOMETHING}. Additionally, several large-effect
-domestication loci \cite{SOMETHINGS} and putative domestication
+Americas \cite{wilkes1967}. Additionally, several large-effect
+domestication loci \cite{doebley1995, wills2013, wang2015} and putative domestication
 regions \cite{hufford2012} have been identified. But despite all that is
@@
 Previous research of the maize domestication process has relied upon
-observations drawn from genic DNA \textcolor{red}{(several references
-  here)}. Our data were generated through whole genome sequencing,
+observations drawn from genic DNA \cite{wright2005,wang1999,eyre1998}. Our data were generated through whole genome sequencing,
 which eliminated this constraint. Importantly, we observed substantial
@@
 dearth of low-frequency alleles and Tajima's D
-\textcolor{red}{(reference here)} was therefore shifted
-to more positive values \textcolor{red}{(figure here)}. This is consistent with the aforementioned purifying
+\cite{tajima1989} was therefore shifted
+to more positive values \ref{fig:sffs}. This is consistent with the aforementioned purifying
 selection and indicates that genic regions are not evolving
@@
 to sites at least 5kb from gene start/stop postions.
+
+\begin{figure}[b]
+\centering
+\includegraphics[width=.5\textwidth]{FigsAndFiles/SFS_and_Tajima.png}
+\caption{The folded SFS and Tajimas D were calculated both inside and outside of genes for maize and teosinte.
+In both taxa, a dearth of rare alleles was observed within genes relative to outside of genes, which led to higher values of Tajima's D outside of genes than inside.{\bf A:} Folded SFS for maize; {\bf B:} Folded SFS for teosinte; {\bf C:} Maize Tajima's D; {\bf D:} Teosinte Tajima's D.
+\label{fig:sfs} }
+\end{figure}
 
@@
 
-The most likely model suggested an ancestral for $\theta$ ($4N\mu$) of
-0.014734, which is similar to previous estimates
-\textcolor{red}{(references here)}. Assuming a mutation rate of $\mu =
-3 \times 10^{-8}$ \textcolor{red}{(reference here)}, this suggests an effective
+The most likely model suggested an ancestral for $\theta$ ($4N\mu$) per base pair of
+0.014734, which approximately matches a previous independent estimate \cite{eyre1998}. Assuming a mutation rate of $\mu =
+3 \times 10^{-8}$ \cite{clark2005}, this suggests an effective
 population size of $N_a = 122,783$ individuals, a population split
@@
 archaeological data which suggests maize domestication began
-approximately 9,000 years before present \textcolor{red}{(reference
-  here)}. This could reflect teosinte population structure that was present before
+no more than 10,000 years before present \cite{smith1995}. This could reflect teosinte population structure that was present before
 domestication. Also, note that the genetic time of the population
@@
 By utilizing the two-dimensional site frequency spectrum (SFS) between maize and teosinte for inference \cite{gutenkunst2009} we estimate that the maize effective population size during the bottleneck was approximately 5\% that of teosinte.
-Elevated drift during a bottleneck of this magnitude may have enabled moderately deleterious alleles to reach intermediate frequency in maize, while also increasing the probability that strongly deleterious alleles segregating at low frequency in teosinte were purged from the maize gene pool, as is predicted during a domestication bottleneck \cite{somebody}.
+A bottleneck of this magnitude may have enabled moderately deleterious alleles to reach intermediate frequency in maize, while also increasing the probability that strongly deleterious alleles segregating at low frequency in teosinte were purged from the maize gene pool, due to elevated drift as is predicted during a domestication bottleneck \cite{tajima1989}.
 
 After the initial bottleneck, our analysis suggests that maize $N_e$ expanded to reach levels much greater than that of teosinte.
-Our WGS-based estimate shows that the $N_e$ of landrace maize is at least 3X that of ancestral teosinte, while our singleton-based estimate from the the SeeDs project data \cite{somebody} implies an $N_e$ of close to 1 million individuals, or $\sim$8X that of ancestral teosinte.
+Our WGS-based estimate shows that the $N_e$ of landrace maize is at least 3X that of ancestral teosinte, while our singleton-based estimate from the the SeeDs project data \textcolor{red}{(citation)} implies an $N_e$ of close to 1 million individuals, or $\sim$8X that of ancestral teosinte.
 From the perspective of a new mutation, both of these estimates are biased downward, since they represent past reductions in diversity due to the aforementioned bottleneck.
@@
 The observation that maize pairwise diversity is less impacted by the distance to the nearest gene than is teosinte pairwise diversity is reasonable from a long-term evolutionary standpoint.
-Theory demonstrates that purifying selection is more efficient in a large population than in a small one \cite{somebody}, and this observation most likely reflects that prediction.
+Theory has established that purifying selection is more efficient in a large population than in a small one \cite{kimura1984}, and this observation most likely reflects that prediction.
 More specifically, if teosinte $N_e$ remained relatively constant while maize bottlenecked and recovered exponentially, this provides that the average $N_e$ of maize over the previous several thousand generations is much smaller than that of teosinte, regardless of how much maize has ballooned in the recent past.
@@
 A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more deleterious alleles segregating at intermediate frequency than does teosinte.
-This could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{something}; if deleterious alleles tend to be recessive and are particularly frequent, they will have increased opportunities to display their phenotypes in inbreds.
+This could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{meghji1984}; if deleterious alleles tend to be recessive and are particularly frequent, they will have increased opportunities to display their phenotypes in inbreds.
 These results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously selected.
@@
 units using a mutation rate estimate of 3e-8
-\textcolor{red}{(REFERENCE)}. Among the 1,000 dadi runs, parameters
+\cite{clark2005}. Among the 1,000 dadi runs, parameters
 from that with the lowest log-likelihood were taken as estimates of

### 2015-07-31 manuscript.tex
many updates

@@
 To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene tga1, one of the few known instances of a hard sweep corresponding to a causal nonsynonymous mutuation in maize \cite{wang2015}.
-In the case of this known sweep, we observed a reduction in diversity surrounding the causitive substitution to levels much lower than seen for synonymous substitutions \textcolor{red}{(SUPP FIG)}, confirming that this method should be informative if hard sweeps were the prevaling mode of selection in maize.
+In the case of this known sweep, we observed a reduction in diversity surrounding the causitive substitution to levels much lower than seen for synonymous substitutions (Figure S1), confirming that this method should be informative if hard sweeps were the prevaling mode of selection in maize.
 Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize.
-When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score, putatively corresponding to genes undergoing the strongest and weakest purifying selection \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites \textcolor{red}{(SUPP FIG)}.
+When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score, putatively corresponding to genes undergoing the strongest and weakest purifying selection \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites (Figure S2).
 
@@
   alleles and thus depicting evolutionary patterns in the recent
-  past.
-Bootstrap-based 95\% confidence intervals are depicted via shading.
-Inset plots depict a smaller range on the x-axis.
-\label{fig:purify} \jri{should also add the singleton vs pi comparison within species as a supplemental figure.
+  past. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a smaller range on the x-axis. \label{fig:purify}
+  %\jri{should also add the singleton vs pi comparison within species as a supplemental figure.
 if $\pi$/singleton curves {\bf don't} match, will need to explain this suggests teosinte not in equilibrium.
-good justification i think for msmc model, which we then use to say ``gee whiz, not perfect equilibrium but who cares because the big interesting demography is in maize.'' }}
+good justification i think for msmc model, which we then use to say ``gee whiz, not perfect equilibrium but who cares because the big interesting demography is in maize.''}
+  }
 \end{figure*}
@@
 \cite{tajima1989} was therefore shifted
-to more positive values \ref{fig:sffs}. This is consistent with the aforementioned purifying
+to more positive values \ref{fig:tajima1989}. This is consistent with the aforementioned purifying
 selection and indicates that genic regions are not evolving
@@
 implies that rare alleles will most accurately reflect current maize
-demography, we obtained the estimate \textcolor{red}{$N_e =
-  992,713.5$}. Although less biased than the previously mentioned
+demography, we obtained the estimate $N_e =
+  992,713.5$. Although less biased than the previously mentioned
 estimate, we note that ascertainment bias of the
@@
 
-Similarly, to complement our investigation of pairwise diversity surrounding substitutions, we studied patterns of singleton diversity surrounding synonymous and nonsynonymous substitutions in maize \textcolor{red}{(Supplemental Fig)}.
+Similarly, to complement our investigation of pairwise diversity surrounding substitutions, we studied patterns of singleton diversity surrounding synonymous and nonsynonymous substitutions in maize (Figure S3).
 A nearly identical pattern to that shown in Figure \ref{fig:hardSweeps} was observed.
@@
 H12 is sensitive to both hard and soft sweeps.
-Sites nearest to genes that displayed among the top 20\% of H12 values in either maize or teosinte were removed from our dataset, and both pairwise and singleton diversity at remaining sites, as a function of the distance to the nearest gene, was calculated \textcolor{red}{(Supp Fig)}.
+Sites nearest to genes that displayed among the top 20\% of H12 values in either maize or teosinte were removed from our dataset, and both pairwise and singleton diversity at remaining sites, as a function of the distance to the nearest gene, was calculated (Figure S4).
 This subset of the data in no longer subject to the potentially confounding effects of soft-sweeps, and nearly identical patterns of pairwise and singleton diversity surrounding genes are observed.
@@
 
-\jri{either we need to provide all scripts in a single github repo with a readme, or provide links to individual scripts in the methods as we go along}
+  \subsection{BASH, R, and Python scripts}
+All analysis scripts are available in an online repository at \textcolor{red}{REPO ADDRESS HERE.}
 
@@
 panel \cite{chia2012, lemmon2014}.
-Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results \textcolor{red}{(maybe a supplemental figure here)}.
-\jri{yes supp.
-figure good idea}
+Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results \textcolor{red}{(Figure S-PCA)}.
 Ultimately, 23 maize inbreds derived from a diverse assortment of landraces were selected for inclusion.
@@
 ssp. \emph{parviglumis}, were utilized.
-Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} \url(ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/).
+Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{chia2012}.
 
 \subsection{Interpolating genetic position}
-For many of the following analyses, physical position along a chromosome was a less relevant measure of map location than was genetic position.
-Therefore, physical positions were converted to genetic positions by interpolating from the NAM genetic map \citex \jri{Wallace 2014 no?}, which provides a 1 cM resolution for physical to genetic conversion.
+For most of our analyses, physical position along a chromosome was a less relevant measure of map location than genetic position.
+Physical positions were converted to genetic positions by interpolating from the NAM genetic map \cite{glaubitz2014}, which provides a 1 cM resolution for physical to genetic conversion.
 Within R \cite{R2014}, physical positions with corresponding genetic positions in the NAM map were used as anchors.
@@
 maize.The free parameters of our model were $N_a, N_b, N_{maize}, T_b,
-M_{mt}, \text{and } M_{tm} $. Python code specifying this model is
-provided in \textcolor{red}{(SUPPLEMENT)}. We implimented 1,000 dadi
+M_{mt}, \text{and } M_{tm} $. We implimented 1,000 dadi
 optimizations of this model using the ``dadi.Inference.optimize\_log\_fmin''
@@
 \cite{clark2005}. Among the 1,000 dadi runs, parameters
-from that with the lowest log-likelihood were taken as estimates of
+from the run with the lowest log-likelihood were taken as estimates of
 the parameters of the domestication process.
@@
 \cite{elshire2011}, a protocol that suffers from a high rate of
-missing data \cite{beissinger2013}. To minimiaze biases imposed from
+missing data \cite{beissinger2013}. To minimize biases imposed from
 missing data, we used R to isolate only those SNPs that
@@
 SFS. Next, SNPs and genotypes among maize, teosinte, and \emph{Tripsacum} were called. \emph{Tripsacum} bam files
-were downloaded from \url{}(TRIPSICUM FILES), and then all SNPs with a
+were downloaded from \textcolor{red}{(TRIPSICUM FILES)}, and then all SNPs with a
 p-value less than 1e-6 were called using ANGSD. Quality filters were
@@
 no more than 20\% missing data for which every maize or teosinte
-allele differed from the observed \emph{Tripsacum} allele. At each class of
+allele differed from the observed \emph{Tripsacum} allele. At each
 substitution, effects were estimated using the ensembl variant effects
@@
 For each diversity window with at least 100 bps observed, the distance from the window center to the
-nearest synonymous and nonsynonymous (missense) substitution was
+nearest synonymous and missense substitution was
 computed. Then, following the methods of \cite{hernandez2011}, a loess
@@
 nearest synonymous or nonsynonymous substitution. A span of 0.01 was
-utilized. Unlike
-\cite{sattath2011}, we did not fit separate loess curves in the up- and
-down-stream directions, but instead fit single curves encompassing
-both directions.
+utilized.
+
+To verify that this approach is robust to the criticisms of \cite{enard2014}, which involve reduced power at highly conserved sites, we applied the method to subsets of sites that included only positions nearest to the  most- or least-conserved 10\% of genes data.
+This was achieved by assigning every gene a GERP score corresponding to the average score observed across it \cite{rodgers2015}.
+Next, genes ing the top and bottom 10\% quantiles of GERP score were isolated.
+Two datasets were created, each including all diversity windows who's nearest gene was in the high or low GERP score sets.
+Finally, Loess fits of diversity vs.
+distance to gene were computed from sites within each GERP score.
 
@@
 
+Additionally, we performed a parallel implementation of the above analysis after excluding sites potentially experiencing positive selection in the form of either hard- or soft-sweeps.
+To isolate these sites, we performed genome-scans separately for maize and teosinte based on the H12 statistic \cite{garud2015}.
+SNPs within each taxa were called in ANGSD \cite{korneliussen2014} using the previously specified SNP calling parameters.
+H12 was computed with a window size of 200 SNPs and a jump of 25 SNPs.
+H12 windows in the highest 20\% for maize or teosinte were identified and considered as regions most likely to be under positive selection.
+Finally, levels of pairwise and singleton diversity vs.
+distance to gene was computed using only sites that were not nearest to selected sites in either maize or teosinte.
+
 %JRI IDEAS:
@@
 \begin{acknowledgments}
-Various thankyous will be in order. \jri{Coop and his lab. funding from NSF Biology of Rare alleles -- need grant number.}
+We are indebted to Graham Coop and Simon Aeshbacher for their constructive input during this study. Funding was provided by \textcolor{red}{NSF Grant number ZZZZZZZZZ}.
 \end{acknowledgments}

### 2015-07-31 manuscript.tex
Draft to edit

@@
 \author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
-    California, Davis, CA, USA}, Li Wang\affil{2}{Iowa State University, Ames, IA, USA}, Arun
-  Durvasula\affil{1}{}, Kate Crosby\affil{1}{}, Matthew Hufford\affil{2}{}, \and Jeffrey
-  Ross-Ibrarra\affil{1}{}\affil{3}{Genome Center and Center for population biology, University of
+    California, Davis, CA, USA}\affil{2}{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}\affil{3}{Division of Plant Sciences, University of Missouri, Columbia, MO, USA} Li Wang\affil{4}{Iowa State University, Ames, IA, USA}, Kate Crosby\affil{1}{}, Arun
+  Durvasula\affil{1}{}, Matthew Hufford\affil{4}{}, \and Jeffrey
+  Ross-Ibrarra\affil{1}{}\affil{5}{Genome Center and Center for population biology, University of
     California, Davis, CA, USA} }
 
-\significancetext{Patterns of linked selection, or the impact of selection on sites that neighbor the functional variant, have been carefully evaluated in only a few species.
+\significancetext{Patterns of linked selection, or the impact of selection on sites that neighbor a functional variant, have been carefully evaluated in only a few species.
 In this work, we demonstrate that selection against deleterious mutations leaves a pronounced signature on the maize genome, reducing diversity in and immediately around genes.
 We show how demography interacts with selection to impact genome-wide patterns of diversity, including the important observation that rapid population expansion can increase the efficiency of selection as much as a sudden population bottleneck can weaken it.
-Along the way, we develop the first estimate the domegraphic parameters of the maize domestication from whole genome sequence data.}
+Along the way, we develop the first estimate the demographic parameters of the maize domestication from whole genome sequence data.}
 
@@
 \begin{abstract}
-  Unique selective and demographic forces operate on plants during domestication.
-These forces interact to generate the patterns of DNA variability that are persistent in crop species today.
-To quantify this interplay, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide.
+  Unique selective and demographic operate on domesticated plant species.
+These forces interact during and after domestication to generate the patterns of DNA variability that are persistent in crop species today.
+To quantify the interplay between demography and selection, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide.
 We utilized whole genome sequence data from 23 maize and 13 teosinte individuals to make inferences.
 We obtained a complete estimate of the population size fluctuations and other demographic parameters experienced by maize as it was domesticated from teosinte.
-Here, we show that maize experienced a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid population size expansion post-domestication.
+Here, we show that maize went through a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid population size expansion post-domestication.
 We observe that hard sweeps, specifically positive selection on new genic mutations, are not the primary force driving maize evolution.
-Also, we find that a reduced population size during domestication decreased the efficiency of purifying selection at purging deleterious alleles from maize.
+We find that a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize.
 However, expansion after domestication has since increased the efficiency of purifying selection to levels superior to those seen in teosinte.
-Our results demonstrate that in domesticated species, demography and selection are constantly fluctuating through time.
+Our results demonstrate that in domesticated species, demographic and selective history in the ancient and recent past both contribute to genetic variability that is present today, providing substantial implications for the continued improvement of domesticated species.
+
 \end{abstract}
@@
 important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
-selection have on whole-genome patterns of genetic variability therefore remains limited.
+selection have on whole-genome patterns of genetic variability remains limited.
 
@@
 relative importance of different forms of selection on whole-genome
-variability in both maize and teosinte, as different forms of selection affect DNA
-variability remarkably differently; 2) research the impact that the
+variability in both maize and teosinte 2) research the impact that the
 domestication process has had on genetic variability in maize, and how
@@
 \subsection{Patterns of diversity differ between genic and
-  nongenic regions of the genome}
+  non-genic regions of the genome}
 Previous research of the maize domestication process has relied upon
@@
 diversity ($\pi$) within genes was significantly lower than at
-positions at least 5kb away from genes (0.00668 within, 0.00691 away, p$<$2e-44). The same
+positions at least 5 kb away from genes (0.00668 within, 0.00691 away, p$<$2e-44). The same
 pattern of significantly greater diversity outside of genes was observed in
 teosinte, but to a larger extent. For teosinte we observed $\pi$
-within genes to be 0.0088 and at least 5kb from genes it was 0.0115
+within genes to be 0.0088 and at least 5 kb from genes it was 0.0115
 (p$\approx$0). These observations suggest that genes are not evolving
@@
 sweeps was evaluated by comparing diversity
-surrounding non-synonymous and synonymous substitutions in maize and
+surrounding non-synonymous (specifically missense) and synonymous substitutions in maize and
 teosinte since divergence from \emph{Tripsacum}, roughly 1-1.2 million years before
@@
 
-To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene tga1, one of the few known instances of a hard sweep corresponding to a causal nonsynonymous mutuation in maize \cite{wang2015}.
-In the case of this known sweep, we observed a reduction in diversity surrounding the causitive substitution to levels much lower than seen for synonymous substitutions (Figure S1), confirming that this method should be informative if hard sweeps were the prevaling mode of selection in maize.
+To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene \emph{tga1}, one of the few known instances of a hard sweep corresponding to a causal missense substitution in maize \cite{wang2015}.
+In the case of this known sweep, we observed a reduction in diversity surrounding the causative substitution to levels much lower than seen for synonymous substitutions (Figure S1), confirming that this method should be informative if hard sweeps were the prevailing mode of selection in maize.
 Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize.
@@
 
-\begin{figure*}[b]
+\begin{figure*}[!htb]
 %\vspace*{.05in}
@@
 \hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_June}
-\caption{Pairwise diversity surrounding synonymous and nonsynonymous
+\caption{Pairwise diversity surrounding synonymous and non-synonymous
   substitutions in maize and teosinte. {\bf A:} Maize diversity; {\bf B:} Teosinte
@@
 When this form of selection is operating, it can serve to reduce genetic variability at linked neutral sites, a phenomenon often called background selection \cite{charlesworth1993}.
-Purifying and background selection lead to lower diversity within genes and other functional sites relative to neutral regions \textcolor{red}{(ref here)}.
+Purifying and background selection lead to lower diversity within genes and other functional sites relative to neutral regions.
 We investigated purifying selection in maize and teosinte by evaluating the average magnitude of reduced diversity within genes and recovery away from genes in both taxa.
-To enable comparisons between maize and teosinte, we standardized by neutral levels within each taxa, where neutral levels were defined as mean diversity at positions far from genes at least 0.01 cM distal from genes.
+To enable comparisons between maize and teosinte, we standardized by neutral levels within each taxa, where neutral levels were defined as mean diversity at positions at least 0.01 cM distal from genes.
 For both maize and teosinte, pairwise diversity was reduced within genes and a recovery of diversity away from genes was observed.
@@
 
-\begin{figure*}[t]
+\begin{figure*}[!htb]
 %\vspace*{.05in}
@@
 landraces and 13 teosinte inbred lines included in the HapMap 2 panel
-\cite{chia2012} \textcolor{red}{(supplemental table here)}. The maize
+\cite{chia2012}. The maize
 lines were collected from across the Americas and the teosinte lines
@@
 \cite{tajima1989} was therefore shifted
-to more positive values \ref{fig:tajima1989}. This is consistent with the aforementioned purifying
+to more positive values, as previously shown (Figure \ref{fig:purify}). This is consistent with the aforementioned purifying
 selection and indicates that genic regions are not evolving
-neutrally. This observation is consistent with previous research \cite{hufford2012}. Therefore, for demographic modeling we restricted analysis
-to sites at least 5kb from gene start/stop postions.
+neutrally. Therefore, for demographic modeling we restricted analysis
+to sites at least 5 kb from gene start/stop positions.
 
-\begin{figure}[b]
+\begin{figure}[!htb]
 \centering
@@
 $N_b$ individuals, followed by exponential growth to size
-$N_m$. Additionally, since the split a $M_{tm}$ individuals migrated
+$N_m$. Additionally, since the split $M_{tm}$ individuals migrated
 from teosinte into maize and $M_{mt}$ individuals migrated from maize
@@
 
-The most likely model suggested an ancestral for $\theta$ ($4N\mu$) per base pair of
+The most likely model suggested an ancestral $\theta$ ($4N\mu$) per base pair of
 0.014734, which approximately matches a previous independent estimate \cite{eyre1998}. Assuming a mutation rate of $\mu =
@@
 from
-archaeological data which suggests maize domestication began
+archaeological data, which suggest maize domestication began
 no more than 10,000 years before present \cite{smith1995}. This could reflect teosinte population structure that was present before
 domestication. Also, note that the genetic time of the population
-split must precede morphological changes that could be identified
-morphologically.   Additionally, because recent expansion is
-most evidenced by rare alleles, and since these data provide low
-power to detect rare alleles, we expect that the estimate of  $N_m =
-366,973$, or $\sim 3N_a$, is likely an underestimate.
+split must precede physiological changes that could be identified
+archaelogically.   Additionally, because recent expansion is
+most evidenced by rare alleles, and since our limited sample size provides low
+power to detect rare alleles, we expect that the estimate of  $N_m = 366,973$, or $\sim 3N_a$, is likely an underestimate.
 
-We utilized a complementary dataset of 4,021 maize landrace individuals collected
+Therefore, we utilized a complementary data set of 4,021 maize land-race individuals collected
 from across the Americas \textcolor{red}{(SeeDs reference here)} to
@@
 
-\begin{figure}[b]
+\begin{figure}[!htb]
 %\vspace*{.05in}
@@
 
-\section{Post-domestiction expansion is apparent in patterns of singleton diversity}
-Motivated by the rapid post-domestication expansion in maize is seen in our demographic analysis, we conducted an analysis of genome-wide patterns of singleton diversity.
+\section{Post-domestication expansion is apparent in patterns of singleton diversity}
+Motivated by the rapid post-domestication expansion in maize that is evident from our demographic analysis, we conducted an analysis of genome-wide patterns of singleton diversity.
 Singleton diversity measures the abundance of alleles that are only observed in one individual among the sample.
@@
 diversity reflect recent patterns of evolution. Since our sample size
-for maize (n=23) was larger than teosite (n=13), maize singletons were
-analyzed directly as well as downsampled to reflect the sample size of
+for maize (n=23) was larger than teosinte (n=13), maize singletons were
+analyzed directly as well as down-sampled to reflect the sample size of
 teosinte. For singletons, our definition of
-neutral diversity was modified to be mean diversity at positions at
+neutral diversity was defined as diversity at positions at
 least 0.02 cM distal from genes. When
@@
 
-Similarly, to complement our investigation of pairwise diversity surrounding substitutions, we studied patterns of singleton diversity surrounding synonymous and nonsynonymous substitutions in maize (Figure S3).
+Similarly, to complement our investigation of pairwise diversity surrounding substitutions, we studied patterns of singleton diversity surrounding synonymous and non-synonymous substitutions in maize (Figure S3).
 A nearly identical pattern to that shown in Figure \ref{fig:hardSweeps} was observed.
@@
 
-\subsection{Purifying seleciton results are robust to soft-sweeps}
-As a category, instances of soft-sweeps, or cases of selection on alleles that have already reached intermediate frequency before becoming advantageous, are difficult to identify \textcolor{red}{(refs here)}.
-However, this form of selection is likely to be frequent during crop domestication \textcolor{red}{(refs here)}.
+\subsection{Purifying selection results are robust to soft-sweeps}
+As a category, instances of soft-sweeps, or cases of selection on alleles that have already reached intermediate frequency before becoming advantageous, are difficult to identify \cite{wilson2014}.
+However, this form of selection is likely to be frequent during crop domestication \cite{innan2004}.
 We were concerned that the disparate patterns of pairwise and singleton diversity in maize and teosinte, which seem to reflect differences in the efficiency of purifying selection due to demographic history, may instead reflect differences in the location and abundance of soft-sweeps.
@@
 H12 is sensitive to both hard and soft sweeps.
-Sites nearest to genes that displayed among the top 20\% of H12 values in either maize or teosinte were removed from our dataset, and both pairwise and singleton diversity at remaining sites, as a function of the distance to the nearest gene, was calculated (Figure S4).
-This subset of the data in no longer subject to the potentially confounding effects of soft-sweeps, and nearly identical patterns of pairwise and singleton diversity surrounding genes are observed.
+Sites nearest to genes that displayed among the top 20\% of H12 values in either maize or teosinte were removed from our data set, and both pairwise and singleton diversity at remaining sites, as a function of the distance to the nearest gene, was calculated (Figure S4).
+This subset of the data is no longer subject to the potentially confounding effects of soft-sweeps, and nearly identical patterns of pairwise and singleton diversity surrounding genes are observed.
 This provides further evidence that the patterns of diversity we observe in maize as compared to teosinte are the result of rapid population expansion as opposed to differing selective patterns in the domesticated and wild taxa.
@@
 One known example of positive selection on a non-genic mutation involves the \emph{tb1} locus of maize, which is one of the best characterized examples of positive selection on a ``domestication gene'' in any crop \cite{clark2006}.
-However, the maize \emph{tb1} allele was already present in teosinte before domestication \cite{studer2011}, so this even this example demonstrates supports our finding that hard sweeps are rare.
+However, the maize \emph{tb1} allele was already present in teosinte before domestication \cite{studer2011}, so this this example is in agreement with our finding that hard sweeps are rare.
 The \emph{gt1} locus is another well-characterized case of positive selection operating on standing variation at an enhancer region \cite{wills2013}.
@@
 We should additionally note that our observations do not exclude the possibility of infrequent hard sweeps having taken place during maize evolution.
-In fact, there is one maize locus, \emph{tga1}, that has been shown to correspond to a hard sweep at an amino-acid changing mutation \cite{wang2015}.
+In fact, the maize locus \emph{tga1} that has been shown to correspond to a hard sweep at an amino-acid changing mutation \cite{wang2015}.
 Surrounding \emph{tga1},  our data demonstrate a pattern consistent with a hard sweep.
-But, our data do demonstrate that instances such as this are infrequent.
-This contrasts sharply with \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, where differences in diversity surrounding synonymous and nonsynonymous substitutions are clear.
+But, our data demonstrate that instances such as this are infrequent.
+This contrasts sharply with \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, where differences in diversity surrounding synonymous and non-synonymous substitutions are clear suggesting hard-sweeps are abundant.
 However, it agrees with what has been seen for humans \cite{hernandez2011}.
@@
 This calculation implies a modern $N_e \geq  2$ billion individuals, and incorporates no individuals from hybrid breeding programs.
-If for some reason this back-of-the-envelope calculation is an overestimate, even by several orders of magnitude, it is clear that the post-domesticaiton expansion in maize $N_e$ is enormous.
+If for some reason this back-of-the-envelope calculation is an overestimate, even by several orders of magnitude, it is clear that the post-domestication expansion in maize $N_e$ is enormous.
 Patterns of singleton diversity in maize, as discussed in more detail in the next section, demonstrate that this recent expansion has a notable impact on the maize genome.
@@
 More specifically, if teosinte $N_e$ remained relatively constant while maize bottlenecked and recovered exponentially, this provides that the average $N_e$ of maize over the previous several thousand generations is much smaller than that of teosinte, regardless of how much maize has ballooned in the recent past.
-Therefore, our observation shows that purifying selection in maize has not purged deleterious alleles, or the neutral alleles they are linked to, as effectively in maize as in teosinte.
+Therefore, our observation shows that purifying selection in maize has not purged deleterious alleles, or the neutral alleles they are linked to, as effectively as in teosinte.
 
 The reversal of this trend when we analyze only singleton diversity instead of pairwise diversity, however, stands out as a notable observation.
-Every mutation begins as a singleton (an allele present in only one individual), and therefore singletons are, on average, the youngest class of alleles that we can observe.
+Every mutation begins as a singleton (an allele present in only one individual), and therefore singletons are, on average, the youngest class of alleles that can be observed.
 Therefore unlike pairwise patterns of diversity, which are most heavily influenced by intermediate frequency alleles based on the definition of $\pi$ \cite{nei1979}, singleton diversity is most influenced by recent patterns of evolution.
@@
 A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more deleterious alleles segregating at intermediate frequency than does teosinte.
-This could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{meghji1984}; if deleterious alleles tend to be recessive and are particularly frequent, they will have increased opportunities to display their phenotypes in inbreds.
-These results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously selected.
-The sample we studied contained only open-pollinated varieties of both maize and teosinte, so we cannot explicitly test this prediction.
+This could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{meghji1984}; if deleterious alleles tend to be recessive and are particularly frequent, they will have ample opportunities to display their phenotypes in inbreds.
+Our results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously select
 
 Ultimately, we have shown that purifying selection in maize has operated very differently than purifying selection in teosinte.
-This observation along with the potential for soft sweeps appear to explain the phenotypic divergence between maize and teosinte much more completely than do hard sweeps on coding variation.
+This observation, along with the potential for soft sweeps, appear to explain the phenotypic divergence between maize and teosinte much more completely than dooes positive selection generating hard-sweeps at protien-coding mutations.
 Importantly, our estimation of the parameters of the maize domestication bottleneck contribute to the understanding of how the deomography of crop domestication can impact crop diversity.
 The bottleneck-effects from a sudden collapse in population size have been well studied and are known to impact crops for thousands of generations.
-Our observations demonstrate that the rapid expansion experienced by many crops after domestication can also have a profound influence on patterns of diversity, and the effects of this expansion should be accounted for as important contributors to long-term evolution.
+Complementing this knowledge, our results demonstrate that the rapid expansion experienced by many crops after domestication can also have a profound influence on patterns of diversity, and the effects of this expansion should be accounted for as important contributors to long-term evolution.
 
@@
 panel \cite{chia2012, lemmon2014}.
-Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results \textcolor{red}{(Figure S-PCA)}.
+Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results (Figure S5).
 Ultimately, 23 maize inbreds derived from a diverse assortment of landraces were selected for inclusion.
 Thirteen teosinte inbred lines, all members of  subspecies Z. \emph{mays}
-ssp. \emph{parviglumis}, were utilized.
+ssp. \emph{parviglumis}, were utilized. Included lines are depicted in (Table S1).
 Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{chia2012}.

### 2015-07-31 supportingInfo.tex
Draft to edit

\documentclass{article}
\usepackage{cite}
\usepackage{color}
\usepackage{fullpage}
\usepackage{pgf}
%\usepackage{hyperref}
\usepackage[normalem]{ulem}
\newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }
\newcommand{\citex}{\textcolor{blue}{\bf CITE}}
\newcommand{\X}{\textcolor{blue}{\bf X}}

\begin{document}

\title{Supporting Information: Demographic history and patterns of selection in maize during and since domestication}
\author{Timothy M. Beissinger, Li Wang, Arun Durvasula, Kate Crosby, \\ Matthew Hufford, Jeffrey Ross-Ibarra}

\maketitle
\section*{}

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=.85\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_tga1Supp_June.png} \\
    \end{center}
{\bf Figure S1:} Diversity surrounding the causitive polymorphism at the \emph{tga1} locus is plotted. Since this is only one gene, the large amount of noise compared to our average plots is expected. However, notice that diversity precisely at the causitive polymorphism is reduced and a recovery of diversity is observed away from that site.
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Conserved_Significance_June.png}
  \includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Unconserved_Significance_June.png}
{\bf Figure S2:} Pairwise diversity surrounding synonymous and nonsynonymous
  substitutions in maize at highly conserved (A) or unconserved (B) sites.  Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis.
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{FigsAndFiles/plotDiversity_TvM_Singletons.png}
{\bf Figure S3:} Singleton diversity surrounding synonymous and nonsynonymous
  substitutions in maize.
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_Unselected_manuscript.png}
    \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_unselected_Singletons_manuscript.png}
{\bf Figure S4:} Relative level of diversity versus distance to the nearest gene, in maize and teosinte, based on only sites that do not show evidence of hard or soft sweeps according to H12. Two measures of diversity were investigated. {\bf A} displays pairwise diversity,
which is most influenced by intermediate frequency alleles and therefore depicts more ancient evolutionary patterns, and {\bf B} depicts singleton diversity, influenced by rare alleles
and thus depicting evolutionary patterns in the recent past. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a smaller range on the x-axis.
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=.8\textwidth]{FigsAndFiles/tilPCA_july.pdf}\\
    \includegraphics[width=.8\textwidth]{FigsAndFiles/bknPCA_july}\\
{\bf Figure S5:}Principal component analysis of teosinte and maize individuals to ensure that no close relatives were inadvertantly included in our study. Plots are based on a random sample of 10,000 SNPs. {\bf A:} Percentage of total variance explained by each principal component for teosinte. {\bf B:} PC1 vs PC2 for all 13 teosinte individuals. {\bf C:} Percentage of total variance explained by each principal component for maize. {\bf D:} PC1 vs PC2 for all 23 maize individuals.
\end{figure}

\begin{figure}[h!]
  \begin{center}
  \begin{tabular}{c|c}
    \bf Maize & \bf Teosinte \\ \hline \hline
    BKN009 &  TIL01 \\
    BKN010 & TIL02 \\
    BKN011 & TIL03 \\
    BKN014 & TIL04-TIP454 \\
    BKN015 & TIL07 \\
    BKN016 & TIL09 \\
    BKN017 & TIL10 \\
    BKN018 & TIL11 \\
    BKN019 & TIL12 \\
    BKN020 & TIL14-TIP498 \\
    BKN022 & TIL15 \\
    BKN023 & TIL16 \\
    BKN025 & TIL17 \\
    BKN026 & \\
    BKN027 & \\
    BKN029 & \\
    BKN030 & \\
    BKN031 & \\
    BKN032 & \\
    BKN033 & \\
    BKN034 & \\
    BKN035 & \\
    BKN040 & \\
  \end{tabular} \\
  \end{center}
      {\bf Table S1:} A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by Chia et al (2012).
\end{figure}

\end{document}

### 2015-08-01 manuscript.tex
first forays

@@
 \usepackage[normalem]{ulem}
-\newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }
-\newcommand{\citex}{\textcolor{blue}{\bf CITE}}
-\newcommand{\X}{\textcolor{blue}{\bf X}}
+\newcommand{\jri}[1]{\textcolor{red}{\emph{#1}} }
+\newcommand{\citex}{\textcolor{red}{\bf CITE}}
+\newcommand{\X}{\textcolor{red}{\bf X}}
 
@@
 
-\title{Demographic history and patterns of selection in maize during and since domestication}
+\title{Domestication and the impacts of linked selection on the maize genome}
 \author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
@@
   Durvasula\affil{1}{}, Matthew Hufford\affil{4}{}, \and Jeffrey
-  Ross-Ibrarra\affil{1}{}\affil{5}{Genome Center and Center for population biology, University of
+  Ross-Ibarra\affil{1}{}\affil{5}{Genome Center and Center for population biology, University of
     California, Davis, CA, USA} }
@@
 \significancetext{Patterns of linked selection, or the impact of selection on sites that neighbor a functional variant, have been carefully evaluated in only a few species.
-In this work, we demonstrate that selection against deleterious mutations leaves a pronounced signature on the maize genome, reducing diversity in and immediately around genes.
+\jri{not really.
+we need to cite the Corbett-Detig plos-bio paper here -- he looked at a lot of species including maize!} In this work, we demonstrate that selection against deleterious mutations leaves a pronounced signature on the maize genome, reducing diversity in and immediately around genes.
 We show how demography interacts with selection to impact genome-wide patterns of diversity, including the important observation that rapid population expansion can increase the efficiency of selection as much as a sudden population bottleneck can weaken it.
-Along the way, we develop the first estimate the demographic parameters of the maize domestication from whole genome sequence data.}
+Along the way, we develop the first estimate the demographic parameters of the maize domestication from whole genome sequence data.
+}
 
@@
 A mutation that is immediately beneficial and positively selected leaves a classical hard
-sweep signature in the genome, whereby genetic diversity surrounding
-that mutation is reduced as the haplotype where the mutation first
+sweep signature in the genome, whereby surrounding genetic diversity
+is reduced as the haplotype where the mutation first
 arose increases in frequency to fixation. The prevalence of such hard
@@
 To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene \emph{tga1}, one of the few known instances of a hard sweep corresponding to a causal missense substitution in maize \cite{wang2015}.
-In the case of this known sweep, we observed a reduction in diversity surrounding the causative substitution to levels much lower than seen for synonymous substitutions (Figure S1), confirming that this method should be informative if hard sweeps were the prevailing mode of selection in maize.
+In the case of this known sweep, we observed a reduction in diversity surrounding the causative substitution to levels much lower than seen for synonymous substitutions (Figure S1) \jri{please add refs to link to supp.
+figs if possible}, confirming that this method should be informative if hard sweeps were the prevailing mode of selection in maize.
 Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize.
@@
 \caption{Pairwise diversity surrounding synonymous and non-synonymous
-  substitutions in maize and teosinte. {\bf A:} Maize diversity; {\bf B:} Teosinte
+  substitutions in maize and teosinte. \jri{a bit too much vertical white space} {\bf A:} Maize diversity; {\bf B:} Teosinte
 diversity. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis. \label{fig:hardSweeps}}
@@
 \subsection{Pairwise diversity is heavily influenced by purifying selection}
-% \jri{we will need to define background selection and be careful about ``purifying selection'' vs ``background selection'', as technically background selection does not include very weak deleterious alleles.}
+
 Purifying selection refers to the situation where deleterious mutations arising in a population are continuously selected against.
@@
 We used diffusion approximation as implemented in
-dadi \cite{gutenkunst2009} to find the domestication parameters that best explain
+$\delta\alpha\delta{i}$ \cite{gutenkunst2009} to find the domestication parameters that best explain
 the joint site frequency spectrum of maize and teosinte.  The model we optimized began with an
@@
 \includegraphics[width=.4\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
-\caption{Parameters of domestication as estimated by dadi. Notably,
+\caption{Parameters of domestication as estimated by $\delta\alpha\delta{i}$. Notably,
   the maize effective population size ($N_e$) during the domestication
@@
 Ultimately, we have shown that purifying selection in maize has operated very differently than purifying selection in teosinte.
-This observation, along with the potential for soft sweeps, appear to explain the phenotypic divergence between maize and teosinte much more completely than dooes positive selection generating hard-sweeps at protien-coding mutations.
-Importantly, our estimation of the parameters of the maize domestication bottleneck contribute to the understanding of how the deomography of crop domestication can impact crop diversity.
+This observation, along with the potential for soft sweeps, appear to explain the phenotypic divergence \jri{pretty hard to argue that purifying selection explains phenotypic divergence!} between maize and teosinte much more completely than does positive selection generating hard-sweeps at protein-coding mutations.
+Importantly, our estimation of the parameters of the maize domestication bottleneck contribute to the understanding of how the demography of crop domestication can impact crop diversity.
 The bottleneck-effects from a sudden collapse in population size have been well studied and are known to impact crops for thousands of generations.
@@
 panel \cite{chia2012, lemmon2014}.
-Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results (Figure S5).
+\jri{this needs slight clarification.
+cite chia for hapmap, but that we used higher-coverage of the teosinte from lemmon.
+if possible, include link to fastq or bamfiles used} Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results (Figure S5).
 Ultimately, 23 maize inbreds derived from a diverse assortment of landraces were selected for inclusion.
 Thirteen teosinte inbred lines, all members of  subspecies Z. \emph{mays}
-ssp. \emph{parviglumis}, were utilized. Included lines are depicted in (Table S1).
-Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{chia2012}.
+ssp. \emph{parviglumis}, were utilized. Included lines are listed in (Table S1).
+Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{chia2012}.\jri{did we use their bams?
+I thought we used bwa bams, which are not listed and weren't used for hapmap2.
+if that's the case, we need to mention this difference.}
 
 \subsection{Interpolating genetic position}
-For most of our analyses, physical position along a chromosome was a less relevant measure of map location than genetic position.
-Physical positions were converted to genetic positions by interpolating from the NAM genetic map \cite{glaubitz2014}, which provides a 1 cM resolution for physical to genetic conversion.
+
+Physical positions were converted to genetic positions by interpolating from the NAM genetic map \cite{glaubitz2014}, \jri{does jeff's paper really have the NAM map?
+I didn't think so and quick look didn't find it.
+i thought this came from wallace?} which provides a 1 cM resolution for physical to genetic conversion.
 Within R \cite{R2014}, physical positions with corresponding genetic positions in the NAM map were used as anchors.
-Physical positions in our dataset wihout corresponding genetic position were assigned genetic positions by scaling the anchored genetic positions according to the physical distance between the unlabeled position and the flanking anchors.
+Physical positions in our dataset without corresponding genetic position were assigned genetic positions by scaling the anchored genetic positions according to the physical distance between the unlabeled position and the flanking anchors.
 
@@
 the biomaRt package \cite{durinck2009,durinck2005} of R \cite{R2014}. Genic
-regions were defined as DNA between the start and stop position of a
+regions were defined as DNA between the start and stop position of a \jri{you meant transcription start/stop not translation, right?}
 gene, while intergenic regions were required to be at least 5kb up- or
@@
 maize reference genome was used for polarization and then unfolded spectra
-were folded using dadi \cite{gutenkunst2009}.
+were folded using $\delta\alpha\delta{i}$ \cite{gutenkunst2009}.
 
 \subsection{Demographic inference}
-We used dadi \cite{gutenkunst2009} to estimate the parameters of maize
+We used $\delta\alpha\delta{i}$ \cite{gutenkunst2009} to estimate the parameters of maize
 domestication based on diffusion approximation of the 2-dimensional
@@
 maize.The free parameters of our model were $N_a, N_b, N_{maize}, T_b,
-M_{mt}, \text{and } M_{tm} $. We implimented 1,000 dadi
-optimizations of this model using the ``dadi.Inference.optimize\_log\_fmin''
-optimizer. In dadi optimizations,
+M_{mt}, \text{and } M_{tm} $. We implimented 1,000 $\delta\alpha\delta{i}$
+optimizations of this model using the ``$\delta\alpha\delta{i}$.Inference.optimize\_log\_fmin''
+optimizer. In $\delta\alpha\delta{i}$ optimizations,
 initial values for $N_b, N_{maize}, T_b,
@@
 lower bounds because it is estimated from equilibrium ancestral theta
-($\theta = 4N_a\mu$), which dadi determines automatically regardless
-of model. Population sizes were converted from dadi's
+($\theta = 4N_a\mu$), which $\delta\alpha\delta{i}$ determines automatically regardless
+of model. Population sizes were converted from $\delta\alpha\delta{i}$'s
 units using a mutation rate estimate of 3e-8
-\cite{clark2005}. Among the 1,000 dadi runs, parameters
+\cite{clark2005}. Among the 1,000 $\delta\alpha\delta{i}$ runs, parameters
 from the run with the lowest log-likelihood were taken as estimates of
@@
 the abundance of rare alleles. However, our demographic parameter
-estimation using dadi suffered from a relatively small sample size of
+estimation using $\delta\alpha\delta{i}$ suffered from a relatively small sample size of
 23 maize individuals. Therefore, our approach likely underestimated

### 2015-08-01 supportingInfo.tex
first attempt to include supp

@@
-\documentclass{article}
-\usepackage{cite}
-\usepackage{color}
-\usepackage{fullpage}
-\usepackage{pgf}
-%\usepackage{hyperref}
-\usepackage[normalem]{ulem}
-\newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }
-\newcommand{\citex}{\textcolor{blue}{\bf CITE}}
-\newcommand{\X}{\textcolor{blue}{\bf X}}
-
-\begin{document}
-
-\title{Supporting Information: Demographic history and patterns of selection in maize during and since domestication}
-\author{Timothy M. Beissinger, Li Wang, Arun Durvasula, Kate Crosby, \\ Matthew Hufford, Jeffrey Ross-Ibarra}
-
-\maketitle
-\section*{}
-
-\begin{figure}[h!]
+\section*{Supporting Information}
+\renewcommand\thefigure{S\arabic{figure}}    
+\setcounter{figure}{0} 
+\begin{figure}[h]
   \begin{center}
@@
     \end{center}
-{\bf Figure S1:} Diversity surrounding the causitive polymorphism at the \emph{tga1} locus is plotted.
+\caption{Diversity surrounding the causitive polymorphism at the \emph{tga1} locus is plotted.
 Since this is only one gene, the large amount of noise compared to our average plots is expected.
-However, notice that diversity precisely at the causitive polymorphism is reduced and a recovery of diversity is observed away from that site.
+However, notice that diversity precisely at the causitive polymorphism is reduced and a recovery of diversity is observed away from that site.}
 \end{figure}
+\clearpage
 
-\begin{figure}[h!]
+\begin{figure}
   \includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Conserved_Significance_June.png}
   \includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Unconserved_Significance_June.png}
-{\bf Figure S2:} Pairwise diversity surrounding synonymous and nonsynonymous
-  substitutions in maize at highly conserved (A) or unconserved (B) sites.  Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis.
+\caption{ Pairwise diversity surrounding synonymous and nonsynonymous
+  substitutions in maize at highly conserved (A) or unconserved (B) sites.  Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis.}
 \end{figure}
+\clearpage
 
-\begin{figure}[h!]
+\begin{figure}
   \includegraphics[width=\textwidth]{FigsAndFiles/plotDiversity_TvM_Singletons.png}
-{\bf Figure S3:} Singleton diversity surrounding synonymous and nonsynonymous
-  substitutions in maize.
+\caption{ Singleton diversity surrounding synonymous and nonsynonymous
+  substitutions in maize.}
 \end{figure}
+\clearpage
 
-\begin{figure}[h!]
+\begin{figure}
   \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_Unselected_manuscript.png}
     \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_unselected_Singletons_manuscript.png}
-{\bf Figure S4:} Relative level of diversity versus distance to the nearest gene, in maize and teosinte, based on only sites that do not show evidence of hard or soft sweeps according to H12.
+\caption{ Relative level of diversity versus distance to the nearest gene, in maize and teosinte, based on only sites that do not show evidence of hard or soft sweeps according to H12.
 Two measures of diversity were investigated.
 {\bf A} displays pairwise diversity,
-which is most influenced by intermediate frequency alleles and therefore depicts more ancient evolutionary patterns, and {\bf B} depicts singleton diversity, influenced by rare alleles
-and thus depicting evolutionary patterns in the recent past. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a smaller range on the x-axis.
+which is most influenced by intermediate frequency alleles and therefore depicts more ancient evolutionary patterns, and {\bf B} depicts singleton diversity, influenced by rare alleles and thus depicting evolutionary patterns in the recent past.
+Bootstrap-based 95\% confidence intervals are depicted via shading.
+Inset plots depict a smaller range on the x-axis.}
 \end{figure}
+\clearpage
 
-\begin{figure}[h!]
-  \includegraphics[width=.8\textwidth]{FigsAndFiles/tilPCA_july.pdf}\\
-    \includegraphics[width=.8\textwidth]{FigsAndFiles/bknPCA_july}\\
-{\bf Figure S5:}Principal component analysis of teosinte and maize individuals to ensure that no close relatives were inadvertantly included in our study.
+\begin{figure}
+  \includegraphics[width=.5\textwidth]{FigsAndFiles/tilPCA_july.pdf}\\
+    \includegraphics[width=.5\textwidth]{FigsAndFiles/bknPCA_july}\\
+\caption{Principal component analysis of teosinte and maize individuals to ensure that no close relatives were inadvertantly included in our study.
 Plots are based on a random sample of 10,000 SNPs.
@@
 {\bf D:} PC1 vs PC2 for all 23 maize individuals.
+\jri{need to fix letter label positions}}
 \end{figure}
+\clearpage
 
-\begin{figure}[h!]
+\begin{table}
   \begin{center}
@@
     BKN040 & \\
-  \end{tabular} \\
+  \end{tabular} 
   \end{center}
-      {\bf Table S1:} A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by Chia et al (2012).
-\end{figure}
-
-\end{document}
+     \caption{ A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by \jri{cite chia and lemmon}    }
+\end{table}

### 2015-08-11 manuscript.tex
first attempt to include supp
updates

@@
 \usepackage[normalem]{ulem}
-\newcommand{\jri}[1]{\textcolor{red}{\emph{#1}} }
+\newcommand{\jri}[1]{\textcolor{red}{\scriptsize #1}}
 \newcommand{\citex}{\textcolor{red}{\bf CITE}}
 \newcommand{\X}{\textcolor{red}{\bf X}}
+\usepackage{multibib}
 
@@
 To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene \emph{tga1}, one of the few known instances of a hard sweep corresponding to a causal missense substitution in maize \cite{wang2015}.
-In the case of this known sweep, we observed a reduction in diversity surrounding the causative substitution to levels much lower than seen for synonymous substitutions (Figure S1) \jri{please add refs to link to supp.
-figs if possible}, confirming that this method should be informative if hard sweeps were the prevailing mode of selection in maize.
+In the case of this known sweep, we observed a reduction in diversity surrounding the causative substitution to levels much lower than seen for synonymous substitutions (Figure \ref{sFig:tga1}), confirming that this method should be informative if hard sweeps were the prevailing mode of selection in maize.
 Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize.
-When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score, putatively corresponding to genes undergoing the strongest and weakest purifying selection \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites (Figure S2).
+When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score, putatively corresponding to genes undergoing the strongest and weakest purifying selection \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites (Figure \ref{sFig:consUncons}).
 
@@
 
-Similarly, to complement our investigation of pairwise diversity surrounding substitutions, we studied patterns of singleton diversity surrounding synonymous and non-synonymous substitutions in maize (Figure S3).
+Similarly, to complement our investigation of pairwise diversity surrounding substitutions, we studied patterns of singleton diversity surrounding synonymous and non-synonymous substitutions in maize (Figure \ref{sFig:singleton}).
 A nearly identical pattern to that shown in Figure \ref{fig:hardSweeps} was observed.
@@
 H12 is sensitive to both hard and soft sweeps.
-Sites nearest to genes that displayed among the top 20\% of H12 values in either maize or teosinte were removed from our data set, and both pairwise and singleton diversity at remaining sites, as a function of the distance to the nearest gene, was calculated (Figure S4).
+Sites nearest to genes that displayed among the top 20\% of H12 values in either maize or teosinte were removed from our data set, and both pairwise and singleton diversity at remaining sites, as a function of the distance to the nearest gene, was calculated (Figure \ref{sFig:H12}).
 This subset of the data is no longer subject to the potentially confounding effects of soft-sweeps, and nearly identical patterns of pairwise and singleton diversity surrounding genes are observed.
@@
 cite chia for hapmap, but that we used higher-coverage of the teosinte from lemmon.
-if possible, include link to fastq or bamfiles used} Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results (Figure S5).
+if possible, include link to fastq or bamfiles used} Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results (Figure \ref{sFig:PCA}).
 Ultimately, 23 maize inbreds derived from a diverse assortment of landraces were selected for inclusion.
 Thirteen teosinte inbred lines, all members of  subspecies Z. \emph{mays}
-ssp. \emph{parviglumis}, were utilized. Included lines are listed in (Table S1).
+ssp. \emph{parviglumis}, were utilized. Included lines are listed in (Table \ref{sTab:list}).
 Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{chia2012}.\jri{did we use their bams?
@@
 \begin{acknowledgments}
-We are indebted to Graham Coop and Simon Aeshbacher for their constructive input during this study. Funding was provided by \textcolor{red}{NSF Grant number ZZZZZZZZZ}.
+We are indebted to Graham Coop and Simon Aeshbacher for their constructive input during this study. Funding was provided by NSF Plant Genome Research Project 1238014.
 \end{acknowledgments}
@@
 \bibliographystyle{pnas}
+
+\onecolumn
+\input{supportingInfo} 
 %\nocite{*}

### 2015-08-11 supportingInfo.tex
updates

@@
 \setcounter{figure}{0}
-\begin{figure}[h]
+
+\begin{figure}
   \begin{center}
@@
 Since this is only one gene, the large amount of noise compared to our average plots is expected.
-However, notice that diversity precisely at the causitive polymorphism is reduced and a recovery of diversity is observed away from that site.}
+However, notice that diversity precisely at the causitive polymorphism is reduced and a recovery of diversity is observed away from that site.
+\label{sFig:tga1}}
 \end{figure}
@@
 \caption{ Pairwise diversity surrounding synonymous and nonsynonymous
-  substitutions in maize at highly conserved (A) or unconserved (B) sites.  Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis.}
+  substitutions in maize at highly conserved (A) or unconserved (B) sites.
+Bootstrap-based 95\% confidence intervals are depicted via shading.
+Inset plots depict a larger range on the x-axis.
+\label{sFig:consUncons}}
 \end{figure}
@@
 \caption{ Singleton diversity surrounding synonymous and nonsynonymous
-  substitutions in maize.}
+  substitutions in maize. \label{sFig:singleton}}
 \end{figure}
@@
 Bootstrap-based 95\% confidence intervals are depicted via shading.
-Inset plots depict a smaller range on the x-axis.}
+Inset plots depict a smaller range on the x-axis.
+\label{sFig:H12}}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=.5\textwidth]{FigsAndFiles/tilPCA_july.pdf}\\
-    \includegraphics[width=.5\textwidth]{FigsAndFiles/bknPCA_july}\\
+  \begin{center}
+  \includegraphics[width=.75\textwidth]{FigsAndFiles/tilPCA_july.pdf}\\
+  \includegraphics[width=.75\textwidth]{FigsAndFiles/bknPCA_july.pdf}\\
+  \end{center}
   \caption{Principal component analysis of teosinte and maize individuals to ensure that no close relatives were inadvertantly included in our study.
@@
 {\bf D:} PC1 vs PC2 for all 23 maize individuals.
-\jri{need to fix letter label positions}}
+\jri{need to fix letter label positions} \label{sFig:PCA}}
 \end{figure}
@@
 
-\begin{table}
-  \begin{center}
+\begin{figure}
+%  \begin{center}
   \begin{tabular}{c|c}
@@
   \end{tabular} 
-  \end{center}
-     \caption{ A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by \jri{cite chia and lemmon}    }
-\end{table}
+%  \end{center}
+  \caption{ A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by \jri{cite chia and lemmon}   \label{sTab:list} }
+\end{figure}

### 2015-08-11 manuscript.tex
updates

@@
 \centering
-\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_June}
-\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_June}
+\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_Aug}
+\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_Aug}
 \caption{Pairwise diversity surrounding synonymous and non-synonymous
-  substitutions in maize and teosinte. \jri{a bit too much vertical white space} {\bf A:} Maize diversity; {\bf B:} Teosinte
+  substitutions in maize and teosinte. {\bf A:} Maize diversity; {\bf B:} Teosinte
 diversity. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis. \label{fig:hardSweeps}}
 \end{figure*}
-
-%%%%
-%%%% This outline no longer needed in-text because results headers follow
-%%%%
-% \jri{
-% {\bf Outline}
-% \begin{enumerate}
-% \item Diversity differs between genes, not genes
-% \item No hard sweeps
-% \item soft sweeps? (putting it here means if we find a lot we may have to redo stuff below on the subset of genes w/o evidence of soft sweeps.)
-% \item purifying selection differs ($\pi$)
-% \item Estimating the bneck
-% \item purifying selection (singletons)
-% \item fit the B curve
-% \item simulations
-% \end{enumerate}
-% }
 

### 2015-08-12 manuscript.tex
starting in on methods

--- 2015-08-01 manuscript.tex
@@
 
-\title{Domestication and the impacts of linked selection on the maize genome}
+\title{Demography and linked selection in wild and domesticated maize}
 \author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
@@
 \jri{not really.
-we need to cite the Corbett-Detig plos-bio paper here -- he looked at a lot of species including maize!} In this work, we demonstrate that selection against deleterious mutations leaves a pronounced signature on the maize genome, reducing diversity in and immediately around genes.
+we need to rephrase here and cite the Corbett-Detig plos-bio paper somewhere -- they looked at a lot of species including maize!} In this work, we demonstrate that selection against deleterious mutations leaves a pronounced signature on the maize genome, reducing diversity in and immediately around genes.
 We show how demography interacts with selection to impact genome-wide patterns of diversity, including the important observation that rapid population expansion can increase the efficiency of selection as much as a sudden population bottleneck can weaken it.
@@
   \subsection{BASH, R, and Python scripts}
-All analysis scripts are available in an online repository at \textcolor{red}{REPO ADDRESS HERE.}
+All scripts used for analysis are available in an online repository at \textcolor{red}{REPO ADDRESS HERE.}
 
 \subsection{Plant materials}
-Accessions studied were selected from the Maize HapMap2
-panel \cite{chia2012, lemmon2014}.
-\jri{this needs slight clarification.
-cite chia for hapmap, but that we used higher-coverage of the teosinte from lemmon.
-if possible, include link to fastq or bamfiles used} Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results (Figure S5).
-Ultimately, 23 maize inbreds derived from a diverse assortment of landraces were selected for inclusion.
-Thirteen teosinte inbred lines, all members of  subspecies Z. \emph{mays}
-ssp. \emph{parviglumis}, were utilized. Included lines are listed in (Table S1).
+We made use of published sequence of inbred accessions of teosinte (Z. \emph{mays} ssp. \emph{parviglumis}) and maize landraces from the Maize HapMap2 panel \cite{chia2012, lemmon2014}. 
+We removed \X teosinte and \X maize individuals that appeared as outliers in an initial principal component analysis  (Figure S5), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table S1).
+
+\subsection{Physical and genetic maps}
 Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{chia2012}.\jri{did we use their bams?
 I thought we used bwa bams, which are not listed and weren't used for hapmap2.
-if that's the case, we need to mention this difference.}
-
-\subsection{Interpolating genetic position}
-
-Physical positions were converted to genetic positions by interpolating from the NAM genetic map \cite{glaubitz2014}, \jri{does jeff's paper really have the NAM map?
-I didn't think so and quick look didn't find it.
-i thought this came from wallace?} which provides a 1 cM resolution for physical to genetic conversion.
-Within R \cite{R2014}, physical positions with corresponding genetic positions in the NAM map were used as anchors.
-Physical positions in our dataset without corresponding genetic position were assigned genetic positions by scaling the anchored genetic positions according to the physical distance between the unlabeled position and the flanking anchors.
+if that's the case, we need to mention this difference.
+either way, we need to include a link to the bamfiles too.}
+All analyses made use of uniquely mapping reads with mapping quality score $\geq 30$ and bases with base quality score $\geq 20$, adjusting quality scores around indels. 
+We converted physical coordinates to genetic coordinates via linear interpolation of the 1cM resolution NAM genetic map \cite{glaubitz2014}, \jri{i thought this should be wallace not glaubitz?
+need to specify what you do with physical positions before and after last marker on each chrom and markers, if any, that have physical and genetic order that disagree}.
 
 \subsection{Estimating the site frequency spectrum}
-To estimate individual  and joint site frequency spectra (SFS) for maize and teosinte from inbred lines, each inbred individual was treated as representing a single haplotype from its population.
-SFS were
-separately computed for genic and intergenic regions, as well as for
-the whole genome together. First, genic and intergenic regions were isolated using
-the biomaRt package \cite{durinck2009,durinck2005} of R \cite{R2014}. Genic
-regions were defined as DNA between the start and stop position of a \jri{you meant transcription start/stop not translation, right?}
-gene, while intergenic regions were required to be at least 5kb up- or
-down-stream from a gene start/stop. With regions defined, SFS were
-estimated with ANGSD \cite{korneliussen2014}. Individual population SFS were
-estimated using all positions observed in at least 80\% of the
-individuals in the population, and joint SFS were estimated using all
-positions observed in at least 80\% of individuals in both
-populations. Individuals were
-assumed to be fully inbred (-doSaf 2). Quality filters were
-employed such that reads with quality score below 30 and bases with
-quality score below 20 were discarded (-minMapq 30 and -minQ 20), as were reads that didn't map
-uniquely (-uniqueOnly 0). Quality scores around indels were adjusted
-as in Samtools (-baq 0). Genotype likelihoods were estimated using the samtools
-method (-GL 1). Major and minor alleles were inferred from the data
-(-doMaf 1). Because ANGSD cannot calculate a folded joint SFS, the
-maize reference genome was used for polarization and then unfolded spectra
-were folded using $\delta\alpha\delta{i}$ \cite{gutenkunst2009}.
+We estimated both the genome-wide site frequency spectrum (SFS) as well as a separate SFS for genic (within annotated transcript) and intergenic ($\geq 5kb$ from a transcript) regions. 
+We used the biomaRt package \cite{durinck2009,durinck2005} of R \cite{R2014} to parse annotations from the \X annotation \jri{include annotation version number and link to ENSEMBL annotation used}. 
+We estimated single population and joint SFS with the software ANGSD \cite{korneliussen2014}, including all positions with at least one aligned read in $\geq 80\%$ of samples in one or both populations.
+We assumed individuals were fully inbred and treated each line as a single haplotype.
+\jri{is doSaf -2 equivalent to using the inbreeding coefficient?
+in general i removed the options because they don't make sense to casual reader and the interested reader can find the script}
+Because ANGSD cannot calculate a folded joint SFS, we first polarized SNPs using the maize reference genome and then folded spectra using $\delta\alpha\delta{i}$ \cite{gutenkunst2009}.
 
 \subsection{Demographic inference}
-We used $\delta\alpha\delta{i}$ \cite{gutenkunst2009} to estimate the parameters of maize
-domestication based on diffusion approximation of the 2-dimensional
-SFS for maize and teosinte. Our first step was to specify a model, as
-shown in (Figure \ref{fig:bottleneck}). In words, our model began with an ancestral teosinte
-population with effective population size $N_a$ individuals. The
-ancestral population split into
-two distinct populations, one of maize and the other of teosinte,
-$T_b$ generations in the past. The maize population immediately shrank
-to size $N_b$, and then grew exponentially to size $N_{maize}$
-today. From the split until today, the teosinte effective population
-size remained constant. Each year, $M_{mt}$ individuals migrated from
-maize to teosinte, and $M_{tm}$ individuals migrated from teosinte to
-maize.The free parameters of our model were $N_a, N_b, N_{maize}, T_b,
-M_{mt}, \text{and } M_{tm} $. We implimented 1,000 $\delta\alpha\delta{i}$
-optimizations of this model using the ``$\delta\alpha\delta{i}$.Inference.optimize\_log\_fmin''
-optimizer. In $\delta\alpha\delta{i}$ optimizations,
-initial values for $N_b, N_{maize}, T_b,
-M_{mt}, \text{and } M_{tm}$, respectively, were randomly perturbed up
-to a factor of two from 0.02, 3, 0.04, 1e-5, and 1e-5. Lower
-bounds were set at 1e-7, 1e-7,  0, 1e-10, and
-1e-10, respecively, while upper bounds were respectively set at 2,
-200, 1, .001, and .001. $N_a$ did not have an initial value, upper, or
-lower bounds because it is estimated from equilibrium ancestral theta
-($\theta = 4N_a\mu$), which $\delta\alpha\delta{i}$ determines automatically regardless
-of model. Population sizes were converted from $\delta\alpha\delta{i}$'s
-units using a mutation rate estimate of 3e-8
-\cite{clark2005}. Among the 1,000 $\delta\alpha\delta{i}$ runs, parameters
-from the run with the lowest log-likelihood were taken as estimates of
-the parameters of the domestication process.
+We used the software $\delta\alpha\delta{i}$ \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS.
+We modeled an a teosinte of constant effective population size $N_a$ individuals that at time $T_b$ generations in the past gives rise to a maize population of size $N_b$ which grows exponentially to size $N_{maize}$ in the present (Figure \ref{fig:bottleneck})
+The model includes migration each generation of $M_{mt}$ individuals from maize to teosinte and $M_{tm}$ individuals from teosinte to maize.  
+We fixed $N_a$ using $\delta\alpha\delta{i}$'s estimation of the population mutation rate $\theta=4N_a\mu$ and a mutation rate of $3 \times 10^{-8}$ \cite{clark2005}. 
+We estimated all other parameters using 1,000 $\delta\alpha\delta{i}$ optimizations of the model using ``$\delta\alpha\delta{i}$.Inference.optimize\_log\_fmin'' optimizer \jri{is this last bit about the optimizer in your code?
+if so, i'd ditch} setting initial parameter values at $\frac{N_b}{N_a}=0.02$, $\frac{N_{maize}}{N_a}=3$, $\frac{T_b}{2N_a}=0.04$, and $\frac{M_{mt}}{N_c}=\frac{M_{tm}}{N_a}=10^{-5}$, where the maize population size $Nc$ at time $i$ generations in the past is $N_c=N_ae^{\frac{ln(\frac{N_{maize}}{N_a})(T_b-i)}{T_b}}$.
+\jri{make sure these are right (e.g.
+is it 2Na or 4Na for Tb?).
+note that we have to redo this in fractions or words because above you call Nb a population size when in reality we are estimating a proportion}
+We allowed initial values to be randomly purterbed by a factor of 2, and set upper and lower bounds such that $10^{-7} \eq \frac{N_b}{N_a} \leq 2$, $10^{-7} \leq \frac{N_{maize}}{N_a} \leq{200}$, $0 \leq \frac{T_b}{2N_a} \leq 1$, $10^{-10}\leq \frac{M_{tm}}{N_a}\leq 10^{-3}$, and $10^{-10}\leq \frac{M_{mt}}{N_c}\leq 10^{-3}$
+We report parameter estimates from the optimization run with the highest log-likelihood.
 
-Recent population expansion, as is observed for maize, most influences
-the abundance of rare alleles. However, our demographic parameter
-estimation using $\delta\alpha\delta{i}$ suffered from a relatively small sample size of
-23 maize individuals. Therefore, our approach likely underestimated
-current maize effective population size. To address this, we utilized
-a complimentary dataset of 4,021 maize landrace individuals collected
-from across the Americas \textcolor{red}{(SeeDs reference here)} to
-generate an independent estimate of
-$N_{maize}$. These individuals were genotyped using GBS
-\cite{elshire2011}, a protocol that suffers from a high rate of
-missing data \cite{beissinger2013}. To minimize biases imposed from
-missing data, we used R to isolate only those SNPs that
-were observed in at least 1,500 individuals and subsequently projected the
+We additionally made use of the large genotyping data set of more than 4,000 maize landraces \citex to calculate a singleton-base estimate of the modern maize effective population size.   
+We filtered these data to include only SNPs observed in $\geq 1,500$ individuals, and then projected the SFS down to a sample of 500 individuals. \jri{cite or mention how projection was done}. 
 SFS down to a sample of 500 individuals. Then, we utilized the singleton-based estimate of

### 2015-08-12 manuscript.tex

@@
 We made use of published sequence of inbred accessions of teosinte (Z. \emph{mays} ssp. \emph{parviglumis}) and maize landraces from the Maize HapMap2 panel \cite{chia2012, lemmon2014}. 
-We removed \X teosinte and \X maize individuals that appeared as outliers in an initial principal component analysis  (Figure S5), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table S1).
+We removed \X teosinte and \X maize individuals that appeared as outliers in an initial principal component analysis  (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table sTab:list}).
 

### 2015-08-13 manuscript.tex
more edits

@@
   the ancestral $N_e$, before recovering two at least 2.98
-  times as large as the ancestral $N_e$. \label{fig:bottleneck} }
+  times as large as the ancestral $N_e$ \jri{$M_{mt}$ and $tm$ are the reverse of what is described in text}. \label{fig:bottleneck} }
 \end{figure}
@@
 either way, we need to include a link to the bamfiles too.}
-All analyses made use of uniquely mapping reads with mapping quality score $\geq 30$ and bases with base quality score $\geq 20$, adjusting quality scores around indels. 
+All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following \cite{li2011statistical}.
 We converted physical coordinates to genetic coordinates via linear interpolation of the 1cM resolution NAM genetic map \cite{glaubitz2014}, \jri{i thought this should be wallace not glaubitz?
@@
 We assumed individuals were fully inbred and treated each line as a single haplotype.
-\jri{is doSaf -2 equivalent to using the inbreeding coefficient?
-in general i removed the options because they don't make sense to casual reader and the interested reader can find the script}
+\jri{i removed options details because they don't make sense to casual reader and the interested reader can find the script}
 Because ANGSD cannot calculate a folded joint SFS, we first polarized SNPs using the maize reference genome and then folded spectra using $\delta\alpha\delta{i}$ \cite{gutenkunst2009}.
@@
 We used the software $\delta\alpha\delta{i}$ \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS.
-We modeled an a teosinte of constant effective population size $N_a$ individuals that at time $T_b$ generations in the past gives rise to a maize population of size $N_b$ which grows exponentially to size $N_{maize}$ in the present (Figure \ref{fig:bottleneck})
-The model includes migration each generation of $M_{mt}$ individuals from maize to teosinte and $M_{tm}$ individuals from teosinte to maize.  
-We fixed $N_a$ using $\delta\alpha\delta{i}$'s estimation of the population mutation rate $\theta=4N_a\mu$ and a mutation rate of $3 \times 10^{-8}$ \cite{clark2005}. 
-We estimated all other parameters using 1,000 $\delta\alpha\delta{i}$ optimizations of the model using ``$\delta\alpha\delta{i}$.Inference.optimize\_log\_fmin'' optimizer \jri{is this last bit about the optimizer in your code?
-if so, i'd ditch} setting initial parameter values at $\frac{N_b}{N_a}=0.02$, $\frac{N_{maize}}{N_a}=3$, $\frac{T_b}{2N_a}=0.04$, and $\frac{M_{mt}}{N_c}=\frac{M_{tm}}{N_a}=10^{-5}$, where the maize population size $Nc$ at time $i$ generations in the past is $N_c=N_ae^{\frac{ln(\frac{N_{maize}}{N_a})(T_b-i)}{T_b}}$.
+We modeled a teosinte population of constant effective size $N_a$, that at time $T_b$ generations in the past gives rise to a maize population of size $N_b$ which grows exponentially to size $N_m$ in the present (Figure \ref{fig:bottleneck})
+The model includes migration of $M_{mt}$ individuals each generation from maize to teosinte and $M_{tm}$ individuals from teosinte to maize.  
+We fixed $N_a$ using $\delta\alpha\delta{i}$'s estimation of $\theta=4N_a\mu$ from the data and a mutation rate of $3 \times 10^{-8}$ \cite{clark2005}. 
+We estimated all other parameters using 1,000 $\delta\alpha\delta{i}$ optimizations, setting initial parameter values at $\frac{N_b}{N_a}=0.02$, $\frac{N_m}{N_a}=3$, $\frac{T_b}{2N_a}=0.04$, and $\frac{M_{mt}}{N_c}=\frac{M_{tm}}{N_a}=10^{-5}$, where the maize population size $Nc$ at time $i$ generations in the past is $N_c=N_ae^{\frac{ln(\frac{N_m}{N_a})(T_b-i)}{T_b}}$.
 \jri{make sure these are right (e.g.
@@
 note that we have to redo this in fractions or words because above you call Nb a population size when in reality we are estimating a proportion}
-We allowed initial values to be randomly purterbed by a factor of 2, and set upper and lower bounds such that $10^{-7} \eq \frac{N_b}{N_a} \leq 2$, $10^{-7} \leq \frac{N_{maize}}{N_a} \leq{200}$, $0 \leq \frac{T_b}{2N_a} \leq 1$, $10^{-10}\leq \frac{M_{tm}}{N_a}\leq 10^{-3}$, and $10^{-10}\leq \frac{M_{mt}}{N_c}\leq 10^{-3}$
+We allowed initial values to be randomly purterbed by a factor of 2, and set upper and lower bounds such that $10^{-7} \leq \frac{N_b}{N_a} \leq 2$, $10^{-7} \leq \frac{N_m}{N_a} \leq{200}$, $0 \leq \frac{T_b}{2N_a} \leq 1$, $10^{-10}\leq \frac{M_{tm}}{N_a}\leq 10^{-3}$, and $10^{-10}\leq \frac{M_{mt}}{N_c}\leq 10^{-3}$
 We report parameter estimates from the optimization run with the highest log-likelihood.
 
-We additionally made use of the large genotyping data set of more than 4,000 maize landraces \citex to calculate a singleton-base estimate of the modern maize effective population size.   
-We filtered these data to include only SNPs observed in $\geq 1,500$ individuals, and then projected the SFS down to a sample of 500 individuals. \jri{cite or mention how projection was done}. 
+We additionally made use of a large genotyping data set of more than 4,000 maize landraces \cite{Hearne2015} to calculate the modern maize $N_e$ from singleton counts \cite{fu1993}.
+We filtered these data to include only SNPs observed in $\geq 1,500$ individuals, and then projected the SFS down to a sample of 500 individuals.\jri{cite or mention how projection was done}. 
 SFS down to a sample of 500 individuals. Then, we utilized the singleton-based estimate of

### 2015-08-14 manuscript.tex
methods pretty darned reduced

@@
 \caption{Pairwise diversity surrounding synonymous and non-synonymous
-  substitutions in maize and teosinte. {\bf A:} Maize diversity; {\bf B:} Teosinte
-diversity. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis. \label{fig:hardSweeps}}
+  substitutions in maize and teosinte.
+{\bf A:} Maize diversity; {\bf B:} Teosinte diversity.
+Diversity axes show both absolute values and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution.
+\jri{is this correct or did i confuse things?}
+ Lines depict a loess curve (span of 0.01)  and shading represents bootstrap-based 95\% confidence intervals. Inset plots depict a larger range on the x-axis. \label{fig:hardSweeps}}
 \end{figure*}
@@
 \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_Downsampled_threeLines_manuscript.png}
-\caption{Relative level of diversity versus distance to the nearest
-  gene, in maize and teosinte. Two measures of diversity were
-  investigated. \textbf{A} displays pairwise
-  diversity, which is most influenced by intermediate frequency
-  alleles and therefore depicts more ancient evolutionary patterns,
-  and \textbf{B} depicts singleton diversity, influenced by rare
-  alleles and thus depicting evolutionary patterns in the recent
-  past. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a smaller range on the x-axis. \label{fig:purify}
+\caption{Relative diversity versus distance to nearest gene in maize and teosinte. 
+Shown are \textbf{A} pairwise nucleotide diversity and \textbf{B} singleton diversity.
+Relative diversity is calculated compared to the mean diversity in windows $\geq 0.01 cM$ or $\geq 0.02 cM$ from the nearest
+gene for pairwise diversity and singletons, respectively. 
+  Lines depict a loess curve (span of 0.01) and shading represents bootstrap-based 95\% confidence intervals. 
+  Inset plots depict a smaller range on the x-axis. \label{fig:purify}
   %\jri{should also add the singleton vs pi comparison within species as a supplemental figure.
@@
 We made use of published sequence of inbred accessions of teosinte (Z. \emph{mays} ssp. \emph{parviglumis}) and maize landraces from the Maize HapMap2 panel \cite{chia2012, lemmon2014}. 
-We removed \X teosinte and \X maize individuals that appeared as outliers in an initial principal component analysis  (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table sTab:list}).
+We removed \X teosinte and \X maize individuals that appeared as outliers in an initial principal component analysis  (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table \ref{sTab:list}).
 
@@
 I thought we used bwa bams, which are not listed and weren't used for hapmap2.
-if that's the case, we need to mention this difference.
-either way, we need to include a link to the bamfiles too.}
+if that's the case, we need to mention this difference and acknowledge Robert for providing those bams.
+either way, we need to include a link to the bamfiles too.
+somewhere here or above mention URL for Tripsacum bams.
+did we align those too or download someone else's?}
 All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following \cite{li2011statistical}.
@@
 We assumed individuals were fully inbred and treated each line as a single haplotype.
-\jri{i removed options details because they don't make sense to casual reader and the interested reader can find the script}
+\jri{i removed options details because they don't make sense to casual reader and the interested reader can find the script.
+}
 Because ANGSD cannot calculate a folded joint SFS, we first polarized SNPs using the maize reference genome and then folded spectra using $\delta\alpha\delta{i}$ \cite{gutenkunst2009}.
@@
 We fixed $N_a$ using $\delta\alpha\delta{i}$'s estimation of $\theta=4N_a\mu$ from the data and a mutation rate of $3 \times 10^{-8}$ \cite{clark2005}. 
-We estimated all other parameters using 1,000 $\delta\alpha\delta{i}$ optimizations, setting initial parameter values at $\frac{N_b}{N_a}=0.02$, $\frac{N_m}{N_a}=3$, $\frac{T_b}{2N_a}=0.04$, and $\frac{M_{mt}}{N_c}=\frac{M_{tm}}{N_a}=10^{-5}$, where the maize population size $Nc$ at time $i$ generations in the past is $N_c=N_ae^{\frac{ln(\frac{N_m}{N_a})(T_b-i)}{T_b}}$.
-\jri{make sure these are right (e.g.
+We estimated all other parameters using 1,000 $\delta\alpha\delta{i}$ optimizations and allowing initial values to be randomly perturbed by a factor of 2.  
+Initial values and upper and lower bounds can be found in table \X.
+\jri{this looked terrible in text, please make a supp.
+table.
+keep track of notation as e.g.
+we explain $N_b$ as population size of bneck, but in reality parameter we fit is $\frac{N_b}{N_a}$}
+%, setting initial parameter values at $\frac{N_b}{N_a}=0.02$, $\frac{N_m}{N_a}=3$, $\frac{T_b}{2N_a}=0.04$, and $\frac{M_{mt}}{N_c}=\frac{M_{tm}}{N_a}=10^{-5}$, where the maize population size $Nc$ at time $i$ generations in the past is $N_c=N_ae^{\frac{ln(\frac{N_m}{N_a})(T_b-i)}{T_b}}$.
+%\jri{make sure these are right (e.g.
 is it 2Na or 4Na for Tb?).
 note that we have to redo this in fractions or words because above you call Nb a population size when in reality we are estimating a proportion}
-We allowed initial values to be randomly purterbed by a factor of 2, and set upper and lower bounds such that $10^{-7} \leq \frac{N_b}{N_a} \leq 2$, $10^{-7} \leq \frac{N_m}{N_a} \leq{200}$, $0 \leq \frac{T_b}{2N_a} \leq 1$, $10^{-10}\leq \frac{M_{tm}}{N_a}\leq 10^{-3}$, and $10^{-10}\leq \frac{M_{mt}}{N_c}\leq 10^{-3}$
+%We allowed initial values to be , and set upper and lower bounds such that $10^{-7} \leq \frac{N_b}{N_a} \leq 2$, $10^{-7} \leq \frac{N_m}{N_a} \leq{200}$, $0 \leq \frac{T_b}{2N_a} \leq 1$, $10^{-10}\leq \frac{M_{tm}}{N_a}\leq 10^{-3}$, and $10^{-10}\leq \frac{M_{mt}}{N_c}\leq 10^{-3}$
 We report parameter estimates from the optimization run with the highest log-likelihood.
 
-We additionally made use of a large genotyping data set of more than 4,000 maize landraces \cite{Hearne2015} to calculate the modern maize $N_e$ from singleton counts \cite{fu1993}.
-We filtered these data to include only SNPs observed in $\geq 1,500$ individuals, and then projected the SFS down to a sample of 500 individuals.\jri{cite or mention how projection was done}. 
-SFS down to a sample of 500 individuals. Then, we utilized the singleton-based estimate of
-theta ($\theta = 4N_e\mu = \frac{S}{L}$) to estimate
-modern maize $N_e$ \cite{fu1993}, where $S$ is the total number of
-singletons and $L$ is the total length of DNA sequenced. Based on the
-GBS protocol implemented \cite{elshire2011}, $L$ is equal to the total
-number of SNPs in the dataset.
+We further made use of a large genotyping data set of more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize $N_e$ from singleton counts.
+We filtered these data to include only SNPs with data in $\geq 1,500$ individuals, and then projected the SFS down to a sample of 500 individuals.\jri{cite or mention how projection was done}. 
+We then estimated $N_e$ from the data assuming $\mu = 3 \times 10^{-8}$ \cite{clark2005} and the relation  $4N_e\mu = \frac{S}{L}$ \cite{fu1993}, where where $S$ is the total number of singleton SNPs and $L$ is the total number of SNPs in the dataset.
 
-\subsection{Evaluating diversity around substitutions}
-To investigate diversity around substitutions, maize and teosinte pairwise diversity was first
-calculated in 1,000 kb non-overlapping windows using ANGSD
-\cite{korneliussen2014}. This was performed separately for both maize and
-teosinte, using the same filters as employed for estimating the
-SFS. Next, SNPs and genotypes among maize, teosinte, and \emph{Tripsacum} were called. \emph{Tripsacum} bam files
-were downloaded from \textcolor{red}{(TRIPSICUM FILES)}, and then all SNPs with a
-p-value less than 1e-6 were called using ANGSD. Quality filters were
-as the same as before, and genotypes were only called when the
-posterior probability was above 0.95. From the set of called SNPs and
-genotypes, substitutions between maize and tripsicum, as well as
-between teosinte and \emph{Tripsacum} were identified using R \cite{R2014} as all positions with
-no more than 20\% missing data for which every maize or teosinte
-allele differed from the observed \emph{Tripsacum} allele. At each
-substitution, effects were estimated using the ensembl variant effects
-predictor \cite{mclaren2010}.
+Finally, \jri{MSMC text here}
 
-For each diversity window with at least 100 bps observed, the distance from the window center to the
-nearest synonymous and missense substitution was
-computed. Then, following the methods of \cite{hernandez2011}, a loess
-curve was plotted for diversity values against the distance to the
-nearest synonymous or nonsynonymous substitution. A span of 0.01 was
-utilized.
+\subsection{Diversity}
+We made use of the software ANGSD \cite{korneliussen2014} for diversity calculations and genotype calling. 
+We calculated diversity statistics in maize and teosinte in 100bp non-overlapping windows using filters as described above for the SFS.
+\jri{this said 1,000 kb which is 1Mb, but should say 100bp right?}
+We used allele counts to estimate the number of singleton polymorphisms in each window, and used binomial sampling to create a second maize data set down-sampled to have the same number of samples as teosinte.
+We called genotypes in maize, teosinte, and \emph{Tripsacum} at sites with a SNP p-value $<10^{-6}$ and when the genotype posterior probability $>0.95$.
+We identified substitutions in maize and teosinte as all sites with a fixed difference with \emph{Tripsacum} and $\leq 20\%$ missing data. 
+Substitutions were classified as synonymous, missense, or noncoding using the ensembl variant effects predictor \cite{mclaren2010}.
+For each window with $\geq 100bp$ of data we computed the genetic distance between the window center and the nearest synonymous and missense substitution as well as the genetic distance to the nearest gene transcript.
+\jri{this said gene center, but we're not using gene center anymore, right?}
 
-To verify that this approach is robust to the criticisms of \cite{enard2014}, which involve reduced power at highly conserved sites, we applied the method to subsets of sites that included only positions nearest to the  most- or least-conserved 10\% of genes data.
+%Teosinte diversity is
+%generally higher then maize diversity. Therefore, to enable comparisons between
+%the reduction of diversity around genes in maize and teosinte, a
+%neutral measure of pairwise and singleton diversity for each taxa was
+%estimated. For pairwise diversity, this neutral measure was defined according to
+%mean pairwise diversity at windows greater than 0.01 cM from the nearest
+%gene. For singletons, both maize and teosinte still showed reduced
+%diversity at a distance of 1 cM, so neutral diversity in this case was
+%defined as mean singleton diversity at positions 0.02 cM or greater
+%from genes. Then, pairwise and singleton diversity at each window was
+%standardized by dividing by the corresponding neutral
+%measure. Separately for pairwise and singleton diversity in maize and
+%teosinte, . Significant differences were assessed by taking 100 bootstrap
+%samples from each set of diversity windows and re-fitting the cubic smoothing spline to each. Then, the
+%2.5\% and 97.5\% quantiles of values along the bootstrapped splines
+%were identified.
+
+%To verify that this approach is robust to reduced power at highly conserved sites \cite{enard2014}, we applied the method to subsets of sites that included only positions nearest to the  most- or least-conserved 10\% of genes data.
 This was achieved by assigning every gene a GERP score corresponding to the average score observed across it \cite{rodgers2015}.
-Next, genes ing the top and bottom 10\% quantiles of GERP score were isolated.
+Next, genes in the top and bottom 10\% quantiles of GERP score were isolated.
 Two datasets were created, each including all diversity windows who's nearest gene was in the high or low GERP score sets.
@@
 
-\subsection{Evaluating diversity around genes}
-Two types of diversity surrounding genes were investigated. The first
-was pairwise diversity in 1kb windows, as described previously. The
-second was singleton diversity in 1kb windows. Singletons represent
-the rarest class of alleles that this dataset can identify, and
-collectively demonstrate the most recent patterns of evolution. Minor
-allele frequencies were estimated with ANGSD \cite{korneliussen2014} using the
-same quality filters previously described. Then, the number of
-singletons in each non-overlapping 1kb window was calculated with R
-for both maize and teosinte \cite{R2014}. For maize, we also generated a parallel set of downsampled
-singleton data, for which binomial sampling within R was conducted
-based on allele frequencies at each site, to generate a singleton
-psuedo dataset of sample size 13, equivalent to our sample size for
-teosinte. BiomaRt \cite{durinck2009, durinck2005} was then used to identify
-the center of each gene. Next, the distance from each diversity window center
-to the nearest gene center was computed.
+%In addition  next Two types of diversity surrounding genes were investigated. The first
+%was pairwise diversity in 1kb windows, as described previously. The
+%second was singleton diversity in 1kb windows. Singletons represent
+%the rarest class of alleles that this dataset can identify, and
+%collectively demonstrate the most recent patterns of evolution. Minor
+%allele frequencies were estimated with ANGSD \cite{korneliussen2014} using the
+%same quality filters previously described. Then, the number of
+%singletons in each non-overlapping 1kb window was calculated with R
+%for both maize and teosinte \cite{R2014}. For maize, we also generated a parallel set of downsampled
+%singleton data, for which binomial sampling within R was conducted
+%based on allele frequencies at each site, to generate a singleton
+%psuedo dataset of sample size 13, equivalent to our sample size for
+%teosinte. BiomaRt \cite{durinck2009, durinck2005} was then used to identify
+%the center of each gene. Next, the distance from each diversity window center
+%to the nearest gene center was computed.
 
-Teosinte diversity is
-generally higher then maize diversity. Therefore, to enable comparisons between
-the reduction of diversity around genes in maize and teosinte, a
-neutral measure of pairwise and singleton diversity for each taxa was
-estimated. For pairwise diversity, this nuetral measure was defined according to
-mean pairwise diversity at windows greater than 0.01 cM from the nearest
-gene. For singletons, both maize and teosinte still showed reduced
-diversity at a distance of 1 cM, so neutral diversity in this case was
-defined as mean singleton diversity at positions 0.02 cM or greater
-from genes. Then, pairwise and singleton diversity at each window was
-standardized by dividing by the corresponding neutral
-measure. Separately for pairwise and singleton diversity in maize and
-teosinte, cubic smoothing splines were fit to
-describe diversity levels according to the distance to the nearest
-gene. Significant differences were assessed by taking 100 bootstrap
-samples from each set of diversity windows and re-fitting the cubic smoothing spline to each. Then, the
-2.5\% and 97.5\% quantiles of values along the bootstrapped splines
-were identified.
-
-Additionally, we performed a parallel implementation of the above analysis after excluding sites potentially experiencing positive selection in the form of either hard- or soft-sweeps.
-To isolate these sites, we performed genome-scans separately for maize and teosinte based on the H12 statistic \cite{garud2015}.
-SNPs within each taxa were called in ANGSD \cite{korneliussen2014} using the previously specified SNP calling parameters.
-H12 was computed with a window size of 200 SNPs and a jump of 25 SNPs.
-H12 windows in the highest 20\% for maize or teosinte were identified and considered as regions most likely to be under positive selection.
-Finally, levels of pairwise and singleton diversity vs.
-distance to gene was computed using only sites that were not nearest to selected sites in either maize or teosinte.
+\subsection{Selection scan}
+We scanned the genome to identify sites that have experienced recent positive selection using the H12 statistic \cite{garud2015} in sliding windows of 200 SNPs with a step of 25 SNPs.
 

### 2015-08-14 manuscript.tex
first section of results

@@
   non-genic regions of the genome}
-Previous research of the maize domestication process has relied upon
-observations drawn from genic DNA \cite{wright2005,wang1999,eyre1998}. Our data were generated through whole genome sequencing,
-which eliminated this constraint. Importantly, we observed substantial
-differences in patterns of diversity between genic and non-genic regions
-of the genome for both maize and teosinte. For maize, mean pairwise
-diversity ($\pi$) within genes was significantly lower than at
-positions at least 5 kb away from genes (0.00668 within, 0.00691 away, p$<$2e-44). The same
-pattern of significantly greater diversity outside of genes was observed in
-teosinte, but to a larger extent. For teosinte we observed $\pi$
-within genes to be 0.0088 and at least 5 kb from genes it was 0.0115
-(p$\approx$0). These observations suggest that genes are not evolving
-neutrally in maize or teosinte. Instead, some form of selection is likely
-reducing diversity within genes. Additionally, these observations
-suggest that demographic inference, which relies on the assumption of
-neutrally evolving DNA, will be more accurate if it is based on
-observations from non-genic genetic material.
+Our reanalysis of the maize HapMap 2 data \cite{chia2012} is consistent with earlier findings that patterns of diversity differ between genic and non-genic regions of the genome in both maize and teosinte \cite{hufford2012} (Figure \ref{fig:diversity}).
+In maize, mean pairwise diversity ($\pi$) within genes was significantly lower than at positions at least 5 kb away from genes (0.00668 vs 0.00691, $p<2\times 10^{-44}$). 
+Diversity differences in teosinte are even more pronounced (0.0088 vs. 0.0115, $p\approx 0$). 
+Differences were also apparent in the site frequency spectrum, with mean Tajima's D positive in genic regions in both maize (0.4) and teosinte (0.013) but negative outside of genes (-0.087 in maize and -0.25 in teosinte, $p<X$ for both comparisons).
+These observations suggest that diversity in genes is not evolving neutrally, but instead reduced by the impacts of selection on linked sites.
 
-\subsection{Hard sweeps do not shape maize (or teosinte) diversity}
-A mutation that is immediately beneficial and positively selected leaves a classical hard
-sweep signature in the genome, whereby surrounding genetic diversity
-is reduced as the haplotype where the mutation first
-arose increases in frequency to fixation. The prevalence of such hard
-sweeps was evaluated by comparing diversity
-surrounding non-synonymous (specifically missense) and synonymous substitutions in maize and
-teosinte since divergence from \emph{Tripsacum}, roughly 1-1.2 million years before
-present \cite{ross2009}. For both taxa, no difference in diversity
-surrounding these classes of substitutions was identified (Figure
-\ref{fig:hardSweeps}). This observation suggests that hard sweeps are not a primary form of
-selection for either maize or teosinte.
+\begin{figure}
+\centering
+\includegraphics[width=.5\textwidth]{FigsAndFiles/SFS_and_Tajima.png}
+\caption{The folded SFS and Tajimas D were calculated both inside and outside of genes for maize and teosinte.
+In both taxa, a dearth of rare alleles was observed within genes relative to outside of genes, which led to higher values of Tajima's D outside of genes than inside.{\bf A:} Folded SFS for maize; {\bf B:} Folded SFS for teosinte; {\bf C:} Maize Tajima's D; {\bf D:} Teosinte Tajima's D.
+\label{fig:diversity} \jri{skip the SFS and make this a panel plot of pi and TajD in maize and teosinte} }
+\end{figure}
 
-To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene \emph{tga1}, one of the few known instances of a hard sweep corresponding to a causal missense substitution in maize \cite{wang2015}.
+\subsection{Hard sweeps do not explain diversity differences}
+Selection acting to increase the frequency of a new beneficial mutation will leave a signature of reduced diversity at surrounding linked sites \citex.
+To evaluate whether patterns of such ``hard sweeps'' could explain observed differences in diversity between genic and non-genic regions of the genome, we compared diversity around missense and synonymous substitutions between \emph{Tripsacum} and either maize or teosinte (Figure
+\ref{fig:hardSweeps}).
+If a proportion of missense mutations have been fixed due to hard sweeps, diversity around these substitutions should be lower than around synonymous substitutions. 
+We observe this pattern around the causative amino acid substitution in the the domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005origin, wang2015}.
+Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte, suggesting that hard sweeps do not play a major role in patterning diversity in genes.
+Previous analyses in humans have suggested that this approach is limited by fact that a higher proportion of nonsynonymous substitutions will be found in genes under weak purifying selection and thus higher genetic diversity \cite{enard2014}.
+To address this concern we took advantage of a measure of evolutionary constraint using genomic evolutionary rate profile (GERP) scores \cite{davydov2010} calculated across the maize genome \cite{rodgers2015}.
+We re-analyzed substitutions in subsets of genes with the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection  (Figure \ref{sFig:consUncons}).
+As expected, we see higher overall diversity around substitutions in genes under weak purifying selection, but we still see no difference between synonymous and missense substitutions in either subset of the data.
+
+%To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene \emph{tga1}, one of the few known instances of a hard sweep corresponding to a causal missense substitution in maize \cite{wang2015}.
 In the case of this known sweep, we observed a reduction in diversity surrounding the causative substitution to levels much lower than seen for synonymous substitutions (Figure \ref{sFig:tga1}), confirming that this method should be informative if hard sweeps were the prevailing mode of selection in maize.
-Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize.
+%Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize.
 When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score, putatively corresponding to genes undergoing the strongest and weakest purifying selection \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites (Figure \ref{sFig:consUncons}).
@@
 \caption{Pairwise diversity surrounding synonymous and non-synonymous
-  substitutions in maize and teosinte.
-{\bf A:} Maize diversity; {\bf B:} Teosinte diversity.
-Diversity axes show both absolute values and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution.
+  substitutions in {\bf A} maize and {\bf B} teosinte.
+Axes show both absolute diversity values and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution.
 \jri{is this correct or did i confuse things?}
@@
 to sites at least 5 kb from gene start/stop positions.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=.5\textwidth]{FigsAndFiles/SFS_and_Tajima.png}
-\caption{The folded SFS and Tajimas D were calculated both inside and outside of genes for maize and teosinte.
-In both taxa, a dearth of rare alleles was observed within genes relative to outside of genes, which led to higher values of Tajima's D outside of genes than inside.{\bf A:} Folded SFS for maize; {\bf B:} Folded SFS for teosinte; {\bf C:} Maize Tajima's D; {\bf D:} Teosinte Tajima's D.
-\label{fig:sfs} }
-\end{figure}
 
@@
 
-\section{Post-domestication expansion is apparent in patterns of singleton diversity}
+\subsection{Post-domestication expansion is apparent in patterns of singleton diversity}
 Motivated by the rapid post-domestication expansion in maize that is evident from our demographic analysis, we conducted an analysis of genome-wide patterns of singleton diversity.
@@
 \subsection{Little positive selection on new genic mutations}
-Our findings indicate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize.
+\jri{mention divergence time from \emph{Tripsacum}, roughly 1-1.2 million years before present \cite{ross2009}.}Our findings indicate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize.
 More specifically, we observe that positive selection on new genic mutations in maize is not common compared to other drivers of diversity.
@@
 \subsection{Demography of domestication}
+\jri{add note that demography inference will require regions far from genes, compare to \cite{wright2005,wang1999,eyre1998}.
+cite some theory or empirical paper showing this to be the case in humans/drosophila}
 Since we observed that classical hard sweeps do not explain patterns of maize diversity, it is perhaps unsurprising that we instead found evidence of demographic history contributing substantially to genetic diversity.

### 2015-08-14 manuscript.tex
tweaks

@@
 We observe this pattern around the causative amino acid substitution in the the domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005origin, wang2015}.
-Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte, suggesting that hard sweeps do not play a major role in patterning diversity in genes.
+Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte.
 Previous analyses in humans have suggested that this approach is limited by fact that a higher proportion of nonsynonymous substitutions will be found in genes under weak purifying selection and thus higher genetic diversity \cite{enard2014}.
@@
 As expected, we see higher overall diversity around substitutions in genes under weak purifying selection, but we still see no difference between synonymous and missense substitutions in either subset of the data.
+Taken together, these data suggest hard sweeps do not play a major role in patterning genic diversity in either maize or teosinte.
 

### 2015-08-14 manuscript.tex
tweaks
almost through results

@@
 Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte.
-Previous analyses in humans have suggested that this approach is limited by fact that a higher proportion of nonsynonymous substitutions will be found in genes under weak purifying selection and thus higher genetic diversity \cite{enard2014}.
-To address this concern we took advantage of a measure of evolutionary constraint using genomic evolutionary rate profile (GERP) scores \cite{davydov2010} calculated across the maize genome \cite{rodgers2015}.
-We re-analyzed substitutions in subsets of genes with the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection  (Figure \ref{sFig:consUncons}).
-As expected, we see higher overall diversity around substitutions in genes under weak purifying selection, but we still see no difference between synonymous and missense substitutions in either subset of the data.
-Taken together, these data suggest hard sweeps do not play a major role in patterning genic diversity in either maize or teosinte.
 
-%To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene \emph{tga1}, one of the few known instances of a hard sweep corresponding to a causal missense substitution in maize \cite{wang2015}.
-In the case of this known sweep, we observed a reduction in diversity surrounding the causative substitution to levels much lower than seen for synonymous substitutions (Figure \ref{sFig:tga1}), confirming that this method should be informative if hard sweeps were the prevailing mode of selection in maize.
-%Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize.
-When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score, putatively corresponding to genes undergoing the strongest and weakest purifying selection \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites (Figure \ref{sFig:consUncons}).
-
-\begin{figure*}[!htb]
+\begin{figure*}
 %\vspace*{.05in}
@@
 
-\subsection{Pairwise diversity is heavily influenced by purifying selection}
+Previous analyses have suggested that this approach may have limited power because a higher proportion of nonsynonymous substitutions will be found in genes under weak purifying selection and thus with higher genetic diversity \cite{enard2014}.
+To address this concern, we took advantage of genomic evolutionary rate profile (GERP) scores \cite{davydov2010}, a measure of evolutionary constraint, calculated across the maize genome \cite{rodgers2015}.
+We re-analyzed substitutions in subsets of genes with the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection  (Figure \ref{sFig:consUncons}).
+As expected, we see higher diversity around substitutions in genes under weak purifying selection, but we still see no difference between synonymous and missense substitutions in either subset of the data.
+Taken together, these data suggest hard sweeps do not play a major role in patterning genic diversity in either maize or teosinte.
 
-Purifying selection refers to the situation where deleterious mutations arising in a population are continuously selected against.
-When this form of selection is operating, it can serve to reduce genetic variability at linked neutral sites, a phenomenon often called background selection \cite{charlesworth1993}.
-Purifying and background selection lead to lower diversity within genes and other functional sites relative to neutral regions.
-We investigated purifying selection in maize and teosinte by evaluating the average magnitude of reduced diversity within genes and recovery away from genes in both taxa.
-To enable comparisons between maize and teosinte, we standardized by neutral levels within each taxa, where neutral levels were defined as mean diversity at positions at least 0.01 cM distal from genes.
-For both maize and teosinte, pairwise diversity was reduced within genes and a recovery of diversity away from genes was observed.
-Interestingly, a stronger reduction of diversity and slower recovery was observed for teosinte than for maize, implying that purifying selection has left a more pronounced signature on pairwise diversity in the teosinte genome (Figure \ref{fig:purify}).
+\subsection{Diversity is strongly influenced by purifying selection}
 
-\begin{figure*}[!htb]
+Selection can also reduce diversity in functional regions of the genome via removal of deleterious mutations, a process known to as purifying or background selection \cite{charlesworth1993}.
+We investigated purifying selection in maize and teosinte by evaluating the reduction of diversity within genes.
+Pairwise diversity is strongly reduced within genes for both maize and teosinte (Figure \ref{fig:purify}A) but recovers quickly at sites outside of genes, consistent with the low levels of linkage disequilibrium generally observed in maize \citex.
+The reduction in relative diversity is more pronounced in teosinte, however, reaching lower levels in genes and occurring over a wider region.
+\jri{make sure we have tense consistent throughout.
+is/was etc.}
+We interpret this difference to imply that purifying selection has left a more pronounced signature in the teosinte genome .
+\jri{not sure this is quite right.
+is ``more pronounced signature'' same as ``ne is bigger and selection is stronger''?
+do we want to say this differently?}
+
+\begin{figure*}
 %\vspace*{.05in}
@@
   Inset plots depict a smaller range on the x-axis. \label{fig:purify}
-  %\jri{should also add the singleton vs pi comparison within species as a supplemental figure.
-if $\pi$/singleton curves {\bf don't} match, will need to explain this suggests teosinte not in equilibrium.
-good justification i think for msmc model, which we then use to say ``gee whiz, not perfect equilibrium but who cares because the big interesting demography is in maize.''}
   }
@@
 \subsection{Demography of maize domestication}
-To explore whether differences in purifying selection between maize
-and teosinte can be explained by demographic processes, we
-first estimated the parameters of maize domestication using large-scale
-sequencing data involving 23 inbred maize
-landraces and 13 teosinte inbred lines included in the HapMap 2 panel
-\cite{chia2012}. The maize
-lines were collected from across the Americas and the teosinte lines
-came from central Mexico. Before estimating demography, we compared
-the site frequency spectrum (SFS) in genic and non-genic regions, and
-observed substantial differences in the evolution of these classes of
-sites. For both maize and teosinte, the SFS within genes showed a
-dearth of low-frequency alleles and Tajima's D
-\cite{tajima1989} was therefore shifted
-to more positive values, as previously shown (Figure \ref{fig:purify}). This is consistent with the aforementioned purifying
-selection and indicates that genic regions are not evolving
-neutrally. Therefore, for demographic modeling we restricted analysis
-to sites at least 5 kb from gene start/stop positions.
+To explore whether differences in the efficacy of purifying selection between maize and teosinte can be explained by demographic processes, we estimated the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}).
+The most likely model estimates an ancestral $\theta$ of 0.0147 per bp, which translates to an effective population size of , which approximately matches a previous independent estimate \cite{eyre1998}.
+Assuming a mutation rate of $N_a \approx 123,000$ individuals given the mutation rate \cite{clark2005}.
+The maize population splits from teosinte $\approx 15,000$ generations in the paste with an initial size of only $\approx 5\%$ of ancestral $N_a$. 
+Following domestication we estimate considerable gene flow between the populations, estimating $M_{tm} = 1.35$  migrants per generation from teosinte to maize and $M_{mt} = 1.72$ migrants from maize to teosinte.
+Our model posits exponential growth in maize following its split from teosinte, and estimates a final modern effective population size of $N_m \approx 370,00$.
 
-We used diffusion approximation as implemented in
-$\delta\alpha\delta{i}$ \cite{gutenkunst2009} to find the domestication parameters that best explain
-the joint site frequency spectrum of maize and teosinte.  The model we optimized began with an
-equilibrium ancestral population of size $N_a$
-splitting into separate maize and teosinte populations $T_B$ generations in the past. Moving forward in
-time from the split, teosinte maintained the ancestral population size, while
-maize experienced an immediate effective population size change to
-$N_b$ individuals, followed by exponential growth to size
-$N_m$. Additionally, since the split $M_{tm}$ individuals migrated
-from teosinte into maize and $M_{mt}$ individuals migrated from maize
-to teosinte.
-
-The most likely model suggested an ancestral $\theta$ ($4N\mu$) per base pair of
-0.014734, which approximately matches a previous independent estimate \cite{eyre1998}. Assuming a mutation rate of $\mu =
-3 \times 10^{-8}$ \cite{clark2005}, this suggests an effective
-population size of $N_a = 122,783$ individuals, a population split
-$T_B = 15,523$ generations in the past, a maize bottleneck
-effective populations size of $N_b = 6,455$ individuals (5.26\% of
-$N_a$), a modern maize effective population size of $N_m = 366,973$
-individuals, $M_{tm} = 1.35$ migrants per generation from teosinte to
-maize, and $M_{mt} = 1.72$ migrants per generation from maize to
-teosinte (Figure \ref{fig:bottleneck}). We note that a population
-split over 15 thousand generations before present precedes estimates
-from
-archaeological data, which suggest maize domestication began
-no more than 10,000 years before present \cite{smith1995}. This could reflect teosinte population structure that was present before
-domestication. Also, note that the genetic time of the population
-split must precede physiological changes that could be identified
-archaelogically.   Additionally, because recent expansion is
-most evidenced by rare alleles, and since our limited sample size provides low
-power to detect rare alleles, we expect that the estimate of  $N_m = 366,973$, or $\sim 3N_a$, is likely an underestimate.
-
-Therefore, we utilized a complementary data set of 4,021 maize land-race individuals collected
-from across the Americas \textcolor{red}{(SeeDs reference here)} to
-obtain a less downward-biased estimate of the current maize effective
-population size. According to a singleton-based estimate of $\theta$
-\cite{fu1993}, which was utilized since rapid post-bottleneck expansion
-implies that rare alleles will most accurately reflect current maize
-demography, we obtained the estimate $N_e =
-  992,713.5$. Although less biased than the previously mentioned
-estimate, we note that ascertainment bias of the
-genotyping platform used to generate these data again biases this
-estimate downward.
-
-\begin{figure}[!htb]
+\begin{figure}
 %\vspace*{.05in}
@@
 \includegraphics[width=.4\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
-\caption{Parameters of domestication as estimated by $\delta\alpha\delta{i}$. Notably,
-  the maize effective population size ($N_e$) during the domestication
-  bottleneck appears to have consisted of approximately 5.26\%
-  the ancestral $N_e$, before recovering two at least 2.98
-  times as large as the ancestral $N_e$ \jri{$M_{mt}$ and $tm$ are the reverse of what is described in text}. \label{fig:bottleneck} }
+\caption{Parameter estimates for a simple bottleneck model of maize domestication. See methods for details. \label{fig:bottleneck} }
 \end{figure}
+
+Because our relatively small sample size limits our ability to characterize the rare variants that are most informative of population expansion, we took advantage of a complementary data set of more than 4,000 maize landraces collected from across the Americas \cite{Hearne2015} to estimate the modern maize effective population size from low frequency variants.
+This analysis yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
 
@@
 
+\jri{add note here that We note that a population split over 15 thousand generations before present precedes estimates from archaeological data, which suggest maize domestication began no more than 10,000 years before present \cite{smith1995}.
+This could reflect teosinte population  structure that was present before domestication. 
+Also, note that the genetic time of the population split must precede physiological changes that could be identified archaelogically. }
+
 After the initial bottleneck, our analysis suggests that maize $N_e$ expanded to reach levels much greater than that of teosinte.
-Our WGS-based estimate shows that the $N_e$ of landrace maize is at least 3X that of ancestral teosinte, while our singleton-based estimate from the the SeeDs project data \textcolor{red}{(citation)} implies an $N_e$ of close to 1 million individuals, or $\sim$8X that of ancestral teosinte.
-From the perspective of a new mutation, both of these estimates are biased downward, since they represent past reductions in diversity due to the aforementioned bottleneck.
+Our WGS-based estimate shows that the $N_e$ of landrace maize is at least 3X that of ancestral teosinte, while our singleton-based estimate from the the SeeDs project data \textcolor{red}{(citation)} implies an $N_e$ of close to 1 million individuals, or $\approx 8X$ that of ancestral teosinte.
+\jri{note ascertainment of GBS biases it downward} From the perspective of a new mutation, both of these estimates are biased downward, since they represent past reductions in diversity due to the aforementioned bottleneck.
 A back-of-the-envelope calculation of modern $N_e$ can be made by assuming there are 47.9 million ha of landrace maize in production \cite{cimmyt1999}, 42,000 plants per ha for open pollinated maize varieties \cite{van2010}, and conservatively that no more than one in 1,000 plants contribute gametes to the next generation.
@@
 
-\subsection{Demography strongly impacts efficiency of purifying selection}
+\subsection{Demography influences the efficiency of purifying selection}
 The observation that maize pairwise diversity is less impacted by the distance to the nearest gene than is teosinte pairwise diversity is reasonable from a long-term evolutionary standpoint.
@@
 \subsection{Demographic inference}
-We used the software $\delta\alpha\delta{i}$ \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS.
+We used the software $\delta\alpha\delta{i}$ \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS, using only sites $>5 kb$ from a gene to ameliorate the effects of linked selection.
 We modeled a teosinte population of constant effective size $N_a$, that at time $T_b$ generations in the past gives rise to a maize population of size $N_b$ which grows exponentially to size $N_m$ in the present (Figure \ref{fig:bottleneck})
 The model includes migration of $M_{mt}$ individuals each generation from maize to teosinte and $M_{tm}$ individuals from teosinte to maize.
+\jri{$M_{mt}$ and $tm$ are the reverse of what are in figure}
 We fixed $N_a$ using $\delta\alpha\delta{i}$'s estimation of $\theta=4N_a\mu$ from the data and a mutation rate of $3 \times 10^{-8}$ \cite{clark2005}. 
@@
 keep track of notation as e.g.
-we explain $N_b$ as population size of bneck, but in reality parameter we fit is $\frac{N_b}{N_a}$}
-%, setting initial parameter values at $\frac{N_b}{N_a}=0.02$, $\frac{N_m}{N_a}=3$, $\frac{T_b}{2N_a}=0.04$, and $\frac{M_{mt}}{N_c}=\frac{M_{tm}}{N_a}=10^{-5}$, where the maize population size $Nc$ at time $i$ generations in the past is $N_c=N_ae^{\frac{ln(\frac{N_m}{N_a})(T_b-i)}{T_b}}$.
-%\jri{make sure these are right (e.g.
-is it 2Na or 4Na for Tb?).
-note that we have to redo this in fractions or words because above you call Nb a population size when in reality we are estimating a proportion}
-%We allowed initial values to be , and set upper and lower bounds such that $10^{-7} \leq \frac{N_b}{N_a} \leq 2$, $10^{-7} \leq \frac{N_m}{N_a} \leq{200}$, $0 \leq \frac{T_b}{2N_a} \leq 1$, $10^{-10}\leq \frac{M_{tm}}{N_a}\leq 10^{-3}$, and $10^{-10}\leq \frac{M_{mt}}{N_c}\leq 10^{-3}$
+we explain $N_b$ as population size of bneck, but in reality parameter we fit is $\frac{N_b}{N_a}$.
+my stab at notation is $\frac{N_b}{N_a}=0.02$, $\frac{N_m}{N_a}=3$, $\frac{T_b}{2N_a}=0.04$, and $\frac{M_{mt}}{N_c}=\frac{M_{tm}}{N_a}=10^{-5}$, where the maize population size $Nc$ at time $i$ generations in the past is $N_c=N_ae^{\frac{ln(\frac{N_m}{N_a})(T_b-i)}{T_b}}$ but if you use these make sure they are right (e.g.
+is it 2Na or 4Na for Tb?).}
 We report parameter estimates from the optimization run with the highest log-likelihood.
@@
 
-%Teosinte diversity is
-%generally higher then maize diversity. Therefore, to enable comparisons between
-%the reduction of diversity around genes in maize and teosinte, a
-%neutral measure of pairwise and singleton diversity for each taxa was
-%estimated. For pairwise diversity, this neutral measure was defined according to
-%mean pairwise diversity at windows greater than 0.01 cM from the nearest
-%gene. For singletons, both maize and teosinte still showed reduced
-%diversity at a distance of 1 cM, so neutral diversity in this case was
-%defined as mean singleton diversity at positions 0.02 cM or greater
-%from genes. Then, pairwise and singleton diversity at each window was
-%standardized by dividing by the corresponding neutral
-%measure. Separately for pairwise and singleton diversity in maize and
-%teosinte, . Significant differences were assessed by taking 100 bootstrap
-%samples from each set of diversity windows and re-fitting the cubic smoothing spline to each. Then, the
-%2.5\% and 97.5\% quantiles of values along the bootstrapped splines
-%were identified.
-
-%To verify that this approach is robust to reduced power at highly conserved sites \cite{enard2014}, we applied the method to subsets of sites that included only positions nearest to the  most- or least-conserved 10\% of genes data.
-This was achieved by assigning every gene a GERP score corresponding to the average score observed across it \cite{rodgers2015}.
-Next, genes in the top and bottom 10\% quantiles of GERP score were isolated.
-Two datasets were created, each including all diversity windows who's nearest gene was in the high or low GERP score sets.
-Finally, Loess fits of diversity vs.
-distance to gene were computed from sites within each GERP score.
-
-%In addition  next Two types of diversity surrounding genes were investigated. The first
-%was pairwise diversity in 1kb windows, as described previously. The
-%second was singleton diversity in 1kb windows. Singletons represent
-%the rarest class of alleles that this dataset can identify, and
-%collectively demonstrate the most recent patterns of evolution. Minor
-%allele frequencies were estimated with ANGSD \cite{korneliussen2014} using the
-%same quality filters previously described. Then, the number of
-%singletons in each non-overlapping 1kb window was calculated with R
-%for both maize and teosinte \cite{R2014}. For maize, we also generated a parallel set of downsampled
-%singleton data, for which binomial sampling within R was conducted
-%based on allele frequencies at each site, to generate a singleton
-%psuedo dataset of sample size 13, equivalent to our sample size for
-%teosinte. BiomaRt \cite{durinck2009, durinck2005} was then used to identify
-%the center of each gene. Next, the distance from each diversity window center
-%to the nearest gene center was computed.
-
 \subsection{Selection scan}
@@
 
-%JRI IDEAS:
-% look at SFS of SNPs by GERP score in maize vs teosinte. shift to lower freq. should be seen at lower GERP scores in maize
-% discuss our estimates of deleterious mutation rate from B stuff.  these suggest only ~25% of mutations are deleterious, which contrasts results from a standard DFE
-% do a standard DFE! this gives us the proportion of mutations in categories of Ns. Can use Eyre-Walker software.
-% i think we do need to say something about load. add up GERP scores like Li?
+\color{red}
+{\bf Notes:}
+\begin{itemize}
+
+ \item should also add the singleton vs pi comparison within species as a supplemental figure.
+if $\pi$/singleton curves {\bf don't} match, will need to explain this suggests teosinte not in equilibrium.
+good justification i think for msmc model, which we then use to say ``gee whiz, not perfect equilibrium but who cares because the big interesting demography is in maize.''
+
+ \end{itemize}
+\color{black}
 

### 2015-08-14 manuscript.tex
pi plots

@@
 \centering
-\includegraphics[width=.5\textwidth]{FigsAndFiles/SFS_and_Tajima.png}
-\caption{The folded SFS and Tajimas D were calculated both inside and outside of genes for maize and teosinte.
-In both taxa, a dearth of rare alleles was observed within genes relative to outside of genes, which led to higher values of Tajima's D outside of genes than inside.{\bf A:} Folded SFS for maize; {\bf B:} Folded SFS for teosinte; {\bf C:} Maize Tajima's D; {\bf D:} Teosinte Tajima's D.
+\includegraphics[width=.5\textwidth]{FigsAndFiles/Pi_and_Tajima.png}
+\caption{Pairwise diversity (Pi) and Tajimas D were calculated both inside and outside of genes for maize and teosinte.
+In both taxa, a dearth of rare alleles was observed within genes relative to outside of genes, which led to higher values of Tajima's D outside of genes than inside.{\bf A:} Histogram of Pi in 1 kb windows in maize; {\bf B:} Histogram of Pi in 1 kb windows in teosinte; {\bf C:} Histogram of Tajima's D in 1 kb windows in maize; {\bf D:} Histogram of Tajima's D in 1 kb windows in teosinte.
 \label{fig:diversity} \jri{skip the SFS and make this a panel plot of pi and TajD in maize and teosinte} }

### 2015-08-14 manuscript.tex
more deletions, yay

@@
 is/was etc.}
-We interpret this difference to imply that purifying selection has left a more pronounced signature in the teosinte genome .
-\jri{not sure this is quite right.
-is ``more pronounced signature'' same as ``ne is bigger and selection is stronger''?
-do we want to say this differently?}
+
+Our initial comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation. \jri{need a citation?}
+Such ``soft sweeps'', however, would still be expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes.
+To test this possibility, we performed a genome-wide scan for selection using a method expected to be reasonably sensitive to both hard and soft sweeps \cite{garud2015}. 
+Even after removing genes in the top 20\% of the H12 statistic used to identify targets of selection, however, qualitative differences between maize and teosinte in patterns of diversity in and near genes remained unchanged (Figure \ref{sFig:H12}).
+
+We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.
 
@@
 \subsection{Post-domestication expansion is apparent in patterns of singleton diversity}
-Motivated by the rapid post-domestication expansion in maize that is evident from our demographic analysis, we conducted an analysis of genome-wide patterns of singleton diversity.
-Singleton diversity measures the abundance of alleles that are only observed in one individual among the sample.
-As a class, singleton alleles depict the most recent patterns of evolution, but
+Motivated by the rapid post-domestication expansion of maize evident from our demographic analysis, we investigated whether patterns of diversity at low frequency --- and thuse younger --- polymorphisms would show similar results in comparisons of maize and teosinte.
+We thus repeated our analysis of diversity
+ an analysis of genome-wide patterns of diversity using alleles present in only a single individual in the sample. 
+These ingleton diversity measures the abundance of alleles that are only observed in one individual among the sample. As a class, singleton alleles depict the most recent patterns of evolution, but
 they also have the lowest effect on pairwise diversity. Therefore, unlike
@@
 This further demonstrates that even the most recent selection patterns in maize are not dominated by hard sweeps.
-
-\subsection{Purifying selection results are robust to soft-sweeps}
-As a category, instances of soft-sweeps, or cases of selection on alleles that have already reached intermediate frequency before becoming advantageous, are difficult to identify \cite{wilson2014}.
-However, this form of selection is likely to be frequent during crop domestication \cite{innan2004}.
-We were concerned that the disparate patterns of pairwise and singleton diversity in maize and teosinte, which seem to reflect differences in the efficiency of purifying selection due to demographic history, may instead reflect differences in the location and abundance of soft-sweeps.
-Therefore, we re-analyzed patterns of pairwise and singleton diversity in maize and teosinte including only a subset of sites that are least likely to correspond to soft-sweeps.
-To achieve this, we performed a genome-wide scan for soft or hard sweeps in maize or teosinte based on the H12 statistic \cite{garud2015}.
-H12 is sensitive to both hard and soft sweeps.
-Sites nearest to genes that displayed among the top 20\% of H12 values in either maize or teosinte were removed from our data set, and both pairwise and singleton diversity at remaining sites, as a function of the distance to the nearest gene, was calculated (Figure \ref{sFig:H12}).
-This subset of the data is no longer subject to the potentially confounding effects of soft-sweeps, and nearly identical patterns of pairwise and singleton diversity surrounding genes are observed.
-This provides further evidence that the patterns of diversity we observe in maize as compared to teosinte are the result of rapid population expansion as opposed to differing selective patterns in the domesticated and wild taxa.
 
@@
 Substitutions were classified as synonymous, missense, or noncoding using the ensembl variant effects predictor \cite{mclaren2010}.
-For each window with $\geq 100bp$ of data we computed the genetic distance between the window center and the nearest synonymous and missense substitution as well as the genetic distance to the nearest gene transcript.
-\jri{this said gene center, but we're not using gene center anymore, right?}
+For each window with $\geq 100bp$ of data we computed the genetic distance between the window center and the nearest synonymous and missense substitution as well as the genetic distance to the center of the nearest gene transcript.
 

### 2015-08-14 manuscript.tex
new fig 1

@@
 \includegraphics[width=.5\textwidth]{FigsAndFiles/Pi_and_Tajima.png}
-\caption{Pairwise diversity (Pi) and Tajimas D were calculated both inside and outside of genes for maize and teosinte.
-In both taxa, a dearth of rare alleles was observed within genes relative to outside of genes, which led to higher values of Tajima's D outside of genes than inside.{\bf A:} Histogram of Pi in 1 kb windows in maize; {\bf B:} Histogram of Pi in 1 kb windows in teosinte; {\bf C:} Histogram of Tajima's D in 1 kb windows in maize; {\bf D:} Histogram of Tajima's D in 1 kb windows in teosinte.
-\label{fig:diversity} \jri{skip the SFS and make this a panel plot of pi and TajD in maize and teosinte} }
+\caption{Pairwise diversity (Pi) and Tajimas D were calculated inside and outside of genes for maize and teosinte.
+In both taxa, a dearth of rare alleles was observed within genes relative to outside of genes, which led to higher values of Tajima's D outside of genes than inside.{\bf A:} Mean and standard deviation of pi for 1 kb genic and non-genic windows of maize; {\bf B:} Mean and standard deviation of pi for 1 kb genic and non-genic windows of teosinte  {\bf C:} Histogram of Tajima's D in 1 kb windows in maize; {\bf D:} Histogram of Tajima's D in 1 kb windows in teosinte.
+\label{fig:diversity} }
 \end{figure}

### 2015-08-15 manuscript.tex
almost there

@@
 To explore whether differences in the efficacy of purifying selection between maize and teosinte can be explained by demographic processes, we estimated the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}).
-The most likely model estimates an ancestral $\theta$ of 0.0147 per bp, which translates to an effective population size of , which approximately matches a previous independent estimate \cite{eyre1998}.
-Assuming a mutation rate of $N_a \approx 123,000$ individuals given the mutation rate \cite{clark2005}.
+The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ individuals given the mutation rate \cite{clark2005}.
 The maize population splits from teosinte $\approx 15,000$ generations in the paste with an initial size of only $\approx 5\%$ of ancestral $N_a$. 
-Following domestication we estimate considerable gene flow between the populations, estimating $M_{tm} = 1.35$  migrants per generation from teosinte to maize and $M_{mt} = 1.72$ migrants from maize to teosinte.
-Our model posits exponential growth in maize following its split from teosinte, and estimates a final modern effective population size of $N_m \approx 370,00$.
+Following domestication we estimate considerable gene flow between the populations: $M_{tm} = 1.35$  migrants per generation from teosinte to maize and $M_{mt} = 1.72$ migrants from maize to teosinte.
+\jri{do you get these numbers by multiplying estimates by Na and modern N_maize?
+if so we should probably report the proportions not the absolute number of inds.}
+After its split from teosinte our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,00$.
 
@@
 
-\subsection{Post-domestication expansion is apparent in patterns of singleton diversity}
-Motivated by the rapid post-domestication expansion of maize evident from our demographic analysis, we investigated whether patterns of diversity at low frequency --- and thuse younger --- polymorphisms would show similar results in comparisons of maize and teosinte.
-We thus repeated our analysis of diversity
- an analysis of genome-wide patterns of diversity using alleles present in only a single individual in the sample. 
-These ingleton diversity measures the abundance of alleles that are only observed in one individual among the sample. As a class, singleton alleles depict the most recent patterns of evolution, but
-they also have the lowest effect on pairwise diversity. Therefore, unlike
-pairwise diversity, patterns of singleton
-diversity reflect recent patterns of evolution. Since our sample size
-for maize (n=23) was larger than teosinte (n=13), maize singletons were
-analyzed directly as well as down-sampled to reflect the sample size of
-teosinte. For singletons, our definition of
-neutral diversity was defined as diversity at positions at
-least 0.02 cM distal from genes. When
-evaluating the data in this manner, a very different pattern
-emerged compared to what we saw with respect to pairwise diversity. Maize singleton diversity was at least as reduced as teosinte singleton
-diversity near genes, but recovered more slowly
-(Figure \ref{fig:purify}), implying that in the
-recent past maize has been more influenced by purifying selection than has
-teosinte.
+\subsection{Population expansion leads to stronger purifying selection in modern maize}
+Motivated by the rapid post-domestication expansion of maize evident from our demographic analysis, we investigated whether patterns of diversity at low frequency --- and thus on average younger --- polymorphisms would show similar results in comparisons of maize and teosinte.
+We thus repeated our analysis of diversity using alleles present in only a single individual in the sample.
+In direct contrast to patterns observed for pairwise nucleotide diversity (Figure \ref{fig:purify}A), diversity of singletons was more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B).
+This pattern remained after downsampling our maize data to account for differences in sample size. 
+Taken together, these results suggest that demographic history and population size change have impacted the efficacy of purifying selection during maize evolution.
 
-Together, these findings suggest that demographic history has
-a strong influence on the effect of purifying selection. Historically,
-teosinte has had a larger population size than maize, and only
-recently has maize population size overcome that of teosinte. Since
-the efficacy of purifying selection scales with population size,
-these results likely reflect changes in $N_e$ more than they reflect
-underlying changes in selection pressure.
-
-Similarly, to complement our investigation of pairwise diversity surrounding substitutions, we studied patterns of singleton diversity surrounding synonymous and non-synonymous substitutions in maize (Figure \ref{sFig:singleton}).
-A nearly identical pattern to that shown in Figure \ref{fig:hardSweeps} was observed.
-This further demonstrates that even the most recent selection patterns in maize are not dominated by hard sweeps.
+Patterns of diversity around substitutions, however
 

### 2015-08-15 manuscript.tex
done with results

@@
 In direct contrast to patterns observed for pairwise nucleotide diversity (Figure \ref{fig:purify}A), diversity of singletons was more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B).
-This pattern remained after downsampling our maize data to account for differences in sample size. 
-Taken together, these results suggest that demographic history and population size change have impacted the efficacy of purifying selection during maize evolution.
+This pattern remained after downsampling our maize data to account for differences in sample size. \jri{we should test singleton curve for genes not in 20\% highest H12 also!}
 
-Patterns of diversity around substitutions, however
+These results suggest that demographic history and population size change have impacted the efficacy of purifying selection during maize evolution.
+In contrast, analysis of singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) provides a nearly identical pattern to that shown in Figure \ref{fig:hardSweeps}, providing little support for a substantial increase in the number or strength of hard sweeps occurring in maize.
 

### 2015-08-15 manuscript.tex
discussion too

@@
 
-For instance, supposed neutrally-evolving DNA is often used to estimate historical demographic
-parameters of a population such as effective population size, structure, and expansion
-history \cite{luikart2003, gutenkunst2009}. However, researchers have
-called into question whether or not there are reasonable approaches
-for identifying neutral regions of the genome, since the effects of
-selection can be wide-ranging \cite{li2012, slotte2014}. For example, in
-\emph{Drosophila} it appears that the majority of the genome is
-impacted by the effects of selection through linkage
-\cite{sella2009}. A natural next question, therefore, is how can
-demographic parameters be estimated independently of
-selective parameters? Human researchers have attempted to address this
-problem by limiting analyses to only sites far from genes
-\cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is
-difficult to be certain that sites even in gene-poor regions of the
-genome are not influenced by linked selection. Ultimately, an understanding
-of which sites have the potential to be adaptive, and how these affect
-genome-wide patterns of variability through linkage, is required for
-reasonable demographic inference.
-
 Maize represents an excellent organism to study these
@@
 teosinte; and 3) precisely estimate the parameters of the
-maize domestication bottleneck. We show that our third objective, estimating the
-parameters of domestication, is not possible without first completing
- objectives one and two, which demonstrate that as in humans \cite{gazave2014}, the
-majority of maize non-genic DNA may reasonably be treated as
-neutral. We achieve these objectives by utilizing
-whole-genome-sequence information from 23 maize and 13 teosinte lines
-sequenced as part of the Maize HapMap 2 panel \cite{chia2012}.
+maize domestication bottleneck.  \cite{chia2012}.
 
 \section{Results}
-\subsection{Patterns of diversity differ between genic and
-  non-genic regions of the genome}
-Our reanalysis of the maize HapMap 2 data \cite{chia2012} is consistent with earlier findings that patterns of diversity differ between genic and non-genic regions of the genome in both maize and teosinte \cite{hufford2012} (Figure \ref{fig:diversity}).
+\subsection{Patterns of diversity differ between genic and  non-genic regions of the genome}
+To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we reanalyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 project \cite{chia2012}.
+We find broad differences in genic and non-genic diversity consistent with earlier results  \cite{hufford2012} (Figure \ref{fig:diversity}).  
 In maize, mean pairwise diversity ($\pi$) within genes was significantly lower than at positions at least 5 kb away from genes (0.00668 vs 0.00691, $p<2\times 10^{-44}$). 
@@
 The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ individuals given the mutation rate \cite{clark2005}.
-The maize population splits from teosinte $\approx 15,000$ generations in the paste with an initial size of only $\approx 5\%$ of ancestral $N_a$. 
+The maize population splits from teosinte $\approx 15,000$ generations in the past with an initial size of only $\approx 5\% $ of ancestral $N_a$. 
 Following domestication we estimate considerable gene flow between the populations: $M_{tm} = 1.35$  migrants per generation from teosinte to maize and $M_{mt} = 1.72$ migrants from maize to teosinte.
-\jri{do you get these numbers by multiplying estimates by Na and modern N_maize?
+\jri{do you get these numbers by multiplying estimates by Na and modern Nmaize?
 if so we should probably report the proportions not the absolute number of inds.}
-After its split from teosinte our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,00$.
+After its split from teosinte our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,000$.
 
@@
 
+Finally, we reconstructed the demographic history of maize and teosinte using a recently developed coalescent approach \citex that does not require conditioning on a particular model (Figure \X).
+Though this analysis does highlight a potential weakness with our assumption of a constant size for teosinte, it nonetheless is broadly consistent with our simple model,   identifying a clear domestication bottleneck followed by rapid population expansion (Figure \X).
+
 \subsection{Population expansion leads to stronger purifying selection in modern maize}
@@
 We thus repeated our analysis of diversity using alleles present in only a single individual in the sample.
-In direct contrast to patterns observed for pairwise nucleotide diversity (Figure \ref{fig:purify}A), diversity of singletons was more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B).
-This pattern remained after downsampling our maize data to account for differences in sample size. \jri{we should test singleton curve for genes not in 20\% highest H12 also!}
+In direct contrast to patterns observed for pairwise nucleotide diversity (Figure \ref{fig:purify}A), diversity of singletons was more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
+\jri{we should test singleton curve for genes not in 20\% highest H12 also!}
+Finally, while direct comparison of curves for pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, these too reveal much stronger differences in maize (Figure \X).
+\jri{if we see qualitative rather than quantitative difference in these too we should mention}
 
@@
 \section{Discussion}
-\subsection{Little positive selection on new genic mutations}
-\jri{mention divergence time from \emph{Tripsacum}, roughly 1-1.2 million years before present \cite{ross2009}.}Our findings indicate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize.
-More specifically, we observe that positive selection on new genic mutations in maize is not common compared to other drivers of diversity.
-This finding does not exclude the possibility that hard sweeps have taken place at non-genic sites, such as in those in enhancer regions.
-One known example of positive selection on a non-genic mutation involves the \emph{tb1} locus of maize, which is one of the best characterized examples of positive selection on a ``domestication gene'' in any crop \cite{clark2006}.
+\subsection{Hard sweeps do not shape genome-wide diversity in maize}
+
+\color{blue}
+
+no strong evidence for hard sweeps, even considering constraint (GERP).
+although this is not super powerful test as misses sweeps on noncoding and only has power if substantial proportion of missense sites are swept, much previous evidence argues against hard sweeps in maize.
+popgen studies of local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} show importance of standing variation. 
+even in domestication, few fixed differences between maize and teo \cite{hufford2012}, and of domestication loci that have been well studied from a popgen standpoint only one shows hard sweep \cite{wang2015} while there are multiple examples of standing variation \cite{studer2011, wills2013} or selection on multiple mutations \cite{wills2013}.
+\jri{maybe others?
+barren stalk 1 (gallavotti 2004)?
+ramosa 1 (sigmon 2010)?
+}
+
+this lack of hard sweeps patterning diversity differs from drosophila \citex and capsella \citex and many species with high Ne \jri{cite corbet-detig}.
+\jri{add something about corbet-dettig results for maize}
+but is consistent with humans \cite{hernandez2011}
+discussion of whether we expect hard vs soft in maize \cite{messer2013population}.
+$\theta_b = 2N\mu_b$ with $mu_b$ being mutation rate of beneficial alleles of selection coefficient s. 
+if $\theta_b > \frac{1}{log{Ns_b}}$ then adaptation from multiple mutations should be common. 
+if  $\theta_b > \frac{1}{log{2Ns_b}}$ then adaptation from standing neutral variation should be common, or if $\theta > \frac{1}{log(1+\frac{2Ns_b}{2Ns_d + 1}}$ for standing deleterious variation with selection coefficient $s_d$.
+\jri{i think this latter assumes codominance} codominant?)
+
+%Our findings indicate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize.
+%One known example of positive selection on a non-genic mutation involves the \emph{tb1} locus of maize, which is one of the best characterized examples of positive selection on a ``domestication gene'' in any crop \cite{clark2006}.
 However, the maize \emph{tb1} allele was already present in teosinte before domestication \cite{studer2011}, so this this example is in agreement with our finding that hard sweeps are rare.
@@
 
-Unfortunately, our ability to accurately identify soft sweeps, and particularly to distinguish them from hard sweeps, remains limited \cite{innan2004,messer2013}.
+%Unfortunately, our ability to accurately identify soft sweeps, and particularly to distinguish them from hard sweeps, remains limited \cite{innan2004,messer2013}.
 However, we implemented a scan based on the H12 statistic, which is designed to identify both hard and soft sweeps with reasonable power \cite{garud2015}.
@@
 Similarly, previous studies scanning for evidence of positive selection in maize \cite{hufford2012} may have picked up primarily evidence of soft sweeps.
-In domesticated species, artificial selection beginning at the onset of domestication represents a drastic shift in selective pressure.
-One can easily imagine a scenario in which previously neutral or slightly deleterious variants that were segregating in the population when this shift took place suddenly became beneficial, leading to soft sweeps.
-As appealing as this explanation is, however, our data do not speak to the abundance of soft sweeps, beyond demonstrating that they are one possibility for the lack of observed hard sweeps.
 
-We should additionally note that our observations do not exclude the possibility of infrequent hard sweeps having taken place during maize evolution.
+%We should additionally note that our observations do not exclude the possibility of infrequent hard sweeps having taken place during maize evolution.
 In fact, the maize locus \emph{tga1} that has been shown to correspond to a hard sweep at an amino-acid changing mutation \cite{wang2015}.
@@
 
+\color{black}
+
 \subsection{Demography of domestication}
-\jri{add note that demography inference will require regions far from genes, compare to \cite{wright2005,wang1999,eyre1998}.
-cite some theory or empirical paper showing this to be the case in humans/drosophila}
-Since we observed that classical hard sweeps do not explain patterns of maize diversity, it is perhaps unsurprising that we instead found evidence of demographic history contributing substantially to genetic diversity.
-Our estimate of the demographic history of maize, which is the first estimate of maize demographic history to utilize whole-genome-sequence (WGS) data, allowed us to disentangle the confounding parameters of bottleneck strength and bottleneck intensity.
-This confounding limited previous domestication estimates \cite{wright2005}.
-By utilizing the two-dimensional site frequency spectrum (SFS) between maize and teosinte for inference \cite{gutenkunst2009} we estimate that the maize effective population size during the bottleneck was approximately 5\% that of teosinte.
-A bottleneck of this magnitude may have enabled moderately deleterious alleles to reach intermediate frequency in maize, while also increasing the probability that strongly deleterious alleles segregating at low frequency in teosinte were purged from the maize gene pool, due to elevated drift as is predicted during a domestication bottleneck \cite{tajima1989}.
 
-\jri{add note here that We note that a population split over 15 thousand generations before present precedes estimates from archaeological data, which suggest maize domestication began no more than 10,000 years before present \cite{smith1995}.
-This could reflect teosinte population  structure that was present before domestication. 
-Also, note that the genetic time of the population split must precede physiological changes that could be identified archaelogically. }
+Although many other authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied  exclusively genic polymoprhisms and have made a number of limiting assumptions about the demographic model.
+We show that diversity within genes has been strongly reduced by the effects of linked selection, making genic data inappropriate for estimating demography.
+Because levels of diversity alone do not allow independent inference of the strength and duration of a bottleneck \cite{tenaillon2004selection}, previous investigations have instead estimated the ratio of these parameters.
+We have overcome this limitations through the use of the full joint SFS of maize and teosinte and consideration of a more realistic model of exponential growth, enabling us to estimate for the first time the strength of the domestication bottleneck.
+Given that modern maize exhibits $\approx 80\%$ of the diversity of teosinte, our estimate that  the effective population size of the initial maize  population was only $\approx 5\%$ of the ancestral teosinte population seems surprisingly low.
+The impact of a strong domestication bottleneck is likely ameliorated to some degree by  maize-teosinte gene flow and the rapid post-domestication growth estimated in our model.
 
-After the initial bottleneck, our analysis suggests that maize $N_e$ expanded to reach levels much greater than that of teosinte.
-Our WGS-based estimate shows that the $N_e$ of landrace maize is at least 3X that of ancestral teosinte, while our singleton-based estimate from the the SeeDs project data \textcolor{red}{(citation)} implies an $N_e$ of close to 1 million individuals, or $\approx 8X$ that of ancestral teosinte.
-\jri{note ascertainment of GBS biases it downward} From the perspective of a new mutation, both of these estimates are biased downward, since they represent past reductions in diversity due to the aforementioned bottleneck.
-A back-of-the-envelope calculation of modern $N_e$ can be made by assuming there are 47.9 million ha of landrace maize in production \cite{cimmyt1999}, 42,000 plants per ha for open pollinated maize varieties \cite{van2010}, and conservatively that no more than one in 1,000 plants contribute gametes to the next generation.
-This calculation implies a modern $N_e \geq  2$ billion individuals, and incorporates no individuals from hybrid breeding programs.
-If for some reason this back-of-the-envelope calculation is an overestimate, even by several orders of magnitude, it is clear that the post-domestication expansion in maize $N_e$ is enormous.
-Patterns of singleton diversity in maize, as discussed in more detail in the next section, demonstrate that this recent expansion has a notable impact on the maize genome.
+Another surprising result from our model is the estimation of the timing of domestication.
+Although we estimate the maize population split $\approx 5,000$ years earlier than previous molecular \cite{matsuoka2002} or archaelogical \cite{piperno2009starch} estimates, we do not interpret these results as strongly conflicting.
+We first note that the timing of the genetic split between populations likely preceded the anatomical changes that can be identified in the archaeological record. 
+We also caution, however, that our estimate may be inflated by population structure as our geographically diverse sample may include populations diverged from those that gave rise to maize.
+
+Although we estimate that the modern effective size of maize is three times that of teosinte, the small size of our sample dramatically reduces our power to estimate rapid population growth \citex.  
+A better estimate of the effective size of modern maize comes from our analysis of singleton diversity in a large sample of maize landraces \cite{Hearne2015}.
+These data suggest a modern effective size of nearly 1 million, nearly eight times higher than modern teosinte.
+Nonetheless, missing data, sample size, and ascertainment bias in the genotyping data all bias this estimate downward as well.
+We can, however, get reasonable estimates of the census size of modern maize.
+There are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, with a lower limit of 25,000 individuals planter per hectare \cite{baden2001culture}.
+Using this lower bound on planting density, ignoring all industrial maize agriculture, and even assuming the effective size is only 0.1\% of the census size, this still implies a modern effective population size of more than one billion.
+Clearly, the effective size of modern maize is extremely large.
 
@@
 Therefore, our observation shows that purifying selection in maize has not purged deleterious alleles, or the neutral alleles they are linked to, as effectively as in teosinte.
+\jri{shorten this!}
 
@@
 This observation is very clearly demonstrated by the fact that singleton diversity in maize is more impacted by the distance to the nearest gene than it is in teosinte, as was shown in Figure \ref{fig:purify}.
+\jri{shorten!}
 
@@
 This could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{meghji1984}; if deleterious alleles tend to be recessive and are particularly frequent, they will have ample opportunities to display their phenotypes in inbreds.
-Our results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously select
+Our results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously selected.
+\jri{this only works for weakly selected mutations but is an interesting point.
+need to also point out that this means new mutations in maize likely much more strongly selected than new mutations in teosinte, perhaps with implications for breeding?}
 
-Ultimately, we have shown that purifying selection in maize has operated very differently than purifying selection in teosinte.
-This observation, along with the potential for soft sweeps, appear to explain the phenotypic divergence \jri{pretty hard to argue that purifying selection explains phenotypic divergence!} between maize and teosinte much more completely than does positive selection generating hard-sweeps at protein-coding mutations.
  Importantly, our estimation of the parameters of the maize domestication bottleneck contribute to the understanding of how the demography of crop domestication can impact crop diversity.
@@
 Complementing this knowledge, our results demonstrate that the rapid expansion experienced by many crops after domestication can also have a profound influence on patterns of diversity, and the effects of this expansion should be accounted for as important contributors to long-term evolution.
+\jri{this is sorta fluff and doesn't say much}
+
+\jri{need paragraph also talking about how this is important for considering impact of linked selection.
+cite some recent human papers that are considering both demography and linked selection, but also a number of plant papers that e.g.
+estimate the DFE from current data or just use the sny/nonsyn graphs and ignore recent demographic history.
+is capsella a good example?
+}
 
@@
 did we align those too or download someone else's?}
-All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following \cite{li2011statistical}.
-We converted physical coordinates to genetic coordinates via linear interpolation of the 1cM resolution NAM genetic map \cite{glaubitz2014}, \jri{i thought this should be wallace not glaubitz?
+All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following Li \emph{et al.} \cite{li2011statistical}.
+We converted physical coordinates to genetic coordinates via linear interpolation of the 1cM resolution NAM genetic map \cite{glaubitz2014}, \jri{i think this is be wallace not glaubitz?
 need to specify what you do with physical positions before and after last marker on each chrom and markers, if any, that have physical and genetic order that disagree}.
@@
 
-\color{red}
-{\bf Notes:}
-\begin{itemize}
-
- \item should also add the singleton vs pi comparison within species as a supplemental figure.
-if $\pi$/singleton curves {\bf don't} match, will need to explain this suggests teosinte not in equilibrium.
-good justification i think for msmc model, which we then use to say ``gee whiz, not perfect equilibrium but who cares because the big interesting demography is in maize.''
-
- \end{itemize}
-\color{black}
-
 \end{materials}
@@
 \bibliographystyle{pnas}
-
 \onecolumn

### 2015-08-19 manuscript.tex
Working through Jeff's comments

@@
 Diversity differences in teosinte are even more pronounced (0.0088 vs. 0.0115, $p\approx 0$). 
-Differences were also apparent in the site frequency spectrum, with mean Tajima's D positive in genic regions in both maize (0.4) and teosinte (0.013) but negative outside of genes (-0.087 in maize and -0.25 in teosinte, $p<X$ for both comparisons).
-These observations suggest that diversity in genes is not evolving neutrally, but instead reduced by the impacts of selection on linked sites.
+Differences were also apparent in the site frequency spectrum, with mean Tajima's D positive in genic regions in both maize (0.4) and teosinte (0.013) but negative outside of genes (-0.087 in maize and -0.25 in teosinte, $p\approx 0$ for both comparisons).
+These observations suggest that diversity in genes is not evolving neutrally, but instead is reduced by the impacts of selection on linked sites.
 
 \begin{figure}
-\centering
-\includegraphics[width=.5\textwidth]{FigsAndFiles/Pi_and_Tajima.png}
+\begin{center}
+  \includegraphics[width=.5\textwidth] {FigsAndFiles/Pi_and_Tajima.png}
+\end{center}
 \caption{Pairwise diversity (Pi) and Tajimas D were calculated inside and outside of genes for maize and teosinte.
@@
 \subsection{Hard sweeps do not explain diversity differences}
-Selection acting to increase the frequency of a new beneficial mutation will leave a signature of reduced diversity at surrounding linked sites \citex.
+Selection acting to increase the frequency of a new beneficial mutation will leave a signature of reduced diversity at surrounding linked sites \cite{smith1974}.
 To evaluate whether patterns of such ``hard sweeps'' could explain observed differences in diversity between genic and non-genic regions of the genome, we compared diversity around missense and synonymous substitutions between \emph{Tripsacum} and either maize or teosinte (Figure
@@
 \hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_Aug}
-\caption{Pairwise diversity surrounding synonymous and non-synonymous
+\caption{Pairwise diversity surrounding synonymous and non-synonymous (missense)
   substitutions in {\bf A} maize and {\bf B} teosinte.
-Axes show both absolute diversity values and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution.
-\jri{is this correct or did i confuse things?}
- Lines depict a loess curve (span of 0.01) and shading represents bootstrap-based 95\% confidence intervals. Inset plots depict a larger range on the x-axis. \label{fig:hardSweeps}}
+Axes show both absolute diversity values (on right) and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution (on left).
+Lines depict a loess curve (span of 0.01) and shading represents bootstrap-based 95\% confidence intervals.
+Inset plots depict a larger range on the x-axis.
+\label{fig:hardSweeps}}
 \end{figure*}
@@
 We investigated purifying selection in maize and teosinte by evaluating the reduction of diversity within genes.
-Pairwise diversity is strongly reduced within genes for both maize and teosinte (Figure \ref{fig:purify}A) but recovers quickly at sites outside of genes, consistent with the low levels of linkage disequilibrium generally observed in maize \citex.
+Pairwise diversity is strongly reduced within genes for both maize and teosinte (Figure \ref{fig:purify}A) but recovers quickly at sites outside of genes, consistent with the low levels of linkage disequilibrium generally observed in maize \cite{tenaillon2002,chia2012}.
 The reduction in relative diversity is more pronounced in teosinte, however, reaching lower levels in genes and occurring over a wider region.
-\jri{make sure we have tense consistent throughout.
-is/was etc.}
 
 Our initial comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation. \jri{need a citation?}
-Such ``soft sweeps'', however, would still be expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes.
+Such ``soft sweeps'', however, are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes.
 To test this possibility, we performed a genome-wide scan for selection using a method expected to be reasonably sensitive to both hard and soft sweeps \cite{garud2015}. 
@@
 gene for pairwise diversity and singletons, respectively. 
-  Lines depict a loess curve (span of 0.01) and shading represents bootstrap-based 95\% confidence intervals. 
+  Lines depict cubic smoothing splines with smoothing parameters chosen via generalized cross validation and shading depicts bootstrap-based 95\% confidence intervals.
   Inset plots depict a smaller range on the x-axis. \label{fig:purify}
@@
 The maize population splits from teosinte $\approx 15,000$ generations in the past with an initial size of only $\approx 5\% $ of ancestral $N_a$. 
-Following domestication we estimate considerable gene flow between the populations: $M_{tm} = 1.35$  migrants per generation from teosinte to maize and $M_{mt} = 1.72$ migrants from maize to teosinte.
-\jri{do you get these numbers by multiplying estimates by Na and modern Nmaize?
-if so we should probably report the proportions not the absolute number of inds.}
+Following domestication we estimate considerable gene flow between the populations: $M_{tm} = N_a \times 1.1 \times 10^{-5}$  migrants per generation from teosinte to maize and $M_{mt} = N_a \times 1.4 \times 10^{-5}$ migrants from maize to teosinte.
 After its split from teosinte our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,000$.
@@
 In direct contrast to patterns observed for pairwise nucleotide diversity (Figure \ref{fig:purify}A), diversity of singletons was more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
-\jri{we should test singleton curve for genes not in 20\% highest H12 also!}
-Finally, while direct comparison of curves for pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, these too reveal much stronger differences in maize (Figure \X).
-\jri{if we see qualitative rather than quantitative difference in these too we should mention}
+%\jri{we should test singleton curve for genes not in 20\% highest H12 also!} I don
+Finally, while direct comparison of curves for pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, as indicated by our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
 
@@
 \subsection{Plant materials}
-We made use of published sequence of inbred accessions of teosinte (Z. \emph{mays} ssp. \emph{parviglumis}) and maize landraces from the Maize HapMap2 panel \cite{chia2012, lemmon2014}. 
-We removed \X teosinte and \X maize individuals that appeared as outliers in an initial principal component analysis  (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table \ref{sTab:list}).
+We made use of published sequences from inbred accessions of teosinte (Z.
+\emph{mays} ssp.
+\emph{parviglumis}) and maize landraces from the Maize HapMap2 panel as part of the Panzea project (bam files are available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}) \cite{chia2012, lemmon2014}.
+From these data, we removed 4 teosinte individuals that were not ssp.
+\emph{parviglumis} or appeared as outliers in an initial principal component analysis (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table \ref{sTab:list}).
+We also utilized a single tripsacum (\emph{T.
+dactyloides}) individual as an outgroup (bam file available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}).
 
 \subsection{Physical and genetic maps}
-Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{chia2012}.\jri{did we use their bams?
-I thought we used bwa bams, which are not listed and weren't used for hapmap2.
-if that's the case, we need to mention this difference and acknowledge Robert for providing those bams.
-either way, we need to include a link to the bamfiles too.
-somewhere here or above mention URL for Tripsacum bams.
-did we align those too or download someone else's?}
+Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/).
+\textcolor{red}{(I think we need more details here, but Robert did it...
+what to cite?
+also, notice Robert is in acknowledgements.)}
 All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following Li \emph{et al.} \cite{li2011statistical}.
-We converted physical coordinates to genetic coordinates via linear interpolation of the 1cM resolution NAM genetic map \cite{glaubitz2014}, \jri{i think this is be wallace not glaubitz?
-need to specify what you do with physical positions before and after last marker on each chrom and markers, if any, that have physical and genetic order that disagree}.
+We converted physical coordinates to genetic coordinates via linear interpolation of the previously published 1cM resolution NAM genetic map \cite{glaubitz2014}.
 
@@
 We estimated both the genome-wide site frequency spectrum (SFS) as well as a separate SFS for genic (within annotated transcript) and intergenic ($\geq 5kb$ from a transcript) regions. 
-We used the biomaRt package \cite{durinck2009,durinck2005} of R \cite{R2014} to parse annotations from the \X annotation \jri{include annotation version number and link to ENSEMBL annotation used}. 
+We used the biomaRt package \cite{durinck2009,durinck2005} of R \cite{R2014} to parse annotations from genebuild version 5b of AGPv3. 
 We estimated single population and joint SFS with the software ANGSD \cite{korneliussen2014}, including all positions with at least one aligned read in $\geq 80\%$ of samples in one or both populations.
 We assumed individuals were fully inbred and treated each line as a single haplotype.
-\jri{i removed options details because they don't make sense to casual reader and the interested reader can find the script.
-}
 Because ANGSD cannot calculate a folded joint SFS, we first polarized SNPs using the maize reference genome and then folded spectra using $\delta\alpha\delta{i}$ \cite{gutenkunst2009}.
@@
 We used the software $\delta\alpha\delta{i}$ \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS, using only sites $>5 kb$ from a gene to ameliorate the effects of linked selection.
-We modeled a teosinte population of constant effective size $N_a$, that at time $T_b$ generations in the past gives rise to a maize population of size $N_b$ which grows exponentially to size $N_m$ in the present (Figure \ref{fig:bottleneck})
+We modeled a teosinte population of constant effective size $N_a$, that at time $T_b$ generations in the past gave rise to a maize population of size $N_b$ which grew exponentially to size $N_m$ in the present (Figure \ref{fig:bottleneck}).
 The model includes migration of $M_{mt}$ individuals each generation from maize to teosinte and $M_{tm}$ individuals from teosinte to maize.
-\jri{$M_{mt}$ and $tm$ are the reverse of what are in figure}
-We fixed $N_a$ using $\delta\alpha\delta{i}$'s estimation of $\theta=4N_a\mu$ from the data and a mutation rate of $3 \times 10^{-8}$ \cite{clark2005}. 
-We estimated all other parameters using 1,000 $\delta\alpha\delta{i}$ optimizations and allowing initial values to be randomly perturbed by a factor of 2.  
-Initial values and upper and lower bounds can be found in table \X.
-\jri{this looked terrible in text, please make a supp.
-table.
-keep track of notation as e.g.
-we explain $N_b$ as population size of bneck, but in reality parameter we fit is $\frac{N_b}{N_a}$.
-my stab at notation is $\frac{N_b}{N_a}=0.02$, $\frac{N_m}{N_a}=3$, $\frac{T_b}{2N_a}=0.04$, and $\frac{M_{mt}}{N_c}=\frac{M_{tm}}{N_a}=10^{-5}$, where the maize population size $Nc$ at time $i$ generations in the past is $N_c=N_ae^{\frac{ln(\frac{N_m}{N_a})(T_b-i)}{T_b}}$ but if you use these make sure they are right (e.g.
-is it 2Na or 4Na for Tb?).}
+We estimated $N_a$ using $\delta\alpha\delta{i}$'s estimation of $\theta=4N_a\mu$ from the data and a mutation rate of $\mu = 3 \times 10^{-8}$ \cite{clark2005}.
+We estimated all other parameters using 1,000 $\delta\alpha\delta{i}$ optimizations and allowing initial values between runs to be randomly perturbed by a factor of 2.  
+Optimized parameters along with their initial values and upper and lower bounds can be found in table \ref{sTab:dadi}.
 We report parameter estimates from the optimization run with the highest log-likelihood.
@@
 We further made use of a large genotyping data set of more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize $N_e$ from singleton counts.
-We filtered these data to include only SNPs with data in $\geq 1,500$ individuals, and then projected the SFS down to a sample of 500 individuals.\jri{cite or mention how projection was done}. 
+We filtered these data to include only SNPs with data in $\geq 1,500$ individuals, and then projected the SFS down to a sample of 500 individuals by sampling each marker without replacement 1,000 times according to the observed allele frequencies.
 We then estimated $N_e$ from the data assuming $\mu = 3 \times 10^{-8}$ \cite{clark2005} and the relation  $4N_e\mu = \frac{S}{L}$ \cite{fu1993}, where where $S$ is the total number of singleton SNPs and $L$ is the total number of SNPs in the dataset.
@@
 We made use of the software ANGSD \cite{korneliussen2014} for diversity calculations and genotype calling. 
-We calculated diversity statistics in maize and teosinte in 100bp non-overlapping windows using filters as described above for the SFS.
-\jri{this said 1,000 kb which is 1Mb, but should say 100bp right?}
+We calculated diversity statistics in maize and teosinte in 1 kb non-overlapping windows using filters as described above for the SFS. 
 We used allele counts to estimate the number of singleton polymorphisms in each window, and used binomial sampling to create a second maize data set down-sampled to have the same number of samples as teosinte.
@@
 \begin{acknowledgments}
-We are indebted to Graham Coop and Simon Aeshbacher for their constructive input during this study. Funding was provided by NSF Plant Genome Research Project 1238014.
+We are indebted to Graham Coop and Simon Aeshbacher for their constructive input during this study.
+\textcolor{red}{We thank Robert Bukowski for providing Panzea data.} Funding was provided by NSF Plant Genome Research Project 1238014.
 \end{acknowledgments}

### 2015-08-19 supportingInfo.tex
Working through Jeff's comments

@@
 \begin{figure}
+  \includegraphics[width=\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_maizeAndTeoSingleVsPi.png}
+\caption{ Relative diversity versus distance to nearest gene in maize and teosinte.
+Relative diversity is calculated by comparing to the mean diversity in all windows $\geq 0.02 cM$ from the nearest gene.
+Lines depict cubic smoothing splines with smoothing parameters chosen via generalized cross validation and shading depicts bootstrap-based 95\% confidence intervals.
+\label{sFig:singletonPi}}
+\end{figure}
+\clearpage
+
+\begin{figure}
   \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_Unselected_manuscript.png}
@@
   \begin{center}
-  \includegraphics[width=.75\textwidth]{FigsAndFiles/tilPCA_july.pdf}\\
-  \includegraphics[width=.75\textwidth]{FigsAndFiles/bknPCA_july.pdf}\\
+  \includegraphics[width=.75\textwidth]{FigsAndFiles/tilPCA_aug.pdf}\\
+  \includegraphics[width=.75\textwidth]{FigsAndFiles/bknPCA_aug.pdf}\\
   \end{center}
@@
 {\bf D:} PC1 vs PC2 for all 23 maize individuals.
-\jri{need to fix letter label positions} \label{sFig:PCA}}
+\label{sFig:PCA}}
 \end{figure}
@@
 \end{figure}
+\clearpage
+
+\begin{figure}
+  \def\arraystretch{2}
+%  \begin{center}
+  \begin{tabular}{l|c|c|c}
+    \bf Parameter & \bf Initial value & \bf Upper bound & \bf Lower bound\\ \hline 
+    $\frac{N_b}{N_a}$ & 0.02 & $1\times10^{-7}$ & 2 \\ 
+    $\frac{N_{m}}{N_a}$ & 3 & $1\times10^{-7}$ & 200 \\
+    $\frac{T_b}{2N_a}$ & 0.04 & 0  & 1 \\ 
+    $\frac{M_{mt}}{N_a}$ & $1\times10^{-10}$ & $1\times10^{-7}$ & 0.001 \\
+    $\frac{M_{tm}}{N_a}$ & $1\times10^{-10}$ & $1\times10^{-7}$ & 0.001 \\
+  \end{tabular} 
+%  \end{center}
+  \caption{ Parameters, initial values, and boundaries used for model-fitting with $\delta\alpha\delta{i}$.
+Parameters are shown in the units utilized by $\delta\alpha\delta{i}$, although in the text simplified units are reported.
+\label{sTab:dadi} }
+    \def\arraystretch{1} % undo stretching
+\end{figure}
+\clearpage

### 2015-08-20 manuscript.tex
updates

@@
 
-Our initial comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation. \jri{need a citation?}
+Our initial comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because in such cases diversity is not necessarily reduced \cite{innan2004,messer2013}.
 Such ``soft sweeps'', however, are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes.
@@
 \subsection{Population expansion leads to stronger purifying selection in modern maize}
-Motivated by the rapid post-domestication expansion of maize evident from our demographic analysis, we investigated whether patterns of diversity at low frequency --- and thus on average younger --- polymorphisms would show similar results in comparisons of maize and teosinte.
+Motivated by the rapid post-domestication expansion of maize evident from our demographic analysis, we investigated whether patterns of diversity at low frequency --- and therefore on average younger --- polymorphisms would show similar results in comparisons of maize and teosinte.
 We thus repeated our analysis of diversity using alleles present in only a single individual in the sample.
@@
 \section{Discussion}
+
 \subsection{Hard sweeps do not shape genome-wide diversity in maize}
 
-\color{blue}
+Our findings demonstrate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, and our analysis restricting to sites undergoing the strongest and weakest purifying selection confirm this result is robust to previously highlighted limitations \cite{enard2014}.
+Our approach does not address sweeps on noncoding variants, but it agrees with a growing body of previous evidence arguing against hard sweeps shaping maize variability.
+For instance, population genetic studies of local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} demonstrate the importance of selection on standing variation as opposed to on new mutations.
+Even during domestication, very few differences between maize and teosinte were fixed \cite{hufford2012}.
+Moreover, of domestication loci that have been well studied from a population genetics standpoint, only one shows a signal of a hard sweep \cite{wang2015}, although there are several consistent with selection on standing variation \cite{studer2011,wills2013} or selection on multiple mutations \cite{wills2013}.
 
-no strong evidence for hard sweeps, even considering constraint (GERP).
-although this is not super powerful test as misses sweeps on noncoding and only has power if substantial proportion of missense sites are swept, much previous evidence argues against hard sweeps in maize.
-popgen studies of local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} show importance of standing variation. 
-even in domestication, few fixed differences between maize and teo \cite{hufford2012}, and of domestication loci that have been well studied from a popgen standpoint only one shows hard sweep \cite{wang2015} while there are multiple examples of standing variation \cite{studer2011, wills2013} or selection on multiple mutations \cite{wills2013}.
-\jri{maybe others?
-barren stalk 1 (gallavotti 2004)?
-ramosa 1 (sigmon 2010)?
-}
+The lack of hard sweeps patterning diversity differs from drosophila \cite{sattath2011} and capsella \cite{williamson2014},but is consistent with humans \cite{hernandez2011}.
+In a related study, \cite{corbett2015} showed that selection tends to have an elevated impact on reducing diversity in species with large population size compared to small.
+This may explain why maize and humans, both bottlenecked species, show similar patterns while drosophila and capsella are different.
+However, we also do not observe abundant hard sweeps in teosinte, which calls this hypothesis into question.
 
-this lack of hard sweeps patterning diversity differs from drosophila \citex and capsella \citex and many species with high Ne \jri{cite corbet-detig}.
-\jri{add something about corbet-dettig results for maize}
-but is consistent with humans \cite{hernandez2011}
-discussion of whether we expect hard vs soft in maize \cite{messer2013population}.
-$\theta_b = 2N\mu_b$ with $mu_b$ being mutation rate of beneficial alleles of selection coefficient s. 
-if $\theta_b > \frac{1}{log{Ns_b}}$ then adaptation from multiple mutations should be common. 
-if  $\theta_b > \frac{1}{log{2Ns_b}}$ then adaptation from standing neutral variation should be common, or if $\theta > \frac{1}{log(1+\frac{2Ns_b}{2Ns_d + 1}}$ for standing deleterious variation with selection coefficient $s_d$.
-\jri{i think this latter assumes codominance} codominant?)
-
-%Our findings indicate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize.
-%One known example of positive selection on a non-genic mutation involves the \emph{tb1} locus of maize, which is one of the best characterized examples of positive selection on a ``domestication gene'' in any crop \cite{clark2006}.
-However, the maize \emph{tb1} allele was already present in teosinte before domestication \cite{studer2011}, so this this example is in agreement with our finding that hard sweeps are rare.
-The \emph{gt1} locus is another well-characterized case of positive selection operating on standing variation at an enhancer region \cite{wills2013}.
-Instances of selection on standing variation such as these, often called soft sweeps, may be a major contributor to maize patterns of diversity \cite{beissinger2014}, and this could be part of the explanation as to why hard sweeps appear to be so rare, despite obvious morphological differences between maize and teosinte.
-
-%Unfortunately, our ability to accurately identify soft sweeps, and particularly to distinguish them from hard sweeps, remains limited \cite{innan2004,messer2013}.
-However, we implemented a scan based on the H12 statistic, which is designed to identify both hard and soft sweeps with reasonable power \cite{garud2015}.
-Although the goal of this study was not to identify specific sweeps, it may be that the outlier sites distinguished by H12 are primarily composed of soft sweeps instead of hard.
-Similarly, previous studies scanning for evidence of positive selection in maize \cite{hufford2012} may have picked up primarily evidence of soft sweeps.
-
-%We should additionally note that our observations do not exclude the possibility of infrequent hard sweeps having taken place during maize evolution.
-In fact, the maize locus \emph{tga1} that has been shown to correspond to a hard sweep at an amino-acid changing mutation \cite{wang2015}.
-Surrounding \emph{tga1},  our data demonstrate a pattern consistent with a hard sweep.
-But, our data demonstrate that instances such as this are infrequent.
-This contrasts sharply with \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, where differences in diversity surrounding synonymous and non-synonymous substitutions are clear suggesting hard-sweeps are abundant.
-However, it agrees with what has been seen for humans \cite{hernandez2011}.
-
-\color{black}
+Another possibility is that maize and teosinte don't show evidence of abundant hard sweeps because instead positive selection tends to operate in the form of soft sweeps.
+The theory of when soft sweeps are expected was worked out by \cite{messer2013population}.
+They showed that for $\theta_b = 2N\mu_b$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b > 0$, if $\theta_b > \frac{1}{log({Ns_b})}$, then soft sweeps in the form of adaptation from multiple mutations should be common.
+Conversely, for neutral sites that somehow become beneficial (e.g.
+change in environment), if $\theta_b > \frac{1}{log({2Ns_b})}$ then soft sweeps in the form of adaptation from standing neutral variation should be common.
+Alternatively, for previously deleterious sites that become beneficial, if $\theta > \frac{1}{log \left(1+\frac{2Ns_b}{2Ns_d + 1} \right)}$, where $s_d$ is the previous selection coefficient, soft sweeps again the form of adaptation from standing variation should be common.
+The implications of this theory are that soft sweeps are common when $\theta_b \geq 1$, which can be achieved through a large population size or a high mutation rate.
+If $\mu = 3 \times 10^{-8}$  \cite{clark2005}, this implies a population size of $\geq$ 16.67 million, which our demographic analysis shows is larger than that of teosinte but not out of the question for modern maize.
 
@@
 
-Although we estimate that the modern effective size of maize is three times that of teosinte, the small size of our sample dramatically reduces our power to estimate rapid population growth \citex.  
+Although we estimate that the modern effective size of maize is three times that of teosinte, the small size of our sample dramatically reduces our power to estimate rapid population growth since rapid growth is most evidenced by rare alleles\cite{keinan2012}.
 A better estimate of the effective size of modern maize comes from our analysis of singleton diversity in a large sample of maize landraces \cite{Hearne2015}.
@@
 \emph{mays} ssp.
-\emph{parviglumis}) and maize landraces from the Maize HapMap2 panel as part of the Panzea project (bam files are available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}) \cite{chia2012, lemmon2014}.
+\emph{parviglumis}) and maize landraces from the Maize HapMap3 panel as part of the Panzea project (bam files are available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}) \cite{chia2012, lemmon2014,hapmap3}.
 From these data, we removed 4 teosinte individuals that were not ssp.
-\emph{parviglumis} or appeared as outliers in an initial principal component analysis (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table \ref{sTab:list}).
+\emph{parviglumis} or appeared as outliers in an initial principal component analysis conducted with the package adegenet \cite{jombart2011} (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table \ref{sTab:list}).
 We also utilized a single tripsacum (\emph{T.
@@
 \subsection{Physical and genetic maps}
-Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/).
-\textcolor{red}{(I think we need more details here, but Robert did it...
-what to cite?
-also, notice Robert is in acknowledgements.)}
+Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{hapmap3}.
 All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following Li \emph{et al.} \cite{li2011statistical}.
@@
 We are indebted to Graham Coop and Simon Aeshbacher for their constructive input during this study.
-\textcolor{red}{We thank Robert Bukowski for providing Panzea data.} Funding was provided by NSF Plant Genome Research Project 1238014.
+\textcolor{red}{We thank Robert Bukowski and Qi Sun for providing early-access Panzea data.} Funding was provided by NSF Plant Genome Research Project 1238014.
 \end{acknowledgments}

### 2015-08-20 manuscript.tex
updates

@@
 \begin{abstract}
-  Unique selective and demographic operate on domesticated plant species.
-These forces interact during and after domestication to generate the patterns of DNA variability that are persistent in crop species today.
+  Unique selective and demographic forces operate on domesticated plant species.
+These forces interact during and after domestication to generate the patterns of DNA variability that are persistent today.
 To quantify the interplay between demography and selection, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide.
-We utilized whole genome sequence data from 23 maize and 13 teosinte individuals to make inferences.
+Our sample included whole genome sequence data from 23 maize and 13 teosinte individuals.
 We obtained a complete estimate of the population size fluctuations and other demographic parameters experienced by maize as it was domesticated from teosinte.
 Here, we show that maize went through a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid population size expansion post-domestication.
-We observe that hard sweeps, specifically positive selection on new genic mutations, are not the primary force driving maize evolution.
-We find that a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize.
+We observe that hard sweeps on genic mutations are not the primary force driving maize evolution.
+We find that a reduced population size during domestication decreased the efficiency of purifying selection at purging deleterious alleles from maize.
 However, expansion after domestication has since increased the efficiency of purifying selection to levels superior to those seen in teosinte.
-Our results demonstrate that in domesticated species, demographic and selective history in the ancient and recent past both contribute to genetic variability that is present today, providing substantial implications for the continued improvement of domesticated species.
+Our results demonstrate that particularly in domesticated species or bottlenecked species, demographic and selective history in the ancient and recent past both contribute to genetic variability that is present today, providing substantial implications for the continued improvement of domesticated species.
 
@@
 Such ``soft sweeps'', however, are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes.
-To test this possibility, we performed a genome-wide scan for selection using a method expected to be reasonably sensitive to both hard and soft sweeps \cite{garud2015}. 
-Even after removing genes in the top 20\% of the H12 statistic used to identify targets of selection, however, qualitative differences between maize and teosinte in patterns of diversity in and near genes remained unchanged (Figure \ref{sFig:H12}).
+To test this possibility, we performed a genome-wide scan for selection using a method expected to be reasonably sensitive to both hard and soft sweeps \cite{garud2015};
+after removing genes in the top 20\% of the H12 statistic used to identify targets of selection, we observe that qualitative differences between maize and teosinte in patterns of diversity in and near genes remained unchanged (Figure \ref{sFig:H12}A).
 
@@
 
+\color{blue}
 Finally, we reconstructed the demographic history of maize and teosinte using a recently developed coalescent approach \citex that does not require conditioning on a particular model (Figure \X).
 Though this analysis does highlight a potential weakness with our assumption of a constant size for teosinte, it nonetheless is broadly consistent with our simple model,   identifying a clear domestication bottleneck followed by rapid population expansion (Figure \X).
+\color{black}
 
@@
 In direct contrast to patterns observed for pairwise nucleotide diversity (Figure \ref{fig:purify}A), diversity of singletons was more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
-%\jri{we should test singleton curve for genes not in 20\% highest H12 also!} I don
+This relationship was maintained when we analyzed diversity surrounding genes that did not show evidence of hard or soft sweeps according to H12 (Figure \ref{sFig:H12})
 Finally, while direct comparison of curves for pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, as indicated by our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
@@
 Alternatively, for previously deleterious sites that become beneficial, if $\theta > \frac{1}{log \left(1+\frac{2Ns_b}{2Ns_d + 1} \right)}$, where $s_d$ is the previous selection coefficient, soft sweeps again the form of adaptation from standing variation should be common.
-The implications of this theory are that soft sweeps are common when $\theta_b \geq 1$, which can be achieved through a large population size or a high mutation rate.
-If $\mu = 3 \times 10^{-8}$  \cite{clark2005}, this implies a population size of $\geq$ 16.67 million, which our demographic analysis shows is larger than that of teosinte but not out of the question for modern maize.
+The implications of this theory are that soft sweeps are common when $\theta_b \geq 1$ \cite{messer2013population}, which can be achieved through a large population size or a high mutation rate.
+The best estimate of mutation rate for maize is $\mu = 3 \times 10^{-8}$  \cite{clark2005}, which implies that soft sweeps should be common if population size is $\geq$ 16.67 million.
+According to our demographic analysis, this is larger than that of teosinte but not out of the question for modern maize.
 
@@
 The observation that maize pairwise diversity is less impacted by the distance to the nearest gene than is teosinte pairwise diversity is reasonable from a long-term evolutionary standpoint.
-Theory has established that purifying selection is more efficient in a large population than in a small one \cite{kimura1984}, and this observation most likely reflects that prediction.
-More specifically, if teosinte $N_e$ remained relatively constant while maize bottlenecked and recovered exponentially, this provides that the average $N_e$ of maize over the previous several thousand generations is much smaller than that of teosinte, regardless of how much maize has ballooned in the recent past.
+Purifying selection works more efficiently in a large population than in a small one \cite{kimura1984}, so this observation likely reflects that prediction.
+Since maize bottlenecked before recovering exponentially, the average $N_e$ of maize over the previous several thousand generations is much smaller than that of teosinte.
 Therefore, our observation shows that purifying selection in maize has not purged deleterious alleles, or the neutral alleles they are linked to, as effectively as in teosinte.
-\jri{shorten this!}
 
-The reversal of this trend when we analyze only singleton diversity instead of pairwise diversity, however, stands out as a notable observation.
+The reversal of this trend when we analyze only singleton diversity instead of pairwise diversit stands out as a notable.
 Every mutation begins as a singleton (an allele present in only one individual), and therefore singletons are, on average, the youngest class of alleles that can be observed.
-Therefore unlike pairwise patterns of diversity, which are most heavily influenced by intermediate frequency alleles based on the definition of $\pi$ \cite{nei1979}, singleton diversity is most influenced by recent patterns of evolution.
+Unlike pairwise patterns of diversity, which are most heavily influenced by intermediate frequency alleles based on the definition of $\pi$ \cite{nei1979}, singleton diversity is most influenced by recent patterns of evolution.
 Hence, because our demographic estimation indicates dramatic expansion of maize $N_e$ in the recent past, we expect for purifying selection to presently operate more efficiently in maize than in teosinte.
 This observation is very clearly demonstrated by the fact that singleton diversity in maize is more impacted by the distance to the nearest gene than it is in teosinte, as was shown in Figure \ref{fig:purify}.
-\jri{shorten!}
 
-A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more deleterious alleles segregating at intermediate frequency than does teosinte.
+A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more weakly deleterious alleles segregating at intermediate frequency than does teosinte.
 This could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{meghji1984}; if deleterious alleles tend to be recessive and are particularly frequent, they will have ample opportunities to display their phenotypes in inbreds.
 Our results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously selected.
-\jri{this only works for weakly selected mutations but is an interesting point.
-need to also point out that this means new mutations in maize likely much more strongly selected than new mutations in teosinte, perhaps with implications for breeding?}
+Additionally, the large $N_e$ of modern maize compared to modern teosinte implies that for new mutations, selection will operate much more efficiently in maize.
+\textcolor{red}{breeding implications?}
 
@@
 
-Finally, \jri{MSMC text here}
+\color{blue}
+Finally, MSMC text here
+\color{black}
 
@@
 \input{supportingInfo}
-%\nocite{*}
 
 \end{article}
+
 \end{document}

### 2015-08-25 manuscript.tex
Ready for another look

@@
 
-\significancetext{Patterns of linked selection, or the impact of selection on sites that neighbor a functional variant, have been carefully evaluated in only a few species.
-\jri{not really.
-we need to rephrase here and cite the Corbett-Detig plos-bio paper somewhere -- they looked at a lot of species including maize!} In this work, we demonstrate that selection against deleterious mutations leaves a pronounced signature on the maize genome, reducing diversity in and immediately around genes.
+\significancetext{Patterns of linked selection, or the impact of selection on sites that neighbor a functional variant, play a critical role in shaping the genetic diversity of organisms.
+In this work, we demonstrate that selection against deleterious mutations leaves a pronounced signature on the maize genome, reducing diversity in and immediately around genes.
 We show how demography interacts with selection to impact genome-wide patterns of diversity, including the important observation that rapid population expansion can increase the efficiency of selection as much as a sudden population bottleneck can weaken it.
@@
 We observe that hard sweeps on genic mutations are not the primary force driving maize evolution.
-We find that a reduced population size during domestication decreased the efficiency of purifying selection at purging deleterious alleles from maize.
-However, expansion after domestication has since increased the efficiency of purifying selection to levels superior to those seen in teosinte.
+We find that a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize.
+However, expansion after domestication has since increased the efficiency of purifying selection to levels exceeding those seen in teosinte.
 Our results demonstrate that particularly in domesticated species or bottlenecked species, demographic and selective history in the ancient and recent past both contribute to genetic variability that is present today, providing substantial implications for the continued improvement of domesticated species.
@@
 In a related study, \cite{corbett2015} showed that selection tends to have an elevated impact on reducing diversity in species with large population size compared to small.
-This may explain why maize and humans, both bottlenecked species, show similar patterns while drosophila and capsella are different.
-However, we also do not observe abundant hard sweeps in teosinte, which calls this hypothesis into question.
+This may explain why maize and humans, both bottlenecked species, show similar patterns while drosophila and capsella behave differently.
+\textcolor{red}{However, we also do not observe abundant hard sweeps in teosinte, which calls this hypothesis into question.}
 
@@
 The best estimate of mutation rate for maize is $\mu = 3 \times 10^{-8}$  \cite{clark2005}, which implies that soft sweeps should be common if population size is $\geq$ 16.67 million.
-According to our demographic analysis, this is larger than that of teosinte but not out of the question for modern maize.
-
+\textcolor{red}{According to our demographic analysis, this is larger than that of teosinte but, as discussed below, not out of the question for modern maize.}
 \subsection{Demography of domestication}
 
-Although many other authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied  exclusively genic polymoprhisms and have made a number of limiting assumptions about the demographic model.
+Although many other authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied  exclusively genic polymorphisms and have made a number of limiting assumptions about the demographic model.
 We show that diversity within genes has been strongly reduced by the effects of linked selection, making genic data inappropriate for estimating demography.
@@
 
-The reversal of this trend when we analyze only singleton diversity instead of pairwise diversit stands out as a notable.
+The reversal of this trend when we analyze only singleton diversity instead of pairwise diversity stands out as a notable.
 Every mutation begins as a singleton (an allele present in only one individual), and therefore singletons are, on average, the youngest class of alleles that can be observed.
@@
 Additionally, the large $N_e$ of modern maize compared to modern teosinte implies that for new mutations, selection will operate much more efficiently in maize.
-\textcolor{red}{breeding implications?}
 
@@
 is capsella a good example?
-}
+} \textcolor{red}{I agree about this paragraph, but I ran out of time to address.
+Leave like this and I'll add more after Sept.
+11.}
 

### 2015-08-26 manuscript.tex
minor edits

@@
 To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we reanalyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 project \cite{chia2012}.
-We find broad differences in genic and non-genic diversity consistent with earlier results  \cite{hufford2012} (Figure \ref{fig:diversity}).  
+We find broad differences in genic and non-genic diversity consistent with earlier results  \cite{hufford2012} (Figure \ref{fig:diversity}\jri{pi graph should use greek letter $\pi$}).  
 In maize, mean pairwise diversity ($\pi$) within genes was significantly lower than at positions at least 5 kb away from genes (0.00668 vs 0.00691, $p<2\times 10^{-44}$). 
@@
 \begin{center}
-  \includegraphics[width=.5\textwidth] {FigsAndFiles/Pi_and_Tajima.png}
+  \includegraphics[width=.4\textwidth] {FigsAndFiles/Pi_and_Tajima.png}
 \end{center}
-\caption{Pairwise diversity (Pi) and Tajimas D were calculated inside and outside of genes for maize and teosinte.
-In both taxa, a dearth of rare alleles was observed within genes relative to outside of genes, which led to higher values of Tajima's D outside of genes than inside.{\bf A:} Mean and standard deviation of pi for 1 kb genic and non-genic windows of maize; {\bf B:} Mean and standard deviation of pi for 1 kb genic and non-genic windows of teosinte  {\bf C:} Histogram of Tajima's D in 1 kb windows in maize; {\bf D:} Histogram of Tajima's D in 1 kb windows in teosinte.
+\caption{Pairwise diversity $\pi$ (A,B) and Tajimas D (C,D) in 1kb windows from genic and nongenic regions of maize (A,C) and teosinte (B,D).
+Shown in A and B are means and one standard deviation.
 \label{fig:diversity} }
@@
 
-\begin{figure}
+\begin{figure*}
 %\vspace*{.05in}
@@
 \caption{Parameter estimates for a simple bottleneck model of maize domestication. See methods for details. \label{fig:bottleneck} }
-\end{figure}
+\end{figure*}
 

### 2015-08-28 manuscript.tex
minor edits

@@
 Such ``soft sweeps'', however, are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes.
-To test this possibility, we performed a genome-wide scan for selection using a method expected to be reasonably sensitive to both hard and soft sweeps \cite{garud2015};
-after removing genes in the top 20\% of the H12 statistic used to identify targets of selection, we observe that qualitative differences between maize and teosinte in patterns of diversity in and near genes remained unchanged (Figure \ref{sFig:H12}A).
+To test this possibility, we performed a genome-wide scan for selection using a method expected to be reasonably sensitive to both hard and soft sweeps \cite{garud2015}. 
+After removing genes in the top 20\% of the H12 statistic used to identify targets of selection, we observe that qualitative differences between maize and teosinte in patterns of diversity in and near genes remained unchanged (Figure \ref{sFig:H12}A).
 
-We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.
+We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.\jri{weird that this is a 1-sentence paragraph?}
 
@@
 The maize population splits from teosinte $\approx 15,000$ generations in the past with an initial size of only $\approx 5\% $ of ancestral $N_a$. 
-Following domestication we estimate considerable gene flow between the populations: $M_{tm} = N_a \times 1.1 \times 10^{-5}$  migrants per generation from teosinte to maize and $M_{mt} = N_a \times 1.4 \times 10^{-5}$ migrants from maize to teosinte.
+Following domestication we estimate considerable gene flow between the populations: $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation from teosinte to maize and $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants from maize to teosinte.
 After its split from teosinte our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,000$.
@@
 
-\color{blue}
-Finally, we reconstructed the demographic history of maize and teosinte using a recently developed coalescent approach \citex that does not require conditioning on a particular model (Figure \X).
-Though this analysis does highlight a potential weakness with our assumption of a constant size for teosinte, it nonetheless is broadly consistent with our simple model,   identifying a clear domestication bottleneck followed by rapid population expansion (Figure \X).
-\color{black}
+%MSMC results
+%Finally, we reconstructed the demographic history of maize and teosinte using a recently developed coalescent approach \citex that does not require conditioning on a particular model (Figure \X).
+%Though this analysis does highlight a potential weakness with our assumption of a constant size for teosinte, it nonetheless is broadly consistent with our simple model,   identifying a clear domestication bottleneck followed by rapid population expansion (Figure \X).
 
@@
 This relationship was maintained when we analyzed diversity surrounding genes that did not show evidence of hard or soft sweeps according to H12 (Figure \ref{sFig:H12})
-Finally, while direct comparison of curves for pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, as indicated by our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
+Finally, while direct comparison of curves for pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
+%ref.
+here to MSMC if we use: "as indicated by our coalescent-based demographic analysis"
 
@@
 
-Although we estimate that the modern effective size of maize is three times that of teosinte, the small size of our sample dramatically reduces our power to estimate rapid population growth since rapid growth is most evidenced by rare alleles\cite{keinan2012}.
+Although we estimate that the modern effective size of maize is three times that of teosinte, the small size of our sample dramatically reduces our power to identify low frequency alleles most sensitive to rapid population growth \cite{keinan2012}.
 A better estimate of the effective size of modern maize comes from our analysis of singleton diversity in a large sample of maize landraces \cite{Hearne2015}.
@@
 \subsection{Demography influences the efficiency of purifying selection}
-The observation that maize pairwise diversity is less impacted by the distance to the nearest gene than is teosinte pairwise diversity is reasonable from a long-term evolutionary standpoint.
+The observation that pairwise diversity in maize is less affected by the distance to the nearest gene than in teosinte is reasonable from a long-term evolutionary standpoint.
 Purifying selection works more efficiently in a large population than in a small one \cite{kimura1984}, so this observation likely reflects that prediction.
@@
 
-\color{blue}
-Finally, MSMC text here
-\color{black}
+%%%Shorten if we include MSMC
+%Finally, MSMC \cite{schiffels2014} was employed to complement our model-based demographic inference. 
+%Using six each of maize and teosinte
+%The analyses were run for the following two groups: six maize haplotypes (BKN022, BKN025, BKN029, BKN030, BKN031, BKN033) and six teosinte haplotypes (TIL01, TIL03, TIL09, TIL10, TIL11 and TIL14).
+%SNPs were called in ANGSD \cite{korneliussen2014} with the SNP p value set as $1e-6$.
+%As the samples are highly inbred, each individual was treated as one haplotype.
+%Only Sites with mapping quality bigger than $30$, base quality over $20$ and depth between half and twice of its mean depth were chosen for the analyses, and the residual heterozygous sites were removed from the analyses.
+%In addition, as the maize genome contains a large body of repetitive regions \cite{schnable2009}, we used Heng Li's software SNPable (\href{<url>}{<http://lh3lh3.users.sourceforge.net/snpable.shtml>}) to mask out genomic regions, on which short sequencing reads are not uniquely mapped.
+%The mappability mask file was generated by cutting the maize genome with $1bp$ shift each time into the $100bp$ single-end reads, which were then mapped back to the maize B73 reference genome \cite{schnable2009}.
+%The sites with the majority of overlapping $100$-mers mapped uniquely without one mismatch were called the "SNPable" sites and used for the MSMC analyses.
+%The pattern parameter $20\times2+20\times4+10\times2$ was used for the population size inference.
 
@@
 We are indebted to Graham Coop and Simon Aeshbacher for their constructive input during this study.
-\textcolor{red}{We thank Robert Bukowski and Qi Sun for providing early-access Panzea data.} Funding was provided by NSF Plant Genome Research Project 1238014.
+We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.
+Funding was provided by NSF Plant Genome Research Project 1238014.
 \end{acknowledgments}

### 2015-08-28 manuscript.tex
edits

@@
 
+%JRI suggestions
+%no strong evidence for hard sweeps, even considering constraint (GERP).
+%although this is not super powerful test as misses sweeps on noncoding and only has power if substantial proportion of missense sites are swept, much previous evidence argues against hard sweeps in maize.
+%popgen studies of local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} show importance of standing variation. 
+%even in domestication, few fixed differences between maize and teo \cite{hufford2012}, and of domestication loci that have been well studied from a popgen standpoint only one shows hard sweep \cite{wang2015} while there are multiple examples of standing variation \cite{studer2011, wills2013} or selection on multiple mutations \cite{wills2013}.
+\jri{maybe others?
+barren stalk 1 (gallavotti 2004)?
+ramosa 1 (sigmon 2010)?
+}
+%
+%this lack of hard sweeps patterning diversity differs from drosophila \citex and capsella \citex and many species with high Ne \jri{cite corbet-detig}.
+\jri{add something about corbet-dettig results for maize}
+%but is consistent with humans \cite{hernandez2011}
+%discussion of whether we expect hard vs soft in maize \cite{messer2013population}.
+%$\theta_b = 2N\mu_b$ with $mu_b$ being mutation rate of beneficial alleles of selection coefficient s. 
+%if $\theta_b > \frac{1}{log{Ns_b}}$ then adaptation from multiple mutations should be common. 
+%if  $\theta_b > \frac{1}{log{2Ns_b}}$ then adaptation from standing neutral variation should be common, or if $\theta > \frac{1}{log(1+\frac{2Ns_b}{2Ns_d + 1}}$ for standing deleterious variation with selection coefficient $s_d$.
+\jri{i think this latter assumes codominance} codominant?)
+
+%Our findings indicate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize.
+%One known example of positive selection on a non-genic mutation involves the \emph{tb1} locus of maize, which is one of the best characterized examples of positive selection on a ``domestication gene'' in any crop \cite{clark2006}.
+However, the maize \emph{tb1} allele was already present in teosinte before domestication \cite{studer2011}, so this this example is in agreement with our finding that hard sweeps are rare.
+The \emph{gt1} locus is another well-characterized case of positive selection operating on standing variation at an enhancer region \cite{wills2013}.
+Instances of selection on standing variation such as these, often called soft sweeps, may be a major contributor to maize patterns of diversity \cite{beissinger2014}, and this could be part of the explanation as to why hard sweeps appear to be so rare, despite obvious morphological differences between maize and teosinte.
+
+%Unfortunately, our ability to accurately identify soft sweeps, and particularly to distinguish them from hard sweeps, remains limited \cite{innan2004,messer2013}.
+However, we implemented a scan based on the H12 statistic, which is designed to identify both hard and soft sweeps with reasonable power \cite{garud2015}.
+Although the goal of this study was not to identify specific sweeps, it may be that the outlier sites distinguished by H12 are primarily composed of soft sweeps instead of hard.
+Similarly, previous studies scanning for evidence of positive selection in maize \cite{hufford2012} may have picked up primarily evidence of soft sweeps.
+
+%We should additionally note that our observations do not exclude the possibility of infrequent hard sweeps having taken place during maize evolution.
+In fact, the maize locus \emph{tga1} that has been shown to correspond to a hard sweep at an amino-acid changing mutation \cite{wang2015}.
+Surrounding \emph{tga1},  our data demonstrate a pattern consistent with a hard sweep.
+But, our data demonstrate that instances such as this are infrequent.
+This contrasts sharply with \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, where differences in diversity surrounding synonymous and non-synonymous substitutions are clear suggesting hard-sweeps are abundant.
+However, it agrees with what has been seen for humans \cite{hernandez2011}.
+
 \subsection{Hard sweeps do not shape genome-wide diversity in maize}
 
-Our findings demonstrate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, and our analysis restricting to sites undergoing the strongest and weakest purifying selection confirm this result is robust to previously highlighted limitations \cite{enard2014}.
-Our approach does not address sweeps on noncoding variants, but it agrees with a growing body of previous evidence arguing against hard sweeps shaping maize variability.
-For instance, population genetic studies of local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} demonstrate the importance of selection on standing variation as opposed to on new mutations.
-Even during domestication, very few differences between maize and teosinte were fixed \cite{hufford2012}.
-Moreover, of domestication loci that have been well studied from a population genetics standpoint, only one shows a signal of a hard sweep \cite{wang2015}, although there are several consistent with selection on standing variation \cite{studer2011,wills2013} or selection on multiple mutations \cite{wills2013}.
+Our findings demonstrate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show robust to concerns about power due to the effects of background selection  \cite{enard2014}.
+Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability.
+Population genomic studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation as the primary source of adaptive variation.
+Even among well-characterized domestication loci, only a single locus shows evidence of a hard sweep on a missense mutation \cite{wang2015}, but several are consistent with selection on standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
 

### 2015-08-30 manuscript.tex
some changes to discussion

@@
 
-%JRI suggestions
-%no strong evidence for hard sweeps, even considering constraint (GERP).
-%although this is not super powerful test as misses sweeps on noncoding and only has power if substantial proportion of missense sites are swept, much previous evidence argues against hard sweeps in maize.
-%popgen studies of local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} show importance of standing variation. 
-%even in domestication, few fixed differences between maize and teo \cite{hufford2012}, and of domestication loci that have been well studied from a popgen standpoint only one shows hard sweep \cite{wang2015} while there are multiple examples of standing variation \cite{studer2011, wills2013} or selection on multiple mutations \cite{wills2013}.
-\jri{maybe others?
-barren stalk 1 (gallavotti 2004)?
-ramosa 1 (sigmon 2010)?
-}
-%
 %this lack of hard sweeps patterning diversity differs from drosophila \citex and capsella \citex and many species with high Ne \jri{cite corbet-detig}.
@@
 \jri{i think this latter assumes codominance} codominant?)
-
-%Our findings indicate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize.
-%One known example of positive selection on a non-genic mutation involves the \emph{tb1} locus of maize, which is one of the best characterized examples of positive selection on a ``domestication gene'' in any crop \cite{clark2006}.
-However, the maize \emph{tb1} allele was already present in teosinte before domestication \cite{studer2011}, so this this example is in agreement with our finding that hard sweeps are rare.
-The \emph{gt1} locus is another well-characterized case of positive selection operating on standing variation at an enhancer region \cite{wills2013}.
-Instances of selection on standing variation such as these, often called soft sweeps, may be a major contributor to maize patterns of diversity \cite{beissinger2014}, and this could be part of the explanation as to why hard sweeps appear to be so rare, despite obvious morphological differences between maize and teosinte.
 
@@
 
-Our findings demonstrate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show robust to concerns about power due to the effects of background selection  \cite{enard2014}.
+Our findings demonstrate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of background selection  \cite{enard2014}.
 Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability.
 Population genomic studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation as the primary source of adaptive variation.
-Even among well-characterized domestication loci, only a single locus shows evidence of a hard sweep on a missense mutation \cite{wang2015}, but several are consistent with selection on standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
+Even among well-characterized domestication loci, only a single locus shows evidence of a hard sweep on a missense mutation \cite{wang2015}, but several are consistent with ``soft sweeps''  from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
+Such soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.  
+Assuming a mutation rate of $3 \times 10^{-8}$ \citep{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \citep{eyre2007distribution} , this implies that soft sweeps should be prevalent in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \citep[e.g.][]{studer2011}.
 
-The lack of hard sweeps patterning diversity differs from drosophila \cite{sattath2011} and capsella \cite{williamson2014},but is consistent with humans \cite{hernandez2011}.
+\jri{this paragraph needs help} The absence of evidence for a genome-wide impact of hard sweeps differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011}.
 In a related study, \cite{corbett2015} showed that selection tends to have an elevated impact on reducing diversity in species with large population size compared to small.
@@
 
-Another possibility is that maize and teosinte don't show evidence of abundant hard sweeps because instead positive selection tends to operate in the form of soft sweeps.
-The theory of when soft sweeps are expected was worked out by \cite{messer2013population}.
-They showed that for $\theta_b = 2N\mu_b$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b > 0$, if $\theta_b > \frac{1}{log({Ns_b})}$, then soft sweeps in the form of adaptation from multiple mutations should be common.
-Conversely, for neutral sites that somehow become beneficial (e.g.
-change in environment), if $\theta_b > \frac{1}{log({2Ns_b})}$ then soft sweeps in the form of adaptation from standing neutral variation should be common.
-Alternatively, for previously deleterious sites that become beneficial, if $\theta > \frac{1}{log \left(1+\frac{2Ns_b}{2Ns_d + 1} \right)}$, where $s_d$ is the previous selection coefficient, soft sweeps again the form of adaptation from standing variation should be common.
-The implications of this theory are that soft sweeps are common when $\theta_b \geq 1$ \cite{messer2013population}, which can be achieved through a large population size or a high mutation rate.
-The best estimate of mutation rate for maize is $\mu = 3 \times 10^{-8}$  \cite{clark2005}, which implies that soft sweeps should be common if population size is $\geq$ 16.67 million.
-\textcolor{red}{According to our demographic analysis, this is larger than that of teosinte but, as discussed below, not out of the question for modern maize.}
 \subsection{Demography of domestication}
-
+\jri{section needs help} 
 Although many other authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied  exclusively genic polymorphisms and have made a number of limiting assumptions about the demographic model.
@@
 \subsection{Demography influences the efficiency of purifying selection}
+\jri{section needs help} 
 The observation that pairwise diversity in maize is less affected by the distance to the nearest gene than in teosinte is reasonable from a long-term evolutionary standpoint.
@@
 is capsella a good example?
-} \textcolor{red}{I agree about this paragraph, but I ran out of time to address.
-Leave like this and I'll add more after Sept.
-11.}
+}
 

### 2015-08-31 manuscript.tex
edits

@@
 
-\jri{this paragraph needs help} The absence of evidence for a genome-wide impact of hard sweeps differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011}.
+\jri{paragraph needs work. i need to reread C-D}
+The absence of evidence for a genome-wide impact of hard sweeps differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
 In a related study, \cite{corbett2015} showed that selection tends to have an elevated impact on reducing diversity in species with large population size compared to small.
@@
 \subsection{Demography of domestication}
-\jri{section needs help} 
-Although many other authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied  exclusively genic polymorphisms and have made a number of limiting assumptions about the demographic model.
-We show that diversity within genes has been strongly reduced by the effects of linked selection, making genic data inappropriate for estimating demography.
-Because levels of diversity alone do not allow independent inference of the strength and duration of a bottleneck \cite{tenaillon2004selection}, previous investigations have instead estimated the ratio of these parameters.
-We have overcome this limitations through the use of the full joint SFS of maize and teosinte and consideration of a more realistic model of exponential growth, enabling us to estimate for the first time the strength of the domestication bottleneck.
-Given that modern maize exhibits $\approx 80\%$ of the diversity of teosinte, our estimate that  the effective population size of the initial maize  population was only $\approx 5\%$ of the ancestral teosinte population seems surprisingly low.
-The impact of a strong domestication bottleneck is likely ameliorated to some degree by  maize-teosinte gene flow and the rapid post-domestication growth estimated in our model.
+Although a number of  authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
+We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sties.
+By utilizing the full joint SFS, we are also able to, for the first time,  estimate the strength of the domestication bottleneck without requiring assumptions about its duration.  
+Given that  maize landraces exhibit $\approx 80\%$ of the diversity of teosinte \citep{hufford2012}, our estimate of bottleneck of $\approx 5\%$ of the ancestral teosinte population seems surprisingly low.
+The impact of a strong domestication bottleneck is likely ameliorated to some degree by both gene flow and the rapid post-domestication growth estimated in our model.
 
@@
 
-Although we estimate that the modern effective size of maize is three times that of teosinte, the small size of our sample dramatically reduces our power to identify low frequency alleles most sensitive to rapid population growth \cite{keinan2012}.
+Although we estimate that the modern effective size of maize is three times that of teosinte, the small size of our sample dramatically reduces our power to identify the low frequency alleles most sensitive to rapid population growth \cite{keinan2012}.
 A better estimate of the effective size of modern maize comes from our analysis of singleton diversity in a large sample of maize landraces \cite{Hearne2015}.
-These data suggest a modern effective size of nearly 1 million, nearly eight times higher than modern teosinte.
-Nonetheless, missing data, sample size, and ascertainment bias in the genotyping data all bias this estimate downward as well.
-We can, however, get reasonable estimates of the census size of modern maize.
-There are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, with a lower limit of 25,000 individuals planter per hectare \cite{baden2001culture}.
-Using this lower bound on planting density, ignoring all industrial maize agriculture, and even assuming the effective size is only 0.1\% of the census size, this still implies a modern effective population size of more than one billion.
-Clearly, the effective size of modern maize is extremely large.
+These data suggest a modern effective size nearly eight times higher than modern teosinte at $\approx 1$ million.
+Even so, missing data, sample size, and ascertainment bias in the genotyping data likely bias this estimate downward as well.
+An alternative approach is to use estimates of the census size of modern maize.
+There are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, likely planted at a density of $\approx 25,000$ individuals  per hectare \cite{baden2001culture}.
+Assuming the effective size is only 0.4\% of the census size (i.e. 1 female for every 1000 males), this still implies a modern effective population size of more than four billion.
+While these rough estimates are likely quite poor, all of the evidence points to the fact that the effective size of modern maize is extremely large.
 
@@
 \jri{section needs help} 
-The observation that pairwise diversity in maize is less affected by the distance to the nearest gene than in teosinte is reasonable from a long-term evolutionary standpoint.
-Purifying selection works more efficiently in a large population than in a small one \cite{kimura1984}, so this observation likely reflects that prediction.
-Since maize bottlenecked before recovering exponentially, the average $N_e$ of maize over the previous several thousand generations is much smaller than that of teosinte.
-Therefore, our observation shows that purifying selection in maize has not purged deleterious alleles, or the neutral alleles they are linked to, as effectively as in teosinte.
+We find  a more pronounced decrease in nucleotide diversity around genes in teosinte compared to domesticated maize.
+Because purifying selection is more efficient in larger populations  \cite{kimura1984},  this observation is consistent with our demographic model as well as previous work showing a smaller long-term effective population size in maize.
+This pattern of diversity is reversed, however, when we assess diversity at singleton polymorphisms.  
+Under a simple neutral model, lower frequency polymorphisms will be younger and patterns of diversity of singleton polymorphisms should thus represent much more recent evolutionary history. 
+We argue that the stronger decline in singleton diversity in and around genes in maize is the result of the increased efficacy of purifying selection  associated with the dramatic population expansion in maize following domestication.
 
-The reversal of this trend when we analyze only singleton diversity instead of pairwise diversity stands out as a notable.
-Every mutation begins as a singleton (an allele present in only one individual), and therefore singletons are, on average, the youngest class of alleles that can be observed.
-Unlike pairwise patterns of diversity, which are most heavily influenced by intermediate frequency alleles based on the definition of $\pi$ \cite{nei1979}, singleton diversity is most influenced by recent patterns of evolution.
-Hence, because our demographic estimation indicates dramatic expansion of maize $N_e$ in the recent past, we expect for purifying selection to presently operate more efficiently in maize than in teosinte.
+This difference in recent $N_e$ should lead to an increased efficacy of purifying selection in the recent past of  Differences in diversity observed at  demographic estimation indicates dramatic the recent past, we expect for purifying selection to presently operate more efficiently in maize than in teosinte.
 This observation is very clearly demonstrated by the fact that singleton diversity in maize is more impacted by the distance to the nearest gene than it is in teosinte, as was shown in Figure \ref{fig:purify}.

### 2015-08-31 manuscript.tex
more edits

@@
 \jri{section needs help} 
-We find  a more pronounced decrease in nucleotide diversity around genes in teosinte compared to domesticated maize.
+We find  a more pronounced decrease in nucleotide diversity around genes in teosinte compared to domesticated maize (Figure \ref{fig:purify}A.
 Because purifying selection is more efficient in larger populations  \cite{kimura1984},  this observation is consistent with our demographic model as well as previous work showing a smaller long-term effective population size in maize.
-This pattern of diversity is reversed, however, when we assess diversity at singleton polymorphisms.  
-Under a simple neutral model, lower frequency polymorphisms will be younger and patterns of diversity of singleton polymorphisms should thus represent much more recent evolutionary history. 
+Under a simple neutral model, lower frequency polymorphisms will be younger and patterns of diversity of singleton polymorphisms should  represent  more recent evolutionary history. 
+In contrast to the pattern seen with pairwise diversity, we see a stronger decline of singleton diversity around genes in maize (Figure \ref{fig:purify}B.
 We argue that the stronger decline in singleton diversity in and around genes in maize is the result of the increased efficacy of purifying selection  associated with the dramatic population expansion in maize following domestication.
-
-This difference in recent $N_e$ should lead to an increased efficacy of purifying selection in the recent past of  Differences in diversity observed at  demographic estimation indicates dramatic the recent past, we expect for purifying selection to presently operate more efficiently in maize than in teosinte.
-This observation is very clearly demonstrated by the fact that singleton diversity in maize is more impacted by the distance to the nearest gene than it is in teosinte, as was shown in Figure \ref{fig:purify}.
 
@@
 Additionally, the large $N_e$ of modern maize compared to modern teosinte implies that for new mutations, selection will operate much more efficiently in maize.
+\jri{inbreds stuff OK, but need more.
+Chia, Rogers-Melnick, Mezmouk all looked at load.
+does this have implications for large effect load (i.e.
+should bneck have purged those?), for heterosis, architecture of quant.
+traits, etc.}
 
- Importantly, our estimation of the parameters of the maize domestication bottleneck contribute to the understanding of how the demography of crop domestication can impact crop diversity.
+% Importantly, our estimation of the parameters of the maize domestication bottleneck contribute to the understanding of how the demography of crop domestication can impact crop diversity.
 The bottleneck-effects from a sudden collapse in population size have been well studied and are known to impact crops for thousands of generations.
@@
 
-\jri{need paragraph also talking about how this is important for considering impact of linked selection.
-cite some recent human papers that are considering both demography and linked selection, but also a number of plant papers that e.g.
-estimate the DFE from current data or just use the sny/nonsyn graphs and ignore recent demographic history.
-is capsella a good example?
-}
+\jri{these results have implications for lots of studies.
+recent demographic history often ignored (examples).
+in crops, bneck is considered and loss of diversity, but expansion also likely ubiquitous and consequences not well studied.
+in humans XYZ, and we see some of that already here.
+this argues that a full understanding of diversity in both wild and cultivated plants will require recent demography.}
 

### 2015-08-31 manuscript.tex
edits

@@
 
-%this lack of hard sweeps patterning diversity differs from drosophila \citex and capsella \citex and many species with high Ne \jri{cite corbet-detig}.
-\jri{add something about corbet-dettig results for maize}
-%but is consistent with humans \cite{hernandez2011}
-%discussion of whether we expect hard vs soft in maize \cite{messer2013population}.
-%$\theta_b = 2N\mu_b$ with $mu_b$ being mutation rate of beneficial alleles of selection coefficient s. 
-%if $\theta_b > \frac{1}{log{Ns_b}}$ then adaptation from multiple mutations should be common. 
-%if  $\theta_b > \frac{1}{log{2Ns_b}}$ then adaptation from standing neutral variation should be common, or if $\theta > \frac{1}{log(1+\frac{2Ns_b}{2Ns_d + 1}}$ for standing deleterious variation with selection coefficient $s_d$.
-\jri{i think this latter assumes codominance} codominant?)
-
 %Unfortunately, our ability to accurately identify soft sweeps, and particularly to distinguish them from hard sweeps, remains limited \cite{innan2004,messer2013}.
@@
 
-Our findings demonstrate that hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of background selection  \cite{enard2014}.
+Our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of background selection  \cite{enard2014}.
 Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability.
-Population genomic studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation as the primary source of adaptive variation.
-Even among well-characterized domestication loci, only a single locus shows evidence of a hard sweep on a missense mutation \cite{wang2015}, but several are consistent with ``soft sweeps''  from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
-Such soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.  
-Assuming a mutation rate of $3 \times 10^{-8}$ \citep{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \citep{eyre2007distribution} , this implies that soft sweeps should be prevalent in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \citep[e.g.][]{studer2011}.
+Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with ``soft sweeps''  from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
+Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation as the primary source of adaptive variation.
+Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.  
+Assuming a mutation rate of $3 \times 10^{-8}$ \citep{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \citep{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \citep[e.g.][]{studer2011}.
 
@@
 We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sties.
-By utilizing the full joint SFS, we are also able to, for the first time,  estimate the strength of the domestication bottleneck without requiring assumptions about its duration.  
-Given that  maize landraces exhibit $\approx 80\%$ of the diversity of teosinte \citep{hufford2012}, our estimate of bottleneck of $\approx 5\%$ of the ancestral teosinte population seems surprisingly low.
-The impact of a strong domestication bottleneck is likely ameliorated to some degree by both gene flow and the rapid post-domestication growth estimated in our model.
+By utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without requiring assumptions about its duration.
 
-Another surprising result from our model is the estimation of the timing of domestication.
-Although we estimate the maize population split $\approx 5,000$ years earlier than previous molecular \cite{matsuoka2002} or archaelogical \cite{piperno2009starch} estimates, we do not interpret these results as strongly conflicting.
-We first note that the timing of the genetic split between populations likely preceded the anatomical changes that can be identified in the archaeological record. 
-We also caution, however, that our estimate may be inflated by population structure as our geographically diverse sample may include populations diverged from those that gave rise to maize.
+One surprising result from our model is the estimation of the timing of domestication at $\approx 15,000$ years before present.
+While this appears to conflict with archaeological \cite{piperno2009starch} estimates, we instead argue this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
+We also note that our result may also be inflated due to population structure, as our geographically diverse sample of teosinte may include populations diverged from those that gave rise to maize.
 
-Although we estimate that the modern effective size of maize is three times that of teosinte, the small size of our sample dramatically reduces our power to identify the low frequency alleles most sensitive to rapid population growth \cite{keinan2012}.
-A better estimate of the effective size of modern maize comes from our analysis of singleton diversity in a large sample of maize landraces \cite{Hearne2015}.
-These data suggest a modern effective size nearly eight times higher than modern teosinte at $\approx 1$ million.
-Even so, missing data, sample size, and ascertainment bias in the genotyping data likely bias this estimate downward as well.
-An alternative approach is to use estimates of the census size of modern maize.
-There are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, likely planted at a density of $\approx 25,000$ individuals  per hectare \cite{baden2001culture}.
-Assuming the effective size is only 0.4\% of the census size (i.e. 1 female for every 1000 males), this still implies a modern effective population size of more than four billion.
-While these rough estimates are likely quite poor, all of the evidence points to the fact that the effective size of modern maize is extremely large.
+The estimated bottleneck of $\approx 5\%$ of the ancestral teosinte population seems low given that maize landraces exhibit $\approx 80\%$ of the diversity of teosinte \citep{hufford2012}, but our model suggests that the effects of the bottleneck on diversity are likely ameliorated to some extent by both gene flow and rapid post-domestication growth (Figure \ref{fig:bottleneck}).
+Although we estimate that the modern effective size of maize is larger than teosinte, the small size of our sample reduces our power to identify the low frequency alleles most sensitive to rapid population growth \cite{keinan2012}.
+A better estimate comes from our analysis of singleton diversity in a large sample of maize landraces \cite{Hearne2015}, which suggests a modern effective size of $\approx 1$ million, or nearly eight times higher than modern teosinte.
+Even this estimate, however, is likely downwardly biased due to missing data, sample size, and ascertainment in the genotyping data.
+An alternative approach is to use estimates of the census size of modern maize: there are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, likely planted at a density of $\approx 25,000$ individuals  per hectare \cite{baden2001culture}.
+Assuming the effective size is only $\approx 0.4\%$ of the census size (i.e. 1 ear for every 1000 plants), this still implies a modern effective population size of more than four billion.
+While these rough estimates are likely poor, all of the evidence points to the fact that the effective size of modern maize is extremely large.
 
 \subsection{Demography influences the efficiency of purifying selection}
-\jri{section needs help} 
-We find  a more pronounced decrease in nucleotide diversity around genes in teosinte compared to domesticated maize (Figure \ref{fig:purify}A.
-Because purifying selection is more efficient in larger populations  \cite{kimura1984},  this observation is consistent with our demographic model as well as previous work showing a smaller long-term effective population size in maize.
-Under a simple neutral model, lower frequency polymorphisms will be younger and patterns of diversity of singleton polymorphisms should  represent  more recent evolutionary history. 
-In contrast to the pattern seen with pairwise diversity, we see a stronger decline of singleton diversity around genes in maize (Figure \ref{fig:purify}B.
-We argue that the stronger decline in singleton diversity in and around genes in maize is the result of the increased efficacy of purifying selection  associated with the dramatic population expansion in maize following domestication.
+\jri{section needs help -- i think all the points are there but i'm not happy with the writing} 
+We find a more pronounced decrease in nucleotide diversity around genes in teosinte compared to domesticated maize (Figure \ref{fig:purify}A).
+This difference in diversity is likely due to the effects of purifying selection.
+Purifying selection is more efficient in larger populations  \cite{kimura1984}, and our observation is thus consistent with our demographic model and previous work \cite{eyre1998, tenaillon2004selection, wright2005, ross2009} showing a smaller long-term effective population size in maize.
+To understand how recent population growth has changed the effects of linked selection, we also analyzed diversity of singleton polymorphisms.
+Under a simple neutral model, lower frequency polymorphisms are expected to be younger, and patterns of diversity at singleton SNPs should thus represent recent evolutionary history.
+Consistent with its larger current size and thus more efficient purifying selection, we see a stronger decline of singleton diversity around genes in maize (Figure \ref{fig:purify}B than teosinte.
 
@@
 Additionally, the large $N_e$ of modern maize compared to modern teosinte implies that for new mutations, selection will operate much more efficiently in maize.
-\jri{inbreds stuff OK, but need more.
+\jri{inbreds stuff OK, but should be shortened.
+and we need more/broader.
 Chia, Rogers-Melnick, Mezmouk all looked at load.
@@
 traits, etc.}
-
-% Importantly, our estimation of the parameters of the maize domestication bottleneck contribute to the understanding of how the demography of crop domestication can impact crop diversity.
-The bottleneck-effects from a sudden collapse in population size have been well studied and are known to impact crops for thousands of generations.
-Complementing this knowledge, our results demonstrate that the rapid expansion experienced by many crops after domestication can also have a profound influence on patterns of diversity, and the effects of this expansion should be accounted for as important contributors to long-term evolution.
-\jri{this is sorta fluff and doesn't say much}
 

### 2015-09-01 manuscript.tex
edits

@@
 
-%Unfortunately, our ability to accurately identify soft sweeps, and particularly to distinguish them from hard sweeps, remains limited \cite{innan2004,messer2013}.
-However, we implemented a scan based on the H12 statistic, which is designed to identify both hard and soft sweeps with reasonable power \cite{garud2015}.
-Although the goal of this study was not to identify specific sweeps, it may be that the outlier sites distinguished by H12 are primarily composed of soft sweeps instead of hard.
-Similarly, previous studies scanning for evidence of positive selection in maize \cite{hufford2012} may have picked up primarily evidence of soft sweeps.
-
-%We should additionally note that our observations do not exclude the possibility of infrequent hard sweeps having taken place during maize evolution.
-In fact, the maize locus \emph{tga1} that has been shown to correspond to a hard sweep at an amino-acid changing mutation \cite{wang2015}.
-Surrounding \emph{tga1},  our data demonstrate a pattern consistent with a hard sweep.
-But, our data demonstrate that instances such as this are infrequent.
-This contrasts sharply with \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, where differences in diversity surrounding synonymous and non-synonymous substitutions are clear suggesting hard-sweeps are abundant.
-However, it agrees with what has been seen for humans \cite{hernandez2011}.
-
 \subsection{Hard sweeps do not shape genome-wide diversity in maize}
@@
 Assuming a mutation rate of $3 \times 10^{-8}$ \citep{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \citep{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \citep[e.g.][]{studer2011}.
+\jri{no mention of our H12 results.
+I think that's OK?}
 
@@
 Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
-We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sties.
+We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites.
 By utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without requiring assumptions about its duration.
@@
 
-%%%Shorten if we include MSMC
+%%%Shorten text here if we include MSMC
 %Finally, MSMC \cite{schiffels2014} was employed to complement our model-based demographic inference. 

### 2015-09-03 manuscript.tex
edits

@@
 
-%MSMC results
-%Finally, we reconstructed the demographic history of maize and teosinte using a recently developed coalescent approach \citex that does not require conditioning on a particular model (Figure \X).
-%Though this analysis does highlight a potential weakness with our assumption of a constant size for teosinte, it nonetheless is broadly consistent with our simple model,   identifying a clear domestication bottleneck followed by rapid population expansion (Figure \X).
+Finally, we reconstructed the demographic history of maize and teosinte using a recently developed coalescent approach \cite{schiffels2014}  that does not require conditioning on a particular model.
+Though this analysis suggests the potential for nonequilibrium dynamics in teosinte, it nonetheless is broadly consistent with our simple model,   identifying a clear domestication bottleneck followed by rapid population expansion (Figure \ref{sFig:msmc}).
 
@@
 This relationship was maintained when we analyzed diversity surrounding genes that did not show evidence of hard or soft sweeps according to H12 (Figure \ref{sFig:H12})
-Finally, while direct comparison of curves for pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
-%ref.
-here to MSMC if we use: "as indicated by our coalescent-based demographic analysis"
+Finally, while direct comparison of curves for pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, as indicated by our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
 
@@
 
-\jri{paragraph needs work. i need to reread C-D}
+\jri{paragraph needs work. maybe eliminate/move? i do think we need some comparison to other stuff, but the broader Ne-selection efficiency argument should come later.}
 The absence of evidence for a genome-wide impact of hard sweeps differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
@@
 This may explain why maize and humans, both bottlenecked species, show similar patterns while drosophila and capsella behave differently.
-\textcolor{red}{However, we also do not observe abundant hard sweeps in teosinte, which calls this hypothesis into question.}
+\textcolor{red}{However, we also do not observe abundant hard sweeps in teosinte, which calls this explanation into question.}
 
@@
 \subsection{Demography influences the efficiency of purifying selection}
-\jri{section needs help -- i think all the points are there but i'm not happy with the writing} 
+\jri{section needs help --  i'm not happy with the writing} 
 We find a more pronounced decrease in nucleotide diversity around genes in teosinte compared to domesticated maize (Figure \ref{fig:purify}A).
@@
 
+\jri{I would like to see a supplemental figure or maybe just number here on the partial correlation in both teosinte and maize of diversity with recombination after accounting for gene density.
+do this in 500kb windows and then in 10kb windows.
+we should do the correlation with both pi and singletons.
+This let's us compare more explictly with studies like C-D who evaluate overall patterns of linked selection.
+will be especially cool if we don't see a big difference in big windows but we do in 10kb windows.
+}
 A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more weakly deleterious alleles segregating at intermediate frequency than does teosinte.
@@
 in humans XYZ, and we see some of that already here.
-this argues that a full understanding of diversity in both wild and cultivated plants will require recent demography.}
+this argues that a full understanding of diversity in both wild and cultivated plants will require recent demography.
+mention C-D ignored recent demography -- here's where rec/pi windows could help.}
 
@@
 dactyloides}) individual as an outgroup (bam file available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}).
+\jri{some clarification here.
+did we use Lemmon's higher depth teosinte?
+If so, why does Li say we only had 7 high-depth bam files?
+there should be more}
 
@@
 
-%%%Shorten text here if we include MSMC
-%Finally, MSMC \cite{schiffels2014} was employed to complement our model-based demographic inference. 
-%Using six each of maize and teosinte
-%The analyses were run for the following two groups: six maize haplotypes (BKN022, BKN025, BKN029, BKN030, BKN031, BKN033) and six teosinte haplotypes (TIL01, TIL03, TIL09, TIL10, TIL11 and TIL14).
-%SNPs were called in ANGSD \cite{korneliussen2014} with the SNP p value set as $1e-6$.
-%As the samples are highly inbred, each individual was treated as one haplotype.
-%Only Sites with mapping quality bigger than $30$, base quality over $20$ and depth between half and twice of its mean depth were chosen for the analyses, and the residual heterozygous sites were removed from the analyses.
-%In addition, as the maize genome contains a large body of repetitive regions \cite{schnable2009}, we used Heng Li's software SNPable (\href{<url>}{<http://lh3lh3.users.sourceforge.net/snpable.shtml>}) to mask out genomic regions, on which short sequencing reads are not uniquely mapped.
-%The mappability mask file was generated by cutting the maize genome with $1bp$ shift each time into the $100bp$ single-end reads, which were then mapped back to the maize B73 reference genome \cite{schnable2009}.
-%The sites with the majority of overlapping $100$-mers mapped uniquely without one mismatch were called the "SNPable" sites and used for the MSMC analyses.
-%The pattern parameter $20\times2+20\times4+10\times2$ was used for the population size inference.
+As a final estimate of demography, we employed MSMC \cite{schiffels2014}  to complement our model-based demographic inference. 
+We used six each of maize and teosinte (BKN022, BKN025, BKN029, BKN030, BKN031, BKN033, TIL01, TIL03, TIL09, TIL10, TIL11 and TIL14), treating each inbred genome as a single haplotype.
+We called SNPs in ANGSD \cite{korneliussen2014} using a SNP p-value of $1e-6$ against a reference genome masked using SNPable (\url{http://lh3lh3.users.sourceforge.net/snpable.shtml}).
+We then removed heterozygous genotypes and filtered sites with a mapping quality $<30$, a base quality $<20$, or a $|\emph{log}_2(\emph{depth})|<1$.
+We ran MSMC with pattern parameters $20\times2+20\times4+10\times2$.
 

### 2015-09-09 manuscript.tex
results edits

@@
 \section{Results}
-\subsection{Patterns of diversity differ between genic and  non-genic regions of the genome}
-To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we reanalyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 project \cite{chia2012}.
-We find broad differences in genic and non-genic diversity consistent with earlier results  \cite{hufford2012} (Figure \ref{fig:diversity}\jri{pi graph should use greek letter $\pi$}).  
-In maize, mean pairwise diversity ($\pi$) within genes was significantly lower than at positions at least 5 kb away from genes (0.00668 vs 0.00691, $p<2\times 10^{-44}$). 
+\subsection{Patterns of diversity differ between genic and  intergenic regions of the genome}
+To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012} \jri{add cite to HM3}.
+We find broad differences in genic and intergenic diversity consistent with earlier results  \cite{hufford2012}(Figure \ref{fig:diversity}):  \jri{pi graph should use greek letter $\pi$} in maize, mean pairwise diversity ($\pi$) within genes was significantly lower than at site at least 5 kb away from genes (0.00668 vs 0.00691, $p<2\times 10^{-44}$).
 Diversity differences in teosinte are even more pronounced (0.0088 vs. 0.0115, $p\approx 0$). 
@@
 Selection acting to increase the frequency of a new beneficial mutation will leave a signature of reduced diversity at surrounding linked sites \cite{smith1974}.
-To evaluate whether patterns of such ``hard sweeps'' could explain observed differences in diversity between genic and non-genic regions of the genome, we compared diversity around missense and synonymous substitutions between \emph{Tripsacum} and either maize or teosinte (Figure
-\ref{fig:hardSweeps}).
+To evaluate whether patterns of such ``hard sweeps'' could explain observed differences in diversity between genic and intergenic regions of the genome, we compared diversity around missense and synonymous substitutions between \emph{Tripsacum} and either maize or teosinte.
 If a proportion of missense mutations have been fixed due to hard sweeps, diversity around these substitutions should be lower than around synonymous substitutions. 
 We observe this pattern around the causative amino acid substitution in the the domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005origin, wang2015}.
-Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte.
+Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte (Figure
+\ref{fig:hardSweeps}).
 
@@
 \hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_Aug}
-\caption{Pairwise diversity surrounding synonymous and non-synonymous (missense)
-  substitutions in {\bf A} maize and {\bf B} teosinte.
-Axes show both absolute diversity values (on right) and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution (on left).
+\caption{Pairwise diversity surrounding synonymous and missense  substitutions in {\bf A} maize and {\bf B} teosinte.
+Axes show absolute diversity values (right) and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution (left).
 Lines depict a loess curve (span of 0.01) and shading represents bootstrap-based 95\% confidence intervals.
@@
 
-Previous analyses have suggested that this approach may have limited power because a higher proportion of nonsynonymous substitutions will be found in genes under weak purifying selection and thus with higher genetic diversity \cite{enard2014}.
-To address this concern, we took advantage of genomic evolutionary rate profile (GERP) scores \cite{davydov2010}, a measure of evolutionary constraint, calculated across the maize genome \cite{rodgers2015}.
-We re-analyzed substitutions in subsets of genes with the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection  (Figure \ref{sFig:consUncons}).
-As expected, we see higher diversity around substitutions in genes under weak purifying selection, but we still see no difference between synonymous and missense substitutions in either subset of the data.
+Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes under weak purifying selection and thus higher genetic diversity \cite{enard2014}.
+To address this concern, we took advantage of genome-wide estimates of evolutionary constraint \cite{rodgers2015}, calculated using genomic evolutionary rate profile (GERP) scores \cite{davydov2010}. 
+We then evaluated substitutions only in subsets of genes in the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection  (Figure \ref{sFig:consUncons}).
+As expected, we see higher diversity around substitutions in genes under weak purifying selection \jri{there a table or figure we can reference?}, but we still find no difference between synonymous and missense substitutions in either subset of the data.
 Taken together, these data suggest hard sweeps do not play a major role in patterning genic diversity in either maize or teosinte.
@@
 Our initial comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because in such cases diversity is not necessarily reduced \cite{innan2004,messer2013}.
-Such ``soft sweeps'', however, are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes.
-To test this possibility, we performed a genome-wide scan for selection using a method expected to be reasonably sensitive to both hard and soft sweeps \cite{garud2015}. 
+Such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes. 
+To test this possibility, we performed a genome-wide scan for selection using a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
 After removing genes in the top 20\% of the H12 statistic used to identify targets of selection, we observe that qualitative differences between maize and teosinte in patterns of diversity in and near genes remained unchanged (Figure \ref{sFig:H12}A).
-
-We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.\jri{weird that this is a 1-sentence paragraph?}
+We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.
 
@@
 Assuming a mutation rate of $3 \times 10^{-8}$ \citep{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \citep{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \citep[e.g.][]{studer2011}.
-\jri{no mention of our H12 results.
-I think that's OK?}
+Indeed, many adaptive traits in both maize \citex and teosinte \citex are highly quantitative \jri{cite wallace for maize?
+weber and/or lauter for teosinte?} and adaptation in both maize \citep{hufford2012} and teosinte \citep{pyhajarvi2013complex} has involved selection on regulatory variation.
+\jri{if number is nice, add something like "\X of the windows of top H12 score genome-wide occur outside of genes"}
 
@@
 We find a more pronounced decrease in nucleotide diversity around genes in teosinte compared to domesticated maize (Figure \ref{fig:purify}A).
-This difference in diversity is likely due to the effects of purifying selection.
+This difference is likely due to the effects of purifying selection.
 Purifying selection is more efficient in larger populations  \cite{kimura1984}, and our observation is thus consistent with our demographic model and previous work \cite{eyre1998, tenaillon2004selection, wright2005, ross2009} showing a smaller long-term effective population size in maize.

### 2015-09-09 manuscript.tex
significance

@@
 
-\significancetext{Patterns of linked selection, or the impact of selection on sites that neighbor a functional variant, play a critical role in shaping the genetic diversity of organisms.
-In this work, we demonstrate that selection against deleterious mutations leaves a pronounced signature on the maize genome, reducing diversity in and immediately around genes.
-We show how demography interacts with selection to impact genome-wide patterns of diversity, including the important observation that rapid population expansion can increase the efficiency of selection as much as a sudden population bottleneck can weaken it.
-Along the way, we develop the first estimate the demographic parameters of the maize domestication from whole genome sequence data.
+\significancetext{Both selection and demographic change play important roles in shaping diversity across the genome, but clear empirical examples of the interplay of these two forces are lacking.
+Here we document the combined effects of demography and linked selection on genome-wide diversity in domesticated maize and its wild ancestor teosinte.
+We estimate that maize underwent a bottleneck of 5\% of the size of the ancestral teosinte population, but that recent expansion has resulted in a maize population perhaps orders of magnitude larger than teosinte.
+We show that positive selection on new mutations has had relatively little effect on genetic diversity, but that selection against deleterious mutations has dramatically reduced diversity in and immediately around genes in both taxa.
+We find that the relative effect of selection depend qualitatively on the age of the polymorphisms evaluated: while older polymorphisms in maize show more limited effects of linked selection, new mutations instead reflect its larger current size and more efficient purifying selection.
+Our results demonstrate that a complete understanding of genome-wide patterns of diversity will require careful assessment of both demographic history and the effects of linked selection.
 }
@@
 
-Selection can also reduce diversity in functional regions of the genome via removal of deleterious mutations, a process known to as purifying or background selection \cite{charlesworth1993}.
-We investigated purifying selection in maize and teosinte by evaluating the reduction of diversity within genes.
+Selection can also reduce diversity in functional regions of the genome via removal of deleterious mutations, a process known as purifying or background selection \cite{charlesworth1993}.
+We investigated purifying selection in maize and teosinte by evaluating the reduction of diversity around genes.
 Pairwise diversity is strongly reduced within genes for both maize and teosinte (Figure \ref{fig:purify}A) but recovers quickly at sites outside of genes, consistent with the low levels of linkage disequilibrium generally observed in maize \cite{tenaillon2002,chia2012}.
-The reduction in relative diversity is more pronounced in teosinte, however, reaching lower levels in genes and occurring over a wider region.
+The reduction in relative diversity is more pronounced in teosinte, reaching lower levels in genes and occurring over a wider region.
 
@@
 Such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes. 
-To test this possibility, we performed a genome-wide scan for selection using a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
-After removing genes in the top 20\% of the H12 statistic used to identify targets of selection, we observe that qualitative differences between maize and teosinte in patterns of diversity in and near genes remained unchanged (Figure \ref{sFig:H12}A).
+To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
+Qualitative differences between maize and teosinte remained unchanged even after removing genes in the top 20\% of H12 (Figure \ref{sFig:H12}A).
 We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.
@@
 \subsection{Population expansion leads to stronger purifying selection in modern maize}
-Motivated by the rapid post-domestication expansion of maize evident from our demographic analysis, we investigated whether patterns of diversity at low frequency --- and therefore on average younger --- polymorphisms would show similar results in comparisons of maize and teosinte.
+Motivated by the rapid post-domestication expansion of maize evident in our demographic analysis, we investigated whether diversity at low frequency --- and therefore on average younger --- polymorphisms would show similar results in comparisons of maize and teosinte.
 We thus repeated our analysis of diversity using alleles present in only a single individual in the sample.
-In direct contrast to patterns observed for pairwise nucleotide diversity (Figure \ref{fig:purify}A), diversity of singletons was more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
+In contrast to our restults for pairwise nucleotide diversity (Figure \ref{fig:purify}A), diversity of singletons was more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
 This relationship was maintained when we analyzed diversity surrounding genes that did not show evidence of hard or soft sweeps according to H12 (Figure \ref{sFig:H12})
@@
 These results suggest that demographic history and population size change have impacted the efficacy of purifying selection during maize evolution.
-In contrast, analysis of singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) provides a nearly identical pattern to that shown in Figure \ref{fig:hardSweeps}, providing little support for a substantial increase in the number or strength of hard sweeps occurring in maize.
+In contrast, singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) appears nearly identical to results from pairwise diversity (Figure \ref{fig:hardSweeps}), providing little support for a substantial increase in the number or strength of hard sweeps occurring in maize.
 

### 2015-09-09 manuscript.tex
estimate

@@
 
+\jri{we need a starting paragraph on linked selection and demography.
+check out slotte 2014 briefings functional genomics for a great review of ideas.
+then we do a paragraph on why domesticates are great for this (modified version of your below)}
+
 \dropcap{D}omesticated plant species evolve in a unique fashion
@@
 
-Maize represents an excellent organism to study these
-phenomena. Maize is a species of tremendous importance worldwide as 
-both a staple crop \cite{shiferaw2011} and as a model for
-understanding plant evolution \cite{strable2009}. Broadly speaking, archaeological and genetic studies have
+%Maize represents an excellent organism to study these phenomena.
+% Maize is a species of tremendous importance worldwide as both a staple crop \cite{shiferaw2011} and as a model for understanding plant evolution \cite{strable2009}. 
+Broadly speaking, archaeological and genetic studies have
 established that maize domestication is likely to have taken place in
@@
 \section{Results}
-\subsection{Patterns of diversity differ between genic and  intergenic regions of the genome}
+%\subsection{Patterns of diversity differ between genic and  intergenic regions of the genome} %JRI
 To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012} \jri{add cite to HM3}.
@@
 
-\subsection{Hard sweeps do not explain diversity differences}
+\subsection{Hard sweeps do not explain diversity differences} %JRI
 Selection acting to increase the frequency of a new beneficial mutation will leave a signature of reduced diversity at surrounding linked sites \cite{smith1974}.
@@
 
-\subsection{Diversity is strongly influenced by purifying selection}
+\subsection{Diversity is strongly influenced by purifying selection} %JRI
 
@@
 
-\subsection{Demography of maize domestication}
+\subsection{Demography of maize domestication} %JRI
 To explore whether differences in the efficacy of purifying selection between maize and teosinte can be explained by demographic processes, we estimated the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}).
-The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ individuals given the mutation rate \cite{clark2005}.
-The maize population splits from teosinte $\approx 15,000$ generations in the past with an initial size of only $\approx 5\% $ of ancestral $N_a$. 
+The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ individuals using a mutation rate of $3 \times 10^{-8}$ \cite{clark2005}.
+In our model, the maize population split from teosinte $\approx 15,000$ generations in the past with an initial size of only $\approx 5\% $ of ancestral $N_a$. 
 Following domestication we estimate considerable gene flow between the populations: $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation from teosinte to maize and $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants from maize to teosinte.

### 2015-09-09 manuscript.tex
results

@@
 To explore whether differences in the efficacy of purifying selection between maize and teosinte can be explained by demographic processes, we estimated the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}).
-The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ individuals using a mutation rate of $3 \times 10^{-8}$ \cite{clark2005}.
-In our model, the maize population split from teosinte $\approx 15,000$ generations in the past with an initial size of only $\approx 5\% $ of ancestral $N_a$. 
-Following domestication we estimate considerable gene flow between the populations: $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation from teosinte to maize and $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants from maize to teosinte.
-After its split from teosinte our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,000$.
+The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ individuals.
+We estimate that the maize population split from teosinte $\approx 15,000$ generations in the past, with an initial size of only $\approx 5\% $ of the ancestral $N_a$. 
+After its split from teosinte, our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,000$.
+Maize and teosinte continue to exchange migrants after the population split, with gene flow between the populations estimated at $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation from teosinte to maize and $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants from maize to teosinte.
 
@@
 
-Because our relatively small sample size limits our ability to characterize the rare variants that are most informative of population expansion, we took advantage of a complementary data set of more than 4,000 maize landraces collected from across the Americas \cite{Hearne2015} to estimate the modern maize effective population size from low frequency variants.
+In addition to our simple bottleneck model, we investigated two alternative approaches for demographic inference.
+First, we took advantage of genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using low frequency variants  informative of population expansion.
 This analysis yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
+Finally, we applied a model-free coalescent approach \cite{schiffels2014} using a subset of our samples. 
+Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
 
-Finally, we reconstructed the demographic history of maize and teosinte using a recently developed coalescent approach \cite{schiffels2014}  that does not require conditioning on a particular model.
-Though this analysis suggests the potential for nonequilibrium dynamics in teosinte, it nonetheless is broadly consistent with our simple model,   identifying a clear domestication bottleneck followed by rapid population expansion (Figure \ref{sFig:msmc}).
-
-\subsection{Population expansion leads to stronger purifying selection in modern maize}
-Motivated by the rapid post-domestication expansion of maize evident in our demographic analysis, we investigated whether diversity at low frequency --- and therefore on average younger --- polymorphisms would show similar results in comparisons of maize and teosinte.
-We thus repeated our analysis of diversity using alleles present in only a single individual in the sample.
-In contrast to our restults for pairwise nucleotide diversity (Figure \ref{fig:purify}A), diversity of singletons was more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
-This relationship was maintained when we analyzed diversity surrounding genes that did not show evidence of hard or soft sweeps according to H12 (Figure \ref{sFig:H12})
-Finally, while direct comparison of curves for pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, as indicated by our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
-
-These results suggest that demographic history and population size change have impacted the efficacy of purifying selection during maize evolution.
-In contrast, singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) appears nearly identical to results from pairwise diversity (Figure \ref{fig:hardSweeps}), providing little support for a substantial increase in the number or strength of hard sweeps occurring in maize.
+\subsection{Population expansion leads to stronger purifying selection in modern maize} %JRI
+Motivated by the rapid post-domestication expansion of maize evident in our demographic analysis, we reasoned that low-frequency --- and thus younger --- polymorphisms might show patterns distinct from pairwise diversity.
+Singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) appears nearly identical to results from pairwise diversity (Figure \ref{fig:hardSweeps}), providing little support for a substantial increase in the number or strength of hard sweeps occurring in maize.
+In contrast, we observe a qualitative shift in the effects of purifying selection: singleton polymorphisms are more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
+This relationship again remained after removing the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12})
+Finally, while direct comparison of pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, as indicated by our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
 

### 2015-09-09 manuscript.tex
ok. good with methods and results and significance and all but last section of discussion. last section of discussion, abstract, and intro need work.

@@
 
-\subsection{Hard sweeps do not shape genome-wide diversity in maize}
+\subsection{Hard sweeps do not shape genome-wide diversity in maize} %JRI
 
@@
 weber and/or lauter for teosinte?} and adaptation in both maize \citep{hufford2012} and teosinte \citep{pyhajarvi2013complex} has involved selection on regulatory variation.
-\jri{if number is nice, add something like "\X of the windows of top H12 score genome-wide occur outside of genes"}
+\jri{what \% of the windows of top H12 score genome-wide occur outside of genes?
+is that a number to include here?}
 
-\jri{paragraph needs work. maybe eliminate/move? i do think we need some comparison to other stuff, but the broader Ne-selection efficiency argument should come later.}
-The absence of evidence for a genome-wide impact of hard sweeps differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
-In a related study, \cite{corbett2015} showed that selection tends to have an elevated impact on reducing diversity in species with large population size compared to small.
-This may explain why maize and humans, both bottlenecked species, show similar patterns while drosophila and capsella behave differently.
-\textcolor{red}{However, we also do not observe abundant hard sweeps in teosinte, which calls this explanation into question.}
-
-\subsection{Demography of domestication}
+\subsection{Demography of domestication} %JRI
 Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
 We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites.
-By utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without requiring assumptions about its duration.
+Furthermore, by utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without requiring assumptions about its duration.
 
-One surprising result from our model is the estimation of the timing of domestication at $\approx 15,000$ years before present.
+One surprising result from our model is the estimated timing of domestication at $\approx 15,000$ years before present.
 While this appears to conflict with archaeological \cite{piperno2009starch} estimates, we instead argue this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
@@
 
-The estimated bottleneck of $\approx 5\%$ of the ancestral teosinte population seems low given that maize landraces exhibit $\approx 80\%$ of the diversity of teosinte \citep{hufford2012}, but our model suggests that the effects of the bottleneck on diversity are likely ameliorated to some extent by both gene flow and rapid post-domestication growth (Figure \ref{fig:bottleneck}).
-Although we estimate that the modern effective size of maize is larger than teosinte, the small size of our sample reduces our power to identify the low frequency alleles most sensitive to rapid population growth \cite{keinan2012}.
-A better estimate comes from our analysis of singleton diversity in a large sample of maize landraces \cite{Hearne2015}, which suggests a modern effective size of $\approx 1$ million, or nearly eight times higher than modern teosinte.
-Even this estimate, however, is likely downwardly biased due to missing data, sample size, and ascertainment in the genotyping data.
-An alternative approach is to use estimates of the census size of modern maize: there are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, likely planted at a density of $\approx 25,000$ individuals  per hectare \cite{baden2001culture}.
-Assuming the effective size is only $\approx 0.4\%$ of the census size (i.e. 1 ear for every 1000 plants), this still implies a modern effective population size of more than four billion.
-While these rough estimates are likely poor, all of the evidence points to the fact that the effective size of modern maize is extremely large.
+The estimated bottleneck of $\approx 5\%$ of the ancestral teosinte population seems low given that maize landraces exhibit $\approx 80\%$ of the diversity of teosinte \citep{hufford2012}, but our model suggests that the effects of the bottleneck on diversity are likely ameliorated by both gene flow and rapid population growth (Figure \ref{fig:bottleneck}).
+Although we estimate that the modern effective size of maize is larger than teosinte, the small size of our sample reduces our power to identify the low frequency alleles most sensitive to rapid population growth \cite{keinan2012}, and our model is unable to incorporate growth faster than exponential.
+Both alternative approaches we employ estimate much a larger modern effective size of maize in the range of $\approx 10^6 - 10^9$, an order of magnitude or more than the current size of teosinte.
+Census data suggest these estimates are plausible: there are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, likely planted at a density of $\approx 25,000$ individuals per hectare \cite{baden2001culture}.
+Assuming the effective size is only $\approx 0.4\%$ of the census size (i.e. 1 ear for every 1000 male plants), this still implies a modern effective population size of more than four billion.
+While these genetic and census estimates are likely inaccurate, all of the evidence points to the fact that the effective size of modern maize is extremely large.
 
@@
 This let's us compare more explictly with studies like C-D who evaluate overall patterns of linked selection.
-will be especially cool if we don't see a big difference in big windows but we do in 10kb windows.
-}
+will be especially cool if we don't see a big difference in big windows but we do in 10kb windows -- siggests C-D looked at wrong scale  }
 A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more weakly deleterious alleles segregating at intermediate frequency than does teosinte.
@@
 Additionally, the large $N_e$ of modern maize compared to modern teosinte implies that for new mutations, selection will operate much more efficiently in maize.
-\jri{inbreds stuff OK, but should be shortened.
-and we need more/broader.
+\jri{inbreds stuff is OK, but should be shortened.
+we are aiming at a broader audience that doesn't care about maize.
+what are more/broader.
 Chia, Rogers-Melnick, Mezmouk all looked at load.
@@
 traits, etc.}
+
+The absence of evidence for a genome-wide impact of hard sweeps differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
+In a related study, \cite{corbett2015} showed that selection tends to have an elevated impact on reducing diversity in species with large population size compared to small.
+This may explain why maize and humans, both bottlenecked species, show similar patterns while drosophila and capsella behave differently.
+\textcolor{red}{However, we also do not observe abundant hard sweeps in teosinte, which calls this explanation into question.}
 
@@
 
-\begin{materials}
+\begin{materials} %JRI
 
@@
 \emph{mays} ssp.
-\emph{parviglumis}) and maize landraces from the Maize HapMap3 panel as part of the Panzea project (bam files are available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}) \cite{chia2012, lemmon2014,hapmap3}.
+\emph{parviglumis}) and maize landraces from the Maize HapMap3 panel as part of the Panzea project  \cite{chia2012, lemmon2014, hapmap3}.
 From these data, we removed 4 teosinte individuals that were not ssp.
 \emph{parviglumis} or appeared as outliers in an initial principal component analysis conducted with the package adegenet \cite{jombart2011} (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table \ref{sTab:list}).
-We also utilized a single tripsacum (\emph{T.
-dactyloides}) individual as an outgroup (bam file available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}).
+We also utilized a single individual of (\emph{Tripsacum dactyloides}) as an outgroup.
+All bam files are available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}).
 \jri{some clarification here.
@@
 If so, why does Li say we only had 7 high-depth bam files?
-there should be more}
+shouldn't there be more?}
 

### 2015-09-24 manuscript.tex
intro and results comments and edits

@@
 \newcommand{\jri}[1]{\textcolor{red}{\scriptsize #1}}
+\newcommand{\mbh}[1]{\textcolor{blue}{\scriptsize #1}}
 \newcommand{\citex}{\textcolor{red}{\bf CITE}}
@@
 \title{Demography and linked selection in wild and domesticated maize}
-\author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
-    California, Davis, CA, USA}\affil{2}{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}\affil{3}{Division of Plant Sciences, University of Missouri, Columbia, MO, USA} Li Wang\affil{4}{Iowa State University, Ames, IA, USA}, Kate Crosby\affil{1}{}, Arun
-  Durvasula\affil{1}{}, Matthew Hufford\affil{4}{}, \and Jeffrey
-  Ross-Ibarra\affil{1}{}\affil{5}{Genome Center and Center for population biology, University of
+\author{Timothy M.
+Beissinger\affil{1}{Dept.
+of Plant Sciences, University of California, Davis, CA, USA}\affil{2}{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}\affil{3}{Division of Plant Sciences, University of Missouri, Columbia, MO, USA}, Li Wang\affil{4}{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA}\affil{5}{Genome Informatics Facility, Iowa State University, Ames, IA, USA}, Kate Crosby\affil{1}{}, Arun Durvasula\affil{1}{}, Matthew B.
+Hufford\affil{4}{}, \and Jeffrey Ross-Ibarra\affil{1}{}\affil{6}{Genome Center and Center for Population Biology, University of
     California, Davis, CA, USA} }
@@
   Unique selective and demographic forces operate on domesticated plant species.
-These forces interact during and after domestication to generate the patterns of DNA variability that are persistent today.
+\mbh{I don't agree with this statement; characteristics of demography and selection vary across crops and I wouldn't be surprised to find similar demographies and patterns of selection between crop and wild species} These forces interact during and after domestication to generate the patterns of DNA variability that are persistent today.
 To quantify the interplay between demography and selection, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide.
@@
 However, expansion after domestication has since increased the efficiency of purifying selection to levels exceeding those seen in teosinte.
-Our results demonstrate that particularly in domesticated species or bottlenecked species, demographic and selective history in the ancient and recent past both contribute to genetic variability that is present today, providing substantial implications for the continued improvement of domesticated species.
+Our results demonstrate that particularly in domesticated species or bottlenecked species \mbh{right, observations can be generalized to wild species as well; I don't think there's any need to say crops are a special case}, demographic and selective history in the ancient and recent past both contribute to genetic variability that is present today, providing substantial implications for the continued improvement of domesticated species.
+\mbh{I'd end this more generally, expressing the importance of knowing demography when characterizing selection genome-wide}
 
@@
 then we do a paragraph on why domesticates are great for this (modified version of your below)}
+\mbh{Yes, agree here...think the topics addressed in the manuscript need to be set up as generalizable and not just specific to domesticates.
+Crops are a great system because we know a fair bit more about demography and the process of selection than in wild species, but they are not entirely unique.}
 
 \dropcap{D}omesticated plant species evolve in a unique fashion
-compared to their wild counterparts \cite{doebley2006}. This
-is a result of both the anthropomorphic nature of artificial selection on
+compared to their wild counterparts \cite{doebley2006}.
+\mbh{Again, I'm not a big fan of the distinction between natural and artificial selection.
+While I think it's fair to say that crops and their wild relatives have experienced distinct evolutionary forces/histories (like any other pair of taxa), I don't like the broad domesticated/wild distinction.
+Like domesticated species which encounter the human, agronomic environment, wild species can be confronted with novel environmental conditions and experience a bottleneck during adaptation to these conditions.
+Here it seems the argument is that human selection and domestication demography are unlike any selection/demography plants have previously encountered.
+If this is indeed the point we want to make, I think we need to be much more explicit in explaining how and why they are different}
+This is a result of both the anthropomorphic nature of artificial selection on
 domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
@@
 population size of modern maize. Sequence information from maize and
-teosinte plants may therefore be utilized to address these questions.
+teosinte may therefore be utilized to address these questions.
 
@@
 variability in both maize and teosinte 2) research the impact that the
-domestication process has had on genetic variability in maize, and how
+domestication process \mbh{I think this is referring specifically to demography (i.e., not selection during domestication)...might want to say that more clearly to distinguish this objective from objective 1} has had on genetic variability in maize, and how
 this compares to the impact of a different demographic history in
 teosinte; and 3) precisely estimate the parameters of the
-maize domestication bottleneck.  \cite{chia2012}.
+maize domestication bottleneck.
+\cite{chia2012}.
+\mbh{A quick sentence here about why this matters broadly would help...what kind of basic inferences will this research allow us to make about the interplay of demography and selection?}
 
@@
 To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012} \jri{add cite to HM3}.
-We find broad differences in genic and intergenic diversity consistent with earlier results  \cite{hufford2012}(Figure \ref{fig:diversity}):  \jri{pi graph should use greek letter $\pi$} in maize, mean pairwise diversity ($\pi$) within genes was significantly lower than at site at least 5 kb away from genes (0.00668 vs 0.00691, $p<2\times 10^{-44}$).
+We find broad differences in genic and intergenic diversity consistent with earlier results  \cite{hufford2012}(Figure \ref{fig:diversity}).  \jri{pi graph should use greek letter $\pi$} 
+In maize, mean pairwise diversity ($\pi$) within genes was significantly lower than at sites at least 5 kb away from genes (0.00668 vs 0.00691, $p<2\times 10^{-44}$). 
 Diversity differences in teosinte are even more pronounced (0.0088 vs. 0.0115, $p\approx 0$). 
@@
 \subsection{Hard sweeps do not explain diversity differences} %JRI
-Selection acting to increase the frequency of a new beneficial mutation will leave a signature of reduced diversity at surrounding linked sites \cite{smith1974}.
+When selection increases the frequency of a new beneficial mutation, a signature of reduced diversity is left at surrounding linked sites \cite{smith1974}.
 To evaluate whether patterns of such ``hard sweeps'' could explain observed differences in diversity between genic and intergenic regions of the genome, we compared diversity around missense and synonymous substitutions between \emph{Tripsacum} and either maize or teosinte.
@@
 We observe this pattern around the causative amino acid substitution in the the domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005origin, wang2015}.
-Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte (Figure
-\ref{fig:hardSweeps}).
+\mbh{Awesome}
+Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
+\mbh{I'm curious whether linked selection is affecting the diversity around a sizable proportion of synonymous substitutions.
+Would it be worthwhile to look at diversity around the subset of synonymous substitutions that are the least linked to missense substitutions just to make sure it is not substantially higher?
+Or perhaps generate a plot of diversity around synonymous substitutions vs.
+distance of these subs from missense subs?}
 
@@
 
-Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes under weak purifying selection and thus higher genetic diversity \cite{enard2014}.
+Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes that are under weak purifying selection and thus have higher genetic diversity \cite{enard2014}.
 To address this concern, we took advantage of genome-wide estimates of evolutionary constraint \cite{rodgers2015}, calculated using genomic evolutionary rate profile (GERP) scores \cite{davydov2010}. 
@@
 We investigated purifying selection in maize and teosinte by evaluating the reduction of diversity around genes.
-Pairwise diversity is strongly reduced within genes for both maize and teosinte (Figure \ref{fig:purify}A) but recovers quickly at sites outside of genes, consistent with the low levels of linkage disequilibrium generally observed in maize \cite{tenaillon2002,chia2012}.
+Pairwise diversity is strongly reduced within genes for both maize and teosinte (Figure \ref{fig:purify}A) but recovers quickly at sites outside of genes, consistent with the low levels of linkage disequilibrium generally observed in these subspecies \cite{tenaillon2002,chia2012}.
 The reduction in relative diversity is more pronounced in teosinte, reaching lower levels in genes and occurring over a wider region.
 
-Our initial comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because in such cases diversity is not necessarily reduced \cite{innan2004,messer2013}.
+Our comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because in such cases diversity is not necessarily reduced \cite{innan2004,messer2013}.
 Such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes. 
-To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
+To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
 Qualitative differences between maize and teosinte remained unchanged even after removing genes in the top 20\% of H12 (Figure \ref{sFig:H12}A).
+\mbh{This makes me wonder about the top 20\%. Is the pattern of diversity reduction different in this fraction of the genome?  If so, that seems like something that should be mentioned.}
 We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.
@@
 Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.  
-Assuming a mutation rate of $3 \times 10^{-8}$ \citep{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \citep{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \citep[e.g.][]{studer2011}.
-Indeed, many adaptive traits in both maize \citex and teosinte \citex are highly quantitative \jri{cite wallace for maize?
-weber and/or lauter for teosinte?} and adaptation in both maize \citep{hufford2012} and teosinte \citep{pyhajarvi2013complex} has involved selection on regulatory variation.
+Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.][]{studer2011}.
+Indeed, many adaptive traits in both maize \cite and teosinte \cite are highly quantitative \jri{cite wallace for maize?
+weber and/or lauter for teosinte?} and adaptation in both maize \cite{hufford2012} and teosinte \cite{pyhajarvi2013complex} has involved selection on regulatory variation.
 \jri{what \% of the windows of top H12 score genome-wide occur outside of genes?
@@
 
-The estimated bottleneck of $\approx 5\%$ of the ancestral teosinte population seems low given that maize landraces exhibit $\approx 80\%$ of the diversity of teosinte \citep{hufford2012}, but our model suggests that the effects of the bottleneck on diversity are likely ameliorated by both gene flow and rapid population growth (Figure \ref{fig:bottleneck}).
+The estimated bottleneck of $\approx 5\%$ of the ancestral teosinte population seems low given that maize landraces exhibit $\approx 80\%$ of the diversity of teosinte \cite{hufford2012}, but our model suggests that the effects of the bottleneck on diversity are likely ameliorated by both gene flow and rapid population growth (Figure \ref{fig:bottleneck}).
 Although we estimate that the modern effective size of maize is larger than teosinte, the small size of our sample reduces our power to identify the low frequency alleles most sensitive to rapid population growth \cite{keinan2012}, and our model is unable to incorporate growth faster than exponential.

### 2015-09-24 manuscript.tex
additional edits, discussion

@@
 Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation as the primary source of adaptive variation.
+\mbh{But Sho's paper and Tanja's paper suggest non-coding variants are important for adaptation...should we mention this caveat?}
 Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.  
@@
 \jri{what \% of the windows of top H12 score genome-wide occur outside of genes?
-is that a number to include here?}
+is that a number to include here?} \mbh{additional consideration of non-genic regions in both the results and discussion would probably be a good idea given Tanja and Sho's results}
 
@@
 Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
+\mbh{keep an eye out for Brandon's ninjas}
 We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites.
+\mbh{weird to follow up a statement about incomplete demographic model with selection result}
 Furthermore, by utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without requiring assumptions about its duration.
@@
 Although we estimate that the modern effective size of maize is larger than teosinte, the small size of our sample reduces our power to identify the low frequency alleles most sensitive to rapid population growth \cite{keinan2012}, and our model is unable to incorporate growth faster than exponential.
-Both alternative approaches we employ estimate much a larger modern effective size of maize in the range of $\approx 10^6 - 10^9$, an order of magnitude or more than the current size of teosinte.
+Both alternative approaches we employ estimate a much larger modern effective size of maize in the range of $\approx 10^6 - 10^9$, an order of magnitude or more than the current size of teosinte.
 Census data suggest these estimates are plausible: there are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, likely planted at a density of $\approx 25,000$ individuals per hectare \cite{baden2001culture}.

### 2015-10-06 manuscript.tex
updating introduction.

@@
 
-\jri{we need a starting paragraph on linked selection and demography.
+%\jri{we need a starting paragraph on linked selection and demography.
 check out slotte 2014 briefings functional genomics for a great review of ideas.
 then we do a paragraph on why domesticates are great for this (modified version of your below)}
-\mbh{Yes, agree here...think the topics addressed in the manuscript need to be set up as generalizable and not just specific to domesticates.
+%\mbh{Yes, agree here...think the topics addressed in the manuscript need to be set up as generalizable and not just specific to domesticates.
 Crops are a great system because we know a fair bit more about demography and the process of selection than in wild species, but they are not entirely unique.}
 
-\dropcap{D}omesticated plant species evolve in a unique fashion
-compared to their wild counterparts \cite{doebley2006}.
-\mbh{Again, I'm not a big fan of the distinction between natural and artificial selection.
-While I think it's fair to say that crops and their wild relatives have experienced distinct evolutionary forces/histories (like any other pair of taxa), I don't like the broad domesticated/wild distinction.
-Like domesticated species which encounter the human, agronomic environment, wild species can be confronted with novel environmental conditions and experience a bottleneck during adaptation to these conditions.
-Here it seems the argument is that human selection and domestication demography are unlike any selection/demography plants have previously encountered.
-If this is indeed the point we want to make, I think we need to be much more explicit in explaining how and why they are different}
-This is a result of both the anthropomorphic nature of artificial selection on
-domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
-bottleneck(s) that they tend to have experienced
-\cite{ross2007}. However, the
-complex interplay between selective pressures and demographic
-limitations, and the impact that this interplay has on identifying
-selection and understanding demography, is not fully understood. Although a large body of
-research that involves searching the genomes of domesticated species for evidence
-of positive selection exists \cite{hufford2012, he2011, vigouroux2002, chapman2008}, these studies tend to focus on
-identifying or mapping particular genes or regions that play an
-important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
-selection have on whole-genome patterns of genetic variability remains limited.
+\dropcap{T}he genetic properties of populations are determined by a constant interplay between genetic drift and selection.
+Drift is the consequence of a finite population size, such that random sampling of gametes every generation will incrementally lead to genetic changes.
+In other words, the demographic history of a population, which includes its population size fluctuations and mating patterns, has a long-lasting impact on the patterns of DNA variability exhibited by that population.
+Conversely, selection systematically alters a population by favoring particular alleles at the expense of others.
+In contrast to drift, which results from demographic history, selection is the result of the relationship between genotypes, phenoytpes, and ultimately fitness.
+Importantly, demography and selection do not operate independently to determine genetic variability.
+Instead, these forces are engaged in a continuous battle by which each masks and obfuscates the outcomes of the other, ultimately meaning that the selective history of a species cannot be properly evaluated without an thorough estimate of demographic history, and demography cannot be evaluated without an understanding of selection.
 
-%Maize represents an excellent organism to study these phenomena.
-% Maize is a species of tremendous importance worldwide as both a staple crop \cite{shiferaw2011} and as a model for understanding plant evolution \cite{strable2009}. 
+For instance, supposed neutrally-evolving DNA is often used to estimate historical demographic
+parameters of a population such as effective population size, structure, and expansion
+history \cite{luikart2003, gutenkunst2009}. However, researchers have
+called into question whether or not there are reasonable approaches
+for identifying neutral regions of the genome, since the effects of
+selection can be wide-ranging \cite{li2012, slotte2014}. For example, in
+\emph{Drosophila} it appears that the majority of the genome is
+impacted by the effects of selection through linkage
+\cite{sella2009}. A natural next question, therefore, is how can
+demographic parameters be estimated independently of
+selective parameters? One technique utilized to address this
+problem is limiting analyses to only sites far from genes
+\cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is
+difficult to be certain that sites even in gene-poor regions of the
+genome are not influenced by linked selection. Ultimately, an understanding
+of which sites have the potential to be adaptive, and how these affect
+genome-wide patterns of variability through linkage, is required for
+reasonable demographic inference.
+
+Maize represents an excellent organism to study these phenomena.
+Maize is a species of tremendous importance worldwide as both a staple crop \cite{shiferaw2011} and as a model for understanding plant evolution \cite{strable2009}.
 Broadly speaking, archaeological and genetic studies have
@@
 relative importance of different forms of selection on whole-genome
-variability in both maize and teosinte 2) research the impact that the
-domestication process \mbh{I think this is referring specifically to demography (i.e., not selection during domestication)...might want to say that more clearly to distinguish this objective from objective 1} has had on genetic variability in maize, and how
+variability, using maize and teosinte as a model 2) research the impact that the
+demography history has had on genetic variability in maize, and how
 this compares to the impact of a different demographic history in
@@
 \mbh{A quick sentence here about why this matters broadly would help...what kind of basic inferences will this research allow us to make about the interplay of demography and selection?}
+
+\dropcap{D}omesticated plant species evolve in a unique fashion
+compared to their wild counterparts \cite{doebley2006}.
+This is a result of both the anthropomorphic nature of artificial selection on
+domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
+bottleneck(s) that they tend to have experienced
+\cite{ross2007}. However, the
+complex interplay between selective pressures and demographic
+limitations, and the impact that this interplay has on identifying
+selection and understanding demography, is not fully understood. Although a large body of
+research that involves searching the genomes of domesticated species for evidence
+of positive selection exists \cite{hufford2012, he2011, vigouroux2002, chapman2008}, these studies tend to focus on
+identifying or mapping particular genes or regions that play an
+important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
+selection have on whole-genome patterns of genetic variability remains limited.
 
@@
 Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.  
-Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.][]{studer2011}.
-Indeed, many adaptive traits in both maize \cite and teosinte \cite are highly quantitative \jri{cite wallace for maize?
+Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.]{studer2011}.
+Indeed, many adaptive traits in both maize \citex and teosinte \citex are highly quantitative \jri{cite wallace for maize?
 weber and/or lauter for teosinte?} and adaptation in both maize \cite{hufford2012} and teosinte \cite{pyhajarvi2013complex} has involved selection on regulatory variation.

### 2015-10-06 manuscript.tex
updates

@@
 Maize is a species of tremendous importance worldwide as both a staple crop \cite{shiferaw2011} and as a model for understanding plant evolution \cite{strable2009}.
-Broadly speaking, archaeological and genetic studies have
+Archaeological and genetic studies have
 established that maize domestication is likely to have taken place in
@@
 the bottleneck's duration, nor are there sequence-based estimates of the effective
-population size of modern maize. Sequence information from maize and
-teosinte may therefore be utilized to address these questions.
+population size of modern maize.
+At the same time, the predominant form or forms of selection that have taken place during and since domestication remain unknown.
+Sequence information from maize and teosinte may therefore be utilized to address these questions.
 
@@
 this compares to the impact of a different demographic history in
-teosinte; and 3) precisely estimate the parameters of the
+its wild progenitor teosinte; and 3) precisely estimate the parameters of the
 maize domestication bottleneck.
-\cite{chia2012}.
-\mbh{A quick sentence here about why this matters broadly would help...what kind of basic inferences will this research allow us to make about the interplay of demography and selection?}
+Gaining an improved understanding of the interaction between demography and selection in maize will help to increase the understanding of these phenomenon across a wide array of organisms.
 
-\dropcap{D}omesticated plant species evolve in a unique fashion
-compared to their wild counterparts \cite{doebley2006}.
-This is a result of both the anthropomorphic nature of artificial selection on
-domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
-bottleneck(s) that they tend to have experienced
-\cite{ross2007}. However, the
-complex interplay between selective pressures and demographic
-limitations, and the impact that this interplay has on identifying
-selection and understanding demography, is not fully understood. Although a large body of
-research that involves searching the genomes of domesticated species for evidence
-of positive selection exists \cite{hufford2012, he2011, vigouroux2002, chapman2008}, these studies tend to focus on
-identifying or mapping particular genes or regions that play an
-important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
-selection have on whole-genome patterns of genetic variability remains limited.
+%% \dropcap{D}omesticated plant species evolve in a unique fashion
+%% compared to their wild counterparts \cite{doebley2006}.
+%% This is a result of both the anthropomorphic nature of artificial selection on
+%% domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
+%% bottleneck(s) that they tend to have experiencedtu
+%% \cite{ross2007}. However, the
+%% complex interplay between selective pressures and demographic
+%% limitations, and the impact that this interplay has on identifying
+%% selection and understanding demography, is not fully understood. Although a large body of
+%% research that involves searching the genomes of domesticated species for evidence
+%% of positive selection exists \cite{hufford2012, he2011, vigouroux2002, chapman2008}, these studies tend to focus on
+%% identifying or mapping particular genes or regions that play an
+%% important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
+%% selection have on whole-genome patterns of genetic variability remains limited.
 

### 2015-10-08 manuscript.tex
updates

@@
 \usepackage{pgf}
-%\usepackage{hyperref}
+%\usepackage{hyperref}scho
 \usepackage[normalem]{ulem}
@@
 \dropcap{T}he genetic properties of populations are determined by a constant interplay between genetic drift and selection.
-Drift is the consequence of a finite population size, such that random sampling of gametes every generation will incrementally lead to genetic changes.
+Drift is the consequence of a finite population size, such that random sampling of gametes every generation will incrementally lead to genetic changes \cite{dobzhansky1957}.
 In other words, the demographic history of a population, which includes its population size fluctuations and mating patterns, has a long-lasting impact on the patterns of DNA variability exhibited by that population.
@@
 In contrast to drift, which results from demographic history, selection is the result of the relationship between genotypes, phenoytpes, and ultimately fitness.
-Importantly, demography and selection do not operate independently to determine genetic variability.
+Importantly, demography and selection do not operate independently to determine genetic variability \cite{li2012}.
 Instead, these forces are engaged in a continuous battle by which each masks and obfuscates the outcomes of the other, ultimately meaning that the selective history of a species cannot be properly evaluated without an thorough estimate of demographic history, and demography cannot be evaluated without an understanding of selection.
@@
 %\subsection{Patterns of diversity differ between genic and  intergenic regions of the genome} %JRI
-To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012} \jri{add cite to HM3}.
-We find broad differences in genic and intergenic diversity consistent with earlier results  \cite{hufford2012}(Figure \ref{fig:diversity}).  \jri{pi graph should use greek letter $\pi$} 
+To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012, hapmap3}.
+We find broad differences in genic and intergenic diversity consistent with earlier results  \cite{hufford2012}(Figure \ref{fig:diversity}).
 In maize, mean pairwise diversity ($\pi$) within genes was significantly lower than at sites at least 5 kb away from genes (0.00668 vs 0.00691, $p<2\times 10^{-44}$).
@@
 \begin{center}
-  \includegraphics[width=.4\textwidth] {FigsAndFiles/Pi_and_Tajima.png}
+  \includegraphics[width=.45\textwidth] {FigsAndFiles/Pi_and_Tajima.png}
 \end{center}
@@
 To address this concern, we took advantage of genome-wide estimates of evolutionary constraint \cite{rodgers2015}, calculated using genomic evolutionary rate profile (GERP) scores \cite{davydov2010}. 
-We then evaluated substitutions only in subsets of genes in the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection  (Figure \ref{sFig:consUncons}).
-As expected, we see higher diversity around substitutions in genes under weak purifying selection \jri{there a table or figure we can reference?}, but we still find no difference between synonymous and missense substitutions in either subset of the data.
+We then evaluated substitutions only in subsets of genes in the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection. 
+As expected, we see higher diversity around substitutions in genes under weak purifying selection, but we still find no difference between synonymous and missense substitutions in either subset of the data (Figure \ref{sFig:consUncons}).
 Taken together, these data suggest hard sweeps do not play a major role in patterning genic diversity in either maize or teosinte.

### 2015-10-08 manuscript.tex
updates

@@
 
-\begin{figure}
+\begin{figure}[!tb]
 \begin{center}
@@
 We observe this pattern around the causative amino acid substitution in the the domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005origin, wang2015}.
-\mbh{Awesome}
 Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
@@
 
-\begin{figure*}
+\begin{figure*}[!tb]
 %\vspace*{.05in}
@@
 Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
-\mbh{keep an eye out for Brandon's ninjas}
 We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites.
-\mbh{weird to follow up a statement about incomplete demographic model with selection result}
+This implies that genic polymorphism data are unable to tell the complete or accurate demographic history of maize.
 Furthermore, by utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without requiring assumptions about its duration.
@@
 One surprising result from our model is the estimated timing of domestication at $\approx 15,000$ years before present.
-While this appears to conflict with archaeological \cite{piperno2009starch} estimates, we instead argue this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
+While this appears to conflict with archaeological estimates \cite{piperno2009starch}, we instead argue this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
 We also note that our result may also be inflated due to population structure, as our geographically diverse sample of teosinte may include populations diverged from those that gave rise to maize.

### 2015-10-17 manuscript.tex
updates

@@
 
-For instance, supposed neutrally-evolving DNA is often used to estimate historical demographic
-parameters of a population such as effective population size, structure, and expansion
-history \cite{luikart2003, gutenkunst2009}. However, researchers have
-called into question whether or not there are reasonable approaches
+Therefore, to estimate the historical demographic parameters of a population, researchers often utilize supposed neutrally-evolving DNA  \cite{luikart2003, gutenkunst2009}.
+However, it has been called into question whether reasonable approaches exist
 for identifying neutral regions of the genome, since the effects of
@@
 impacted by the effects of selection through linkage
-\cite{sella2009}. A natural next question, therefore, is how can
-demographic parameters be estimated independently of
-selective parameters? One technique utilized to address this
-problem is limiting analyses to only sites far from genes
+\cite{sella2009}. One technique that has been utilized to estimate demographic parameters independently of selective parameters is to limit analyses to only sites far from genes
 \cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is
-difficult to be certain that sites even in gene-poor regions of the
+difficult to be certain, without species-specific knowledge, that sites even in gene-poor regions of the
 genome are not influenced by linked selection. Ultimately, an understanding
@@
 relative importance of different forms of selection on whole-genome
-variability, using maize and teosinte as a model 2) research the impact that the
-demography history has had on genetic variability in maize, and how
-this compares to the impact of a different demographic history in
-its wild progenitor teosinte; and 3) precisely estimate the parameters of the
+variability, using maize and teosinte as a model; 2) research the impact that the
+demographic history has had on genetic variability in maize, and how
+this compares to the impact of a different demographic history in teosinte; and 3) precisely estimate the parameters of the
 maize domestication bottleneck.
@@
 Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
-\mbh{I'm curious whether linked selection is affecting the diversity around a sizable proportion of synonymous substitutions.
+
+%\mbh{I'm curious whether linked selection is affecting the diversity around a sizable proportion of synonymous substitutions.
 Would it be worthwhile to look at diversity around the subset of synonymous substitutions that are the least linked to missense substitutions just to make sure it is not substantially higher?
@@
 Qualitative differences between maize and teosinte remained unchanged even after removing genes in the top 20\% of H12 (Figure \ref{sFig:H12}A).
-\mbh{This makes me wonder about the top 20\%. Is the pattern of diversity reduction different in this fraction of the genome?  If so, that seems like something that should be mentioned.}
+%\mbh{This makes me wonder about the top 20\%. Is the pattern of diversity reduction different in this fraction of the genome?  If so, that seems like something that should be mentioned.}
+%TB Says: Sites in the top 20% of H12 are unlikely to be under background selection, so maize vs teo background selection pattern at these sites seems strange to investigate.
 We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.
@@
 To explore whether differences in the efficacy of purifying selection between maize and teosinte can be explained by demographic processes, we estimated the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}).
-The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ individuals.
-We estimate that the maize population split from teosinte $\approx 15,000$ generations in the past, with an initial size of only $\approx 5\% $ of the ancestral $N_a$. 
+The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ teosinte individuals.
+We estimate that maize split from teosinte $\approx 15,000$ generations in the past, with an initial size of only $\approx 5\% $ of the ancestral $N_a$. 
 After its split from teosinte, our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,000$.
-Maize and teosinte continue to exchange migrants after the population split, with gene flow between the populations estimated at $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation from teosinte to maize and $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants from maize to teosinte.
+Maize and teosinte have continued to exchange migrants after the population split, with gene flow between the populations estimated at $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation from teosinte to maize and $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants from maize to teosinte.
 
@@
 In addition to our simple bottleneck model, we investigated two alternative approaches for demographic inference.
-First, we took advantage of genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using low frequency variants  informative of population expansion.
+First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using low frequency variants  informative of population expansion.
 This analysis yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
@@
 Singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) appears nearly identical to results from pairwise diversity (Figure \ref{fig:hardSweeps}), providing little support for a substantial increase in the number or strength of hard sweeps occurring in maize.
-In contrast, we observe a qualitative shift in the effects of purifying selection: singleton polymorphisms are more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
-This relationship again remained after removing the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12})
+In contrast, we observe a significant shift in the effects of purifying selection: singleton polymorphisms are more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
+This result is the opposite of the pattern observed for $\pi$, where teosinte demonstrated a stronger reduction of diversity in and around genes than did maize.
+Again, this relationship remained after removing the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12}).
 Finally, while direct comparison of pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, as indicated by our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
@@
 Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with ``soft sweeps''  from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
-Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation as the primary source of adaptive variation.
-\mbh{But Sho's paper and Tanja's paper suggest non-coding variants are important for adaptation...should we mention this caveat?}
+Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation and/or non-coding variants as primary sources of adaptive variation.
 Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.  
 Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.]{studer2011}.
-Indeed, many adaptive traits in both maize \citex and teosinte \citex are highly quantitative \jri{cite wallace for maize?
-weber and/or lauter for teosinte?} and adaptation in both maize \cite{hufford2012} and teosinte \cite{pyhajarvi2013complex} has involved selection on regulatory variation.
-\jri{what \% of the windows of top H12 score genome-wide occur outside of genes?
-is that a number to include here?} \mbh{additional consideration of non-genic regions in both the results and discussion would probably be a good idea given Tanja and Sho's results}
+Indeed, many adaptive traits in both maize \cite{wallace2014} and teosinte \cite{weber2008} are highly quantitative,  and adaptation in both maize \cite{hufford2012} and teosinte \cite{pyhajarvi2013complex} has involved selection on regulatory variation.
+%\jri{what \% of the windows of top H12 score genome-wide occur outside of genes? is that a number to include here?}
+% TB Says: No, because windows were 200 SNPs so even windows that don't correspond to a gene could include a gene.
+
+%\mbh{additional consideration of non-genic regions in both the results and discussion would probably be a good idea given Tanja and Sho's results}
+% TB Says: I agree, but I don't think we have the data to really get at the relative contribution of non-genic regions.
 
@@
 \subsection{Demography influences the efficiency of purifying selection}
-\jri{section needs help --  i'm not happy with the writing} 
 We find a more pronounced decrease in nucleotide diversity around genes in teosinte compared to domesticated maize (Figure \ref{fig:purify}A).
@@
 
-\jri{I would like to see a supplemental figure or maybe just number here on the partial correlation in both teosinte and maize of diversity with recombination after accounting for gene density.
-do this in 500kb windows and then in 10kb windows.
-we should do the correlation with both pi and singletons.
-This let's us compare more explictly with studies like C-D who evaluate overall patterns of linked selection.
-will be especially cool if we don't see a big difference in big windows but we do in 10kb windows -- siggests C-D looked at wrong scale  }
 A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more weakly deleterious alleles segregating at intermediate frequency than does teosinte.
@@
 Additionally, the large $N_e$ of modern maize compared to modern teosinte implies that for new mutations, selection will operate much more efficiently in maize.
-\jri{inbreds stuff is OK, but should be shortened.
-we are aiming at a broader audience that doesn't care about maize.
-what are more/broader.
-Chia, Rogers-Melnick, Mezmouk all looked at load.
-does this have implications for large effect load (i.e.
-should bneck have purged those?), for heterosis, architecture of quant.
-traits, etc.}
 
@@
 mention C-D ignored recent demography -- here's where rec/pi windows could help.}
+
+\jri{I would like to see a supplemental figure or maybe just number here on the partial correlation in both teosinte and maize of diversity with recombination after accounting for gene density.
+do this in 500kb windows and then in 10kb windows.
+we should do the correlation with both pi and singletons.
+This let's us compare more explictly with studies like C-D who evaluate overall patterns of linked selection.
+will be especially cool if we don't see a big difference in big windows but we do in 10kb windows -- siggests C-D looked at wrong scale  }
+
+\jri{inbreds stuff is OK, but should be shortened.
+we are aiming at a broader audience that doesn't care about maize.
+what are more/broader.
+Chia, Rogers-Melnick, Mezmouk all looked at load.
+does this have implications for large effect load (i.e.
+should bneck have purged those?), for heterosis, architecture of quant.
+traits, etc.}
 

### 2015-10-17 manuscript.tex
updates

@@
 \newcommand{\mbh}[1]{\textcolor{blue}{\scriptsize #1}}
+\newcommand{\tmb}[1]{\textcolor{brown}{\scriptsize #1}}
 \newcommand{\citex}{\textcolor{red}{\bf CITE}}
@@
 Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.
+\tmb{Jeff, you put this here, but as written I don't think we need to mention $s_b$}
 Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.]{studer2011}.
 Indeed, many adaptive traits in both maize \cite{wallace2014} and teosinte \cite{weber2008} are highly quantitative,  and adaptation in both maize \cite{hufford2012} and teosinte \cite{pyhajarvi2013complex} has involved selection on regulatory variation.
+
+The absence of evidence for a genome-wide impact of hard sweeps differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
+\cite{corbett2015} recently showed that selection tends to have an elevated impact reducing diversity in species with large population size compared to small, which may explain why maize and humans, both bottlenecked species, show similar patterns while drosophila and capsella behave differently.
+
 %\jri{what \% of the windows of top H12 score genome-wide occur outside of genes? is that a number to include here?}
@@
 We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites.
-This implies that genic polymorphism data are unable to tell the complete or accurate demographic history of maize.
+This implies that genic polymorphism data are unable to tell the complete or accurate demographic history of maize, but the rapid recovery of diversity outside of genes demonstrates that sites far from genes can be reasonably used for demographic inference.
 Furthermore, by utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without requiring assumptions about its duration.
@@
 One surprising result from our model is the estimated timing of domestication at $\approx 15,000$ years before present.
-While this appears to conflict with archaeological estimates \cite{piperno2009starch}, we instead argue this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
+While this appears to conflict with archaeological estimates \cite{piperno2009starch}, we emphasize that this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
 We also note that our result may also be inflated due to population structure, as our geographically diverse sample of teosinte may include populations diverged from those that gave rise to maize.
@@
 \subsection{Demography influences the efficiency of purifying selection}
-We find a more pronounced decrease in nucleotide diversity around genes in teosinte compared to domesticated maize (Figure \ref{fig:purify}A).
-This difference is likely due to the effects of purifying selection.
-Purifying selection is more efficient in larger populations  \cite{kimura1984}, and our observation is thus consistent with our demographic model and previous work \cite{eyre1998, tenaillon2004selection, wright2005, ross2009} showing a smaller long-term effective population size in maize.
-To understand how recent population growth has changed the effects of linked selection, we also analyzed diversity of singleton polymorphisms.
-Under a simple neutral model, lower frequency polymorphisms are expected to be younger, and patterns of diversity at singleton SNPs should thus represent recent evolutionary history.
-Consistent with its larger current size and thus more efficient purifying selection, we see a stronger decline of singleton diversity around genes in maize (Figure \ref{fig:purify}B than teosinte.
+One of our more striking findings is that the impact of purifying selection on maize and teosinte variability has varied through time.
+Although we observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), we see a greater decrease in singleton diversity in maize than teosinte (Figure \ref{fig:purify}B).
+Purifying selection is more efficient in larger populations  \cite{kimura1984}, and our observations using $\pi$  are thus consistent with our demographic model and previous work \cite{eyre1998, tenaillon2004selection, wright2005, ross2009} which shows a smaller long-term effective population size in maize.
+Similarly, our observations based on singleton diversity, which thereby reflect more recent patterns of selection, again are in agreement with maize and teosinte demography; from the viewpoint of a recent mutation, the rapid expansion of maize post-domestication means that today maize has a vastly larger effective population size than does teosinte.
 
-A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more weakly deleterious alleles segregating at intermediate frequency than does teosinte.
-This could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{meghji1984}; if deleterious alleles tend to be recessive and are particularly frequent, they will have ample opportunities to display their phenotypes in inbreds.
+A consequence of the previously-inefficient purifying selection in maize is that it likely harbors more weakly deleterious alleles segregating at intermediate frequency than does teosinte.
+Therefore, high genetic load could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{meghji1984}.
 Our results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously selected.
 Additionally, the large $N_e$ of modern maize compared to modern teosinte implies that for new mutations, selection will operate much more efficiently in maize.
+This observation is not restricted to maize.
+Bottlenecks are have been documented or predicted across most domesticated plant and animal species \cite{doebley2006, hyten2006, zhu2007, bovine2009, larson2013 }, but rapid expansion post-bottleneck, which we demonstrate generates ubiquitous consequences relating to genetic diversity, is not well studied.
+In human, model, and other non-domesticated species, these results demonstrate that observed patterns of selection may depend heavily on recent and/or ancient patterns of demography.
 
-The absence of evidence for a genome-wide impact of hard sweeps differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
-In a related study, \cite{corbett2015} showed that selection tends to have an elevated impact on reducing diversity in species with large population size compared to small.
-This may explain why maize and humans, both bottlenecked species, show similar patterns while drosophila and capsella behave differently.
-\textcolor{red}{However, we also do not observe abundant hard sweeps in teosinte, which calls this explanation into question.}
-
-\jri{these results have implications for lots of studies.
+%\jri{these results have implications for lots of studies.
 recent demographic history often ignored (examples).
@@
 
-\jri{I would like to see a supplemental figure or maybe just number here on the partial correlation in both teosinte and maize of diversity with recombination after accounting for gene density.
+%\jri{I would like to see a supplemental figure or maybe just number here on the partial correlation in both teosinte and maize of diversity with recombination after accounting for gene density.
 do this in 500kb windows and then in 10kb windows.
@@
 will be especially cool if we don't see a big difference in big windows but we do in 10kb windows -- siggests C-D looked at wrong scale  }
-
-\jri{inbreds stuff is OK, but should be shortened.
-we are aiming at a broader audience that doesn't care about maize.
-what are more/broader.
-Chia, Rogers-Melnick, Mezmouk all looked at load.
-does this have implications for large effect load (i.e.
-should bneck have purged those?), for heterosis, architecture of quant.
-traits, etc.}
 

### 2015-10-18 manuscript.tex
updates

@@
 We estimate that maize underwent a bottleneck of 5\% of the size of the ancestral teosinte population, but that recent expansion has resulted in a maize population perhaps orders of magnitude larger than teosinte.
-We show that positive selection on new mutations has had relatively little effect on genetic diversity, but that selection against deleterious mutations has dramatically reduced diversity in and immediately around genes in both taxa.
+We show that positive selection on new genic mutations has had relatively little effect on genetic diversity, but that selection against deleterious mutations has dramatically reduced diversity in and immediately around genes in both taxa.
 We find that the relative effect of selection depend qualitatively on the age of the polymorphisms evaluated: while older polymorphisms in maize show more limited effects of linked selection, new mutations instead reflect its larger current size and more efficient purifying selection.
@@
 \begin{abstract}
-  Unique selective and demographic forces operate on domesticated plant species.
-\mbh{I don't agree with this statement; characteristics of demography and selection vary across crops and I wouldn't be surprised to find similar demographies and patterns of selection between crop and wild species} These forces interact during and after domestication to generate the patterns of DNA variability that are persistent today.
-To quantify the interplay between demography and selection, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide.
+  The genetic variability present in a population is increases through mutation and drift, and generally reduced via selection.
+But these forces do not act in a vacuum, and instead interact continuously to shape patterns of genetic diversity.
+To quantify the interplay between demography (which drives drift) and selection, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide.
 Our sample included whole genome sequence data from 23 maize and 13 teosinte individuals.
@@
 However, expansion after domestication has since increased the efficiency of purifying selection to levels exceeding those seen in teosinte.
-Our results demonstrate that particularly in domesticated species or bottlenecked species \mbh{right, observations can be generalized to wild species as well; I don't think there's any need to say crops are a special case}, demographic and selective history in the ancient and recent past both contribute to genetic variability that is present today, providing substantial implications for the continued improvement of domesticated species.
-\mbh{I'd end this more generally, expressing the importance of knowing demography when characterizing selection genome-wide}
+Our results demonstrate that demographic and selective history in the ancient and recent past both contribute to the levels of genetic variability that are present today.
+Our results also show that for bottlenecked species, rapid post-bottleneck expansion can have wide-ranging impacts on diversity that conflict with the expected loss of diversity that occurs during the bottleneck itself.
 
@@
 Drift is the consequence of a finite population size, such that random sampling of gametes every generation will incrementally lead to genetic changes \cite{dobzhansky1957}.
-In other words, the demographic history of a population, which includes its population size fluctuations and mating patterns, has a long-lasting impact on the patterns of DNA variability exhibited by that population.
+In other words, the demographic history of a population, which includes its population size fluctuations and mating patterns, has a long-lasting impact on the patterns of DNA variability displayed by that population.
 Conversely, selection systematically alters a population by favoring particular alleles at the expense of others.
@@
 Importantly, demography and selection do not operate independently to determine genetic variability \cite{li2012}.
-Instead, these forces are engaged in a continuous battle by which each masks and obfuscates the outcomes of the other, ultimately meaning that the selective history of a species cannot be properly evaluated without an thorough estimate of demographic history, and demography cannot be evaluated without an understanding of selection.
+Instead, these forces are engaged in a continuous battle by which each masks and obfuscates the outcomes of the other, ultimately meaning that the selective history of a species cannot be properly evaluated without a thorough estimate of demographic history, and demography cannot be evaluated without an understanding of selection.
 
-Therefore, to estimate the historical demographic parameters of a population, researchers often utilize supposed neutrally-evolving DNA  \cite{luikart2003, gutenkunst2009}.
+For instance, to estimate the historical demographic parameters of a population, researchers often utilize supposed neutrally-evolving DNA  \cite{luikart2003, gutenkunst2009}.
 However, it has been called into question whether reasonable approaches exist
 for identifying neutral regions of the genome, since the effects of
-selection can be wide-ranging \cite{li2012, slotte2014}. For example, in
-\emph{Drosophila} it appears that the majority of the genome is
+selection can be wide-ranging \cite{li2012, slotte2014}. In \emph{Drosophila}, for example, the majority of the genome is
 impacted by the effects of selection through linkage
@@
 reasonable demographic inference.
+At the same time, since the effects of demography alter genetic variability genome-wide \cite{akey2009}, and because the efficiency of selection depends on demographic characteristics such as population size \cite{kimura1963}, selection is equally difficult to quantify without properly accounting for demographic history.
 
-Maize represents an excellent organism to study these phenomena.
+Maize (\emph{Zea mays}) represents an excellent organism to study this interplay.
 Maize is a species of tremendous importance worldwide as both a staple crop \cite{shiferaw2011} and as a model for understanding plant evolution \cite{strable2009}.
@@
 regions \cite{hufford2012} have been identified. But despite all that is
-known about maize domestication, the parameters of the
+known about maize domestication, the demographic parameters of the
 domestication process remain uncertain. Specifically, the size of the
@@
 At the same time, the predominant form or forms of selection that have taken place during and since domestication remain unknown.
-Sequence information from maize and teosinte may therefore be utilized to address these questions.
+Sequence information from maize and teosinte may therefore be utilized to address these questions, and to better explain the interaction between selection and demography in a system with complex but quantifiable selective and demographic histories.
 
@@
 \includegraphics[width=.4\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
-\caption{Parameter estimates for a simple bottleneck model of maize domestication. See methods for details. \label{fig:bottleneck} }
+\caption{Parameter estimates for a basic bottleneck model of maize domestication. See methods for details. \label{fig:bottleneck} }
 \end{figure*}
@@
 Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.
-\tmb{Jeff, you put this here, but as written I don't think we need to mention $s_b$}
+\tmb{Jeff, rewrote this part, and I like your writing here, but as written I don't think we need to mention $s_b?$}
 Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.]{studer2011}.
@@
 Bottlenecks are have been documented or predicted across most domesticated plant and animal species \cite{doebley2006, hyten2006, zhu2007, bovine2009, larson2013 }, but rapid expansion post-bottleneck, which we demonstrate generates ubiquitous consequences relating to genetic diversity, is not well studied.
-In human, model, and other non-domesticated species, these results demonstrate that observed patterns of selection may depend heavily on recent and/or ancient patterns of demography.
+In human, model, and other species, these results demonstrate that observed patterns of selection depend heavily on recent and/or ancient patterns of demography.
 
@@
 All bam files are available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}).
-\jri{some clarification here.
-did we use Lemmon's higher depth teosinte?
-If so, why does Li say we only had 7 high-depth bam files?
-shouldn't there be more?}
+
+%\jri{some clarification here. did we use Lemmon's higher depth teosinte? If so, why does Li say we only had 7 high-depth bam files?  shouldn't there be more?}
+
+% TB Says: Most bam files were not high-depth, but a few were.
+I used what was made available from Panzea group via iplant in summer 2014, and precise files that were used are linked and described in table.
+ANGSD GL approach makes it OK that depth varied.
 

### 2015-10-19 manuscript.tex
typos and little things.

@@
 \begin{abstract}
-  The genetic variability present in a population is increases through mutation and drift, and generally reduced via selection.
-But these forces do not act in a vacuum, and instead interact continuously to shape patterns of genetic diversity.
+  The genetic variability present in a population results from a constant interaction between drift and selection.
 To quantify the interplay between demography (which drives drift) and selection, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide.
@@
 \dropcap{T}he genetic properties of populations are determined by a constant interplay between genetic drift and selection.
-Drift is the consequence of a finite population size, such that random sampling of gametes every generation will incrementally lead to genetic changes \cite{dobzhansky1957}.
+Drift is the consequence of a finite population size, such that random sampling of gametes every generation will incrementally lead to changes in genetic variabiligy \cite{dobzhansky1957}.
 In other words, the demographic history of a population, which includes its population size fluctuations and mating patterns, has a long-lasting impact on the patterns of DNA variability displayed by that population.
@@
 Importantly, demography and selection do not operate independently to determine genetic variability \cite{li2012}.
-Instead, these forces are engaged in a continuous battle by which each masks and obfuscates the outcomes of the other, ultimately meaning that the selective history of a species cannot be properly evaluated without a thorough estimate of demographic history, and demography cannot be evaluated without an understanding of selection.
+Instead, each of these forces is constantly masking and obfuscating the other, ultimately meaning that the selective history of a species cannot be properly evaluated without a thorough estimate of demographic history, and demography cannot be evaluated without an understanding of selection.
 
@@
 relative importance of different forms of selection on whole-genome
-variability, using maize and teosinte as a model; 2) research the impact that the
+variability, using maize and teosinte as a model; 2) research the impact that
 demographic history has had on genetic variability in maize, and how
@@
 maize domestication bottleneck.
-Gaining an improved understanding of the interaction between demography and selection in maize will help to increase the understanding of these phenomenon across a wide array of organisms.
+Gaining an improved understanding of the interaction between demography and selection in maize will help to increase the understanding of these phenomena across a wide array of organisms.
 
@@
 If a proportion of missense mutations have been fixed due to hard sweeps, diversity around these substitutions should be lower than around synonymous substitutions. 
-We observe this pattern around the causative amino acid substitution in the the domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005origin, wang2015}.
+We observe this pattern around the causative amino acid substitution in the the maize domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005origin, wang2015}.
 Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
@@
 
-Selection can also reduce diversity in functional regions of the genome via removal of deleterious mutations, a process known as purifying or background selection \cite{charlesworth1993}.
+In the case of purifying or background selection, diversity can be reduced in functional regions of the genome via removal of deleterious mutations \cite{charlesworth1993}.
 We investigated purifying selection in maize and teosinte by evaluating the reduction of diversity around genes.
@@
 This result is the opposite of the pattern observed for $\pi$, where teosinte demonstrated a stronger reduction of diversity in and around genes than did maize.
-Again, this relationship remained after removing the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12}).
+As was the case when we analyzed $\pi$, this relationship remained after we removed the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12}).
 Finally, while direct comparison of pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, as indicated by our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
@@
 The absence of evidence for a genome-wide impact of hard sweeps differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
-\cite{corbett2015} recently showed that selection tends to have an elevated impact reducing diversity in species with large population size compared to small, which may explain why maize and humans, both bottlenecked species, show similar patterns while drosophila and capsella behave differently.
+\cite{corbett2015} recently showed that selection tends to have an elevated impact reducing diversity in species with large population size compared to small, which may explain why maize and humans, both bottlenecked species, show similar patterns while \emph{drosophila} and \emph{capsella} behave differently.
 
@@
 This observation is not restricted to maize.
-Bottlenecks are have been documented or predicted across most domesticated plant and animal species \cite{doebley2006, hyten2006, zhu2007, bovine2009, larson2013 }, but rapid expansion post-bottleneck, which we demonstrate generates ubiquitous consequences relating to genetic diversity, is not well studied.
+Bottlenecks have been documented or predicted across most domesticated plant and animal species \cite{doebley2006, hyten2006, zhu2007, bovine2009, larson2013 }, but rapid expansion post-bottleneck, which we demonstrate generates ubiquitous consequences relating to genetic diversity, is not well studied.
 In human, model, and other species, these results demonstrate that observed patterns of selection depend heavily on recent and/or ancient patterns of demography.

### 2015-10-28 manuscript.tex
intro updated

@@
 
-\title{Demography and linked selection in wild and domesticated maize}
+\title{Recent demography drives changes in linked selection across the maize genome}
 \author{Timothy M.
-Beissinger\affil{1}{Dept.
-of Plant Sciences, University of California, Davis, CA, USA}\affil{2}{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}\affil{3}{Division of Plant Sciences, University of Missouri, Columbia, MO, USA}, Li Wang\affil{4}{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA}\affil{5}{Genome Informatics Facility, Iowa State University, Ames, IA, USA}, Kate Crosby\affil{1}{}, Arun Durvasula\affil{1}{}, Matthew B.
+Beissinger\affil{1}{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}\affil{2}{Division of Plant Sciences, University of Missouri, Columbia, MO, USA}\affil{3}{Dept.
+of Plant Sciences, University of California, Davis, CA, USA}, Li Wang\affil{4}{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA}\affil{5}{Genome Informatics Facility, Iowa State University, Ames, IA, USA}, Kate Crosby\affil{1}{}, Arun Durvasula\affil{1}{}, Matthew B.
 Hufford\affil{4}{}, \and Jeffrey Ross-Ibarra\affil{1}{}\affil{6}{Genome Center and Center for Population Biology, University of
@@
 In contrast to drift, which results from demographic history, selection is the result of the relationship between genotypes, phenoytpes, and ultimately fitness.
-Importantly, demography and selection do not operate independently to determine genetic variability \cite{li2012}.
-Instead, each of these forces is constantly masking and obfuscating the other, ultimately meaning that the selective history of a species cannot be properly evaluated without a thorough estimate of demographic history, and demography cannot be evaluated without an understanding of selection.
+A classical approach for studying selection is to remove the genome-wide effects of demography and assume that selection will leave locus-specific evidence \cite{Wall2002, akey2009}.
+Conversely, demography is often researched by excluding potentially selected sites \cite{voight2005, luikart2003, gutenkunst2009}.
 
-For instance, to estimate the historical demographic parameters of a population, researchers often utilize supposed neutrally-evolving DNA  \cite{luikart2003, gutenkunst2009}.
-However, it has been called into question whether reasonable approaches exist
-for identifying neutral regions of the genome, since the effects of
-selection can be wide-ranging \cite{li2012, slotte2014}. In \emph{Drosophila}, for example, the majority of the genome is
+However, recent work has shown that demography and selection do not operate independently to determine genetic variability, in large part because linkage allows the effects of selection to be wide-ranging \cite{li2012,slotte2014}.
+For example, in \emph{Drosophila} the majority of the genome is
 impacted by the effects of selection through linkage
-\cite{sella2009}. One technique that has been utilized to estimate demographic parameters independently of selective parameters is to limit analyses to only sites far from genes
-\cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is
-difficult to be certain, without species-specific knowledge, that sites even in gene-poor regions of the
-genome are not influenced by linked selection. Ultimately, an understanding
-of which sites have the potential to be adaptive, and how these affect
-genome-wide patterns of variability through linkage, is required for
-reasonable demographic inference.
-At the same time, since the effects of demography alter genetic variability genome-wide \cite{akey2009}, and because the efficiency of selection depends on demographic characteristics such as population size \cite{kimura1963}, selection is equally difficult to quantify without properly accounting for demographic history.
+\cite{sella2009}.
+One technique that has been utilized to estimate demographic parameters independently of selective parameters is to limit analyses to only sites far from genes\cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is
+difficult to be certain, without species-specific knowledge, that sites even in gene-poor regions of the genome are not influenced by linked selection.
+Therefore, characterizing selection, demography or both, first requires an understanding of linked selection.
+It has already been established that patterns of linked selection are heavily influenced by population size ($N_e$) \cite{cutter2013, slotte2014}, but nearly all species including humans \cite{reich1998}, domesticates \cite{hyten2006, bovine2009}, ecological \cite{storfer2014}, and model species \cite{duchen2013} have experienced recent or ancient changes in $N_e$.
+Although much is known about how the long-term average $N_e$ affects linked selection, much less is understood about the immediate effects and dynamics of changing in $N_e$ on patterns of linked selection.
 
-Maize (\emph{Zea mays}) represents an excellent organism to study this interplay.
-Maize is a species of tremendous importance worldwide as both a staple crop \cite{shiferaw2011} and as a model for understanding plant evolution \cite{strable2009}.
-Archaeological and genetic studies have
-established that maize domestication is likely to have taken place in
- Mexico approximately 9,000 years bp
-\cite{smith1995,matsuoka2002}. Teosinte, the most
-recent wild ancestor to maize, remains extant throughout much of the
-Americas \cite{wilkes1967}. Additionally, several large-effect
+Because of its relatively simple demographic history and abundant availability of genomic tools, maize (\emph{Zea mays}) represents an excellent organism to study this phenomenon.
+Maize is a species of tremendous importance worldwide as both a model species \cite{strable2009} and a staple crop \cite{shiferaw2011}.
+Archaeological and genetic studies have established that maize domestication began in
+Central Mexico at least 9,000 years bp
+\cite{smith1995,matsuoka2002}, and involved a population bottleneck followed by recent expansion \cite{wright2005}.
+Because of this simple but dynamic demographic history, patterns of linked selection in wild and domesticated maize can be used to understand the effects of changing $N_e$ on linked selection.
+Teosinte, which remains extant throughout much of the Americas \cite{wilkes1967}, is the wild ancestor to maize.
+Additionally, several large-effect
 domestication loci \cite{doebley1995, wills2013, wang2015} and putative domestication
-regions \cite{hufford2012} have been identified. But despite all that is
+regions \cite{hufford2012} have been identified in maize. But despite all that is
 known about maize domestication, the demographic parameters of the
@@
 population size of modern maize.
-At the same time, the predominant form or forms of selection that have taken place during and since domestication remain unknown.
-Sequence information from maize and teosinte may therefore be utilized to address these questions, and to better explain the interaction between selection and demography in a system with complex but quantifiable selective and demographic histories.
+At the same time, the predominant form or forms of selection that have taken place during and since domestication, and how linkage impacts the evidence and patterns of selection across the genome, remain unknown.
 
-To that end, the objectives of our study were to 1) investigate the
-relative importance of different forms of selection on whole-genome
-variability, using maize and teosinte as a model; 2) research the impact that
-demographic history has had on genetic variability in maize, and how
-this compares to the impact of a different demographic history in teosinte; and 3) precisely estimate the parameters of the
-maize domestication bottleneck.
-Gaining an improved understanding of the interaction between demography and selection in maize will help to increase the understanding of these phenomena across a wide array of organisms.
-
-%% \dropcap{D}omesticated plant species evolve in a unique fashion
-%% compared to their wild counterparts \cite{doebley2006}.
-%% This is a result of both the anthropomorphic nature of artificial selection on
-%% domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
-%% bottleneck(s) that they tend to have experiencedtu
-%% \cite{ross2007}. However, the
-%% complex interplay between selective pressures and demographic
-%% limitations, and the impact that this interplay has on identifying
-%% selection and understanding demography, is not fully understood. Although a large body of
-%% research that involves searching the genomes of domesticated species for evidence
-%% of positive selection exists \cite{hufford2012, he2011, vigouroux2002, chapman2008}, these studies tend to focus on
-%% identifying or mapping particular genes or regions that play an
-%% important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
-%% selection have on whole-genome patterns of genetic variability remains limited.
+For this study, we begin by estimating the parameters of maize domestication based on whole-genome sequencing data.
+Next, we utilized these estimates to 1) investigate the relative importance of different forms of selection on genetic
+variability in the recent and acient past; and 2) research the impact that demographic history, specifically changes in $N_e$, has had on linked selection in maize, and how this compares to the impact of a different demographic history in teosinte.
+Gaining an improved understanding of the impacts of a dynamic demography on whole-genome variability in maize will help to increase the understanding of these phenomena across all organisms, since changes in $N_e$ are pervasive.
 
 \section{Results}
-%\subsection{Patterns of diversity differ between genic and  intergenic regions of the genome} %JRI
+
+\subsection{Demography of maize domestication} %JRI
+To explore whether differences in the efficacy of purifying selection between maize and teosinte can be explained by demographic processes, we estimated the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}).
+The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ teosinte individuals.
+We estimate that maize split from teosinte $\approx 15,000$ generations in the past, with an initial size of only $\approx 5\% $ of the ancestral $N_a$. 
+After its split from teosinte, our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,000$.
+Maize and teosinte have continued to exchange migrants after the population split, with gene flow between the populations estimated at $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation from teosinte to maize and $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants from maize to teosinte.
+
+\begin{figure*}[!tb]
+%\vspace*{.05in}
+\centering
+\includegraphics[width=.4\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
+\caption{Parameter estimates for a basic bottleneck model of maize domestication. See methods for details. \label{fig:bottleneck} }
+\end{figure*}
+
+In addition to our simple bottleneck model, we investigated two alternative approaches for demographic inference.
+First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using low frequency variants  informative of population expansion.
+This analysis yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
+Finally, we applied a model-free coalescent approach \cite{schiffels2014} using a subset of our samples. 
+Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
+
+\subsection{Patterns of diversity differ between genic and  intergenic regions of the genome} %JRI
 To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012, hapmap3}.
@@
 
-\subsection{Demography of maize domestication} %JRI
-To explore whether differences in the efficacy of purifying selection between maize and teosinte can be explained by demographic processes, we estimated the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}).
-The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ teosinte individuals.
-We estimate that maize split from teosinte $\approx 15,000$ generations in the past, with an initial size of only $\approx 5\% $ of the ancestral $N_a$. 
-After its split from teosinte, our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,000$.
-Maize and teosinte have continued to exchange migrants after the population split, with gene flow between the populations estimated at $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation from teosinte to maize and $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants from maize to teosinte.
-
-\begin{figure*}[!tb]
-%\vspace*{.05in}
-\centering
-\includegraphics[width=.4\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
-\caption{Parameter estimates for a basic bottleneck model of maize domestication. See methods for details. \label{fig:bottleneck} }
-\end{figure*}
-
-In addition to our simple bottleneck model, we investigated two alternative approaches for demographic inference.
-First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using low frequency variants  informative of population expansion.
-This analysis yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
-Finally, we applied a model-free coalescent approach \cite{schiffels2014} using a subset of our samples. 
-Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
-
 \subsection{Population expansion leads to stronger purifying selection in modern maize} %JRI
@@
 Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.
-\tmb{Jeff, rewrote this part, and I like your writing here, but as written I don't think we need to mention $s_b?$}
 Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.]{studer2011}.

### 2015-10-29 manuscript.tex
comments

@@
 A classical approach for studying selection is to remove the genome-wide effects of demography and assume that selection will leave locus-specific evidence \cite{Wall2002, akey2009}.
-Conversely, demography is often researched by excluding potentially selected sites \cite{voight2005, luikart2003, gutenkunst2009}.
+Conversely, demography is often researched by excluding potentially selected sites \cite{voight2005, luikart2003, gutenkunst2009}.\jri{pretty good first paragraph.
+i will edit later, but i think more or less OK}
 
 However, recent work has shown that demography and selection do not operate independently to determine genetic variability, in large part because linkage allows the effects of selection to be wide-ranging \cite{li2012,slotte2014}.
-For example, in \emph{Drosophila} the majority of the genome is
-impacted by the effects of selection through linkage
+\jri{need to cite older stuff.
+this isn't new.
+what about maynard smith and haigh?
+gillespie?
+lots of popgen on this}
+For example, in \emph{Drosophila} the majority of the genome is impacted by the effects of selection through linkage
 \cite{sella2009}.
@@
 Although much is known about how the long-term average $N_e$ affects linked selection, much less is understood about the immediate effects and dynamics of changing in $N_e$ on patterns of linked selection.
+\jri{main points are there.
+tempted to split into two paragraphs.
+1) selection/drift not independent because of linked selection.
+very briefly explain it -- bgs vs purifying vs hard/soft.
+linked selection impacted by Ne.
+BGS models, hard sweep models, standing variaton (hermisson/pennings petrov/messer) etc.
+2) but always assume long-term Ne.
+pops have gone through demography, this has potential to impact both selection and drift}
 
@@
 At the same time, the predominant form or forms of selection that have taken place during and since domestication, and how linkage impacts the evidence and patterns of selection across the genome, remain unknown.
-
 For this study, we begin by estimating the parameters of maize domestication based on whole-genome sequencing data.
@@
 Gaining an improved understanding of the impacts of a dynamic demography on whole-genome variability in maize will help to increase the understanding of these phenomena across all organisms, since changes in $N_e$ are pervasive.
+\jri{pretty good, but will edit.}
 

### 2015-10-30 manuscript.tex
incorporate
Incorporated Jeff's latest comments into the introduction.

--- 2015-10-28 manuscript.tex
@@
 
-%\jri{we need a starting paragraph on linked selection and demography.
-check out slotte 2014 briefings functional genomics for a great review of ideas.
-then we do a paragraph on why domesticates are great for this (modified version of your below)}
-%\mbh{Yes, agree here...think the topics addressed in the manuscript need to be set up as generalizable and not just specific to domesticates.
-Crops are a great system because we know a fair bit more about demography and the process of selection than in wild species, but they are not entirely unique.}
-
 \dropcap{T}he genetic properties of populations are determined by a constant interplay between genetic drift and selection.
-Drift is the consequence of a finite population size, such that random sampling of gametes every generation will incrementally lead to changes in genetic variabiligy \cite{dobzhansky1957}.
+Drift is the consequence of a finite population size, such that random sampling of gametes every generation will incrementally lead to changes in genetic variability \cite{dobzhansky1957}.
 In other words, the demographic history of a population, which includes its population size fluctuations and mating patterns, has a long-lasting impact on the patterns of DNA variability displayed by that population.
@@
 
-However, recent work has shown that demography and selection do not operate independently to determine genetic variability, in large part because linkage allows the effects of selection to be wide-ranging \cite{li2012,slotte2014}.
-For example, in \emph{Drosophila} the majority of the genome is
-impacted by the effects of selection through linkage
-\cite{sella2009}.
-One technique that has been utilized to estimate demographic parameters independently of selective parameters is to limit analyses to only sites far from genes\cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is
-difficult to be certain, without species-specific knowledge, that sites even in gene-poor regions of the genome are not influenced by linked selection.
+However, demography and selection do not operate independently to determine genetic variability, in large part because linkage allows the effects of selection to be wide-ranging \cite{smith1974, li2012,slotte2014}.
+Linked selection can take the form of hitch-hiking, for which the frequency of neutral alleles changes as a result of positive selection at a physically linked site \cite{smith1974}, or background selection, where case diversity is reduced at loci physically linked to a site undergoing purifying selection \cite{charlesworth1993}.
+\emph{Drosophila} is an example of an organism for which linked selection is extremely important, as the majority of the genome has been impacted by it \cite{sella2009}.
+One technique that has been utilized to disentangle selection from demography is to limit analyses to only sites far from genes\cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is difficult to be certain, without species-specific knowledge, that sites even in gene-poor regions of the genome are not influenced by linked selection.
 Therefore, characterizing selection, demography or both, first requires an understanding of linked selection.
-It has already been established that patterns of linked selection are heavily influenced by population size ($N_e$) \cite{cutter2013, slotte2014}, but nearly all species including humans \cite{reich1998}, domesticates \cite{hyten2006, bovine2009}, ecological \cite{storfer2014}, and model species \cite{duchen2013} have experienced recent or ancient changes in $N_e$.
+
+One factor that heavily influences patterns of linked selection is population size ($N_e$) \cite{cutter2013, slotte2014}.
+but nearly all species including humans \cite{reich1998}, domesticates \cite{hyten2006, bovine2009}, ecological \cite{storfer2014}, and model species \cite{duchen2013} have experienced recent or ancient changes in $N_e$.
 Although much is known about how the long-term average $N_e$ affects linked selection, much less is understood about the immediate effects and dynamics of changing in $N_e$ on patterns of linked selection.
+
+%\jri{main points are there.
+tempted to split into two paragraphs.
+1) selection/drift not independent because of linked selection.
+very briefly explain it -- bgs vs purifying vs hard/soft.
+linked selection impacted by Ne.
+BGS models, hard sweep models, standing variaton (hermisson/pennings petrov/messer) etc.
+2) but always assume long-term Ne.
+pops have gone through demography, this has potential to impact both selection and drift}
 

### 2015-11-02 manuscript.tex
intro

--- 2015-10-28 manuscript.tex
@@
 Here we document the combined effects of demography and linked selection on genome-wide diversity in domesticated maize and its wild ancestor teosinte.
-We estimate that maize underwent a bottleneck of 5\% of the size of the ancestral teosinte population, but that recent expansion has resulted in a maize population perhaps orders of magnitude larger than teosinte.
+We estimate that maize underwent a bottleneck to  $\approx 5\%$ of the size of the ancestral teosinte population, but that recent expansion has resulted in a maize population perhaps orders of magnitude larger than teosinte.
 We show that positive selection on new genic mutations has had relatively little effect on genetic diversity, but that selection against deleterious mutations has dramatically reduced diversity in and immediately around genes in both taxa.
-We find that the relative effect of selection depend qualitatively on the age of the polymorphisms evaluated: while older polymorphisms in maize show more limited effects of linked selection, new mutations instead reflect its larger current size and more efficient purifying selection.
+We find that the relative effect of selection depends qualitatively on the age of the polymorphisms evaluated: while older polymorphisms in maize show more limited effects of linked selection, new mutations instead reflect its larger current size and more efficient selection.
 Our results demonstrate that a complete understanding of genome-wide patterns of diversity will require careful assessment of both demographic history and the effects of linked selection.
@@
 
-%\jri{we need a starting paragraph on linked selection and demography.
-check out slotte 2014 briefings functional genomics for a great review of ideas.
-then we do a paragraph on why domesticates are great for this (modified version of your below)}
-%\mbh{Yes, agree here...think the topics addressed in the manuscript need to be set up as generalizable and not just specific to domesticates.
-Crops are a great system because we know a fair bit more about demography and the process of selection than in wild species, but they are not entirely unique.}
+\dropcap{T}he genetic diversity of populations is determined by a constant interplay between genetic drift and selection. 
+Drift is a consequence of a finite population size and the random sampling of gametes every generation \cite{dobzhansky1957}. 
+In contrast to the stochastic effects of drift, selection systematically alters allele frequencies by favoring particular alleles at the expense of others as a result of their effects on fitness.
+Researchers often seek to study drift by excluding potentially selected sites \cite{voight2005, luikart2003, gutenkunst2009}, or to study selection by focusing on site-specific patterns under the assumption that genome-wide diversity reflects only the action of drift \cite{Wall2002, akey2009}.
+\jri{careful here with these citations. we're effectively calling these people out, so make sure they're really doing what you say they are!}
 
-\dropcap{T}he genetic properties of populations are determined by a constant interplay between genetic drift and selection.
-Drift is the consequence of a finite population size, such that random sampling of gametes every generation will incrementally lead to changes in genetic variabiligy \cite{dobzhansky1957}.
-In other words, the demographic history of a population, which includes its population size fluctuations and mating patterns, has a long-lasting impact on the patterns of DNA variability displayed by that population.
-Conversely, selection systematically alters a population by favoring particular alleles at the expense of others.
-In contrast to drift, which results from demographic history, selection is the result of the relationship between genotypes, phenoytpes, and ultimately fitness.
-A classical approach for studying selection is to remove the genome-wide effects of demography and assume that selection will leave locus-specific evidence \cite{Wall2002, akey2009}.
-Conversely, demography is often researched by excluding potentially selected sites \cite{voight2005, luikart2003, gutenkunst2009}.
-
-However, recent work has shown that demography and selection do not operate independently to determine genetic variability, in large part because linkage allows the effects of selection to be wide-ranging \cite{li2012,slotte2014}.
-For example, in \emph{Drosophila} the majority of the genome is
-impacted by the effects of selection through linkage
-\cite{sella2009}.
-One technique that has been utilized to estimate demographic parameters independently of selective parameters is to limit analyses to only sites far from genes\cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is
-difficult to be certain, without species-specific knowledge, that sites even in gene-poor regions of the genome are not influenced by linked selection.
-Therefore, characterizing selection, demography or both, first requires an understanding of linked selection.
-It has already been established that patterns of linked selection are heavily influenced by population size ($N_e$) \cite{cutter2013, slotte2014}, but nearly all species including humans \cite{reich1998}, domesticates \cite{hyten2006, bovine2009}, ecological \cite{storfer2014}, and model species \cite{duchen2013} have experienced recent or ancient changes in $N_e$.
+Drift and selection do not operate independently to determine genetic variability, however, in large part because linkage allows the effects of selection to be wide-ranging \cite{smith1974, li2012,slotte2014}.
+Linked selection can take the form of hitch-hiking, when the frequency of a neutral allele changes as a result of positive selection at a physically linked site \cite{smith1974}, or background selection, where diversity is reduced at loci linked to a site undergoing selection against deleterious alleles \cite{charlesworth1993}.
+Recent work in a \emph{Drosophila}, for example, has shown that essentially the entire genome is impacted by the combined effects of these processes \cite{sella2009}.
+\jri{add citations for el yashiv biorxiv and andolfatto nature 2005}
+The impact of linked selection, in turn, is heavily influenced by the effective population size ($N_e$), as the  efficiency of natural selection is proportional to the product $N_es$, where $s$ is the strength of selection on a variant \cite{cutter2013, slotte2014}.
+\jri{add corbett-detig and leffler DOI: 10.1371/journal.pbio.1001388}.
+The effective size of a population is not static, and nearly all species, including flies \cite{duchen2013}, humans \cite{reich1998}, domesticates \cite{hyten2006, bovine2009}, and non-model species \citex \jri{this storfer ref didn't seem so great.
+any review on demography in non-models?
+or maybe list of several examples?} have experienced recent or ancient changes in $N_e$.
 Although much is known about how the long-term average $N_e$ affects linked selection, much less is understood about the immediate effects and dynamics of changing in $N_e$ on patterns of linked selection.
 
-Because of its relatively simple demographic history and abundant availability of genomic tools, maize (\emph{Zea mays}) represents an excellent organism to study this phenomenon.
+Because of its relatively simple demographic history and the abundant availability of genomic tools, maize (\emph{Zea mays}) represents an excellent organism to study these effects. 
 Maize is a species of tremendous importance worldwide as both a model species \cite{strable2009} and a staple crop \cite{shiferaw2011}. 
 Archaeological and genetic studies have established that maize domestication began in
-Central Mexico at least 9,000 years bp
-\cite{smith1995,matsuoka2002}, and involved a population bottleneck followed by recent expansion \cite{wright2005}.
-Because of this simple but dynamic demographic history, patterns of linked selection in wild and domesticated maize can be used to understand the effects of changing $N_e$ on linked selection.
-Teosinte, which remains extant throughout much of the Americas \cite{wilkes1967}, is the wild ancestor to maize.
-Additionally, several large-effect
-domestication loci \cite{doebley1995, wills2013, wang2015} and putative domestication
-regions \cite{hufford2012} have been identified in maize. But despite all that is
-known about maize domestication, the demographic parameters of the
-domestication process remain uncertain. Specifically, the size of the
-maize domestication bottleneck has not been estimated independently of
-the bottleneck's duration, nor are there sequence-based estimates of the effective
-population size of modern maize.
-At the same time, the predominant form or forms of selection that have taken place during and since domestication, and how linkage impacts the evidence and patterns of selection across the genome, remain unknown.
-
-For this study, we begin by estimating the parameters of maize domestication based on whole-genome sequencing data.
-Next, we utilized these estimates to 1) investigate the relative importance of different forms of selection on genetic
-variability in the recent and acient past; and 2) research the impact that demographic history, specifically changes in $N_e$, has had on linked selection in maize, and how this compares to the impact of a different demographic history in teosinte.
-Gaining an improved understanding of the impacts of a dynamic demography on whole-genome variability in maize will help to increase the understanding of these phenomena across all organisms, since changes in $N_e$ are pervasive.
+Central Mexico at least 9,000 years bp \cite{smith1995,matsuoka2002}, and involved a population bottleneck followed by recent expansion \cite{wright2005}.\jri{cite eyre-walker 1996 and tenaillon 2004 too}
+Because of this simple but dynamic demographic history, patterns of linked selection domesticated maize and its wild ancestor teosinte can be used to understand the effects of changing $N_e$ on linked selection.
+Here, we take advantage of this system to study how recent demographic change impacts patterns of linked selection genome-wide.
+We begin by estimating the parameters of the maize domestication bottleneck using whole-genome resequencing data, then investigate the relative importance of different forms of linked selection on diversity in the recent and ancient past.
+We show that while patterns of overall nucleotide diversity reflect long-term differences in $Ne$, recent growth following domestication qualitatively changes these effects, illustrating the importance of understanding demographic history when considering the effects of selection on the genome.
 

### 2015-11-03 manuscript.tex
through beginning of results

@@
 We show that positive selection on new genic mutations has had relatively little effect on genetic diversity, but that selection against deleterious mutations has dramatically reduced diversity in and immediately around genes in both taxa.
-We find that the relative effect of selection depends qualitatively on the age of the polymorphisms evaluated: while older polymorphisms in maize show more limited effects of linked selection, new mutations instead reflect its larger current size and more efficient selection.
+We find that the relative impact of selection depends on the age of the polymorphisms evaluated: while older polymorphisms in maize show more limited effects of linked selection, new mutations instead reflect its larger current size and more efficient selection.
 Our results demonstrate that a complete understanding of genome-wide patterns of diversity will require careful assessment of both demographic history and the effects of linked selection.
@@
 In contrast to the stochastic effects of drift, selection systematically alters allele frequencies by favoring particular alleles at the expense of others as a result of their effects on fitness.
-Researchers often seek to study drift by excluding potentially selected sites \cite{voight2005, luikart2003, gutenkunst2009}, or to study selection by focusing on site-specific patterns under the assumption that genome-wide diversity reflects only the action of drift \cite{Wall2002, akey2009}.
-\jri{careful here with these citations. we're effectively calling these people out, so make sure they're really doing what you say they are!}
+Researchers often seek to study drift by excluding potentially selected sites \cite{voight2005, luikart2003, gutenkunst2009}, or to study selection by focusing on site-specific patterns under the assumption that genome-wide diversity reflects primarily the action of drift \cite{akey2009}.
+
+%\jri{careful here with these citations. we're effectively calling these people out, so make sure they're really doing what you say they are!}
+%TB says: not so! Akey's is a review where he calls others out. (got rid of the other)
 
@@
 Linked selection can take the form of hitch-hiking, when the frequency of a neutral allele changes as a result of positive selection at a physically linked site \cite{smith1974}, or background selection, where diversity is reduced at loci linked to a site undergoing selection against deleterious alleles \cite{charlesworth1993}.
-Recent work in a \emph{Drosophila}, for example, has shown that essentially the entire genome is impacted by the combined effects of these processes \cite{sella2009}.
-\jri{add citations for el yashiv biorxiv and andolfatto nature 2005}
-The impact of linked selection, in turn, is heavily influenced by the effective population size ($N_e$), as the  efficiency of natural selection is proportional to the product $N_es$, where $s$ is the strength of selection on a variant \cite{cutter2013, slotte2014}.
-\jri{add corbett-detig and leffler DOI: 10.1371/journal.pbio.1001388}.
-The effective size of a population is not static, and nearly all species, including flies \cite{duchen2013}, humans \cite{reich1998}, domesticates \cite{hyten2006, bovine2009}, and non-model species \citex \jri{this storfer ref didn't seem so great.
-any review on demography in non-models?
-or maybe list of several examples?} have experienced recent or ancient changes in $N_e$.
+Recent work in a \emph{Drosophila}, for example, has shown that virtually the entire genome is impacted by the combined effects of these processes \cite{sella2009,elyashiv2014,andolfatto2005}.
+The impact of linked selection, in turn, is heavily influenced by the effective population size ($N_e$), as the  efficiency of natural selection is proportional to the product $N_es$, where $s$ is the strength of selection on a variant \cite{cutter2013, slotte2014, corbett2015,leffler2012}.
+The effective size of a population is not static, and nearly all species, including flies \cite{duchen2013}, humans \cite{reich1998}, domesticates \cite{hyten2006, bovine2009}, and non-model species \cite{ellegren2014} have experienced recent or ancient changes in $N_e$.
 Although much is known about how the long-term average $N_e$ affects linked selection, much less is understood about the immediate effects and dynamics of changing in $N_e$ on patterns of linked selection.
 
-Because of its relatively simple demographic history and the abundant availability of genomic tools, maize (\emph{Zea mays}) represents an excellent organism to study these effects. 
+Because of its relatively simple demographic history andfa
+the abundant availability of genomic tools, maize (\emph{Zea mays}) represents an excellent organism to study these effects. 
 Maize is a species of tremendous importance worldwide as both a model species \cite{strable2009} and a staple crop \cite{shiferaw2011}. 
 Archaeological and genetic studies have established that maize domestication began in
-Central Mexico at least 9,000 years bp \cite{smith1995,matsuoka2002}, and involved a population bottleneck followed by recent expansion \cite{wright2005}.\jri{cite eyre-walker 1996 and tenaillon 2004 too}
+Central Mexico at least 9,000 years bp \cite{smith1995,matsuoka2002}, and involved a population bottleneck followed by recent expansion \cite{wright2005,eyre1998,tenaillon2004}.
 Because of this simple but dynamic demographic history, patterns of linked selection domesticated maize and its wild ancestor teosinte can be used to understand the effects of changing $N_e$ on linked selection.
-Here, we take advantage of this system to study how recent demographic change impacts patterns of linked selection genome-wide.
+Here, we leverage this system to study how recent demographic change impacts patterns of linked selection genome-wide.
 We begin by estimating the parameters of the maize domestication bottleneck using whole-genome resequencing data, then investigate the relative importance of different forms of linked selection on diversity in the recent and ancient past.
@@
 
+\subsection{Patterns of diversity differ between genic and  intergenic regions of the genome} %JRI
+To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012, hapmap3}.
+As a preliminary step, we evaluated levels of diversity inside and outside of genes across the genome.
+We find broad differences in genic and intergenic diversity consistent with earlier results  \cite{hufford2012}(Figure \ref{fig:diversity}).
+In maize, mean pairwise diversity ($\pi$) within genes was significantly lower than at sites at least 5 kb away from genes (0.00668 vs 0.00691, $p<2\times 10^{-44}$).
+Diversity differences in teosinte are even more pronounced (0.0088 vs. 0.0115, $p\approx 0$). 
+Differences were also apparent in the site frequency spectrum, with mean Tajima's D positive in genic regions in both maize (0.4) and teosinte (0.013) but negative outside of genes (-0.087 in maize and -0.25 in teosinte, $p\approx 0$ for both comparisons).
+These observations suggest that diversity in genes is not evolving neutrally, but instead is reduced by the impacts of selection on linked sites.
+
 \subsection{Demography of maize domestication} %JRI
-To explore whether differences in the efficacy of purifying selection between maize and teosinte can be explained by demographic processes, we estimated the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}).
+After establishing that genic sites are not evolving neutrally, we used sites $>5$kb from genes to estimate the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}). 
 The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ teosinte individuals.
@@
 Maize and teosinte have continued to exchange migrants after the population split, with gene flow between the populations estimated at $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation from teosinte to maize and $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants from maize to teosinte.
+
+\begin{figure}[!tb]
+\begin{center}
+  \includegraphics[width=.45\textwidth] {FigsAndFiles/Pi_and_Tajima.png}
+\end{center}
+\caption{Pairwise diversity $\pi$ (A,B) and Tajimas D (C,D) in 1kb windows from genic and nongenic regions of maize (A,C) and teosinte (B,D).
+Shown in A and B are means and one standard deviation.
+\label{fig:diversity} }
+\end{figure}
 
@@
 Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
-
-\subsection{Patterns of diversity differ between genic and  intergenic regions of the genome} %JRI
-To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012, hapmap3}.
-We find broad differences in genic and intergenic diversity consistent with earlier results  \cite{hufford2012}(Figure \ref{fig:diversity}).
-In maize, mean pairwise diversity ($\pi$) within genes was significantly lower than at sites at least 5 kb away from genes (0.00668 vs 0.00691, $p<2\times 10^{-44}$).
-Diversity differences in teosinte are even more pronounced (0.0088 vs. 0.0115, $p\approx 0$). 
-Differences were also apparent in the site frequency spectrum, with mean Tajima's D positive in genic regions in both maize (0.4) and teosinte (0.013) but negative outside of genes (-0.087 in maize and -0.25 in teosinte, $p\approx 0$ for both comparisons).
-These observations suggest that diversity in genes is not evolving neutrally, but instead is reduced by the impacts of selection on linked sites.
-
-\begin{figure}[!tb]
-\begin{center}
-  \includegraphics[width=.45\textwidth] {FigsAndFiles/Pi_and_Tajima.png}
-\end{center}
-\caption{Pairwise diversity $\pi$ (A,B) and Tajimas D (C,D) in 1kb windows from genic and nongenic regions of maize (A,C) and teosinte (B,D).
-Shown in A and B are means and one standard deviation.
-\label{fig:diversity} }
-\end{figure}
-
-\subsection{Hard sweeps do not explain diversity differences} %JRI
-When selection increases the frequency of a new beneficial mutation, a signature of reduced diversity is left at surrounding linked sites \cite{smith1974}.
-To evaluate whether patterns of such ``hard sweeps'' could explain observed differences in diversity between genic and intergenic regions of the genome, we compared diversity around missense and synonymous substitutions between \emph{Tripsacum} and either maize or teosinte.
-If a proportion of missense mutations have been fixed due to hard sweeps, diversity around these substitutions should be lower than around synonymous substitutions. 
-We observe this pattern around the causative amino acid substitution in the the maize domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005origin, wang2015}.
-Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
-
-%\mbh{I'm curious whether linked selection is affecting the diversity around a sizable proportion of synonymous substitutions.
-Would it be worthwhile to look at diversity around the subset of synonymous substitutions that are the least linked to missense substitutions just to make sure it is not substantially higher?
-Or perhaps generate a plot of diversity around synonymous substitutions vs.
-distance of these subs from missense subs?}
-
-\begin{figure*}[!tb]
-%\vspace*{.05in}
-\centering
-\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_Aug}
-\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_Aug}
-\caption{Pairwise diversity surrounding synonymous and missense  substitutions in {\bf A} maize and {\bf B} teosinte.
-Axes show absolute diversity values (right) and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution (left).
-Lines depict a loess curve (span of 0.01) and shading represents bootstrap-based 95\% confidence intervals.
-Inset plots depict a larger range on the x-axis.
-\label{fig:hardSweeps}}
-\end{figure*}
-
-Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes that are under weak purifying selection and thus have higher genetic diversity \cite{enard2014}.
-To address this concern, we took advantage of genome-wide estimates of evolutionary constraint \cite{rodgers2015}, calculated using genomic evolutionary rate profile (GERP) scores \cite{davydov2010}. 
-We then evaluated substitutions only in subsets of genes in the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection. 
-As expected, we see higher diversity around substitutions in genes under weak purifying selection, but we still find no difference between synonymous and missense substitutions in either subset of the data (Figure \ref{sFig:consUncons}).
-Taken together, these data suggest hard sweeps do not play a major role in patterning genic diversity in either maize or teosinte.
 
@@
 
+\subsection{Hard sweeps do not explain diversity differences} %JRI
+When selection increases the frequency of a new beneficial mutation, a signature of reduced diversity is left at surrounding linked sites \cite{smith1974}.
+To evaluate whether patterns of such ``hard sweeps'' could explain observed differences in diversity between genic and intergenic regions of the genome, we compared diversity around missense and synonymous substitutions between \emph{Tripsacum} and either maize or teosinte.
+If a proportion of missense mutations have been fixed due to hard sweeps, diversity around these substitutions should be lower than around synonymous substitutions. 
+We observe this pattern around the causative amino acid substitution in the the maize domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005, wang2015}.
+Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
+
+%\mbh{I'm curious whether linked selection is affecting the diversity around a sizable proportion of synonymous substitutions.
+Would it be worthwhile to look at diversity around the subset of synonymous substitutions that are the least linked to missense substitutions just to make sure it is not substantially higher?
+Or perhaps generate a plot of diversity around synonymous substitutions vs.
+distance of these subs from missense subs?}
+
+\begin{figure*}[!tb]
+%\vspace*{.05in}
+\centering
+\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_Aug}
+\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_Aug}
+\caption{Pairwise diversity surrounding synonymous and missense  substitutions in {\bf A} maize and {\bf B} teosinte.
+Axes show absolute diversity values (right) and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution (left).
+Lines depict a loess curve (span of 0.01) and shading represents bootstrap-based 95\% confidence intervals.
+Inset plots depict a larger range on the x-axis.
+\label{fig:hardSweeps}}
+\end{figure*}
+
+Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes that are under weak purifying selection and thus have higher genetic diversity \cite{enard2014}.
+To address this concern, we took advantage of genome-wide estimates of evolutionary constraint \cite{rodgers2015}, calculated using genomic evolutionary rate profile (GERP) scores \cite{davydov2010}. 
+We then evaluated substitutions only in subsets of genes in the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection. 
+As expected, we see higher diversity around substitutions in genes under weak purifying selection, but we still find no difference between synonymous and missense substitutions in either subset of the data (Figure \ref{sFig:consUncons}).
+Taken together, these data suggest hard sweeps do not play a major role in patterning genic diversity in either maize or teosinte.
+
 \section{Discussion}
@@
 Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation and/or non-coding variants as primary sources of adaptive variation.
-Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013population}.
-Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007distribution}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.]{studer2011}.
-Indeed, many adaptive traits in both maize \cite{wallace2014} and teosinte \cite{weber2008} are highly quantitative,  and adaptation in both maize \cite{hufford2012} and teosinte \cite{pyhajarvi2013complex} has involved selection on regulatory variation.
+Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013}.
+Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.]{studer2011}.
+Indeed, many adaptive traits in both maize \cite{wallace2014} and teosinte \cite{weber2008} are highly quantitative,  and adaptation in both maize \cite{hufford2012} and teosinte \cite{pyhajarvi2013} has involved selection on regulatory variation.
 
@@
 \subsection{Demography of domestication} %JRI
-Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004selection, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
+Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
 We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites.
@@
 One surprising result from our model is the estimated timing of domestication at $\approx 15,000$ years before present.
-While this appears to conflict with archaeological estimates \cite{piperno2009starch}, we emphasize that this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
+While this appears to conflict with archaeological estimates \cite{piperno2009}, we emphasize that this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
 We also note that our result may also be inflated due to population structure, as our geographically diverse sample of teosinte may include populations diverged from those that gave rise to maize.
@@
 Both alternative approaches we employ estimate a much larger modern effective size of maize in the range of $\approx 10^6 - 10^9$, an order of magnitude or more than the current size of teosinte.
-Census data suggest these estimates are plausible: there are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, likely planted at a density of $\approx 25,000$ individuals per hectare \cite{baden2001culture}.
+Census data suggest these estimates are plausible: there are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, likely planted at a density of $\approx 25,000$ individuals per hectare \cite{baden2001}.
 Assuming the effective size is only $\approx 0.4\%$ of the census size (i.e. 1 ear for every 1000 male plants), this still implies a modern effective population size of more than four billion.
@@
 Although we observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), we see a greater decrease in singleton diversity in maize than teosinte (Figure \ref{fig:purify}B).
-Purifying selection is more efficient in larger populations  \cite{kimura1984}, and our observations using $\pi$  are thus consistent with our demographic model and previous work \cite{eyre1998, tenaillon2004selection, wright2005, ross2009} which shows a smaller long-term effective population size in maize.
+Purifying selection is more efficient in larger populations  \cite{kimura1984}, and our observations using $\pi$  are thus consistent with our demographic model and previous work \cite{eyre1998, tenaillon2004, wright2005, ross2009} which shows a smaller long-term effective population size in maize.
 Similarly, our observations based on singleton diversity, which thereby reflect more recent patterns of selection, again are in agreement with maize and teosinte demography; from the viewpoint of a recent mutation, the rapid expansion of maize post-domestication means that today maize has a vastly larger effective population size than does teosinte.

### 2015-11-06 manuscript.tex
Final... (yeah right!)

@@
 \centering
-\includegraphics[width=.4\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
+\includegraphics[width=.5\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
 \caption{Parameter estimates for a basic bottleneck model of maize domestication. See methods for details. \label{fig:bottleneck} }
@@
 Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
-
-\subsection{Diversity is strongly influenced by purifying selection} %JRI
-
-In the case of purifying or background selection, diversity can be reduced in functional regions of the genome via removal of deleterious mutations \cite{charlesworth1993}.
-We investigated purifying selection in maize and teosinte by evaluating the reduction of diversity around genes.
-Pairwise diversity is strongly reduced within genes for both maize and teosinte (Figure \ref{fig:purify}A) but recovers quickly at sites outside of genes, consistent with the low levels of linkage disequilibrium generally observed in these subspecies \cite{tenaillon2002,chia2012}.
-The reduction in relative diversity is more pronounced in teosinte, reaching lower levels in genes and occurring over a wider region.
-
-Our comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because in such cases diversity is not necessarily reduced \cite{innan2004,messer2013}.
-Such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes. 
-To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
-Qualitative differences between maize and teosinte remained unchanged even after removing genes in the top 20\% of H12 (Figure \ref{sFig:H12}A).
-%\mbh{This makes me wonder about the top 20\%. Is the pattern of diversity reduction different in this fraction of the genome?  If so, that seems like something that should be mentioned.}
-%TB Says: Sites in the top 20% of H12 are unlikely to be under background selection, so maize vs teo background selection pattern at these sites seems strange to investigate.
-We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.
-
-\begin{figure*}[!tb]
-%\vspace*{.05in}
-\centering
-\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_Downsampled_threeLines_manuscript.png}
-\caption{Relative diversity versus distance to nearest gene in maize and teosinte. 
-Shown are \textbf{A} pairwise nucleotide diversity and \textbf{B} singleton diversity.  
-Relative diversity is calculated compared to the mean diversity in windows $\geq 0.01 cM$ or $\geq 0.02 cM$ from the nearest
-gene for pairwise diversity and singletons, respectively. 
-  Lines depict cubic smoothing splines with smoothing parameters chosen via generalized cross validation and shading depicts bootstrap-based 95\% confidence intervals.
-  Inset plots depict a smaller range on the x-axis. \label{fig:purify}
-  }
-\end{figure*}
-
-\subsection{Population expansion leads to stronger purifying selection in modern maize} %JRI
-Motivated by the rapid post-domestication expansion of maize evident in our demographic analysis, we reasoned that low-frequency --- and thus younger --- polymorphisms might show patterns distinct from pairwise diversity.
-Singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) appears nearly identical to results from pairwise diversity (Figure \ref{fig:hardSweeps}), providing little support for a substantial increase in the number or strength of hard sweeps occurring in maize.
-In contrast, we observe a significant shift in the effects of purifying selection: singleton polymorphisms are more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
-This result is the opposite of the pattern observed for $\pi$, where teosinte demonstrated a stronger reduction of diversity in and around genes than did maize.
-As was the case when we analyzed $\pi$, this relationship remained after we removed the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12}).
-Finally, while direct comparison of pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, as indicated by our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
 
@@
 
-\begin{figure*}[!tb]
+\begin{figure*}[tb]
 %\vspace*{.05in}
@@
 
+\subsection{Diversity is strongly influenced by purifying selection} %JRI
+
+In the case of purifying or background selection, diversity is reduced in functional regions of the genome via removal of deleterious mutations \cite{charlesworth1993}.
+We investigated purifying selection in maize and teosinte by evaluating the reduction of diversity around genes.
+Pairwise diversity is strongly reduced within genes for both maize and teosinte (Figure \ref{fig:purify}A) but recovers quickly at sites outside of genes, consistent with the low levels of linkage disequilibrium generally observed in these subspecies \cite{tenaillon2002,chia2012}.
+The reduction in relative diversity is more pronounced in teosinte, reaching lower levels in genes and occurring over a wider region.
+
+Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because in such cases diversity is not necessarily reduced \cite{innan2004,messer2013}.
+Such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes. 
+To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
+Qualitative differences between maize and teosinte remained unchanged even after removing genes in the top 20\% of H12 (Figure \ref{sFig:H12}A).
+%\mbh{This makes me wonder about the top 20\%. Is the pattern of diversity reduction different in this fraction of the genome?  If so, that seems like something that should be mentioned.}
+%TB Says: Sites in the top 20% of H12 are unlikely to be under background selection, so maize vs teo background selection pattern at these sites seems strange to investigate.
+We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.
+
+\begin{figure*}[tb]
+%\vspace*{.05in}
+\centering
+\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_Downsampled_threeLines_manuscript.png}
+\caption{Relative diversity versus distance to nearest gene in maize and teosinte. 
+Shown are \textbf{A} pairwise nucleotide diversity and \textbf{B} singleton diversity.  
+Relative diversity is calculated compared to the mean diversity in windows $\geq 0.01 cM$ or $\geq 0.02 cM$ from the nearest
+gene for pairwise diversity and singletons, respectively. 
+  Lines depict cubic smoothing splines with smoothing parameters chosen via generalized cross validation and shading depicts bootstrap-based 95\% confidence intervals.
+  Inset plots depict a smaller range on the x-axis. \label{fig:purify}
+  }
+\end{figure*}
+
+\subsection{Population expansion leads to stronger purifying selection in modern maize} %JRI
+Motivated by the rapid post-domestication expansion of maize evident in our demographic analysis, we reasoned that low-frequency --- and thus younger --- polymorphisms might show patterns distinct from pairwise diversity.
+Singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) appears nearly identical to results from pairwise diversity (Figure \ref{fig:hardSweeps}), providing little support for a substantial recent increase in the number or strength of hard sweeps occurring in maize.
+
+In contrast, we observe a significant shift in the effects of purifying selection: singleton polymorphisms are more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
+This result is the opposite of the pattern observed for $\pi$, where teosinte demonstrated a stronger reduction of diversity in and around genes than did maize.
+As was the case when we analyzed $\pi$, this relationship remained after we removed the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12}).
+Finally, while direct comparison of pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, as indicated by our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
+
 \section{Discussion}
+
+\subsection{Demography of domestication} %JRI
+Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
+We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites.
+This implies that genic polymorphism data are unable to tell the complete or accurate demographic history of maize, but the rapid recovery of diversity outside of genes demonstrates that sites far from genes can be reasonably used for demographic inference.
+Furthermore, by utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without making assumptions about its duration.
+
+One surprising result from our model is the estimated timing of domestication at $\approx 15,000$ years before present.
+While this appears to conflict with archaeological estimates \cite{piperno2009}, we emphasize that this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
+We also note that our result may be inflated due to population structure, as our geographically diverse sample of teosinte may include populations diverged from those that gave rise to maize.
+
+The estimated bottleneck of $\approx 5\%$ of the ancestral teosinte population seems low given that maize landraces exhibit $\approx 80\%$ of the diversity of teosinte \cite{hufford2012}, but our model suggests that the effects of the bottleneck on diversity are likely ameliorated by both gene flow and rapid population growth (Figure \ref{fig:bottleneck}).
+Although we estimate that the modern effective size of maize is larger than teosinte, the small size of our sample reduces our power to identify the low frequency alleles most sensitive to rapid population growth \cite{keinan2012}, and our model is unable to incorporate growth faster than exponential.
+Both alternative approaches we employ estimate a much larger modern effective size of maize in the range of $\approx 10^6 - 10^9$, an order of magnitude or more than the current size of teosinte.
+Census data suggest these estimates are plausible: there are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, likely planted at a density of $\approx 25,000$ individuals per hectare \cite{baden2001}.
+Assuming the effective size is only $\approx 0.4\%$ of the census size (i.e. 1 ear for every 1000 male plants), this still implies a modern effective population size of more than four billion.
+While these genetic and census estimates are likely inaccurate, all of the evidence points to the fact that the effective size of modern maize is extremely large.
 
@@
 
-\subsection{Demography of domestication} %JRI
-Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
-We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites.
-This implies that genic polymorphism data are unable to tell the complete or accurate demographic history of maize, but the rapid recovery of diversity outside of genes demonstrates that sites far from genes can be reasonably used for demographic inference.
-Furthermore, by utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without requiring assumptions about its duration.
-
-One surprising result from our model is the estimated timing of domestication at $\approx 15,000$ years before present.
-While this appears to conflict with archaeological estimates \cite{piperno2009}, we emphasize that this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
-We also note that our result may also be inflated due to population structure, as our geographically diverse sample of teosinte may include populations diverged from those that gave rise to maize.
-
-The estimated bottleneck of $\approx 5\%$ of the ancestral teosinte population seems low given that maize landraces exhibit $\approx 80\%$ of the diversity of teosinte \cite{hufford2012}, but our model suggests that the effects of the bottleneck on diversity are likely ameliorated by both gene flow and rapid population growth (Figure \ref{fig:bottleneck}).
-Although we estimate that the modern effective size of maize is larger than teosinte, the small size of our sample reduces our power to identify the low frequency alleles most sensitive to rapid population growth \cite{keinan2012}, and our model is unable to incorporate growth faster than exponential.
-Both alternative approaches we employ estimate a much larger modern effective size of maize in the range of $\approx 10^6 - 10^9$, an order of magnitude or more than the current size of teosinte.
-Census data suggest these estimates are plausible: there are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, likely planted at a density of $\approx 25,000$ individuals per hectare \cite{baden2001}.
-Assuming the effective size is only $\approx 0.4\%$ of the census size (i.e. 1 ear for every 1000 male plants), this still implies a modern effective population size of more than four billion.
-While these genetic and census estimates are likely inaccurate, all of the evidence points to the fact that the effective size of modern maize is extremely large.
-
 \subsection{Demography influences the efficiency of purifying selection}
@@
 Purifying selection is more efficient in larger populations  \cite{kimura1984}, and our observations using $\pi$  are thus consistent with our demographic model and previous work \cite{eyre1998, tenaillon2004, wright2005, ross2009} which shows a smaller long-term effective population size in maize.
-Similarly, our observations based on singleton diversity, which thereby reflect more recent patterns of selection, again are in agreement with maize and teosinte demography; from the viewpoint of a recent mutation, the rapid expansion of maize post-domestication means that today maize has a vastly larger effective population size than does teosinte.
+Similarly, our observations based on singleton diversity, which reflect more recent patterns of selection, again are in agreement with maize and teosinte demography; from the viewpoint of a recent mutation, the rapid expansion of maize post-domestication means that today maize has a vastly larger effective population size than does teosinte.
 
-A consequence of the previously-inefficient purifying selection in maize is that it likely harbors more weakly deleterious alleles segregating at intermediate frequency than does teosinte.
-Therefore, high genetic load could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{meghji1984}.
-Our results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously selected.
-Additionally, the large $N_e$ of modern maize compared to modern teosinte implies that for new mutations, selection will operate much more efficiently in maize.
-This observation is not restricted to maize.
-Bottlenecks have been documented or predicted across most domesticated plant and animal species \cite{doebley2006, hyten2006, zhu2007, bovine2009, larson2013 }, but rapid expansion post-bottleneck, which we demonstrate generates ubiquitous consequences relating to genetic diversity, is not well studied.
-In human, model, and other species, these results demonstrate that observed patterns of selection depend heavily on recent and/or ancient patterns of demography.
+\section{Demography and selection interact to shape genetic diversity}
+Importantly, our results demonstrate that abstractions such as average long-term $N_e$ cannot capture the dynamic relationship between demography and selection.
+While it is certainly the case that diversity is lost during a bottleneck, it is equally true that from the viewpoint of a mutation that arises later, the bottleneck is irrelevant.
+Although the relatively straight-forward demographic history of maize risks oversimplifying this observation, it is applicable to any species that has undergone demographic change.
+For instance, human populations have experienced recent growth much like maize has, and this growth corresponds to a documented excess of rare genetic variants \cite{keinan2012, tennessen2012, coventry2010}.
+Our results suggest that in such scenarios, ongoing background selection may lead to a dearth of rare variants in and around genes compared to outside of them.
+This conflicts with an alternative possibility; that sudden expansion should lead to a comparable increase in rare variants genome-wide.
 
-%\jri{these results have implications for lots of studies.
-recent demographic history often ignored (examples).
-in crops, bneck is considered and loss of diversity, but expansion also likely ubiquitous and consequences not well studied.
-in humans XYZ, and we see some of that already here.
-this argues that a full understanding of diversity in both wild and cultivated plants will require recent demography.
-mention C-D ignored recent demography -- here's where rec/pi windows could help.}
+The abundance of rare variants across the genome, and the efficiency of purifying selection against them, has notable implications for genetic load.
+One possibility is that the excess of rare variants present in expanding populations implies higher genetic load \cite{keinan2012}.
+However, more efficient purifying selection in larger populations, which we show operates specifically on these rare variants, should decrease genetic load by purging rare deleterious variants.
+Several studies have demonstrated that despite different demographic histories, African and European human populations have a similar genetic load \cite{do2015,simons2014}, due in part to new deleterious mutations being increasingly rare as population size increases \cite{simons2014, lohmueller2014}.
+We propose that this observation could also result from more efficient selection against such mutations.
+This proposition is in agreement not only with our maize observations, but also with the results of \cite{gazave2013}, who showed via simulations that the increased efficiency of selection in a larger population can partially offset the increase in the total number of segregating deleterious variants.
 
-%\jri{I would like to see a supplemental figure or maybe just number here on the partial correlation in both teosinte and maize of diversity with recombination after accounting for gene density.
-do this in 500kb windows and then in 10kb windows.
-we should do the correlation with both pi and singletons.
-This let's us compare more explictly with studies like C-D who evaluate overall patterns of linked selection.
-will be especially cool if we don't see a big difference in big windows but we do in 10kb windows -- siggests C-D looked at wrong scale  }
+Across species, our results demonstrate that the genetic impacts of major demographic events such as bottlenecks depend not only on population size reductions, but also on subsequent, potentially rapid and recent, expansion that often comes afterwards.
+Reduced population sizes do lead to a decreased efficiency of selection \cite[e.g]{renaut2015}, which has long-term impacts on diversity measured via $\pi$.
+But new mutations, which we measure with singleton diversity, are oblivious to past demography, and therefore selection for or against such sites depends primarily on recent demographic processes.
 

### 2015-11-06 manuscript.tex
minor edits to the intro

@@
 The effective size of a population is not static, and nearly all species, including flies \cite{duchen2013}, humans \cite{reich1998}, domesticates \cite{hyten2006, bovine2009}, and non-model species \cite{ellegren2014} have experienced recent or ancient changes in $N_e$.
-Although much is known about how the long-term average $N_e$ affects linked selection, much less is understood about the immediate effects and dynamics of changing in $N_e$ on patterns of linked selection.
+Although much is known about how the long-term average $N_e$ affects linked selection, less is understood about the immediate effects of more recent changes in $N_e$ on patterns of linked selection.
 
-Because of its relatively simple demographic history andfa
-the abundant availability of genomic tools, maize (\emph{Zea mays}) represents an excellent organism to study these effects. 
-Maize is a species of tremendous importance worldwide as both a model species \cite{strable2009} and a staple crop \cite{shiferaw2011}. 
-Archaeological and genetic studies have established that maize domestication began in
-Central Mexico at least 9,000 years bp \cite{smith1995,matsuoka2002}, and involved a population bottleneck followed by recent expansion \cite{wright2005,eyre1998,tenaillon2004}.
-Because of this simple but dynamic demographic history, patterns of linked selection domesticated maize and its wild ancestor teosinte can be used to understand the effects of changing $N_e$ on linked selection.
-Here, we leverage this system to study how recent demographic change impacts patterns of linked selection genome-wide.
-We begin by estimating the parameters of the maize domestication bottleneck using whole-genome resequencing data, then investigate the relative importance of different forms of linked selection on diversity in the recent and ancient past.
-We show that while patterns of overall nucleotide diversity reflect long-term differences in $Ne$, recent growth following domestication qualitatively changes these effects, illustrating the importance of understanding demographic history when considering the effects of selection on the genome.
+Because of its relatively simple demographic history and well-developed genomic resources, maize (\emph{Zea mays}) represents an excellent organism to study these effects. 
+Maize is a species of tremendous importance worldwide as both a model species \cite{strable2009} and a staple crop \cite{shiferaw2011}. \mbh{not sure we need the previous sentence}
+Archaeological and genetic studies have established that maize domestication began in Central Mexico at least 9,000 years bp \cite{smith1995,matsuoka2002}, and involved a population bottleneck followed by recent expansion \cite{wright2005,eyre1998,tenaillon2004}.
+Because of this simple but dynamic demographic history, domesticated maize and its wild ancestor teosinte can be used to understand the effects of changing $N_e$ on linked selection. 
+We leverage the maize-teosinte study system to study these effects by first estimating the parameters of the maize domestication bottleneck using whole-genome resequencing data and then investigating the relative importance of different forms of linked selection on diversity in the recent and ancient past.
+We show that, while patterns of overall nucleotide diversity reflect long-term differences in $Ne$, recent growth following domestication qualitatively changes these effects, thereby illustrating the importance of a comprehensive understanding of demography when considering the effects of selection genome-wide.
 

### 2015-11-08 manuscript.tex
edits

@@
 Our results demonstrate that a complete understanding of genome-wide patterns of diversity will require careful assessment of both demographic history and the effects of linked selection.
-}
+} %JRI NOT DONE
 
@@
 \begin{abstract}
-  The genetic variability present in a population results from a constant interaction between drift and selection. 
+  The genetic variability present in a population results from a constant interaction between drift and natural selection. 
 To quantify the interplay between demography (which drives drift) and selection, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide. 
@@
 Our results also show that for bottlenecked species, rapid post-bottleneck expansion can have wide-ranging impacts on diversity that conflict with the expected loss of diversity that occurs during the bottleneck itself.
+%JRI NOT DONE
 
@@
 
-\dropcap{T}he genetic diversity of populations is determined by a constant interplay between genetic drift and selection. 
+\dropcap{T}he genetic diversity of populations is determined by a constant interplay between genetic drift and natural selection. 
 Drift is a consequence of a finite population size and the random sampling of gametes every generation \cite{dobzhansky1957}. 
@@
 Recent work in a \emph{Drosophila}, for example, has shown that virtually the entire genome is impacted by the combined effects of these processes \cite{sella2009,elyashiv2014,andolfatto2005}.
+
 The impact of linked selection, in turn, is heavily influenced by the effective population size ($N_e$), as the  efficiency of natural selection is proportional to the product $N_es$, where $s$ is the strength of selection on a variant \cite{cutter2013, slotte2014, corbett2015,leffler2012}.
 The effective size of a population is not static, and nearly all species, including flies \cite{duchen2013}, humans \cite{reich1998}, domesticates \cite{hyten2006, bovine2009}, and non-model species \cite{ellegren2014} have experienced recent or ancient changes in $N_e$.
-Although much is known about how the long-term average $N_e$ affects linked selection, less is understood about the immediate effects of more recent changes in $N_e$ on patterns of linked selection.
+Although much is known about how the long-term average $N_e$ affects linked selection \jri{citations needed}, relatively little is understood about the immediate effects of more recent changes in $N_e$ on patterns of linked selection.
 
 Because of its relatively simple demographic history and well-developed genomic resources, maize (\emph{Zea mays}) represents an excellent organism to study these effects. 
-Maize is a species of tremendous importance worldwide as both a model species \cite{strable2009} and a staple crop \cite{shiferaw2011}. \mbh{not sure we need the previous sentence}
 Archaeological and genetic studies have established that maize domestication began in Central Mexico at least 9,000 years bp \cite{smith1995,matsuoka2002}, and involved a population bottleneck followed by recent expansion \cite{wright2005,eyre1998,tenaillon2004}.
 Because of this simple but dynamic demographic history, domesticated maize and its wild ancestor teosinte can be used to understand the effects of changing $N_e$ on linked selection. 
-We leverage the maize-teosinte study system to study these effects by first estimating the parameters of the maize domestication bottleneck using whole-genome resequencing data and then investigating the relative importance of different forms of linked selection on diversity in the recent and ancient past.
+In this study, we leverage the maize-teosinte study system to study these effects by first estimating the parameters of the maize domestication bottleneck using whole-genome resequencing data and then investigating the relative importance of different forms of linked selection on diversity in the recent and ancient past.
 We show that, while patterns of overall nucleotide diversity reflect long-term differences in $Ne$, recent growth following domestication qualitatively changes these effects, thereby illustrating the importance of a comprehensive understanding of demography when considering the effects of selection genome-wide.
@@
 \caption{Pairwise diversity $\pi$ (A,B) and Tajimas D (C,D) in 1kb windows from genic and nongenic regions of maize (A,C) and teosinte (B,D).
-Shown in A and B are means and one standard deviation.
+Shown in A and B are means $\pm$ one standard deviation.
 \label{fig:diversity} }
@@
 To evaluate whether patterns of such ``hard sweeps'' could explain observed differences in diversity between genic and intergenic regions of the genome, we compared diversity around missense and synonymous substitutions between \emph{Tripsacum} and either maize or teosinte.
-If a proportion of missense mutations have been fixed due to hard sweeps, diversity around these substitutions should be lower than around synonymous substitutions. 
+If a substantial proportion of missense mutations have been fixed due to hard sweeps, diversity around these substitutions should be lower than around synonymous substitutions. 
 We observe this pattern around the causative amino acid substitution in the the maize domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005, wang2015}.
@@
 
-Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes that are under weak purifying selection and thus have higher genetic diversity \cite{enard2014}.
-To address this concern, we took advantage of genome-wide estimates of evolutionary constraint \cite{rodgers2015}, calculated using genomic evolutionary rate profile (GERP) scores \cite{davydov2010}. 
+\jri{use of (A) or {\bf A} needs to be consistent in figure legends} Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes that are under weak purifying selection and thus have higher genetic diversity \cite{enard2014}.
+To address this concern, we took advantage of genome-wide estimates of evolutionary constraint \cite{rodgers2015} calculated using genomic evolutionary rate profile (GERP) scores \cite{davydov2010}. 
 We then evaluated substitutions only in subsets of genes in the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection. 
@@
 
-Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because in such cases diversity is not necessarily reduced \cite{innan2004,messer2013}.
+Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because diversity in such cases  is not necessarily reduced \cite{innan2004,messer2013}.
 Such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes. 
 To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
-Qualitative differences between maize and teosinte remained unchanged even after removing genes in the top 20\% of H12 (Figure \ref{sFig:H12}A).
+Qualitative differences between maize and teosinte remained unchanged even after removing genes in the top 20\% quantile of H12 (Figure \ref{sFig:H12}A).
 %\mbh{This makes me wonder about the top 20\%. Is the pattern of diversity reduction different in this fraction of the genome?  If so, that seems like something that should be mentioned.}
@@
 \subsection{Population expansion leads to stronger purifying selection in modern maize} %JRI
-Motivated by the rapid post-domestication expansion of maize evident in our demographic analysis, we reasoned that low-frequency --- and thus younger --- polymorphisms might show patterns distinct from pairwise diversity.
+Motivated by the rapid post-domestication expansion of maize evident in our demographic analyses, we reasoned that low-frequency --- and thus younger --- polymorphisms might show patterns distinct from pairwise diversity.
 Singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) appears nearly identical to results from pairwise diversity (Figure \ref{fig:hardSweeps}), providing little support for a substantial recent increase in the number or strength of hard sweeps occurring in maize.
 
-In contrast, we observe a significant shift in the effects of purifying selection: singleton polymorphisms are more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B), even after downsampling our maize data to account for differences in sample size.
+In contrast, we observe a significant shift in the effects of purifying selection: singleton polymorphisms are more strongly reduced in and near genes in maize than in teosinte, even after downsampling our maize data to account for differences in sample size (Figure \ref{fig:purify}B).
 This result is the opposite of the pattern observed for $\pi$, where teosinte demonstrated a stronger reduction of diversity in and around genes than did maize. 
-As was the case when we analyzed $\pi$, this relationship remained after we removed the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12}).
-Finally, while direct comparison of pairwise and singleton diversity within taxa are consistent with at least some non-equilibrium dynamics in teosinte, as indicated by our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
+As before, this relationship remained after we removed the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12}).
+Finally, while direct comparison of pairwise and singleton diversity within taxa is consistent with   non-equilibrium dynamics in teosinte as indicated in our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
+\jri{tempted to remove phrase "as indicated in our coalescent-based demographic analysis"}
 
@@
 
-\subsection{Hard sweeps do not shape genome-wide diversity in maize} %JRI
+\subsection{Hard sweeps do not shape genome-wide diversity in maize} %JRI NOT DONE
 
@@
 Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with ``soft sweeps''  from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
-Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation and/or non-coding variants as primary sources of adaptive variation.
+Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation and/or \jri{why and/or?
+why not and?} non-coding variants as primary sources of adaptive variation.
 Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013}.
@@
 
-\subsection{Demography influences the efficiency of purifying selection}
+\subsection{Demography influences the efficiency of purifying selection} %JRI NOT DONE
 One of our more striking findings is that the impact of purifying selection on maize and teosinte variability has varied through time.
@@
 
-\section{Demography and selection interact to shape genetic diversity}
+\section{Demography and selection interact to shape genetic diversity} %JRI NOT DONE
 Importantly, our results demonstrate that abstractions such as average long-term $N_e$ cannot capture the dynamic relationship between demography and selection.
@@
 Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{hapmap3}.
+\jri{update ref.
+29 to hapmap 3 paper}
  All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following Li \emph{et al.} \cite{li2011statistical}.
+\jri{did we really only use uniquely mapping?
+or just a MQ $>30$ filter?}
 We converted physical coordinates to genetic coordinates via linear interpolation of the previously published 1cM resolution NAM genetic map \cite{glaubitz2014}.
@@
 
-We further made use of a large genotyping data set of more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize $N_e$ from singleton counts.
+We further made use of a large genotyping data set of more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize $N_e$ from singleton counts. \jri{imputed or nonimputed?}
 We filtered these data to include only SNPs with data in $\geq 1,500$ individuals, and then projected the SFS down to a sample of 500 individuals by sampling each marker without replacement 1,000 times according to the observed allele frequencies.
@@
 We called SNPs in ANGSD \cite{korneliussen2014} using a SNP p-value of $1e-6$ against a reference genome masked using SNPable (\url{http://lh3lh3.users.sourceforge.net/snpable.shtml}).
-We then removed heterozygous genotypes and filtered sites with a mapping quality $<30$, a base quality $<20$, or a $|\emph{log}_2(\emph{depth})|<1$.
+\jri{fix url, not showing up in pdf}
+We then removed heterozygous genotypes and filtered sites with a mapping quality $<30$, a base quality $<20$, or a $|\emph{log}_2(\emph{depth})|<1$. \jri{fix depth, italics is weird}
 We ran MSMC with pattern parameters $20\times2+20\times4+10\times2$.
@@
 We called genotypes in maize, teosinte, and \emph{Tripsacum} at sites with a SNP p-value $<10^{-6}$ and when the genotype posterior probability $>0.95$.
+\jri{in methods, not sure if you should use emph instead of textit or something, as italics text looks weird}
 We identified substitutions in maize and teosinte as all sites with a fixed difference with \emph{Tripsacum} and $\leq 20\%$ missing data. 

### 2015-11-08 supportingInfo.tex
edits
updating introduction.

@@
 
+\jri{first figure should be below header}
 \begin{figure}
@@
     \end{center}
-\caption{Diversity surrounding the causitive polymorphism at the \emph{tga1} locus is plotted.
-Since this is only one gene, the large amount of noise compared to our average plots is expected.
-However, notice that diversity precisely at the causitive polymorphism is reduced and a recovery of diversity is observed away from that site.
-\label{sFig:tga1}}
+\caption{Diversity surrounding the causitive substitution at the \emph{tga1} locus.  \label{sFig:tga1}}
+\end{figure}
+\jri{why blank page before MSMC?}
+\clearpage
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.85\textwidth]{FigsAndFiles/BKN_TIL_msmc.pdf} \\
+    \end{center}
+\caption{\jri{need MSMC caption, remove gray background, fix axis labels and plot legend.} \label{sFig:msmc}}
 \end{figure}
@@
 \end{figure}
+\jri{why no confidence interval? can we add?}
 \clearpage
@@
 \end{figure}
+\jri{use of A: and A etc. needs to be consistent with main text}
 \clearpage
@@
 \end{figure}
+\jri{add citations}
 \clearpage
@@
 \end{figure}
+\jri{why are these shown as figures and not tables?}
 \clearpage

### 2015-11-08 manuscript.tex
first part of discussion

@@
 Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with ``soft sweeps''  from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
-Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{beissinger2014} all support the importance of standing variation and/or \jri{why and/or?
-why not and?} non-coding variants as primary sources of adaptive variation.
+Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{van2012historical, beissinger2014} all support the importance of standing variation as primary sources of adaptive variation.
 Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013}.
@@
 
-The absence of evidence for a genome-wide impact of hard sweeps differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
-\cite{corbett2015} recently showed that selection tends to have an elevated impact reducing diversity in species with large population size compared to small, which may explain why maize and humans, both bottlenecked species, show similar patterns while \emph{drosophila} and \emph{capsella} behave differently.
+The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}, but is consistent with estimates of the percentage of nonsynonmyous substitutions fixed by natural selection \citep{ross2009,sella2009,eyre2009estimating,williamson2014}.
+While larger long-term $N_e$ likely explains some of this difference, we see little change in the importance of hard sweeps in genes even looking at singleton diversity in modern maize \jri{reference appropriate figure}, suggesting other factors may contribute to these differences.
+If mutational target size scales with genome size, the larger genomes of human and maize may offer more possibilities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively small role
+In line  with this argument, studies of local adaptation in taxa with large genomes such as teosinte \citep{pyhajarvi2013} and humans \citep{fraser2013gene} find enrichment for loci in regulatory regions, while similar analyses in the small \emph{Arabidopsis} genome finds an enrichment for nonsynonymous variants.
 
@@
 But new mutations, which we measure with singleton diversity, are oblivious to past demography, and therefore selection for or against such sites depends primarily on recent demographic processes.
+
+%\cite{corbett2015} recently showed that selection tends to have an elevated impact reducing diversity in species with large population size compared to small, which may explain why maize and humans, both bottlenecked species, show similar patterns while \emph{drosophila} and \emph{capsella} behave differently.
 

### 2015-11-09 manuscript.tex
discussion section

@@
 \author{Timothy M.
-Beissinger\affil{1}{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}\affil{2}{Division of Plant Sciences, University of Missouri, Columbia, MO, USA}\affil{3}{Dept.
-of Plant Sciences, University of California, Davis, CA, USA}, Li Wang\affil{4}{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA}\affil{5}{Genome Informatics Facility, Iowa State University, Ames, IA, USA}, Kate Crosby\affil{1}{}, Arun Durvasula\affil{1}{}, Matthew B.
+Beissinger\affil{1}{Dept.
+of Plant Sciences, University of California, Davis, CA, USA}\affil{2}{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}\affil{3}{Division of Plant Sciences, University of Missouri, Columbia, MO, USA}, Li Wang\affil{4}{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA}\affil{5}{Genome Informatics Facility, Iowa State University, Ames, IA, USA}, Kate Crosby\affil{1}{}, Arun Durvasula\affil{1}{}, Matthew B.
 Hufford\affil{4}{}, \and Jeffrey Ross-Ibarra\affil{1}{}\affil{6}{Genome Center and Center for Population Biology, University of
@@
 Our results demonstrate that a complete understanding of genome-wide patterns of diversity will require careful assessment of both demographic history and the effects of linked selection.
-} %JRI NOT DONE
+} %JRI
 
@@
   The genetic variability present in a population results from a constant interaction between drift and natural selection. 
-To quantify the interplay between demography (which drives drift) and selection, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide. 
-Our sample included whole genome sequence data from 23 maize and 13 teosinte individuals. 
-We obtained a complete estimate of the population size fluctuations and other demographic parameters experienced by maize as it was domesticated from teosinte. 
-Here, we show that maize went through a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid population size expansion post-domestication. 
-We observe that hard sweeps on genic mutations are not the primary force driving maize evolution. 
-We find that a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize. 
-However, expansion after domestication has since increased the efficiency of purifying selection to levels exceeding those seen in teosinte. 
-Our results demonstrate that demographic and selective history in the ancient and recent past both contribute to the levels of genetic variability that are present today. 
+To quantify the interplay between drift and selection, we investigated patterns of linked selection in in the genomes of 23 domesticated maize and 13 wild teosinte individuals. 
+We show that maize went through a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid population size expansion post-domestication. 
+We observe that hard sweeps on genic mutations are not the primary force driving maize evolution, and that a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize.
+We find that expansion after domestication has increased the efficiency of purifying selection to levels exceeding those seen in teosinte. 
+Our results demonstrate that demographic and selective history in the ancient and recent past both contribute to the levels of genetic variability that present today. 
 Our results also show that for bottlenecked species, rapid post-bottleneck expansion can have wide-ranging impacts on diversity that conflict with the expected loss of diversity that occurs during the bottleneck itself.
-%JRI NOT DONE
+\jri{the whole abstract still needs work.
+other bottlenecked species?
+not very exciting.
+i think we need to add in the motivation rather than just "to quantify".
+explain that everyone else looks at long-term Ne when discussing linked selection.}
 
@@
 
-\subsection{Hard sweeps do not shape genome-wide diversity in maize} %JRI NOT DONE
+\subsection{Hard sweeps do not shape genome-wide diversity in maize} %JRI
 
@@
 The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}, but is consistent with estimates of the percentage of nonsynonmyous substitutions fixed by natural selection \citep{ross2009,sella2009,eyre2009estimating,williamson2014}.
-While larger long-term $N_e$ likely explains some of this difference, we see little change in the importance of hard sweeps in genes even looking at singleton diversity in modern maize \jri{reference appropriate figure}, suggesting other factors may contribute to these differences.
-If mutational target size scales with genome size, the larger genomes of human and maize may offer more possibilities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively small role
-In line  with this argument, studies of local adaptation in taxa with large genomes such as teosinte \citep{pyhajarvi2013} and humans \citep{fraser2013gene} find enrichment for loci in regulatory regions, while similar analyses in the small \emph{Arabidopsis} genome finds an enrichment for nonsynonymous variants.
-
-%\jri{what \% of the windows of top H12 score genome-wide occur outside of genes? is that a number to include here?}
-% TB Says: No, because windows were 200 SNPs so even windows that don't correspond to a gene could include a gene.
-
-%\mbh{additional consideration of non-genic regions in both the results and discussion would probably be a good idea given Tanja and Sho's results}
-% TB Says: I agree, but I don't think we have the data to really get at the relative contribution of non-genic regions.
+While larger long-term $N_e$ likely explains some of this difference, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}, suggesting other factors may contribute to these differences.
+Another possibility is that, if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
+Support from this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \citep{pyhajarvi2013} and humans \citep{fraser2013gene} but for nonsynonymous variants in the small \emph{Arabidopsis} genome.
 
 \subsection{Demography influences the efficiency of purifying selection} %JRI NOT DONE
-One of our more striking findings is that the impact of purifying selection on maize and teosinte variability has varied through time.
-Although we observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), we see a greater decrease in singleton diversity in maize than teosinte (Figure \ref{fig:purify}B).
-Purifying selection is more efficient in larger populations  \cite{kimura1984}, and our observations using $\pi$  are thus consistent with our demographic model and previous work \cite{eyre1998, tenaillon2004, wright2005, ross2009} which shows a smaller long-term effective population size in maize.
-Similarly, our observations based on singleton diversity, which reflect more recent patterns of selection, again are in agreement with maize and teosinte demography; from the viewpoint of a recent mutation, the rapid expansion of maize post-domestication means that today maize has a vastly larger effective population size than does teosinte.
+One of our more striking findings is that the impact of purifying selection on maize and teosinte qualitatively changed over time. 
+We observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), but the opposite trend when we evaluate diversity using singleton polymorphisms (Figure \ref{fig:purify}B).
+The efficiency of purifying selection is proportional to the effective population size \cite{kimura1984}, and these results are thus consistent with a demestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} and our demographic analyses showing rapid expansion and a much larger modern $N_e.
+
+Although demographic change affects the efficiency of purifying selection, it may have limited implications for genetic load. 
+Recent population bottlenecks and expansions have increased the relative abundance of 
+deleterious variants in domesticated plants \cite{Gunther:2010gt,renaut2015} and human populations out of Africa \jri{is keinan2012 appropriate here?
+others?}, and such variants have been shown to play an important role in phenotypic variation \cite{Mezmouk:2014es} \jri{need some human cites here}.
+Nonetheless, demographic history may have little impact on the overall genetic load of populations \cite{do2015,simons2014}, as decreases in $N_e$ that allow weakly deleterious variants to escape selection also help purge strongly deleterious ones, and the increase of new deleterious mutations in expanding populations is mitigated by their  lower initial frequency and the increasing efficiency of purifying selection \cite{gazave2013, simons2014, lohmueller2014}.
+
+%We propose that this observation could also result from more efficient selection against such mutations.
+This proposition is in agreement not only with our maize observations, but also with the results of \cite{}, who showed via simulations that the increased efficiency of selection in a larger population can partially offset the increase in the total number of segregating deleterious variants.
+%While the large numbers of rare variants present in expanding populations may contribute significantly to the genetic basis of complex disease \cite{keinan2012}, recent analyses suggest that 
+%. However, more efficient purifying selection in larger populations, which we show operates specifically on these rare variants, should decrease genetic load by purging rare deleterious variants.
+
+%Across species, our results demonstrate that the genetic impacts of major demographic events such as bottlenecks depend not only on population size reductions, but also on subsequent, potentially rapid and recent, expansion that often comes afterwards.
+Reduced population sizes do lead to a decreased efficiency of selection \cite[e.g]{renaut2015}, which has long-term impacts on diversity measured via $\pi$.
+But new mutations, which we measure with singleton diversity, are oblivious to past demography, and therefore selection for or against such sites depends primarily on recent demographic processes.}
 
 \section{Demography and selection interact to shape genetic diversity} %JRI NOT DONE
-Importantly, our results demonstrate that abstractions such as average long-term $N_e$ cannot capture the dynamic relationship between demography and selection.
-While it is certainly the case that diversity is lost during a bottleneck, it is equally true that from the viewpoint of a mutation that arises later, the bottleneck is irrelevant.
+Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection.
+
+%While it is certainly the case that diversity is lost during a bottleneck, it is equally true that from the viewpoint of a mutation that arises later, the bottleneck is irrelevant.
 Although the relatively straight-forward demographic history of maize risks oversimplifying this observation, it is applicable to any species that has undergone demographic change.
@@
 This conflicts with an alternative possibility; that sudden expansion should lead to a comparable increase in rare variants genome-wide.
-
-The abundance of rare variants across the genome, and the efficiency of purifying selection against them, has notable implications for genetic load.
-One possibility is that the excess of rare variants present in expanding populations implies higher genetic load \cite{keinan2012}.
-However, more efficient purifying selection in larger populations, which we show operates specifically on these rare variants, should decrease genetic load by purging rare deleterious variants.
-Several studies have demonstrated that despite different demographic histories, African and European human populations have a similar genetic load \cite{do2015,simons2014}, due in part to new deleterious mutations being increasingly rare as population size increases \cite{simons2014, lohmueller2014}.
-We propose that this observation could also result from more efficient selection against such mutations.
-This proposition is in agreement not only with our maize observations, but also with the results of \cite{gazave2013}, who showed via simulations that the increased efficiency of selection in a larger population can partially offset the increase in the total number of segregating deleterious variants.
-
-Across species, our results demonstrate that the genetic impacts of major demographic events such as bottlenecks depend not only on population size reductions, but also on subsequent, potentially rapid and recent, expansion that often comes afterwards.
-Reduced population sizes do lead to a decreased efficiency of selection \cite[e.g]{renaut2015}, which has long-term impacts on diversity measured via $\pi$.
-But new mutations, which we measure with singleton diversity, are oblivious to past demography, and therefore selection for or against such sites depends primarily on recent demographic processes.
 

### 2015-11-09 manuscript.tex
very little

@@
 Researchers often seek to study drift by excluding potentially selected sites \cite{voight2005, luikart2003, gutenkunst2009}, or to study selection by focusing on site-specific patterns under the assumption that genome-wide diversity reflects primarily the action of drift \cite{akey2009}.
-
-%\jri{careful here with these citations. we're effectively calling these people out, so make sure they're really doing what you say they are!}
-%TB says: not so! Akey's is a review where he calls others out. (got rid of the other)
 
@@
 
-The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}, but is consistent with estimates of the percentage of nonsynonmyous substitutions fixed by natural selection \citep{ross2009,sella2009,eyre2009estimating,williamson2014}.
+The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}, but is consistent with estimates of the percentage of nonsynonmyous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014}.
 While larger long-term $N_e$ likely explains some of this difference, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}, suggesting other factors may contribute to these differences.
 Another possibility is that, if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
-Support from this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \citep{pyhajarvi2013} and humans \citep{fraser2013gene} but for nonsynonymous variants in the small \emph{Arabidopsis} genome.
+Support from this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the small \emph{Arabidopsis} genome.
 
@@
 We observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), but the opposite trend when we evaluate diversity using singleton polymorphisms (Figure \ref{fig:purify}B).
-The efficiency of purifying selection is proportional to the effective population size \cite{kimura1984}, and these results are thus consistent with a demestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} and our demographic analyses showing rapid expansion and a much larger modern $N_e.
+The efficiency of purifying selection is proportional to the effective population size \cite{kimura1984}, and these results are thus consistent with a demestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} and our demographic analyses showing rapid expansion and a much larger modern $N_e$.
 
@@
 deleterious variants in domesticated plants \cite{Gunther:2010gt,renaut2015} and human populations out of Africa \jri{is keinan2012 appropriate here?
-others?}, and such variants have been shown to play an important role in phenotypic variation \cite{Mezmouk:2014es} \jri{need some human cites here}.
+others?}, and such variant have been shown to play an important role in phenotypic variation \cite{Mezmouk:2014es} \jri{need some human cites here}.
 Nonetheless, demographic history may have little impact on the overall genetic load of populations \cite{do2015,simons2014}, as decreases in $N_e$ that allow weakly deleterious variants to escape selection also help purge strongly deleterious ones, and the increase of new deleterious mutations in expanding populations is mitigated by their  lower initial frequency and the increasing efficiency of purifying selection \cite{gazave2013, simons2014, lohmueller2014}.

### 2015-11-10 manuscript.tex
abstract, and more

@@
 \begin{abstract}
-  The genetic variability present in a population results from a constant interaction between drift and natural selection. 
-To quantify the interplay between drift and selection, we investigated patterns of linked selection in in the genomes of 23 domesticated maize and 13 wild teosinte individuals. 
-We show that maize went through a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid population size expansion post-domestication. 
-We observe that hard sweeps on genic mutations are not the primary force driving maize evolution, and that a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize.
-We find that expansion after domestication has increased the efficiency of purifying selection to levels exceeding those seen in teosinte. 
-Our results demonstrate that demographic and selective history in the ancient and recent past both contribute to the levels of genetic variability that present today. 
-Our results also show that for bottlenecked species, rapid post-bottleneck expansion can have wide-ranging impacts on diversity that conflict with the expected loss of diversity that occurs during the bottleneck itself.
-\jri{the whole abstract still needs work.
-other bottlenecked species?
-not very exciting.
-i think we need to add in the motivation rather than just "to quantify".
-explain that everyone else looks at long-term Ne when discussing linked selection.}
-
+Genetic drift and selection interact to determine the level of genetic variability present in a population.
+Linked selection, whereby polymporhisms at neutral sites are manipulated by selection at linked sites, plays a primary roll in shaping genetic diversity.
+However, patterns of linked selection depends heavily on a population's propensity to drift, which in turn is shaped by its demographic history.
+Generally, demographic history is accounted for in terms of long-term effective population size ($N_e$).
+In practice, bottlenecks, invasions, or novel selective pressures can cause demography to change rapidly and dramatically, but the immediate genomic effects of a changing demography have not been well studied.
+To address this, We investigated linked selection in in the genomes of 23 domesticated maize and 13 wild maize (teosinte) individuals.
+We show that maize went through a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid expansion post-domestication.
+We observe that hard sweeps on genic mutations are not the primary force driving maize evolution.
+As expected, a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize, but rapid expansion after domestication has increased the efficiency of purifying selection to levels exceeding those seen in teosinte.
+This final observation demonstrates that rapid demographic change can have wide-ranging impacts on diversity that conflict with would-be expectations based on long-term $N_e$.
 \end{abstract}
@@
 The effective size of a population is not static, and nearly all species, including flies \cite{duchen2013}, humans \cite{reich1998}, domesticates \cite{hyten2006, bovine2009}, and non-model species \cite{ellegren2014} have experienced recent or ancient changes in $N_e$.
-Although much is known about how the long-term average $N_e$ affects linked selection \jri{citations needed}, relatively little is understood about the immediate effects of more recent changes in $N_e$ on patterns of linked selection.
+Although much is known about how the long-term average $N_e$ affects linked selection \cite{cutter2013}, relatively little is understood about the immediate effects of more recent changes in $N_e$ on patterns of linked selection.
 
@@
 Because of this simple but dynamic demographic history, domesticated maize and its wild ancestor teosinte can be used to understand the effects of changing $N_e$ on linked selection. 
-In this study, we leverage the maize-teosinte study system to study these effects by first estimating the parameters of the maize domestication bottleneck using whole-genome resequencing data and then investigating the relative importance of different forms of linked selection on diversity in the recent and ancient past.
+In this study, we leverage the maize-teosinte system to study these effects by first estimating the parameters of the maize domestication bottleneck using whole-genome resequencing data and then investigating the relative importance of different forms of linked selection on diversity in the recent and ancient past.
 We show that, while patterns of overall nucleotide diversity reflect long-term differences in $Ne$, recent growth following domestication qualitatively changes these effects, thereby illustrating the importance of a comprehensive understanding of demography when considering the effects of selection genome-wide.
@@
 \subsection{Patterns of diversity differ between genic and  intergenic regions of the genome} %JRI
-To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012, hapmap3}.
+To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012, bukowski2015}.
 As a preliminary step, we evaluated levels of diversity inside and outside of genes across the genome.
@@
 \end{center}
-\caption{Pairwise diversity $\pi$ (A,B) and Tajimas D (C,D) in 1kb windows from genic and nongenic regions of maize (A,C) and teosinte (B,D).
+\caption{\textbf{A} and \textbf{B} Show pairwise diversity $\pi$, while \textbf{C} and \textbf{D} depict and Tajimas D in 1kb windows from genic and nongenic regions of maize (A,C) and teosinte (B,D).
 Shown in A and B are means $\pm$ one standard deviation.
@@
 As before, this relationship remained after we removed the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12}).
-Finally, while direct comparison of pairwise and singleton diversity within taxa is consistent with   non-equilibrium dynamics in teosinte as indicated in our coalescent-based demographic analysis, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
-\jri{tempted to remove phrase "as indicated in our coalescent-based demographic analysis"}
+Finally, while direct comparison of pairwise and singleton diversity within taxa is consistent with   non-equilibrium dynamics in teosinte, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
 
@@
 Although demographic change affects the efficiency of purifying selection, it may have limited implications for genetic load. 
-Recent population bottlenecks and expansions have increased the relative abundance of 
-deleterious variants in domesticated plants \cite{Gunther:2010gt,renaut2015} and human populations out of Africa \jri{is keinan2012 appropriate here?
-others?}, and such variant have been shown to play an important role in phenotypic variation \cite{Mezmouk:2014es} \jri{need some human cites here}.
+Recent population bottlenecks and expansions have increased the relative abundance of rare and 
+deleterious variants in domesticated plants \cite{Gunther:2010gt,renaut2015} and human populations out of Africa \cite{keinan2012,coventry2010}, and such variants may play an important role in phenotypic variation \cite{Mezmouk:2014es, coventry2010, eyre2010}.
 Nonetheless, demographic history may have little impact on the overall genetic load of populations \cite{do2015,simons2014}, as decreases in $N_e$ that allow weakly deleterious variants to escape selection also help purge strongly deleterious ones, and the increase of new deleterious mutations in expanding populations is mitigated by their  lower initial frequency and the increasing efficiency of purifying selection \cite{gazave2013, simons2014, lohmueller2014}.
@@
 We made use of published sequences from inbred accessions of teosinte (Z.
-\emph{mays} ssp.
-\emph{parviglumis}) and maize landraces from the Maize HapMap3 panel as part of the Panzea project  \cite{chia2012, lemmon2014, hapmap3}.
+{\it mays} ssp.
+{\it parviglumis}) and maize landraces from the Maize HapMap3 panel as part of the Panzea project  \cite{chia2012, lemmon2014, bukowski2015}.
 From these data, we removed 4 teosinte individuals that were not ssp.
-\emph{parviglumis} or appeared as outliers in an initial principal component analysis conducted with the package adegenet \cite{jombart2011} (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table \ref{sTab:list}).
-We also utilized a single individual of (\emph{Tripsacum dactyloides}) as an outgroup.
-All bam files are available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}).
-
-%\jri{some clarification here. did we use Lemmon's higher depth teosinte? If so, why does Li say we only had 7 high-depth bam files?  shouldn't there be more?}
-
-% TB Says: Most bam files were not high-depth, but a few were.
-I used what was made available from Panzea group via iplant in summer 2014, and precise files that were used are linked and described in table.
-ANGSD GL approach makes it OK that depth varied.
+\textit{parviglumis} or appeared as outliers in an initial principal component analysis conducted with the package adegenet \cite{jombart2011} (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table \ref{sTab:list}).
+We also utilized a single individual of (\textit{Tripsacum dactyloides}) as an outgroup.
+All bam files are available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}.
 
 \subsection{Physical and genetic maps}
-Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{hapmap3}.
-\jri{update ref.
-29 to hapmap 3 paper}
- All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following Li \emph{et al.} \cite{li2011statistical}.
-\jri{did we really only use uniquely mapping?
-or just a MQ $>30$ filter?}
+Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{bukowski2015}. 
+All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following \cite{li2011statistical}.
+%\jri{did we really only use uniquely mapping? or just a MQ $>30$ filter?}
+%both. quality filter and unique only
 We converted physical coordinates to genetic coordinates via linear interpolation of the previously published 1cM resolution NAM genetic map \cite{glaubitz2014}.
@@
 We used six each of maize and teosinte (BKN022, BKN025, BKN029, BKN030, BKN031, BKN033, TIL01, TIL03, TIL09, TIL10, TIL11 and TIL14), treating each inbred genome as a single haplotype.
-We called SNPs in ANGSD \cite{korneliussen2014} using a SNP p-value of $1e-6$ against a reference genome masked using SNPable (\url{http://lh3lh3.users.sourceforge.net/snpable.shtml}).
-\jri{fix url, not showing up in pdf}
-We then removed heterozygous genotypes and filtered sites with a mapping quality $<30$, a base quality $<20$, or a $|\emph{log}_2(\emph{depth})|<1$. \jri{fix depth, italics is weird}
+We called SNPs in ANGSD \cite{korneliussen2014} using a SNP p-value of $1e-6$ against a reference genome masked using SNPable (http://lh3lh3.users.sourceforge.net/snpable.shtml). 
+We then removed heterozygous genotypes and filtered sites with a mapping quality $<30$, a base quality $<20$, or a $|\textit{log}_2(\text{depth})|<1$.
 We ran MSMC with pattern parameters $20\times2+20\times4+10\times2$.
@@
 We used allele counts to estimate the number of singleton polymorphisms in each window, and used binomial sampling to create a second maize data set down-sampled to have the same number of samples as teosinte.
-We called genotypes in maize, teosinte, and \emph{Tripsacum} at sites with a SNP p-value $<10^{-6}$ and when the genotype posterior probability $>0.95$.
-\jri{in methods, not sure if you should use emph instead of textit or something, as italics text looks weird}
-We identified substitutions in maize and teosinte as all sites with a fixed difference with \emph{Tripsacum} and $\leq 20\%$ missing data. 
+We called genotypes in maize, teosinte, and \textit{ Tripsacum} at sites with a SNP p-value $<10^{-6}$ and when the genotype posterior probability $>0.95$. 
+We identified substitutions in maize and teosinte as all sites with a fixed difference with \textit{Tripsacum} and $\leq 20\%$ missing data. 
 Substitutions were classified as synonymous, missense, or noncoding using the ensembl variant effects predictor \cite{mclaren2010}.
@@
 We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.
-Funding was provided by NSF Plant Genome Research Project 1238014.
+Funding was provided by NSF Plant Genome Research Project 1238014 and the USDA-Agricultural Research Service.
 \end{acknowledgments}

### 2015-11-10 supportingInfo.tex
discussion section
abstract, and more

@@
 \jri{first figure should be below header}
-\begin{figure}
+\begin{figure}[h!]
   \begin{center}
@@
 \end{figure}
-\jri{why blank page before MSMC?}
 \clearpage
@@
 \caption{ Pairwise diversity surrounding synonymous and nonsynonymous
-  substitutions in maize at highly conserved (A) or unconserved (B) sites.
+  substitutions in maize at \textbf{A} highly conserved or \textbf{B}  unconserved sites.
 Bootstrap-based 95\% confidence intervals are depicted via shading.
@@
 \end{figure}
-\jri{why no confidence interval? can we add?}
 \clearpage
@@
 \end{figure}
-\jri{use of A: and A etc. needs to be consistent with main text}
 \clearpage
@@
 Plots are based on a random sample of 10,000 SNPs.
-{\bf A:} Percentage of total variance explained by each principal component for teosinte.
-{\bf B:} PC1 vs PC2 for all 13 teosinte individuals.
-{\bf C:} Percentage of total variance explained by each principal component for maize.
-{\bf D:} PC1 vs PC2 for all 23 maize individuals.
+{\bf A} displays the percentage of total variance explained by each principal component for teosinte, while {\bf B} shows PC1 vs PC2 for all 13 teosinte individuals.
+Simlarly, {\bf C} depicts the percentage of total variance explained by each principal component for maize, and {\bf D} shows PC1 vs PC2 for all 23 maize individuals.
 \label{sFig:PCA}}
@@
 %  \end{center}
-  \caption{ A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by \jri{cite chia and lemmon}   \label{sTab:list} }
+  \caption{ A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by \cite{chia2012}   \label{sTab:list} }
 \end{figure}
-\jri{add citations}
 \clearpage

### 2015-11-10 manuscript.tex
edits

@@
 Genetic drift and selection interact to determine the level of genetic variability present in a population.
-Linked selection, whereby polymporhisms at neutral sites are manipulated by selection at linked sites, plays a primary roll in shaping genetic diversity.
-However, patterns of linked selection depends heavily on a population's propensity to drift, which in turn is shaped by its demographic history.
+Linked selection, whereby polymorphisms at neutral sites are manipulated by selection at linked sites, plays a primary roll in shaping genetic diversity.
+However, patterns of linked selection depend heavily on a population's propensity to drift, which in turn is shaped by its demographic history.
 Generally, demographic history is accounted for in terms of long-term effective population size ($N_e$).
@@
 
-The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}, but is consistent with estimates of the percentage of nonsynonmyous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014}.
-While larger long-term $N_e$ likely explains some of this difference, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}, suggesting other factors may contribute to these differences.
-Another possibility is that, if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
-Support from this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the small \emph{Arabidopsis} genome.
+The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
+Comparisons of estimates of the percentage of nonsynonmyous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
+While larger long-term $N_e$ likely explains some of this difference, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}, perhaps suggesting other factors may contribute to these differences.
+One possibility for example, is that if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
+Support from this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the small \emph{Arabidopsis} genome.
+Our results, for example, are not very different from those of \cite{Halligan:2013ir}, who find no differences in diversity around nonsynonymous and synonymous substitutions and estimate that as many as 80\% of adaptive substitutions occur outside of genes.
+Future comparative analyses using a common statistical framework (e.g. \cite{corbett2015}) should allow explicit testing of this idea.
 
@@
 Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection.
+
+\cite{Elyashiv:2010ga} $N_e$ among populations leads to differences in efficiency of purifying selection
 

### 2015-11-10 manuscript.tex
stupid pnastwo

@@
 Our results, for example, are not very different from those of \cite{Halligan:2013ir}, who find no differences in diversity around nonsynonymous and synonymous substitutions and estimate that as many as 80\% of adaptive substitutions occur outside of genes.
-Future comparative analyses using a common statistical framework (e.g. \cite{corbett2015}) should allow explicit testing of this idea.
+Future comparative analyses using a common statistical framework (e.g.
+\cite{corbett2015}) and considering additional ecological and life history factors (c.f.
+\cite{leffler2012}) should allow explicit testing of this idea.
 
@@
 We observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), but the opposite trend when we evaluate diversity using singleton polymorphisms (Figure \ref{fig:purify}B).
-The efficiency of purifying selection is proportional to the effective population size \cite{kimura1984}, and these results are thus consistent with a demestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} and our demographic analyses showing rapid expansion and a much larger modern $N_e$.
+The efficiency of purifying selection is proportional to the effective population size \cite{kimura1984}, and these results are thus consistent with a domestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} and our demographic analyses showing rapid expansion and a much larger modern $N_e$.
 
@@
 
-%We propose that this observation could also result from more efficient selection against such mutations.
-This proposition is in agreement not only with our maize observations, but also with the results of \cite{}, who showed via simulations that the increased efficiency of selection in a larger population can partially offset the increase in the total number of segregating deleterious variants.
-%While the large numbers of rare variants present in expanding populations may contribute significantly to the genetic basis of complex disease \cite{keinan2012}, recent analyses suggest that 
-%. However, more efficient purifying selection in larger populations, which we show operates specifically on these rare variants, should decrease genetic load by purging rare deleterious variants.
+\section{Rapid changes in linked selection} %JRI NOT DONE
+Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
+While a number of authors have incorporated demographic change into methods for detecting selection \cite{eyre2009estimating} \jri{include other sites.
+maybe zeng and charlesworth 2010 genetics?, XP-CLR?
+others?} and others have compared estimates of the efficiency of selection across species \citex or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size
 
-%Across species, our results demonstrate that the genetic impacts of major demographic events such as bottlenecks depend not only on population size reductions, but also on subsequent, potentially rapid and recent, expansion that often comes afterwards.
-Reduced population sizes do lead to a decreased efficiency of selection \cite[e.g]{renaut2015}, which has long-term impacts on diversity measured via $\pi$.
-But new mutations, which we measure with singleton diversity, are oblivious to past demography, and therefore selection for or against such sites depends primarily on recent demographic processes.}
+ long-term estimates of the effective population size or
 
-\section{Demography and selection interact to shape genetic diversity} %JRI NOT DONE
-Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection.
-
-\cite{Elyashiv:2010ga} $N_e$ among populations leads to differences in efficiency of purifying selection
+ cite e.g. 
+ $N_e$ among populations leads to differences in efficiency of purifying selection
 

### 2015-11-10 manuscript.tex
updates

--- 2015-11-10 manuscript.tex
@@
 
-We further made use of a large genotyping data set of more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize $N_e$ from singleton counts. \jri{imputed or nonimputed?}
+We further made use of a large genotyping data set of more than 4,000 partially imputed maize landraces \cite{Hearne2015} to estimate the modern maize $N_e$ from singleton counts. 
 We filtered these data to include only SNPs with data in $\geq 1,500$ individuals, and then projected the SFS down to a sample of 500 individuals by sampling each marker without replacement 1,000 times according to the observed allele frequencies.

### 2015-11-10 manuscript.tex
minor edits

--- 2015-11-10 manuscript.tex
@@
 
-\subsection{Demography influences the efficiency of purifying selection} %JRI NOT DONE
+\subsection{Demography influences the efficiency of purifying selection} %JRI 
 One of our more striking findings is that the impact of purifying selection on maize and teosinte qualitatively changed over time. 
@@
 others?} and others have compared estimates of the efficiency of selection across species \citex or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size
-
- long-term estimates of the effective population size or
 

### 2015-11-10 manuscript.tex
draft

@@
 
-\section{Rapid changes in linked selection} %JRI NOT DONE
+\section{Rapid changes in linked selection} %JRI 
 Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
-While a number of authors have incorporated demographic change into methods for detecting selection \cite{eyre2009estimating} \jri{include other sites.
+While a number of authors have incorporated demographic change into methods for detecting selection \cite{eyre2009estimating} \jri{include other citations.
 maybe zeng and charlesworth 2010 genetics?, XP-CLR?
-others?} and others have compared estimates of the efficiency of selection across species \citex or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size
-
- cite e.g. 
- $N_e$ among populations leads to differences in efficiency of purifying selection
-
-%While it is certainly the case that diversity is lost during a bottleneck, it is equally true that from the viewpoint of a mutation that arises later, the bottleneck is irrelevant.
-Although the relatively straight-forward demographic history of maize risks oversimplifying this observation, it is applicable to any species that has undergone demographic change.
-For instance, human populations have experienced recent growth much like maize has, and this growth corresponds to a documented excess of rare genetic variants \cite{keinan2012, tennessen2012, coventry2010}.
-Our results suggest that in such scenarios, ongoing background selection may lead to a dearth of rare variants in and around genes compared to outside of them.
-This conflicts with an alternative possibility; that sudden expansion should lead to a comparable increase in rare variants genome-wide.
-
-%\cite{corbett2015} recently showed that selection tends to have an elevated impact reducing diversity in species with large population size compared to small, which may explain why maize and humans, both bottlenecked species, show similar patterns while \emph{drosophila} and \emph{capsella} behave differently.
+others?} and others have compared estimates of the efficiency of adaptive and purifying selection across species \citex or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size \cite{corbett2015,leffler2012}.
+Our results show that demographic change over short periods of time can quickly change the dynamics of linked selection: mutations arising in extant maize populations are much more strongly impacted by the effects of selection on linked sites than would be suggested by analyses using long-term effective population size.
+As many natural and domesticated populations have undergone considerable demographic change in their recent past, long-term comparisons of $N_e$ are likely not informative about current processes affecting allele frequency trajectories.
 
@@
 All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following \cite{li2011statistical}.
-%\jri{did we really only use uniquely mapping? or just a MQ $>30$ filter?}
-%both. quality filter and unique only
 We converted physical coordinates to genetic coordinates via linear interpolation of the previously published 1cM resolution NAM genetic map \cite{glaubitz2014}.

### 2015-11-10 manuscript.tex
tiny update

@@
 
-\jri{use of (A) or {\bf A} needs to be consistent in figure legends} Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes that are under weak purifying selection and thus have higher genetic diversity \cite{enard2014}.
+Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes that are under weak purifying selection and thus have higher genetic diversity \cite{enard2014}.
 To address this concern, we took advantage of genome-wide estimates of evolutionary constraint \cite{rodgers2015} calculated using genomic evolutionary rate profile (GERP) scores \cite{davydov2010}. 

### 2015-11-11 manuscript.tex
edits to intro., results
tweak

@@
 Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
+Our assessment of the historical demography of maize and teosinte provides important context for subsequent analyses of linked selection.
 
@@
 
-Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because diversity in such cases  is not necessarily reduced \cite{innan2004,messer2013}.
-Such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity in genes. 
+Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because in such cases diversity around the substitution may be reduced to a lesser degree \cite{innan2004,messer2013}.
+Nonetheless, such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity at linked sites in genes.
 To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 

### 2015-11-11 manuscript.tex
matt edit

@@
 To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
-Qualitative differences between maize and teosinte remained unchanged even after removing genes in the top 20\% quantile of H12 (Figure \ref{sFig:H12}A).
+Qualitative differences between maize and teosinte in patterns of diversity within and outside of genes remained unchanged even after removing genes in the top 20\% quantile of H12 (Figure \ref{sFig:H12}A).
 %\mbh{This makes me wonder about the top 20\%. Is the pattern of diversity reduction different in this fraction of the genome?  If so, that seems like something that should be mentioned.}

### 2015-11-12 manuscript.tex
small but meaningful changes to intro. and results

@@
 \dropcap{T}he genetic diversity of populations is determined by a constant interplay between genetic drift and natural selection. 
-Drift is a consequence of a finite population size and the random sampling of gametes every generation \cite{dobzhansky1957}. 
+Drift is a consequence of a finite population size and the random sampling of gametes each generation \cite{dobzhansky1957}. 
 In contrast to the stochastic effects of drift, selection systematically alters allele frequencies by favoring particular alleles at the expense of others as a result of their effects on fitness.
-Researchers often seek to study drift by excluding potentially selected sites \cite{voight2005, luikart2003, gutenkunst2009}, or to study selection by focusing on site-specific patterns under the assumption that genome-wide diversity reflects primarily the action of drift \cite{akey2009}.
+Researchers often study drift by excluding potentially selected sites \cite{voight2005, luikart2003, gutenkunst2009}, or selection by focusing on site-specific patterns under the assumption that genome-wide diversity reflects primarily the action of drift \cite{akey2009}.
 
@@
 Because of this simple but dynamic demographic history, domesticated maize and its wild ancestor teosinte can be used to understand the effects of changing $N_e$ on linked selection. 
-In this study, we leverage the maize-teosinte system to study these effects by first estimating the parameters of the maize domestication bottleneck using whole-genome resequencing data and then investigating the relative importance of different forms of linked selection on diversity in the recent and ancient past.
+In this study, we leverage the maize-teosinte system to study these effects by first estimating the parameters of the maize domestication bottleneck using whole-genome resequencing data and then investigating the relative importance of different forms of linked selection on diversity in the ancient and more recent past.
 We show that, while patterns of overall nucleotide diversity reflect long-term differences in $Ne$, recent growth following domestication qualitatively changes these effects, thereby illustrating the importance of a comprehensive understanding of demography when considering the effects of selection genome-wide.
@@
 Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
-Our assessment of the historical demography of maize and teosinte provides important context for subsequent analyses of linked selection.
+Our assessment of the historical demography of maize and teosinte provides context for subsequent analyses of linked selection.
 
@@
 We observe this pattern around the causative amino acid substitution in the the maize domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005, wang2015}.
-Genome-wide, however, we observe no differences in diversity between synonymous and missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
+Genome-wide, however, we observe no differences in diversity at sites near synonymous versus missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
 
@@
 
-Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes that are under weak purifying selection and thus have higher genetic diversity \cite{enard2014}.
+Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes that, due to weak purifying selection, have higher genetic diversity \cite{enard2014}.
 To address this concern, we took advantage of genome-wide estimates of evolutionary constraint \cite{rodgers2015} calculated using genomic evolutionary rate profile (GERP) scores \cite{davydov2010}. 
 We then evaluated substitutions only in subsets of genes in the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection. 
-As expected, we see higher diversity around substitutions in genes under weak purifying selection, but we still find no difference between synonymous and missense substitutions in either subset of the data (Figure \ref{sFig:consUncons}).
+As expected, we see higher diversity around substitutions in genes under weak purifying selection, but we still find no difference in diversity near synonymous and missense substitutions in either subset of the data (Figure \ref{sFig:consUncons}).
 Taken together, these data suggest hard sweeps do not play a major role in patterning genic diversity in either maize or teosinte.
@@
 
-Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple mutations or standing genetic variation, because in such cases diversity around the substitution may be reduced to a lesser degree \cite{innan2004,messer2013}.
-Nonetheless, such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation for the observed reduction of diversity at linked sites in genes.
+Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple beneficial mutations or standing genetic variation, because in such cases diversity around the substitution may be reduced to a lesser degree \cite{innan2004,messer2013}.
+Nonetheless, such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation to background selection for the observed reduction of diversity at linked sites in genes.
 To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
@@
 %TB Says: Sites in the top 20% of H12 are unlikely to be under background selection, so maize vs teo background selection pattern at these sites seems strange to investigate.
-We interpret these combined results as suggesting that purifying selection has left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in effective population size.
+We interpret these combined results as suggesting that purifying selection has predominantly shaped diversity near genes and left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in long-term effective population size.
 

### 2015-11-12 manuscript.tex
updates

--- 2015-11-11 manuscript.tex
@@
 \begin{abstract}
-Genetic drift and selection interact to determine the level of genetic variability present in a population.
-Linked selection, whereby polymorphisms at neutral sites are manipulated by selection at linked sites, plays a primary roll in shaping genetic diversity.
-However, patterns of linked selection depend heavily on a population's propensity to drift, which in turn is shaped by its demographic history.
-Generally, demographic history is accounted for in terms of long-term effective population size ($N_e$).
-In practice, bottlenecks, invasions, or novel selective pressures can cause demography to change rapidly and dramatically, but the immediate genomic effects of a changing demography have not been well studied.
-To address this, We investigated linked selection in in the genomes of 23 domesticated maize and 13 wild maize (teosinte) individuals.
+The interaction between genetic drift and selection in shaping genetic diversity is not fully understood.
+In particular, a population's propensity to drift is typically summarized by its long-term effective population size ($N_e$) but rapidly changing population demographics may complicate this relationship.
+To better understand how changing demography impacts selection, we investigated linked selection in in the genomes of 23 domesticated maize and 13 wild maize (teosinte) individuals.
 We show that maize went through a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid expansion post-domestication.
@@
 Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
-While a number of authors have incorporated demographic change into methods for detecting selection \cite{eyre2009estimating} \jri{include other citations.
-maybe zeng and charlesworth 2010 genetics?, XP-CLR?
-others?} and others have compared estimates of the efficiency of adaptive and purifying selection across species \citex or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size \cite{corbett2015,leffler2012}.
+While a number of authors have tested for selection using methods that explicitly incorporate or are robust to demographic change \cite{eyre2009estimating, chen2010, zeng2010} and others have compared estimates of the efficiency of adaptive and purifying selection across species \cite{popadin2012} or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size \cite{corbett2015,leffler2012}.
 Our results show that demographic change over short periods of time can quickly change the dynamics of linked selection: mutations arising in extant maize populations are much more strongly impacted by the effects of selection on linked sites than would be suggested by analyses using long-term effective population size.
@@
 \subsection{BASH, R, and Python scripts}
-All scripts used for analysis are available in an online repository at \textcolor{red}{REPO ADDRESS HERE.}
+All scripts used for analysis are available in an online repository at https://github.com/timbeissinger/Maize-Teo-Scripts.
 

### 2015-11-13 manuscript.tex
update
edits

@@
   The interaction between genetic drift and selection in shaping genetic diversity is not fully understood.
-In particular, a population's propensity to drift is typically summarized by its long-term effective population size ($N_e$) but rapidly changing population demographics may complicate this relationship.
-To better understand how changing demography impacts selection, we investigated linked selection in in the genomes of 23 domesticated maize and 13 wild maize (teosinte) individuals.
+In particular, a population's propensity to drift is typically summarized by its long-term effective population size ($N_e$), but rapidly changing population demographics may complicate this relationship.
+To better understand how changing demography impacts selection, we investigated linked selection in the genomes of 23 domesticated maize and 13 wild maize (teosinte) individuals.
 We show that maize went through a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid expansion post-domestication.
 We observe that hard sweeps on genic mutations are not the primary force driving maize evolution.
-As expected, a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize, but rapid expansion after domestication has increased the efficiency of purifying selection to levels exceeding those seen in teosinte.
+As expected, a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize, but rapid expansion after domestication has since increased the efficiency of purifying selection to levels exceeding those seen in teosinte.
 This final observation demonstrates that rapid demographic change can have wide-ranging impacts on diversity that conflict with would-be expectations based on long-term $N_e$.
@@
 
+In addition to our simple bottleneck model, we investigated two alternative approaches for demographic inference.
+First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using low frequency variants  informative of population expansion.
+This analysis yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
+Finally, we applied a model-free coalescent approach \cite{schiffels2014} using a subset of our samples. 
+Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
+Our assessment of the historical demography of maize and teosinte provides context for subsequent analyses of linked selection.
+
 \begin{figure}[!tb]
@@
 \end{figure*}
-
-In addition to our simple bottleneck model, we investigated two alternative approaches for demographic inference.
-First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using low frequency variants  informative of population expansion.
-This analysis yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
-Finally, we applied a model-free coalescent approach \cite{schiffels2014} using a subset of our samples. 
-Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
-Our assessment of the historical demography of maize and teosinte provides context for subsequent analyses of linked selection.
 
@@
 Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple beneficial mutations or standing genetic variation, because in such cases diversity around the substitution may be reduced to a lesser degree \cite{innan2004,messer2013}.
-Nonetheless, such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation to background selection for the observed reduction of diversity at linked sites in genes.
+Nonetheless, such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation to purifying selection for the observed reduction of diversity at linked sites in genes.
 To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
@@
 
-Our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of background selection  \cite{enard2014}.
+Our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of purifying selection  \cite{enard2014}.
 Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability.

### 2015-11-13 manuscript.tex
edits to beginning of discussion

--- 2015-11-12 manuscript.tex
@@
 The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
-Comparisons of estimates of the percentage of nonsynonmyous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
-While larger long-term $N_e$ likely explains some of this difference, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}, perhaps suggesting other factors may contribute to these differences.
-One possibility for example, is that if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
-Support from this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the small \emph{Arabidopsis} genome.
-Our results, for example, are not very different from those of \cite{Halligan:2013ir}, who find no differences in diversity around nonsynonymous and synonymous substitutions and estimate that as many as 80\% of adaptive substitutions occur outside of genes.
+Comparisons of the estimated percentage of nonsynonymous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
+While differences in long-term $N_e$ likely explain some of the observed variation across species, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}, perhaps suggesting other factors may contribute to these differences as well.
+One possibility, for example, is that, if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
+Support for this idea comes from numerous cases of adaptive transposable element insertions modifying gene regulation in maize \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the smaller \emph{Arabidopsis} genome.
+Results similar to ours have also been found by \cite{Halligan:2013ir}, who find no differences in diversity around nonsynonymous and synonymous substitutions in murid rodents and estimate that as many as 80\% of adaptive substitutions occur outside of genes.
 Future comparative analyses using a common statistical framework (e.g.

### 2015-11-13 supportingInfo.tex
updates

@@
 
-\jri{first figure should be below header}
 \begin{figure}[h!]
@@
   \begin{center}
-    \includegraphics[width=.85\textwidth]{FigsAndFiles/BKN_TIL_msmc.pdf} \\
+    \includegraphics[width=.85\textwidth]{FigsAndFiles/TIL_BKN_bootstrapping_msmc.pdf}
     \end{center}
-\caption{\jri{need MSMC caption, remove gray background, fix axis labels and plot legend.} \label{sFig:msmc}}
+\caption{Population size estimates from six haplotypes (six highly inbred lines) from both maize and teosinte revealed by MSMC analyses. \label{sFig:msmc}}
 \end{figure}

### 2015-11-13 manuscript.tex
edits to discussion

@@
 Linked selection can take the form of hitch-hiking, when the frequency of a neutral allele changes as a result of positive selection at a physically linked site \cite{smith1974}, or background selection, where diversity is reduced at loci linked to a site undergoing selection against deleterious alleles \cite{charlesworth1993}.
-Recent work in a \emph{Drosophila}, for example, has shown that virtually the entire genome is impacted by the combined effects of these processes \cite{sella2009,elyashiv2014,andolfatto2005}.
+Recent work in \emph{Drosophila}, for example, has shown that virtually the entire genome is impacted by the combined effects of these processes \cite{sella2009,elyashiv2014,andolfatto2005}.
 
@@
 Comparisons of the estimated percentage of nonsynonymous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
-While differences in long-term $N_e$ likely explain some of the observed variation across species, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}, perhaps suggesting other factors may contribute to these differences as well.
+While differences in long-term $N_e$ likely explain some of the observed variation across species, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}), perhaps suggesting other factors may contribute to these differences as well.
 One possibility, for example, is that, if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.

### 2015-11-13 manuscript.tex
murids

--- 2015-11-13 manuscript.tex
@@
 The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
-Comparisons of the estimated percentage of nonsynonymous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
-While differences in long-term $N_e$ likely explain some of the observed variation across species, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}, perhaps suggesting other factors may contribute to these differences as well.
-One possibility, for example, is that, if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
-Support for this idea comes from numerous cases of adaptive transposable element insertions modifying gene regulation in maize \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the smaller \emph{Arabidopsis} genome.
-Results similar to ours have also been found by \cite{Halligan:2013ir}, who find no differences in diversity around nonsynonymous and synonymous substitutions in murid rodents and estimate that as many as 80\% of adaptive substitutions occur outside of genes.
+Comparisons of estimates of the percentage of nonsynonmyous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
+While larger long-term $N_e$ likely explains some of this difference, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}, perhaps suggesting other factors may contribute to these differences.
+One possibility for example, is that if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
+Support from this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the small \emph{Arabidopsis} genome.
+Our results, for example, are not dissimilar to findings in the comparably-sized mouse genome, where no differences are seen in diversity around nonsynonymous and synonymous substitutions in spite of a large $N_e$ and  as many as 80\% of adaptive substitutions occur outside of genes\cite{Halligan:2013ir}.
 Future comparative analyses using a common statistical framework (e.g.

### 2015-11-13 manuscript.tex

--- 2015-11-13 manuscript.tex
@@
 The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
-Comparisons of the estimated percentage of nonsynonymous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
-While differences in long-term $N_e$ likely explain some of the observed variation across species, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}), perhaps suggesting other factors may contribute to these differences as well.
+Comparisons of the estimated percentages of nonsynonmyous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
+While differences in long-term $N_e$ likely explains some of the observed variation across species, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}), perhaps suggesting other factors may contribute to these differences as well.
 One possibility, for example, is that, if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
-Support for this idea comes from numerous cases of adaptive transposable element insertions modifying gene regulation in maize \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the smaller \emph{Arabidopsis} genome.
-Results similar to ours have also been found by \cite{Halligan:2013ir}, who find no differences in diversity around nonsynonymous and synonymous substitutions in murid rodents and estimate that as many as 80\% of adaptive substitutions occur outside of genes.
+Support for this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the smaller \emph{Arabidopsis} genome.
+Our results, for example, are not dissimilar to findings in the comparably-sized mouse genome, where no differences are seen in diversity around nonsynonymous and synonymous substitutions in spite of a large $N_e$ and  as many as 80\% of adaptive substitutions occur outside of genes\cite{Halligan:2013ir}.
 Future comparative analyses using a common statistical framework (e.g.
@@
 We observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), but the opposite trend when we evaluate diversity using singleton polymorphisms (Figure \ref{fig:purify}B).
-The efficiency of purifying selection is proportional to the effective population size \cite{kimura1984}, and these results are thus consistent with a domestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} and our demographic analyses showing rapid expansion and a much larger modern $N_e$.
+The efficiency of purifying selection is proportional to effective population size \cite{kimura1984}, and these results are thus consistent with our demographic analyses which show a domestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} followed by recent rapid expansion and a much larger modern $N_e$.
 

### 2015-11-13 preprint.tex
preprint.tex

--- 2015-11-13 manuscript.tex
@@
-\documentclass{pnastwo}
+\documentclass[12pt,a4paper]{article}
 \usepackage{cite}
@@
 \usepackage{pgf}
-%\usepackage{hyperref}scho
+\usepackage[margin=1in]{geometry}
+%\usepackage{hyperref}
 \usepackage[normalem]{ulem}
@@
 \newcommand{\X}{\textcolor{red}{\bf X}}
-\usepackage{multibib}
+\usepackage{natbib}
+\usepackage{authblk}
+\usepackage[colorlinks=true,
+            linkcolor=red,
+            urlcolor=blue,
+            citecolor=gray]{hyperref}
+
+\title{More than one Author with different Affiliations}
+\author[1,2,3]{Timothy M. Beissinger}
+\author[4]{Li Wang}
+\author[1]{Kate Crosby}
+\author[1]{Arun Durvasula}
+\author[4]{Matthew B. Hufford}
+\author[1,5]{Jeffrey Ross-Ibarra}
+\affil[1]{Dept. of Plant Sciences, University of California, Davis, CA, USA}
+\affil[2]{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}
+\affil[3]{Division of Plant Sciences, University of Missouri, Columbia, MO, USA}
+\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA}
+\affil[5]{Genome Center and Center for Population Biology, University of California, Davis, CA, USA}
+
+\renewcommand\Authands{ and }
 
@@
 \title{Recent demography drives changes in linked selection across the maize genome}
-\author{Timothy M.
-Beissinger\affil{1}{Dept.
-of Plant Sciences, University of California, Davis, CA, USA}\affil{2}{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}\affil{3}{Division of Plant Sciences, University of Missouri, Columbia, MO, USA}, Li Wang\affil{4}{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA}\affil{5}{Genome Informatics Facility, Iowa State University, Ames, IA, USA}, Kate Crosby\affil{1}{}, Arun Durvasula\affil{1}{}, Matthew B.
-Hufford\affil{4}{}, \and Jeffrey Ross-Ibarra\affil{1}{}\affil{6}{Genome Center and Center for Population Biology, University of
-    California, Davis, CA, USA} }
 
-\significancetext{Both selection and demographic change play important roles in shaping diversity across the genome, but clear empirical examples of the interplay of these two forces are lacking.
-Here we document the combined effects of demography and linked selection on genome-wide diversity in domesticated maize and its wild ancestor teosinte.
+%\author{\affil{1}{}\affil{2}{}\affil{3}{}, \affil{4}{}\affil{5}{Genome Informatics Facility, Iowa State University, Ames, IA, USA}, \affil{1}{}, \affil{1}{}, \affil{4}{}, \and Jeffrey Ross-Ibarra\affil{1}{}\affil{6}{} }
+
+%\significancetext{Both selection and demographic change play important roles in shaping diversity across the genome, but clear empirical examples of the interplay of these two forces are lacking.
+%Here we document the combined effects of demography and linked selection on genome-wide diversity in domesticated maize and its wild ancestor teosinte.
 We estimate that maize underwent a bottleneck to  $\approx 5\%$ of the size of the ancestral teosinte population, but that recent expansion has resulted in a maize population perhaps orders of magnitude larger than teosinte.
@@
 \maketitle
-
-\begin{article}
 
@@
 
-\dropcap{T}he genetic diversity of populations is determined by a constant interplay between genetic drift and natural selection. 
+\section*{Introduction}
+The genetic diversity of populations is determined by a constant interplay between genetic drift and natural selection. 
 Drift is a consequence of a finite population size and the random sampling of gametes each generation \cite{dobzhansky1957}. 
@@
 
-\section{Results}
+\section*{Results}
 
-\subsection{Patterns of diversity differ between genic and  intergenic regions of the genome} %JRI
+\subsection*{Patterns of diversity differ between genic and  intergenic regions of the genome} %JRI
 To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012, bukowski2015}.
@@
 
-\subsection{Demography of maize domestication} %JRI
+\subsection*{Demography of maize domestication} %JRI
 After establishing that genic sites are not evolving neutrally, we used sites $>5$kb from genes to estimate the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}). 
@@
 
-\subsection{Hard sweeps do not explain diversity differences} %JRI
+\subsection*{Hard sweeps do not explain diversity differences} %JRI
 When selection increases the frequency of a new beneficial mutation, a signature of reduced diversity is left at surrounding linked sites \cite{smith1974}.
@@
 
-\subsection{Diversity is strongly influenced by purifying selection} %JRI
+\subsection*{Diversity is strongly influenced by purifying selection} %JRI
 
@@
 
-\subsection{Population expansion leads to stronger purifying selection in modern maize} %JRI
+\subsection*{Population expansion leads to stronger purifying selection in modern maize} %JRI
 Motivated by the rapid post-domestication expansion of maize evident in our demographic analyses, we reasoned that low-frequency --- and thus younger --- polymorphisms might show patterns distinct from pairwise diversity.
@@
 
-\section{Discussion}
+\section*{Discussion}
 
-\subsection{Demography of domestication} %JRI
+\subsection*{Demography of domestication} %JRI
 Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
@@
 
-\subsection{Hard sweeps do not shape genome-wide diversity in maize} %JRI
+\subsection*{Hard sweeps do not shape genome-wide diversity in maize} %JRI
 
@@
 
-\subsection{Demography influences the efficiency of purifying selection} %JRI 
+\subsection*{Demography influences the efficiency of purifying selection} %JRI 
 One of our more striking findings is that the impact of purifying selection on maize and teosinte qualitatively changed over time. 
@@
 
-\section{Rapid changes in linked selection} %JRI 
+\subsection*{Rapid changes in linked selection} %JRI 
 Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
@@
 
-\begin{materials} %JRI
+\section*{materials} %JRI
 
-\subsection{BASH, R, and Python scripts}
-All scripts used for analysis are available in an online repository at https://github.com/timbeissinger/Maize-Teo-Scripts.
+\subsection*{BASH, R, and Python scripts}
+All scripts used for analysis are available in an online repository at \\ \url{https://github.com/timbeissinger/Maize-Teo-Scripts}.
 
-\subsection{Plant materials}
+\subsection*{Plant materials}
 We made use of published sequences from inbred accessions of teosinte (Z.
@@
 
-\subsection{Physical and genetic maps}
-Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{bukowski2015}. 
+\subsection*{Physical and genetic maps}
+Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} \\ (\url{ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/}) as described by \cite{bukowski2015}.
 All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following \cite{li2011statistical}.
@@
 
-\subsection{Estimating the site frequency spectrum}
+\subsection*{Estimating the site frequency spectrum}
 We estimated both the genome-wide site frequency spectrum (SFS) as well as a separate SFS for genic (within annotated transcript) and intergenic ($\geq 5kb$ from a transcript) regions. 
@@
 
-\subsection{Demographic inference}
+\subsection*{Demographic inference}
 We used the software $\delta\alpha\delta{i}$ \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS, using only sites $>5 kb$ from a gene to ameliorate the effects of linked selection.
@@
 We used six each of maize and teosinte (BKN022, BKN025, BKN029, BKN030, BKN031, BKN033, TIL01, TIL03, TIL09, TIL10, TIL11 and TIL14), treating each inbred genome as a single haplotype.
-We called SNPs in ANGSD \cite{korneliussen2014} using a SNP p-value of $1e-6$ against a reference genome masked using SNPable (http://lh3lh3.users.sourceforge.net/snpable.shtml). 
+We called SNPs in ANGSD \cite{korneliussen2014} using a SNP p-value of $1e-6$ against a reference genome masked using SNPable (\url{http://lh3lh3.users.sourceforge.net/snpable.shtml}). 
 We then removed heterozygous genotypes and filtered sites with a mapping quality $<30$, a base quality $<20$, or a $|\textit{log}_2(\text{depth})|<1$.
@@
 
-\subsection{Diversity}
+\subsection*{Diversity}
 We made use of the software ANGSD \cite{korneliussen2014} for diversity calculations and genotype calling. 
@@
 
-\subsection{Selection scan}
+\subsection*{Selection scan}
 We scanned the genome to identify sites that have experienced recent positive selection using the H12 statistic \cite{garud2015} in sliding windows of 200 SNPs with a step of 25 SNPs.
-
-\end{materials}
 

### 2015-11-13 supportingInfo.tex
preprint.tex

@@
     \end{center}
-\caption{Population size estimates from six haplotypes (six highly inbred lines) from both maize and teosinte revealed by MSMC analyses. \label{sFig:msmc}}
+\caption{Effective population size estimated over time using MSMC.
+Shown are estimates (solid lines) and boostrap resampling (dotted lines) for both maize (red) and teosinte (blue).
+Time is estimated assuming an annual generation time and a mutation rate of $\mu=3\times 10^{-8}$ \label{sFig:msmc}}
 \end{figure}

### 2015-11-13 manuscript.tex
Preprint DONE

@@
 Support for this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the smaller \emph{Arabidopsis} genome.
-Our results, for example, are not dissimilar to findings in the comparably-sized mouse genome, where no differences are seen in diversity around nonsynonymous and synonymous substitutions in spite of a large $N_e$ and  as many as 80\% of adaptive substitutions occur outside of genes\cite{Halligan:2013ir}.
+Our results, for example, are not dissimilar to findings in the comparably-sized mouse genome, where no differences are seen in diversity around nonsynonymous and synonymous substitutions in spite of a large $N_e$ and  as many as 80\% of adaptive substitutions occuring outside of genes\cite{Halligan:2013ir}.
 Future comparative analyses using a common statistical framework (e.g.
@@
 We assumed individuals were fully inbred and treated each line as a single haplotype.
-Because ANGSD cannot calculate a folded joint SFS, we first polarized SNPs using the maize reference genome and then folded spectra using $\delta\alpha\delta{i}$ \cite{gutenkunst2009}.
+Because ANGSD cannot calculate a folded joint SFS, we first polarized SNPs using the maize reference genome and then folded spectra using $\delta$a$\delta$i \cite{gutenkunst2009}.
 
 \subsection{Demographic inference}
-We used the software $\delta\alpha\delta{i}$ \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS, using only sites $>5 kb$ from a gene to ameliorate the effects of linked selection.
+We used the software $\delta$a$\delta$i \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS, using only sites $>5 kb$ from a gene to ameliorate the effects of linked selection.
 We modeled a teosinte population of constant effective size $N_a$, that at time $T_b$ generations in the past gave rise to a maize population of size $N_b$ which grew exponentially to size $N_m$ in the present (Figure \ref{fig:bottleneck}).
 The model includes migration of $M_{mt}$ individuals each generation from maize to teosinte and $M_{tm}$ individuals from teosinte to maize.
-We estimated $N_a$ using $\delta\alpha\delta{i}$'s estimation of $\theta=4N_a\mu$ from the data and a mutation rate of $\mu = 3 \times 10^{-8}$ \cite{clark2005}.
-We estimated all other parameters using 1,000 $\delta\alpha\delta{i}$ optimizations and allowing initial values between runs to be randomly perturbed by a factor of 2.  
+We estimated $N_a$ using $\delta$a$\delta$i's estimation of $\theta=4N_a\mu$ from the data and a mutation rate of $\mu = 3 \times 10^{-8}$ \cite{clark2005}.
+We estimated all other parameters using 1,000 $\delta$a$\delta$i optimizations and allowing initial values between runs to be randomly perturbed by a factor of 2.  
 Optimized parameters along with their initial values and upper and lower bounds can be found in table \ref{sTab:dadi}.
@@
 We identified substitutions in maize and teosinte as all sites with a fixed difference with \textit{Tripsacum} and $\leq 20\%$ missing data. 
-Substitutions were classified as synonymous, missense, or noncoding using the ensembl variant effects predictor \cite{mclaren2010}.
+Substitutions were classified as synonymous, or missense using the ensembl variant effects predictor \cite{mclaren2010}.
 For each window with $\geq 100bp$ of data we computed the genetic distance between the window center and the nearest synonymous and missense substitution as well as the genetic distance to the center of the nearest gene transcript.

### 2015-11-13 manuscript.tex
updates

@@
 
-\significancetext{Both selection and demographic change play important roles in shaping diversity across the genome, but clear empirical examples of the interplay of these two forces are lacking.
+\significancetext{
+  %Both selection and demographic change play important roles in shaping diversity across the genome, but clear empirical examples of the interplay of these two forces are lacking.
 Here we document the combined effects of demography and linked selection on genome-wide diversity in domesticated maize and its wild ancestor teosinte.
@@
 Our results demonstrate that a complete understanding of genome-wide patterns of diversity will require careful assessment of both demographic history and the effects of linked selection.
+  Both selection and demographic change play important roles in shaping diversity across the genome, but clear empirical examples of the interplay of these two forces are lacking.
+Here we document the combined effects of demography and linked selection on genome-wide diversity in domesticated maize and its wild ancestor teosinte.
+We estimate that although maize underwent a bottleneck to  $\approx 5\%$ of the size of the ancestral teosinte population, recent expansion has resulted in a maize population much larger than teosinte.
+Positive selection on new genic mutations has had relatively little effect on genetic diversity, but that selection against deleterious mutations has dramatically reduced diversity in and immediately around genes in both taxa.
+Importantly, the relative impact of selection depends on the age of the polymorphisms evaluated, which shows that a complete understanding of genome-wide patterns of diversity requires careful assessment of both demographic history and the effects of linked selection.
 } %JRI
@@
 We also utilized a single individual of (\textit{Tripsacum dactyloides}) as an outgroup.
-All bam files are available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}.
+All bam files are available at /iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem.
 
@@
 \bibliographystyle{pnas}
+
 \onecolumn

### 2015-11-13 manuscript.tex
updates

--- 2015-11-13 manuscript.tex
@@
 
-\significancetext{Both selection and demographic change play important roles in shaping diversity across the genome, but clear empirical examples of the interplay of these two forces are lacking.
+\significancetext{
+Both selection and demographic change play important roles in
+shaping diversity across the genome, but there are few clear empirical examples of the interplay of these two forces.
 Here we document the combined effects of demography and linked selection on genome-wide diversity in domesticated maize and its wild ancestor teosinte.
-We estimate that maize underwent a bottleneck to  $\approx 5\%$ of the size of the ancestral teosinte population, but that recent expansion has resulted in a maize population perhaps orders of magnitude larger than teosinte.
-We show that positive selection on new genic mutations has had relatively little effect on genetic diversity, but that selection against deleterious mutations has dramatically reduced diversity in and immediately around genes in both taxa.
-We find that the relative impact of selection depends on the age of the polymorphisms evaluated: while older polymorphisms in maize show more limited effects of linked selection, new mutations instead reflect its larger current size and more efficient selection.
-Our results demonstrate that a complete understanding of genome-wide patterns of diversity will require careful assessment of both demographic history and the effects of linked selection.
+We estimate that although maize underwent a bottleneck to ≈ 5\% of the size of the ancestral teosinte population, recent expansion has resulted in a maize population much larger than teosinte.
+Positive selection on new genic mutations has had relatively little effect on genetic diversity, but selection against deleterious mutations has dramatically reduced diversity in and immediately around genes in both taxa.
+The relative impact of selection depends on the age of the polymorphisms evaluated, revealing that a complete understanding of genome-wide patterns of diversity requires careful assessment of both demographic history and the effects of linked selection.
 } %JRI

### 2015-11-13 supportingInfo.tex
Preprint DONE
updates
omg what crappy hack for table captions

@@
 \renewcommand\thefigure{S\arabic{figure}}
+\renewcommand\thetable{S\arabic{table}}    
 \setcounter{figure}{0}
+\setcounter{table}{0}
 
-\begin{figure}[h!]
+\begin{figure}[h]
   \begin{center}
-    \includegraphics[width=.85\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_tga1Supp_June.png} \\
+    \includegraphics[width=\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_tga1Supp_June.png}
     \end{center}
@@
 \end{figure}
+
+\begin{table}[h]
+\caption{A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by \cite{chia2012}   } \label{sTab:list}
+\end{table}
 \clearpage
 
-\begin{figure}
-%  \begin{center}
+\begin{table}[h]  
   \begin{tabular}{c|c}
@@
     BKN035 & \\
-    BKN040 & \\
+    BKN040 & \\ \hline
   \end{tabular}  
-%  \end{center}
-  \caption{ A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by \cite{chia2012}   \label{sTab:list} }
-\end{figure}
+\end{table}
+
+\begin{table}[h]
+\caption{ Parameters, initial values, and boundaries used for model-fitting with $\delta\alpha\delta{i}$.
+Parameters are shown in the units utilized by $\delta\alpha\delta{i}$, although in the text simplified units are reported.
+}\label{sTab:dadi}
+\end{table}
 \clearpage
 
-\begin{figure}
+\begin{table}
   \def\arraystretch{2}
@@
   \end{tabular} 
-%  \end{center}
-  \caption{ Parameters, initial values, and boundaries used for model-fitting with $\delta\alpha\delta{i}$.
-Parameters are shown in the units utilized by $\delta\alpha\delta{i}$, although in the text simplified units are reported.
-\label{sTab:dadi} }
   \def\arraystretch{1} % undo stretching
-\end{figure}
-\jri{why are these shown as figures and not tables?}
-\clearpage
+\end{table}

### 2015-11-13 manuscript.tex
FINAL FINAL FINAL

@@
 \significancetext{
-Both selection and demographic change play important roles in
-shaping diversity across the genome, but there are few clear empirical examples of the interplay of these two forces.
-Here we document the combined effects of demography and linked selection on genome-wide diversity in domesticated maize and its wild ancestor teosinte.
-We estimate that although maize underwent a bottleneck to ≈ 5\% of the size of the ancestral teosinte population, recent expansion has resulted in a maize population much larger than teosinte.
-Positive selection on new genic mutations has had relatively little effect on genetic diversity, but selection against deleterious mutations has dramatically reduced diversity in and immediately around genes in both taxa.
-The relative impact of selection depends on the age of the polymorphisms evaluated, revealing that a complete understanding of genome-wide patterns of diversity requires careful assessment of both demographic history and the effects of linked selection.
+Both selection and demographic change play important roles in shaping genetic diversity, but there are few empirical examples of the interplay between these forces.
+We document the combined effects of demography and linked selection on maize genome-wide diversity.
+After going through a bottleneck ≈5\% of the size of the ancestral teosinte population, recent expansion resulted in a maize population much larger than teosinte.
+Positive selection on new genic mutations has affected genetic diversity little, but selection against deleterious mutations dramatically reduced diversity within and surrounding genes.
+The impact of selection depends on the age of the polymorphisms evaluated, revealing that a complete understanding of genome-wide diversity requires careful assessment of demographic history and the effects of linked selection.
 } %JRI

### 2015-11-16 manuscript.tex
submitted version.

@@
 \significancetext{
-Both selection and demographic change play important roles in shaping genetic diversity, but there are few empirical examples of the interplay between these forces.
-We document the combined effects of demography and linked selection on maize genome-wide diversity.
-After going through a bottleneck ≈5\% of the size of the ancestral teosinte population, recent expansion resulted in a maize population much larger than teosinte.
-Positive selection on new genic mutations has affected genetic diversity little, but selection against deleterious mutations dramatically reduced diversity within and surrounding genes.
-The impact of selection depends on the age of the polymorphisms evaluated, revealing that a complete understanding of genome-wide diversity requires careful assessment of demographic history and the effects of linked selection.
+Both selection and demographic change play important roles in shaping genetic diversity, but there are few empirical examples of the interplay between them.
+Although it is well known that both diversity and the efficacy of natural selection are reduced in smaller populations, the short-term consequences of demographic change on linked selection have not been well studied.
+We use wild and domesticated maize to demonstrate that rapid changes in population size impact the effect of selection on linked sites in ways not well described by long-term average population size.
+We also produce an unbiased estimate of maize demographic history and show that positive selection on new genic mutations plays a lesser role in shaping diversity than purifying selection against deleterious mutations.
 } %JRI
@@
 \begin{acknowledgments}
-We are indebted to Graham Coop and Simon Aeshbacher for their constructive input during this study.
+We are indebted to Graham Coop and Simon Aeschbacher for their constructive input during this study.
 We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.

### 2015-11-24 NatGen/main.tex
nature Genetics format

%\documentclass[twocolumn,twoside]{article}
\documentclass[twoside, twocolumn, letterpaper]{article}
\usepackage{graphics}
\usepackage{hyperref}
\usepackage{color}
\usepackage{NatGenArtT}
%\usepackage{NatureLetT}
\usepackage{times}
\DeclareMathAlphabet{\msfsl}{OT1}{cmss}{m}{sl}
\DeclareMathAlphabet{\msfrg}{OT1}{cmss}{n}{sl}
\renewcommand{\baselinestretch}{1}
\addtolength{\oddsidemargin}{-.2cm}
\addtolength{\evensidemargin}{-1.2cm}
\addtolength{\textwidth}{1.5cm}
\addtolength{\topmargin}{-1.75cm}
\addtolength{\textheight}{3.5cm}
\renewcommand{\textfraction}{0.01}
\renewcommand{\topfraction}{0.99}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.90}
\renewcommand{\dbltopfraction}{0.95}
\renewcommand{\dblfloatpagefraction}{0.80}
\renewcommand{\sfdefault}{phv}

\usepackage[sort&compress]{natbib}
\bibpunct{}{}{,}{s}{}{\textsuperscript{,}}
\usepackage{amsmath}
\usepackage{graphicx}

%added commands YANG
\usepackage[colorinlistoftodos]{todonotes} % comments in margins
\setlength{\marginparwidth}{1.5cm}
\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
\newcommand{\jri}[1]{\todo[size=\scriptsize, color=flame]{#1}}
\newcommand{\yang}[1]{\todo[size=\scriptsize, color=cyan]{#1}}
\newcommand{\X}{\textcolor{red}{\bf X\,}}
\newcommand{\citex}{\textcolor{red}{\bf (CITE)\,}}
\newcommand{\out}[1]{\textcolor{red}{#1}} %outline placeholder text

%\usepackage[natbib=true]{biblatex}% natbib compatibility mode
%\addbibresource{Diallel.bib}
%\bibliographystyle{NatureSeriesT}

\graphicspath{{Figure_Table/}{SI/}} % Location of the graphics files
%
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\tablename}{Supplementary Table}%
        \setcounter{figure}{0}
        \renewcommand{\figurename}{Supplementary Figure}%
}

\makeatletter
\renewcommand{\footnotesep}{-2pt}
\makeatletter

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}

% fancy for Nature Genetics
\fancyhead[RO]{\begin{picture}(600,1)(10,20)\put(440,32) {\textcolor{Dgreen}{\large{\sfbf{ARTICLES}}}}\end{picture}}
\fancyhead[RE]{\begin{picture}(600,1)(10,20)\put(10,32)  {\textcolor{Dgreen}{\large{\sfbf{ARTICLES}}}}\end{picture}}
\fancyhead[C]{\begin{picture}(600,1)(10,20)\linethickness{50pt}\put(-43,52) {{\color{pTop}{\line(1,0){615}}}}\linethickness{0.5pt}\put(10,-678) {{\line(1,0){515}}}\end{picture}}

\fancyfoot[RO]{\thepage}
\fancyfoot[LE]{\thepage}

\renewcommand{\headrulewidth}{0pt}
\fancypagestyle{plain}{
    \fancyhf{}
}
\setcounter{footnote}{0}%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Incorporation of Evolutionary Constraint Improves Genomic Prediction of Hybrid Phenotypes}

\author{
Jinliang Yang\thanks{Department of Plant Sciences, University of California, Davis, CA 95616, USA} $^,$\thanks{These authors contributed equally to this work} \hspace{0.5mm}, 
Sofiane Mezmouk$^{1,2,}$\thanks{Current address: KWS SAAT AG, Grimsehlstr. 31, 37555 Einbeck, Germany} \hspace{0.5mm}, 
Andy Baumgarten\thanks{DuPont Pioneer, Johnston, IA 50131, USA} \hspace{0.5mm}, 
Edward S. Buckler\thanks{US Department of Agriculture, Agricultural Research Service, Ithaca, NY 14853, USA} \hspace{0.5mm}, 
Katherine E. Guill\thanks{US Department of Agriculture, Agricultural Research Service, Columbia, MO 65211, USA} \hspace{0.5mm},
Michael D. McMullen$^{6,}$\thanks{Division of Plant Sciences, University of Missouri, Columbia, MO 65211, USA} \hspace{0.5mm},
Rita H. Mumm\thanks{Department of Crop Sciences, University of Illinois at Urbana-Champaign, Urbana, IL 61801, USA} \hspace{0.5mm},
and Jeffrey Ross-Ibarra$^{1,}$\thanks{Center for Population Biology and Genome Center, University of California, Davis, CA 95616, USA} $^,$\thanks{Correspondence should be addressed to J.R.-I. (rossibarra@ucdavis.edu).}\hspace{0.5mm}
}
\date{\small Manuscript intended for \emph{Nature Genetics}, \today}

\begin{document} 
\maketitle

\begin{abstract}
\noindent \bf
\noindent
Complementation of deleterious alleles has long been proposed as a major contributor to the hybrid vigor observed in offspring of inbred parents. 
We tested this hypothesis using evolutionary measures of sequence conservation to ask whether incorporating information about putatively deleterious alleles can inform genomic selection (GS) models and improve phenotypic prediction.
We measured a number of agronomic traits in both the inbred parents and hybrids of an elite maize partial diallel population and re-sequenced the parents of the population. 
We identified haplotype blocks using an identity-by-decent (IBD) analysis and scored these blocks on the basis of segregating putatively deleterious variants. 
We implement a genomic prediction model and show that incorporating sequence conservation, especially with an incomplete dominance model, improves prediction accuracy in a five-fold cross-validation experiment for several traits \emph{per se}, as well as heterosis for those traits. 
These results provide strong empirical support for incomplete dominance (or partial complementation) in explaining heterosis, and demonstrates the utility of incorporating functional annotation and its potential in phenotypic prediction and plant breeding.
\end{abstract}

\vspace{6mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent %\textbf{Why we care about deleterious variants. Discuss results of Mezmouk 2014. We are extending this in three ways}  
The phenomenon of heterosis or hybrid vigor has been observed across many species, from yeast \citep{Shapira2014} to plants \citep{shull1908composition} and vertebrates \citep{Gama2013}. 
Hybrid vigor is particularly important in agriculture, where hybrid breeding is fundamental to the production of a number of crops including rice \citep{virmani1982heterosis} and maize \citep{east1936heterosis, shull1946hybrid}.
A number of hypotheses have been put forth to explain the phenomenon, including gene dosage \citep{birchler2003search}, overdominance \citep{east1936heterosis, schwartz1973single, krieger2010flowering, frascaroli2007classical}, 
%pseudo-overdomiance \cite[]{graham1997characterization, McMullen2009}, 
and epistasis \citep{minvielle1987dominance, schnell1992multiplicative}.

Complementation of recessive deleterious alleles \citep{crow199890,Charlesworth2009}, however, remains the simplest genetic explanation, and one that is supported by considerable empirical evidence \citep{garcia2008quantitative, xiao1995dominance}. It remains controversial, however, because... \out{birchler arguments here. tetraploid inbreeding depression, triploid heterosis AAB vs ABB.}\jri{red text placeholders for me to come back and fill in. to minimize comments}

\out{One of the best studied examples of hybrid vigor is that of maize. 
Hybrid maize has formed the basis of modern maize agriculture since the early 20th century \cite[]{crow1998}. maize is rad, here we leverage the system to ask X, Y, Z.
In this study, we characterized genome-wide deleterious variants using genomic evolutionary rate profiling (GERP) \citep{Cooper2005}. Using a genomic selection approach, where the evolutionary constraint information was incorporated into the training model as a biological prior under the Bayesian framework, we demonstrated the prediction accuracy could be improved for heterosis relative to permuted data.}

%Cr the benefits of cross-fertilization or hybridization were not unknown, it was not until the early 20th century that East, Shull, and others demonstrated the utility of hybrid breeding for  in the past, it was not until the early 20th century that hybridwork of Shull, East, and others that it's significance for agriculture was well appreciated \citex.
%Now, hybrid seed makes up the vast majority \jri{in progress, still reading}

%Deleterious alleles were arisen from new mutations during meiosis. In maize, about 90 new mutations were generated per meiosis \cite[]{Clark2005}, majority of which were deleterious according to empirical estimates \cite[]{Joseph2004}. In a natural outcross population, the negative effects on fitness of these deleterious alleles make them subject to be selection against, which lead the deleterious alleles to be maintained in a low frequency \cite[]{Eyre-Walker2007}. But the deleterious alleles could not be completely purged.

%In maize, the total number of mildly deleterious mutations is substantial because of the exponential growth of population size after domestication. The modern breeding probably aims to remove these deleterious mutations and pyramiding beneficial alleles for agronomical purposes. In practice, the relatively homogeneous maize germplasm pool was artificially divided into different heterotic groups \cite[]{Heerwaarden2012}. It enabled the improvement of germplasm pools to be conducted in a parallel fashion, and therefore, facilitated the breeding efficiency. Using this hybrid breeding approach, the maize yield has been steadily improved since the early 20th century \cite[]{duvick2001biotechnology}. However, removing deleterious mutations in low recombination regions or in tightly linked regions become less effective. Studies indicated that residual heterozygosity correlates negatively with recombination \cite[]{Gore2009, McMullen2009} and the low recombination is effective over long period of time \cite[]{Haddrill2007}. As a consequence, the deleterious alleles would be accumulated in the low recombination regions, such as the pericentromeric regions in maize, and the vigorous performance could be realized by combining two sets of non-deleterious or beneficial alleles in repulsion state, thus lead to pesudo-overdominance. A recent QTL study identified loci controlling for heterosis are enriched in centromeric regions \cite[]{Lariepe2012}, which partly support this pesudo-overdominance hypothesis.

%Typical heterosis traits, for example, grain yield or plant height, were highly polygenic \citep{huang2010genetic, peiffer2014genetic} and were believed to be controlled by many large effect loci and thousands of hard-to-be-detected minor effect loci. Traditional QTL or GWAS studies could only estimate gene action at large effect loci. However, the accumulative effects of the many minor effect loci had not been considered. Therefore, the heterosis hypotheses tested using QTL or GWAS approaches might be somewhat biased.
%Despite the importance of deleterious alleles in contributing to heterosis, they have not been systematically investigated probably because of their low frequencies in the population and mostly exhibiting minor effects. Here, we employed a genomic selection (GS) approach to simultaneously estimate genome-wide deleterious variants in a half diallel population. The diallel population was composed of a set of hybrids, which enabled us to explore different modes of inheritance of the deleterious variants. And the study can be conducted with millions of variants but using relative little sequencing efforts. In our previous study, deleterious SNPs were found to be enriched in a SNP set identified by GWAS \cite[]{Mezmouk2014}. The deleterious variants in the study were defined as non-synonymous mutations in the coding regions. Clearly, deleterious variants are not limited to coding regions.

%%%%%%%%%%%%%%%%%%%%---phenotype------%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure*}[tb]   
  \begin{center}
   \vspace{-2mm}
   \includegraphics[width=0.8\linewidth]{Fig1_pheno.pdf}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{
   {\bfseries Phenotypic variability in an elite maize partial diallel.}  (\textbf{a}) Density plots of the BLUE values for the seven phenotypic traits. On the x-axis, the genetic values were normalized. (\textbf{b}) Boxplot of the percent best-parent heterosis (BPH). (\textbf{c}) Pairwise correlation plot of the seven phenotypic traits. On the left, Pearson correlation coefficient ($r$). Stars indicate the correlations are statistically significant. On the right, the dot plots and smoothed regression lines. (\textbf{d}) Heatmap of the specific combining ability for grain yield.  } 
\vspace{-4mm}
    \label{fig:pheno}
  \end{center}
\end{figure*}
%In the plot, ASI was calculated using pBPHmin and the other six traits were calculated using pBPHmax.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% RESULTS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{RESULTS}
\subsection*{Genetic values, heterosis and combining abilities}
\jri{numbers on axes in fig. 1c are not straightforward, can they be fixed? add to this figure a graph of heritability?}
\yang{checked and removed all passive voice.}
FIgure test \ref{fig:gerpibd}
We created a partial diallel population from 12 maize inbred lines including the reference genome line B73 \citep{schnable2009b73}, the important public inbred Mo17, and 10 proprietary inbreds with expired Plant Variety Porection (LH1, LH123HT, LH82, PH207, 4676A, PHG39, PHG47, PHG84, PHJ40, and PHZ51) which together represent much of the lineage of key heterotic germplasm pools used in present-day commercial corn hybrids \citep{nelson2008molecular}.
We collected a number of agronomically important phenotypes from the inbred parents and hybrid crosses of this diallel across three years: anthesis-silking interval (ASI, in days), days to 50\% pollen shed (DTP), days to 50\% silking (DTS),  height of primary ear (EHT, in cm), plant height (PHT, in cm), test weight (TW, in pounds), and grain yield (GY, in bu/A).

\yang{added table and figure refs.}
We derived best linear unbiased estimators (BLUEs) of each phenotype for each genotype from mixed linear models (\textbf{Supplementary Table \ref{table:traits}}).
In the models, all fixed effects were significant (Wald test \emph{P} value $<0.05$) for all traits except ASI, for which the effect of replicates within environments was not significant. \jri{can this sentence be moved to supp?} 
Genetic values were relatively normally distributed for all traits (\textbf{Figure \ref{fig:pheno}A}, Shapiro-Wilk normality test \emph{P} values $>0.05$), and broad sense heritabilities ($H^2$) ranged from 0.65 for ASI to 0.95 for PHT. \jri{can we add $H^2$ as a panel? why not report/plot $h^2$ instead?}
Some of the genetic values were highly correlated (\textbf{Figure \ref{fig:pheno}C}), i.e., DTS and DTP were positively correlated (Pearson correlation coefficient $(r) = 0.93$), PHT and EHT were positively correlated ($r = 0.79$), and GY were negatively correlated with the flowering time related traits, DTP ($r = -0.25$) and DTS ($r = -0.19$). \jri{spearman better than pearson? text says these are genetic. figure cap just says phenotypes? are correls genetic or phenotypic? }

Using  parental phenotypic data, we then estimated best-parent heterosis (BPH) for each trait (\textbf{Figure \ref{fig:pheno}B}) as well as general and specific combining ability (GCA and SCA) of the inbreds.  
Because the selected inbred lines are commercially relevant and fairly elite in performance, hybrids in this population exhibit relatively low hybrid vigor (overall mean percent BPH = 0.3\% $\pm$ 0.4\%). 
However, hybrid vigor varied among traits, from flowering time traits that showed virtually no heterosis to 95\% $\pm$ 16\% for GY.
%Finally, general and specific combining ability (GCA and SCA) were estimated following Falconer and Mackay \citep{Falconer1996}.
General and specific combining ability (GCA and SCA) also varied among traits (\textbf{Figure \ref{fig:pheno}d, Supplementary Table 2}\jri{need figure/table ref}), but B73 and Mo17, founder of two important heterotic groups, not surprisingly exhibited one of the highest SCA for GY.
\jri{need a sentence on estimating GCA/SCA in methods/supp?}

%%%%%%%%%%%%%%%%%%%%---phenotype------%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure*}[tbh]   
  \begin{center}
   \vspace{-2mm}
   \includegraphics[width=0.8\linewidth]{Fig2_gerp.pdf}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{{\bfseries Genomic features of putative deleterious SNPs.} \textbf{(a)} Histogram of GERP scores at $\sim$1.3 million SNP sites. 
%The existing of spikes were because only limited species were used to derive GERP. 
\textbf{(b)} Mean GERP scores across bins of minor allele frequency. Red and grey lines define the regression and its 95\% confidence interval. \textbf{(c)} Mean GERP scores (y-axis) in 1 cM window across 10 maize chromosomes.} 
\vspace{-4mm}
    \label{fig:gerp}
  \end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE

\subsection*{Sequence variation and evolutionary constraint}

All twelve inbreds were resequenced to an average depth of $\sim 10\times$, resulting in a filtered set of 13.8 million SNPs with an mean concordance rate of 99.1\% to SNPs genotyped in previous studies.\jri{put details in supp. if sofiane doesn't respond, do a quick comparison to sequence data from HapMap3 or SNP55K or GBS and call it good.} 
%We estimated the allelic error rate using three independent data sets: for all individuals using 41,292 overlapping SNPs on the maize SNP50 bead chip \citep{Heerwaarden2012}; for all individuals using 180,313 overlapping SNPs identified through genotyping-by-sequencing (GBS) \citep{Romay2013}; and for B73 and Mo17 using the 10,426,715 SNP from the HapMap2 project \citep{Chia2012}.  Compared to corresponding SNPs identified by previous studies, a concordance rate of 99.1\% was observed. %\jri{Sofiane: can we separate those numbers out by study? or just report for one study and mention that similar rates were seen in other studies? either way it would be nice to know what rate went with what data. also is concordance mean identical genotype? do we have minor allele rate (which is a bit more informative)? if not, skip it.} 
We annotated these SNPs using Genomic Evolutionary Rate Profiling (GERP) scores from \citep{rodgers2015recombination}, resulting in 506,898 annotated SNPs.
% than 86 million bp (or $\sim$3.7\%) of the maize reference genome were annotated as conserved, with GERP scores $>0$. Nonetheless, 506,898 (or 0.6\%) of these sites were found to segregate among the 12 inbred parents of our diallel (\textbf{Supplementary Figure 1})\jri{need table ref}.
On average, each inbred parent carried 179,144 putative deleterious variants, ranged from 156,386 (PHG35) to 195,959 (PHG84). %Mo17 carried 189,241 putative deleterious variants, ranked as the third place. 
The minor allele frequencies of these putative deleterious SNPs were negatively correlated with their GERP scores (\textbf{Figure \ref{fig:gerp}B}, $P$ value $< 0.05$, $r = -0.8$), and supporting the use of GERP as a quantitative measure of the deleteriousness of an observed variant.\jri{if n=12 why are there $>>12$ dots on figure 2b??}
Centromeric regions of each chromosome were enriched for SNPs with high GERP scores (\textbf{Figure \ref{fig:gerp}c}), consistent with data from reduced representation genotyping \citep{rodgers2015recombination} and previous observations of retained residual heterozygosity in centromeres during inbreeding \citep{McMullen2009, Gore2009}.

%We calculated the number of complementation of these putative deleterious variants among F1 hybrids, except ones that have B73 as one of their parents. The best combination (Mo17 and PHG35) complemented each other at 412,042 sites, however, it still carried 94,856 homozygous putative deleterious alleles. The correlation of the complmentation significantly (Pearson correlation test, P-value = 0.02) correlated with the SCA of the GY trait and negative correlated with the cumulative deleterious scores (\textbf{Figure 2a}), which indicate hybrid combining ability might be determined by the complementation of deleterious alleles. \yang{correlation of the number of deleterious complementation and SCA, MPH and trait per se? Or just because of the genetic distance?}

%%%%%%%%%%%%%%%%%%%%---GERP SNPs------%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure*}[tbh]   
  \begin{center}
   \vspace{-2mm}
   \includegraphics[width=0.8\linewidth]{Figure_k.pdf}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{{\bfseries Genetic variance, degree of dominance of GERP SNPs and relationships of these values with their GERP scores.} \textbf{(a)} Genetic variance was partitioned into additive and dominant variance. \textbf{(b)} The distribution of degree of dominance (k). \textbf{(c-h)} GERP scores fitted with their corresponding additive effects \textbf{(c)}, dominant effects \textbf{(d)}, dominance (k) of the reference allele \textbf{(e)}, additive variance \textbf{(f)}, dominant variance \textbf{(g)} and total variance \textbf{(h)} for seven phenotypic traits, respectively. Grey color defines their 95\% confidence interval.  } 
\vspace{-4mm}
    \label{fig:k}
  \end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\jri{fig. 3: {\bf a} what is y axis?  {\bf b} don't normalize, just show data, also reflect around 0. filter by effect size? {\bf c-h} loess instead of linear? use more different colors. {\bf c-e} what is y-axis? {\bf h} genetic variance or phenotypic?}
\subsection*{Genetic effects and variance partitioning}

We employed a genomic best linear unbiased prediction (GBLUP) approach \citep{da2014mixed} to estimate the additive and dominance variance components for each trait and effects for each  GERP-annotated SNP.\jri{is this sentence still correct?}
% to partition genetic values into additive (or allele substitution effects) and dominance deviation was employed 
Across traits, the proportion of variance explained by dominance  positively correlated with observed heterosis  (\emph{P} value $<$ 0.01), from 0 for flowering time to 30\% for GY (\textbf{Figue 3a}).\jri{what's the correlation coeff?}
%For traits exhibiting intermediate levels of heterosis, i.e. TW, PHT and EHT, 20\%, 10\% and 5\% of the genetic variance could be explained by dominance. For traits exhibiting almost no heterosis, i.e., ASI, DTS and DTP, genetic variance could be largely  explained by additivity but not dominance. Notably, the accumulative variance explained by dominance positively correlated with their levels of heterosis
We calculated the degree of dominance $k$ for each SNP following \citep{lynch1998genetics}.
Individual SNPs showed patterns of dominance broadly consistent with overall partitioning of variance components (\textbf{Figure \ref{fig:k}b}). \jri{anything else we want to say here? about \% that are overdominant? about effect sizes?}
%According to Lynch and Walsh \citep{lynch1998genetics}, we estimated the degrees of dominance ($k$) of these GERP SNPs by dividing dominant effects with their additive effects. The value of $k > 1$ indicates the GERP SNP shows an overdominant effect; $0 < k < 1$ indicates a partial dominant effect; $-1 < k < 0$ indicates a partial recessive effect; and $k < -1$ indicates a underdominant effect. 
%Note reference alleles at GERP sites were considered as beneficial alleles and the alternative alleles were deemed as deleterious alleles. 
%Consistent with our partitioning of variance components, GERP SNPs exhibit exclusively additive effects for non-heterosis traits (\textbf{Figure \ref{fig:k}b}). 
%On the contrary, an excess of alleles (N= for GY, N= for TW, N= for PHT and N= for EHT)\jri{list as numbers or \% excess?} exhibit dominant or overdominant effects for traits showing intermediate to high levels of heterosis. 
%
%Note the reasons that a large number of recessive loci were detected at GERP sites were: 1) the GERP may no the causal SNPs, the effects they accounted for might be the variants in LD with them; and 2) we did not apply any cutoff on effect detection, by chance a small negative dominant effects will be assigned;  
%
The additive effect size of SNPs correlated positively with GERP score for \out{... explain this somehow.} \jri{are the correlations in 1c genetic? can we interpret them as correlations of each trait with fitness?}
%JRI STOPPED HERE
 observed that SNPs with high GERP scores, on average, tend to have larger additive effects for traits showing low levels of heterosis (\textbf{Figure \ref{fig:k}c}).
However, dominant effects stay constant across the range of GERP score for all the traits (\textbf{Figure \ref{fig:k}d}). 
%The average dominant effects are larger for traits (GY, EHT, PHT) showing high level of heterosis, but the average dominant effects are almost =0 for traits showing mediate or no heterosis (ASI, DTS, DTP, TW).  
The degree of dominance significantly (\emph{P} value $<$ 0.01) increased as the GERP score increase (\textbf{Figure \ref{fig:k}e}).

%%%%%%%% ----- BEAN PLOT using all GERP SNPs-------- %%%%%%%%%%%
\begin{figure*}[tbh]   
  \begin{center}
   \vspace{-2mm}
   \includegraphics[width=0.8\linewidth]{Figure4_beanplots.pdf}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{{\bfseries GERP-enabled genomic prediction accuracy using additive, dominant and incomplete dominant models.} Cross-validation experiments were conducted using GERP SNPs and same set of SNPs with permuted GERP scores for traits \emph{per se} (\textbf{a}, \textbf{b} and \textbf{c}) and BPH (\textbf{d}, \textbf{e} and \textbf{f}) under additive (\textbf{a} and \textbf{d}), dominant (\textbf{b} and \textbf{e}) and incomplete dominant (\textbf{c} and \textbf{f}) models. Prediction accuracy from the real data were plotted on the left (red) and permutation results on the right (green). Horizontal bars indicate mean accuracy for each trait and the gray dashed lines indicate the overall mean accuracy. Stars above the beans indicate significantly (FDR $<$ 0.05) higher cross-validation accuracy.     } 
    \vspace{-4mm}
    \label{fig:beanplots}
  \end{center}
\end{figure*}
%%%%%%%% ------------- %%%%%%%%%%%

\subsection*{GERP-enabled genomic prediction}

The small sample size of our diallel and the general low frequency of deleterious SNPs precludes association-based approaches to evaluate the impact of variants on phenotypic variation.
To alleviate this limitation, we conceived a haplotype-based genomic selection approach in which we use estimates of evolutionary constraint across the genome to sum the individual effects of deleterious alleles within IBD blocks (See \textbf{Methods} and \textbf{Supplementary Figure X}\jri{figure ref}). Three IBD conservation matricies were constructed under additive, dominant and incomplete dominant models. The matrix was solved using a Bayesian-based statistical method (BayesC) \citep{habier2011extension} (\textbf{Methods}).
With 5-fold cross-validation, prediction arracy was derived for real data and circularly shuttled data.

In general, average prediction accuracies were higher using the additive model (mean \emph{r} = 0.81 and 0.49 for traits \emph{per se} and BPH) than the dominant model (mean \emph{r} = 0.70 and 0.42), and accuracies for heterosis traits were lower than for traits \emph{per se} (Table \ref{table:table_s3}).  A student-t test was employed to compare the mean prediction accuracies between data with real GERP information and data with permutated GERP. To account for multiple traits and multiple transformations, the FDR approach was used to correct the obtained P-values. As a result, incorporating evolutionary constraint information improved prediction accuracy for ASI and PHT \emph{per se} under an additive model and for ASI under a dominant model (FDR $<$ 0.05, Figure \ref{fig:gerpall} A and B).
GERP scores also improved prediction accuracies of heterosis (BPH) for GY under the additive model and DTP, DTS and TW under the dominant model (FDR < 0.05, Figure \iffalse\ref{fig:gerpall}\fi C and D). \jri{should these be ref to Figure \ref{fig:beanplots}?}
To rule out the possible confounding of high GERP scores and genic annotations, we re-permuted the data using only deleterious (GERP $> 0$) genic SNPs.  
Though this resulted in fewer SNPs (316, 983), the model prediction accuracies remained significantly improved for GY \emph{per se} under the additive model and for BPH of GY and PHT under the additive model (Figure \iffalse\ref{fig:genicsnp}\fi and \emph{Supplementary Table}). %\ref{table:table_s4}).

%the relevant text:
% pnas201221966 2665..2669 
%However, if complementation of recessive mutations were the sole basis of heterosis, one would predict that the two types of triploid hybrids would be equivalent for their heterotic response because the recessive mutations would be equally complemented. Moreover, the complementation hypothesis would also lead one to predict that heterosis should be basically equivalent between dip- loid and triploid hybrids. Neither of these predictions was met. "
Birchler et al., argued that if the genetic basis of heterosis could be adequately explained by complementation of deleterious mutations, one would expect that the equivalent heterosis for triploid hybrids of AAB and ABB. Under the incomplete dominant model, we simulated a set of triploid hybrids and demonstrated that the two types of hybrids different significantly for GY, but not for other traits exhibiting low to intermediate levels of heterosis (\textbf{Supplementary Figure X}\jri{need figure ref}).

%It was argued that SNPs in genic regions might have higher GERP scores than those in non-genic regions. The circular shuffling permutations may shift the high GERP scores to non-genic regions. If that is the case, the approach tended to weigh more on genic SNPs. To rule out this possibility, we elected SNPs with GERP scores >0 in genic regions only and did the circular shuffling to assign GERP scores to the same set of the selected SNPs. By doing this, the method will not take advantage of genomic positional information any more. Noted that in this study less number of SNPs was selected (N = 316, 983). Nevertheless, model prediction accuracies were significantly improved for traits \emph{per se} of GY under the additive model. For heterosis transformations, prediction accuracies were significantly improved for BPH of GY and PHT under the additive model and the prediction accuracy was significantly improved for pBPH of GY (Figure \ref{fig:genicsnp} and Table S4).

\subsection*{Posterior phenotypic variance explained and model comparisons}

To learn why the prediction performance varied among traits \emph{per se} and heterosis, we obtained the posterior variance explained by our models using the complete set of data. 
As shown in Figure \ref{fig:h2}, additive models explained more phenotypic variance for traits \emph{per se} of DTP, DTS, EHT and PHT; but explained less phenotypic variance for heterosis (BPH) of ASI, GY and TW. 
In contrast, a larger proportion of the phenotypic variance could be explained by the dominant models for heterosis (BPH) of ASI, GY and TW. \jri{is this correct? the figure looks to me to disagree with this statement.}
This difference was particularly striking for grain yield under the dominant model, where only 3\% of the variance in trait \emph{per se} could be explained but 61\% of the variance in BPH was explained. \jri{I think a sentence or two here connecting variance explained with prediction accuracy would be helpful}

Heterosis transformations are largely determined by the accuracies of the parental phenotypes. 
To control for uncertainty of parental phenotypes, we estimated combining ability  directly from the hybrid population itself.
We extracted the breeding values estimated under both additive and dominant models using our haplotype blocks and incorporating GERP scores. 
We then applied the following models:
\begin{equation}
Y_{ij} = \mu + GCA_{i} + GCA_{j} + \varepsilon
\label{eq:refname1}
\end{equation}
\begin{equation}
Y_{ij} = \mu + GCA_{i} + GCA_{j} +  G_{ij} + \varepsilon
\label{eq:refname2}
\end{equation}
\begin{equation}
Y_{ij} = \mu + GCA_{i} + GCA_{j} + SCA_{ij} + \varepsilon
\label{eq:refname3}
\end{equation}
\begin{equation}
Y_{ij} = \mu + GCA_{i} + GCA_{j} + SCA_{ij} + G_{ij} + \varepsilon
\label{eq:refname4}
\end{equation}
where 
$Y_{ij}$ is the BLUE value of the hybrid crossed between the $i^{th}$ inbred and $j^{th}$ inbred; 
$\mu$, the overall mean; 
$GCA_{i}$, the general combining ability of the $i^{th}$ inbred;
$GCA_{j}$, the general combining ability of the $j^{th}$ inbred;
$SCA_{ij}$, the specific combining ability of between the $i^{th}$ and $j^{th}$ inbreds;
$G_{ij}$, breeding values estimated by our GS model for hybrid crossed between the $i^{th}$ inbred and $j^{th}$ inbred; 
$\varepsilon$, the model residuals.

Consistent with the previous analysis, haplotype blocks coded with the dominant mode of inheritance significantly improved the fit of models for heterosis for ASI and GY (equation \ref{eq:refname1} vs. equation \ref{eq:refname2}, ANOVA \emph{P} value $<0.05$, Table \ref{table:table_s5}). \jri{i'm having trouble following here.  GY gives a p of 0.04 which is not significant after bonferonni (maybe after FDR?) under the dominant model, but DOES give a significant p<4E-16 under the additive model. why not mention/discuss that? }
Comparison of models \ref{eq:refname3} and \ref{eq:refname4}, however, show no real difference (ANOVA \emph{P} value = \X), indicating that specific combining ability captures most of the parental interactions and the our haplotype blocks are unable to detect higher order interactions. \jri{ there are also at least three tests that are significant after multiple correction for 4 vs 3. is that not meaningful?}

%%%%%%%% ----- h2 plots-------- %%%%%%%%%%%
\begin{figure}[htbp]
\centering
\includegraphics[width=\linewidth]{Figure_h2.pdf}
\caption{Posterior phenotypic variance explained by deleterious genic SNPs in IBD blocks using additive and dominant models. Dark color indicates trait \emph{per se} and grey color indicates BPH. }  
\label{fig:h2}
\end{figure}
%%%%%%%% ------------- %%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% DISCUSSION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{DISCUSSION}
\jri{mention del. variants here and in results. is there variation? mo17 vs pvps? given comment below about linkage, can we calc. mean del. variants/cM? cool to know and show tons around centromere. more detail on distribution since Eli's paper only uses GBS. then add a few sentences comparing to Eli and McMullen 2009.}
In this study we have identified more than 500,000 evolutionary conserved (GERP $>$ 0) sites in the genome segregating for putatively deleterious alleles in a panel of elite maize lines. 
The non-reference alleles at these SNPs are found at low frequency, consistent with previous observations \citep{Mezmouk2014, rodgers2015recombination} and the role of selection preventing such alleles from reaching high frequency. 
Nonetheless, each inbred line carries a large number of deleterious variants, averaging $\sim 200,000$ per line.

\out{continued improvement of inbreds, constancy of heterosis \citep{Troyer:2009jf}. demography leads us to expect lots of mildy deleterious near-additive. From \citep{Gazave:2013bx}: 'As a consequence, while each individual carries a larger number of deleterious alleles than expected in the absence of growth, the average selection coefficient of each segregating allele is less deleterious.'
Because the low recombination rate in these centromeric regions, deleterious mutations were hard to be purged. The complementation of deleterious mutations in repulsion states may lead to pseudo-overdominance \citep{graham1997characterization, McMullen2009}.   
}

Across lines, however, the majority of these deleterious mutations were maintained at low frequency, consistent with previous observations \citep{rodgers2015recombination}.

\jri{\citep{Phadnis:2005ka,Simmons:1977bx} show correlation between effect size and dominance of mutations in yeast and flies}
The large number of linked deleterious alleles present means that there is likely insufficient recombination in standard breeding programs to completely purge all such alleles. 
Instead, breeders have devised a strategy --- hybrid breeding --- that circumvents much of this problem via complementation.
Consistent with this idea, our results show that prediction accuracies for both traits \emph{per se} as well as heterosis increased when SNPs were weighted by their likelihood of being deleterious.
Because there are likely thousands of deleterious alleles involved in complementation, many with relatively small effects, traditional GWAS approaches with genome-wide thresholds of multiple testing may not have the power to detect such effects.
Using a liberal significance threshold, however, even GWAS methods have identified an enrichment for deleterious genic SNPs among associated markers for a number of traits \citep{Mezmouk2014}.

Our models did not increase the prediction accuracies equally well for traits \emph{per se} and their heterosis transformations. 
This is not surprising, given the variation in genetic architecture of different phenotypic traits --- flowering time, for example, appears primarily determined by many loci of small additive effect \citep{buckler2009genetic}, and prediction accuracy is highest and the vast majority of phenotypic variance can be explained under simple additive models.  \jri{add other ex. from our data and maize lit. sentence or 2 of corr. b/t prediction accuracies and BPH in boxplots -- traits w/ higher BPH have lower $h^2$ and thus harder to predict.}

% Paritcally, with dominant model, up to 20\% of the phenotypic variance could be explained for the heterosis traits. Theoritically, BPH transformation subtracts the joint effects of the additive and dominant alleles in the best parents as residule, the substantial variance of these redidules explained by the additive or dominant models in our studies indicated that genetic components controlling for heterosis might in linked state.

% How to explain the prediction difference?  
 The variation of the prediction accuracies were relative large in this study. First of all, broad sense heritability of the traits are different. Second, from the simulation we learned that different traits may controlled by different proportion of additive, dominant and even recessive gene actions. Our naive model only built the pure additive and pure dominant effects in. For the more complicated cases, the models may not work very well.

\out{paragraph about how additive often beat recessive  consistent with role for partial dom. cite Birchler complete dom is unlikely and doesn't jive with much data. cite Huang showing partial dominance for phenotypes per se among hybrids. can cite \citep{halligan2009spontaneous} reviewing several studies finding evidence of partial dominance. while we don't have power to estimate dominance, that additive model works  suggests values of h $\approx -0.5$. 	 would be consistent with plant height  being mostly additive \citep{peiffer2014genetic}, but plant height correlated with yield (cite) so alleles affecting height will have a weak +  additive effect on yield.  }

%limitation of our current models is that we assume phenotypic traits are determined by complete additive or complete dominant effects. 
%Traits with a mixture of additive and dominant casual loci may thus fail to be predicted. 
Another limitation is likely the population size used; we may simply not have enough power to predict traits with low heritability. \out{missing GERP data b/c of alignment with Tripsacum, missing information on non-SNP variants, etc.}
%where the haplotype was coded with the SNP conservation score as the explainatory variables. 
%Genomic variants occurred at the evolutionary constraint sites were potentially deleterious. The phenotypic effects of these genetic loads and their contributions to heterosis become an interesting area to explore. However, the population size in this study is relative small and SNPs detected at sites containing high GERP scores are generally in low frequencies. The statistical power to detect the separate effects of these putative deleterious alleles becomes very low.

As genotyping costs continue to decline, genomic prediction models are increasing in popularity \citep{desta2014genomic}. 
Most previous work on genomic prediction, however, focuses exclusively on statistical properties of the models, ignoring potentially useful biological information (but see \citex for a recent example). 
In addition to providing evidence in favor of the simple complementation model for heterosis, our work here shows the utility of incorporating functional information about the variants used in genomic prediction models.  
As our functional annotations of genomes improve, we predict that including such information in genomic prediction will be vital to the development of more powerful predictive models for plant breeding.

\section*{METHODS}
Methods and any associated references are available in a separate pdf file.

\section*{ACKNOWLEDGMENTS}
Financial support for this work came from NSF (grants IOS-0820619 and IOS-1238014), USDA (grants 2009-65300-05668 and \X), DuPont Pioneer, and Mars Incorporated. 
We would like to thank Graham Coop, James Holland, \X, and \X reviewers for helpful discussion.

\section*{AUTHOR CONTRIBUTIONS}
J.Y. and J.R.-I. designed this work. E.S.B., K.E.G., M.D.M. and R.H.M. generated the data. J.Y., S.M. and J.R.-I. analyzed data. A.B. provided conceptual advice. J.Y. and J.R.-I. wrote the manuscript.

\section*{COMPETING INTERESTS STATEMENT}
The authors declare no competing financial interests.

{\scriptsize \sf
\renewcommand{\baselinestretch}{2.0}
\bibliography{Diallel}
\bibliographystyle{NatureSeriesT}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section*{ONLINE METHODS} 
\subsection*{Plant materials and phenotypic data}
%-------------------
We selected 12 maize inbred lines, broadly representative of corn belt maize germplasm \citep{mikel2006evolution}\jri{this is a diff. ref than used in introduction. which one is better?}, as parents of a partial diallel population. 
Each parent in a cross was used as both male and female and the resulting seed was bulked  (Figure \ref{fig:diallel}). 
We evaluated the 66 F1 hybrids, 12 inbred parents and two current commercial check hybrids in the field in Urbana, IL over three years (2009-2011) in an incomplete block design with three replicates each year.  
Plots consisted of four rows, with all observations taken from the inside two rows to minimize effects of shading and maturity differences from adjacent plots.  
We measured plant height (PHT, in cm), height of primary ear (EHT, in cm), days to 50\% silking (DTS), days to 50\% pollen shed (DTP), anthesis-silking interval (ASI, in days), grain yield adjusted to 15.5\% moisture (adj GY, in bu/A), and test weight (TW, in pounds). 
Overall mean phenotypic values for each cross can be found at Table \ref{table:table_s1}.

We estimated Best Linear Unbiased Estimates (BLUEs) of the genetic effects in ASReml-R \citep{gilmour2009asreml} with the following linear model: 
%
\[Y_{ijkl} = \mu + \varsigma_{i} + \delta_{ij} + \beta_{jk} + \alpha_{l} +  \varsigma_{i} \cdot \alpha_{l} + \varepsilon\]
%
where 
$Y_{ijkl}$ is the phenotypic value of the $l^{th}$ genotype evaluated in the $k^{th}$ block of the $j^{th}$ replicate within the $i^{th}$ year; 
$\mu$, the overall mean; 
$\varsigma_{i}$, the fixed effect of the $i^{th}$ year;
$\delta_{ij}$, the fixed effect of the $j^{th}$ replicate nested in the $i^{th}$ year; 
$\beta_{jk}$, the random effect of the $k^{th}$ block nested in the $j^{th}$ replicate; 
$\alpha_{l}$, the the fixed genetic effect  of the $l^{th}$ individual; 
$\varsigma_{i} \cdot \alpha_{l}$, the interaction effect of the $l^{th}$ individual with the $i^{th}$ year; 
$\varepsilon$, the model residuals.

We estimated best-parent heterosis (BPH) as:
%
%\[ MPH_{ij}=\hat{G_{ij}}-\frac{1}{2}(\hat{G_{i}}+\hat{G_{j}}) \]
\[ BPH_{min,ij}=\hat{G_{ij}}-min(\hat{G_{i}} ,\hat{G_{j}}) \] 
\[ BPH_{max,ij}=\hat{G_{ij}}-max(\hat{G_{i}} ,\hat{G_{j}}) \]
%
where $\hat{G_{ij}}$, $\hat{G_{i}}$ and $\hat{G_{j}}$ are the genetic values of the hybrid and its two parents $i$ and $j$. $BPH_{min}$ was used instead of $BPH_{max}$ for ASI.

\subsection*{Sequencing and Genotyping}

% wet lab
We extracted DNA from the 12 inbred lines following \citep{Doyle1987} and sheared the DNA on a Covaris (Woburn, Massachusetts) for library preparation. Libraries were prepared using an Illumina paired end libaray protocol with 180 bp fragments. Libraries were then sequenced at Cornell.

%Read mapping
We trimmed raw sequence reads for adapter contamination with Scythe  (\url{https://github.com/vsbuffalo/scythe}) and for quality \jri{Sofiane: what qual score? do we need to say anything about overlapping reads?} and sequence length ($\geq 20$ nucleotides) with Sickle (\url{https://github.com/najoshi/sickle}). 
We mapped filtered reads to the maize B73 reference genome (AGPv2) with bwa-mem \citep{Li2009B}, keeping reads with mapping quality (MAPQ) higher than 10 and with a best alignment score higher than the second best one for further analyses.
%SNP calling
We called single nucleotide polymorphisms (SNPs) using the $mpileup$ function from the samtools utilities \citep{Li2009}. 
To deal with known issues with paralogy in maize \citep{Chia2012}, SNPs were filtered to be heterozygote in less than 3 inbred lines, have a mean minor allele depth of at least 4, have a mean depth over all individuals lower than 30 and have missing/heterozygote alleles in fewer than 6 inbred lines.

%IBD
We used the fastIBD method implemented in BEAGLE \citep{Browning2009} to impute missing data and identify regions of identity by descent (IBD) between the 12 inbred lines. 
We then defined haplotype blocks as contiguous regions within which there were no IBD break points across all pairwise comparisons of the parental lines (Figure \ref{fig:defineibd}). IBD blocks at least 1 Kb in size were kept for further analysis.

%SIFT and MAPP \yang{hide}
%The SNPs were annotated as synonymous and non-synonymous with the software polydNdS from the analysis package of libsequence  \citep{Thornton2003} using the first transcript of each gene in B73 5b filtered gene set. Deleterious effects of amino acid changes were then predicted with both SIFT \citep{Ng2003, Ng2006} and MAPP \citep{Stone2005} software packages as described by \citep{Mezmouk2014}.

%\subsection*{Association mapping} \yang{hide}
%SNP association with heterosis (BPH and MPH) was tested assuming dominance/recessivity of the reference allele or assuming overdominance where only the heterozygote alleles are expected to be significant. For each SNP, root mean square error were used to select the best fitting model. 
%Haplotype association with heterosis were tested comparing the heterozygote alleles to all homozygote ones all confounded.

\subsection*{Genomic selection using IBD blocks incorporated with GERP scores}

%GERP
We used genome-wide estimates of evolutionary constraint (GERP) \citep{Davydov2010} estimated by \citep{rodgers2015recombination}. 
Haplotype blocks were weighted by the summed GERP scores of all deleterious (GERP score $>0$) SNPs; blocks with no deleterious SNPs were excluded from further analysis. 
This estimation was calculated under both additive and dominant modes of inheritance using a custom python script available at (\url{https://github.com/yangjl/zmSNPtools}). 
For a particular SNP with a GERP score $g$, the non-reference homozygote was assigned a value of $2g$, the heterozygote a value of $g$, and the reference homozygote a value of 0.  
Under the dominant model, both the heterozygote and the non-reference homozygote were assigned a value of $g$, with the reference homozygote again assigned a value of 0.
To conduct prediction, a 5-fold cross-validation method was used, dividing the diallel population  randomly  into training (80\%) and validation sets (20\%)  10 times. 
The BayesC option from GenSel4 \citep{habier2011extension} was used for model training, using 41,000 iterations and removing the first 1,000 as burn-in. 
After model training, prediction accuracies were obtained by comparing the predicted breeding values with the observed phenotypes in the corresponding validation sets. 
For comparison, GERP scores were permuted using 50k SNP ($> 100$Mb) windows which were circularly shuffled 10 times to estimate a null conservation score for each IBD blocks. 
Cross-validation experiments using the permuted data were conducted on the same training and validation sets.

{\scriptsize \sf
\renewcommand{\baselinestretch}{2.0}
\bibliography{Diallel}
\bibliographystyle{NatureSeriesT}
}

\onecolumn

%Figure
%----------------------------------------
% SUPPLEMENTARY FIGURES
%----------------------------------------
\pagebreak
\beginsupplement

\section*{Supporting Information}

% \section*{Supporting Table}:
\begin{table}[]
\caption{BLUE values of the seven phenotypic traits. (\url{https://github.com/RILAB/pvpDiallel/blob/master/manuscript/Figure_Table/Table_S1.trait_matrix.csv})}
\label{table:table_s1}
\end{table}

\begin{table}[]
\caption{General combining ability and specific combining ability of the seven phenotypic traits. (\url{https://github.com/RILAB/pvpDiallel/blob/master/manuscript/Figure_Table/Table_S2.CA.csv})}
\label{table:table_s2}
\end{table}

\begin{table}[]
\caption{Cross-validation results using all genome-wide deleterious SNPs. (\url{https://github.com/RILAB/pvpDiallel/blob/master/manuscript/Figure_Table/Table_S3_allsnps_FDR.csv})}
\label{table:table_s3}
\end{table}

\begin{table}[]
\caption{Cross-validation results using deleterious genic SNPs. (\url{https://github.com/RILAB/pvpDiallel/blob/master/manuscript/Figure_Table/Table_S4_genicsnps_FDR.csv})}
\label{table:table_s4}
\end{table}

\begin{table}[]
\caption{ANOVA P values of model comparisons. (\url{https://github.com/RILAB/pvpDiallel/blob/master/manuscript/Figure_Table/Table_S5_model_comp.csv})}
\label{table:table_s5}
\end{table}

%%%%%%%% ----- diallel-------- %%%%%%%%%%%
\begin{figure}[htbp]
\centering
\includegraphics[width=\linewidth]{SFig_diallel.pdf}
\caption{Twelve maize inbred lines were selected and crossed in a partial diallel fashion. Each inbred lines was used as both male and female and the resulting F1s were bulked. }
\label{fig:diallel}
\end{figure}

%%%%%%%% ----- define IBD-------- %%%%%%%%%%%
\begin{figure}[htbp]
\centering
\includegraphics[width=\linewidth]{SFig_define_IBD.pdf}
\caption{Haplotype block identification using an IBD approach. In the upper panel, regions in red are IBD blocks identified by pairwise comparison of the two parental lines of a hybrid. The vertical dashed lines define haplotype blocks. In the lower panel, hybrid genotypes in each block are coded as heterozygotes (0) or homozygotes (1).}
\label{fig:defineibd}
\end{figure}
%%%%%%%% ------------- %%%%%%%%%%%

%%%%%%%% ----- GERP dis1m -------- %%%%%%%%%%%
\begin{figure}[htbp]
\centering
\includegraphics[width=\linewidth]{SFig_gerp_dis1m.pdf}
\caption{GERP score distribution across the genome. Shown are mean GERP scores in a 1-Mb bin region.}
\label{fig:dis1m}
\end{figure}
%%%%%%%% ------------- %%%%%%%%%%%

%%%%%%%% ----- GERP IBD diagram -------- %%%%%%%%%%%
\begin{figure}[htbp]
\includegraphics[width=0.9\textwidth]{SFig_gerpIBD.pdf}
\caption{
\textbf{Incorporation of conservation information into IBD blocks.}
Regions of the genome that are identical by descent (IBD) among the 12 inbreds were identified using Beagle \citep{Browning2009}.  The GERP scores of SNPs in an IBD block were summed up under pure additive and pure dominant assumptions, as well as incomplete dominant assumptions, including under-dominance and over-dominance. For $n$ SNPs in a IBD block, with GERP score $g$, the homozygous non-reference genotype was assigned a value of $2g$, the heterozygote assigned a value of $g$, and the reference homozygote a value of 0.  Under the dominant model, both the heterozygote and the non-reference homozygote were assigned a value of $g$, with the reference homozygote again assigned a value of 0.}
\label{fig:gerpibd}
\end{figure}
%%%%%%%% ------------- %%%%%%%%%%%

%%%%%%%% ----- BEANPLOT-------- %%%%%%%%%%%
\begin{figure}[htbp]
\centering
\includegraphics[width=\linewidth]{SFig_genicsnp_m.pdf}
\caption{Cross-validation accuracies using genic SNPs. Cross-validation experiments were conducted using genic SNPs and compared to circular-shuffled data for traits \emph{per se} (\textbf{A, B}) and pBPH (\textbf{C, D}) under additive (\textbf{A, C}) and dominant (\textbf{B, D}) models. Distirbutions show accuracty of prediction from real data (blue) and permutations (grey), with horizontal bars to indicate mean accuracy.  Stars indicate significantly higher cross-validation accuracy for the real data.  The average accuracy across all traits is shown with the grey dotted line. }
\label{fig:genicsnp}
\end{figure}
\jri{need to explain red points here too.}
%%%%%%%% ------------- %%%%%%%%%%%
\end{document}

### 2015-11-24 NatGen/manuscript.tex
nature Genetics format

--- 2015-11-13 preprint.tex
@@
-\documentclass[12pt,a4paper]{article}
-\usepackage{cite}
+%\documentclass[twocolumn,twoside]{article}
+\documentclass[twoside, twocolumn, letterpaper]{article}
+\usepackage{graphics}
+\usepackage[hidelinks]{hyperref}
 \usepackage{color}
-\usepackage{pgf}
-\usepackage[margin=1in]{geometry}
-%\usepackage{hyperref}
-\usepackage[normalem]{ulem}
-\newcommand{\jri}[1]{\textcolor{red}{\scriptsize #1}}
-\newcommand{\mbh}[1]{\textcolor{blue}{\scriptsize #1}}
-\newcommand{\tmb}[1]{\textcolor{brown}{\scriptsize #1}}
-\newcommand{\citex}{\textcolor{red}{\bf CITE}}
-\newcommand{\X}{\textcolor{red}{\bf X}}
-\usepackage{natbib}
-\usepackage{authblk}
-\usepackage[colorlinks=true,
-            linkcolor=red,
-            urlcolor=blue,
-            citecolor=gray]{hyperref}
+\usepackage{NatGenArtT}
+%\usepackage{NatureLetT}
+\usepackage{times}
+\DeclareMathAlphabet{\msfsl}{OT1}{cmss}{m}{sl}
+\DeclareMathAlphabet{\msfrg}{OT1}{cmss}{n}{sl}
+\renewcommand{\baselinestretch}{1}
+\addtolength{\oddsidemargin}{-.2cm}
+\addtolength{\evensidemargin}{-1.2cm}
+\addtolength{\textwidth}{1.5cm}
+\addtolength{\topmargin}{-1.75cm}
+\addtolength{\textheight}{3.5cm}
+\renewcommand{\textfraction}{0.01}
+\renewcommand{\topfraction}{0.99}
+\renewcommand{\bottomfraction}{0.65}
+\renewcommand{\floatpagefraction}{0.90}
+\renewcommand{\dbltopfraction}{0.95}
+\renewcommand{\dblfloatpagefraction}{0.80}
+\renewcommand{\sfdefault}{phv}
 
-\title{More than one Author with different Affiliations}
-\author[1,2,3]{Timothy M. Beissinger}
-\author[4]{Li Wang}
-\author[1]{Kate Crosby}
-\author[1]{Arun Durvasula}
-\author[4]{Matthew B. Hufford}
-\author[1,5]{Jeffrey Ross-Ibarra}
-\affil[1]{Dept. of Plant Sciences, University of California, Davis, CA, USA}
-\affil[2]{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}
-\affil[3]{Division of Plant Sciences, University of Missouri, Columbia, MO, USA}
-\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA}
-\affil[5]{Genome Center and Center for Population Biology, University of California, Davis, CA, USA}
+\usepackage[sort&compress]{natbib}
+\bibpunct{}{}{,}{s}{}{\textsuperscript{,}}
+\usepackage{amsmath}
+\usepackage{graphicx}
 
-\renewcommand\Authands{ and }
+\newcommand{\beginsupplement}{%
+        \setcounter{table}{0}
+        \renewcommand{\tablename}{Supplementary Table}%
+        \setcounter{figure}{0}
+        \renewcommand{\figurename}{Supplementary Figure}%
+}
+
+\makeatletter
+\renewcommand{\footnotesep}{-2pt}
+\makeatletter
+
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\fancyhf{}
+
+% fancy for Nature Genetics
+\fancyhead[RO]{\begin{picture}(600,1)(10,20)\put(440,32) {\textcolor{Dgreen}{\large{\sfbf{ARTICLES}}}}\end{picture}}
+\fancyhead[RE]{\begin{picture}(600,1)(10,20)\put(10,32)  {\textcolor{Dgreen}{\large{\sfbf{ARTICLES}}}}\end{picture}}
+\fancyhead[C]{\begin{picture}(600,1)(10,20)\linethickness{50pt}\put(-43,52) {{\color{pTop}{\line(1,0){615}}}}\linethickness{0.5pt}\put(10,-678) {{\line(1,0){515}}}\end{picture}}
+
+\fancyfoot[RO]{\thepage}
+\fancyfoot[LE]{\thepage}
+
+\renewcommand{\headrulewidth}{0pt}
+\fancypagestyle{plain}{
+    \fancyhf{}
+}
+\setcounter{footnote}{0}%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{Recent demography drives changes in linked selection across the maize genome}
+
+\author{
+  Timothy M.
+Beissinger\thanks{Dept.
+of Plant Sciences, University of California, Davis, CA, USA} $^,$\thanks{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA} $^,$\thanks{Division of Plant Sciences, University of Missouri, Columbia, MO, USA} \hspace{.5mm},
+  Li Wang\thanks{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA} \hspace{.5mm},
+  Kate Crosby$^{1,}$ \hspace{.5mm},
+  Arun Durvasula$^{1,}$ \hspace{.5mm},
+  Matthew B. Hufford$^{4,}$ \hspace{.5mm},
+  Jeffrey Ross-Ibarra$^{1,}$ \thanks{Genome Center and Center for Population Biology, University of California, Davis, CA, USA}
+  }
+
+\date{\small Manuscript intended for \emph{Nature Genetics}, \today}
 
 \begin{document}
-
-\title{Recent demography drives changes in linked selection across the maize genome}
-
-%\author{\affil{1}{}\affil{2}{}\affil{3}{}, \affil{4}{}\affil{5}{Genome Informatics Facility, Iowa State University, Ames, IA, USA}, \affil{1}{}, \affil{1}{}, \affil{4}{}, \and Jeffrey Ross-Ibarra\affil{1}{}\affil{6}{} }
-
-%\significancetext{Both selection and demographic change play important roles in shaping diversity across the genome, but clear empirical examples of the interplay of these two forces are lacking.
-%Here we document the combined effects of demography and linked selection on genome-wide diversity in domesticated maize and its wild ancestor teosinte.
-We estimate that maize underwent a bottleneck to  $\approx 5\%$ of the size of the ancestral teosinte population, but that recent expansion has resulted in a maize population perhaps orders of magnitude larger than teosinte.
-We show that positive selection on new genic mutations has had relatively little effect on genetic diversity, but that selection against deleterious mutations has dramatically reduced diversity in and immediately around genes in both taxa.
-We find that the relative impact of selection depends on the age of the polymorphisms evaluated: while older polymorphisms in maize show more limited effects of linked selection, new mutations instead reflect its larger current size and more efficient selection.
-Our results demonstrate that a complete understanding of genome-wide patterns of diversity will require careful assessment of both demographic history and the effects of linked selection.
-} %JRI
-
 \maketitle
@@
 \begin{abstract}
+  \noindent \bf
+  \noindent
   The interaction between genetic drift and selection in shaping genetic diversity is not fully understood.
@@
 
-\section*{Introduction}
 The genetic diversity of populations is determined by a constant interplay between genetic drift and natural selection. 
@@
 
-\subsection*{Rapid changes in linked selection} %JRI 
+\section*{Rapid changes in linked selection} %JRI 
 Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
@@
 
-\section*{materials} %JRI
+\section*{METHODS}
+Methods and any associated references are available in a separate pdf file.
 
+\begin{acknowledgments}
+We are indebted to Graham Coop and Simon Aeschbacher for their constructive input during this study.
+We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.
+Funding was provided by NSF Plant Genome Research Project 1238014 and the USDA-Agricultural Research Service.
+\end{acknowledgments}
+
+{\scriptsize \sf
+\renewcommand{\baselinestretch}{2.0}
+\bibliographystyle{NatureSeriesT}
+\bibliography{Reference}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
+\section*{ONLINE METHODS}
 \subsection*{BASH, R, and Python scripts}
-All scripts used for analysis are available in an online repository at \\ \url{https://github.com/timbeissinger/Maize-Teo-Scripts}.
+All scripts used for analysis are available in an online repository at \url{https://github.com/timbeissinger/Maize-Teo-Scripts}.
 
@@
 We also utilized a single individual of (\textit{Tripsacum dactyloides}) as an outgroup.
-All bam files are available at \textcolor{red}{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}.
+All bam files are available at \url{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}.
 
 \subsection*{Physical and genetic maps}
-Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} \\ (\url{ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/}) as described by \cite{bukowski2015}.
+Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{bukowski2015}. 
 All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following \cite{li2011statistical}.
@@
 We assumed individuals were fully inbred and treated each line as a single haplotype.
-Because ANGSD cannot calculate a folded joint SFS, we first polarized SNPs using the maize reference genome and then folded spectra using $\delta\alpha\delta{i}$ \cite{gutenkunst2009}.
+Because ANGSD cannot calculate a folded joint SFS, we first polarized SNPs using the maize reference genome and then folded spectra using $\delta$a$\delta$i \cite{gutenkunst2009}.
 
 \subsection*{Demographic inference}
-We used the software $\delta\alpha\delta{i}$ \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS, using only sites $>5 kb$ from a gene to ameliorate the effects of linked selection.
+We used the software $\delta$a$\delta$i \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS, using only sites $>5 kb$ from a gene to ameliorate the effects of linked selection.
 We modeled a teosinte population of constant effective size $N_a$, that at time $T_b$ generations in the past gave rise to a maize population of size $N_b$ which grew exponentially to size $N_m$ in the present (Figure \ref{fig:bottleneck}).
 The model includes migration of $M_{mt}$ individuals each generation from maize to teosinte and $M_{tm}$ individuals from teosinte to maize.
-We estimated $N_a$ using $\delta\alpha\delta{i}$'s estimation of $\theta=4N_a\mu$ from the data and a mutation rate of $\mu = 3 \times 10^{-8}$ \cite{clark2005}.
-We estimated all other parameters using 1,000 $\delta\alpha\delta{i}$ optimizations and allowing initial values between runs to be randomly perturbed by a factor of 2.  
+We estimated $N_a$ using $\delta$a$\delta$i's estimation of $\theta=4N_a\mu$ from the data and a mutation rate of $\mu = 3 \times 10^{-8}$ \cite{clark2005}.
+We estimated all other parameters using 1,000 $\delta$a$\delta$i optimizations and allowing initial values between runs to be randomly perturbed by a factor of 2.  
 Optimized parameters along with their initial values and upper and lower bounds can be found in table \ref{sTab:dadi}.
@@
 We used six each of maize and teosinte (BKN022, BKN025, BKN029, BKN030, BKN031, BKN033, TIL01, TIL03, TIL09, TIL10, TIL11 and TIL14), treating each inbred genome as a single haplotype.
-We called SNPs in ANGSD \cite{korneliussen2014} using a SNP p-value of $1e-6$ against a reference genome masked using SNPable (\url{http://lh3lh3.users.sourceforge.net/snpable.shtml}). 
+We called SNPs in ANGSD \cite{korneliussen2014} using a SNP p-value of $1e-6$ against a reference genome masked using SNPable (http://lh3lh3.users.sourceforge.net/snpable.shtml). 
 We then removed heterozygous genotypes and filtered sites with a mapping quality $<30$, a base quality $<20$, or a $|\textit{log}_2(\text{depth})|<1$.
@@
 
-\begin{acknowledgments}
-We are indebted to Graham Coop and Simon Aeshbacher for their constructive input during this study.
-We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.
-Funding was provided by NSF Plant Genome Research Project 1238014 and the USDA-Agricultural Research Service.
-\end{acknowledgments}
-
-\bibliography{Reference}
-\bibliographystyle{pnas}
 \onecolumn
@@
 
-\end{article}
-
 \end{document}

### 2015-11-24 NatGen/manuscript.tex
updates

@@
 
-\section*{Results}
+\section*{RESULTS}
 
@@
 \end{center}
-\caption{\textbf{A} and \textbf{B} Show pairwise diversity $\pi$, while \textbf{C} and \textbf{D} depict and Tajimas D in 1kb windows from genic and nongenic regions of maize (A,C) and teosinte (B,D).
-Shown in A and B are means $\pm$ one standard deviation.
+\caption{\textbf{A} and \textbf{B} Show pairwise diversity $\pi$, while \textbf{C} and \textbf{D} depict and Tajimas D in 1kb windows from genic and nongenic regions of maize and teosinte.
+Shown in \textbf{A} and \textbf{B} are means $\pm$ one standard deviation.
 \label{fig:diversity} }
@@
 
-\begin{figure*}[!tb]
+\begin{figure}[!tb]
 %\vspace*{.05in}
@@
 \caption{Parameter estimates for a basic bottleneck model of maize domestication. See methods for details. \label{fig:bottleneck} }
-\end{figure*}
+\end{figure}
 
@@
 
-\section*{Discussion}
+\section*{DISCUSSION}
 
@@
 
-\begin{acknowledgments}
+\section*{ACKNOWLEDGEMENTS}
 We are indebted to Graham Coop and Simon Aeschbacher for their constructive input during this study.
@@
 Funding was provided by NSF Plant Genome Research Project 1238014 and the USDA-Agricultural Research Service.
-\end{acknowledgments}
+
+\section*{AUTHOR CONTRIBUTIONS}
+TMB and JRI devised this study. TB, LW, and KC analyzed the data. AD performed early-stage simulations. MBH provided advice. TB, JRI, and MBH wrote the manuscript.
+
+\section*{COMPETING INTERESTS STATEMENT}
+The authors declare no competing financial interests.
 

### 2015-11-24 NatGen/manuscript.tex
updates

@@
 \subsection*{Diversity is strongly influenced by purifying selection} %JRI
-
 In the case of purifying or background selection, diversity is reduced in functional regions of the genome via removal of deleterious mutations \cite{charlesworth1993}.
@@
 \renewcommand{\baselinestretch}{2.0}
-\bibliographystyle{NatureSeriesT}
+\bibliographystyle{naturemag}
 \bibliography{Reference}

### 2015-11-24 NatGen/manuscript.tex
abstract

@@
   \noindent
-  The interaction between genetic drift and selection in shaping genetic diversity is not fully understood.
-In particular, a population's propensity to drift is typically summarized by its long-term effective population size ($N_e$), but rapidly changing population demographics may complicate this relationship.
-To better understand how changing demography impacts selection, we investigated linked selection in the genomes of 23 domesticated maize and 13 wild maize (teosinte) individuals.
-We show that maize went through a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid expansion post-domestication.
-We observe that hard sweeps on genic mutations are not the primary force driving maize evolution.
-As expected, a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize, but rapid expansion after domestication has since increased the efficiency of purifying selection to levels exceeding those seen in teosinte.
-This final observation demonstrates that rapid demographic change can have wide-ranging impacts on diversity that conflict with would-be expectations based on long-term $N_e$.
+Across species, genetic diversity is shaped by the interaction of drift and selection.
+But the details of this interaction are not well understood.
+Although a population's propensity to drift is typically summarized by its long-term effective population size ($N_e$ ), rapidly changing population demographics complicate this relationship.
+To better understand how changing demography impacts selection, we used whole-genome sequencing data to investigate patterns of linked selection in domesticated and wild maize (teosinte).
+We produce the first unconfounded estimate of $N_e$ at the onset of maize domestication, which shows that maize was reduced to approximately 5\% the population size of teosinte before it experienced rapid expansion post-domestication.
+As expected, a reduced population size during domestication decreased the efficiency selection, but rapid expansion after domestication has since increased the efficiency of selection to levels exceeding those seen in teosinte.
+This final observation is applicable to all species that have undergone recent demographic change, including humans, domesticates, model, and non-model species.
+It demonstrates that rapid demographic change can have immediate and wide-ranging impacts on diversity that conflict with would-be expectations based on $N_e$ alone.
 \end{abstract}

### 2015-11-24 NatGen/manuscript.tex
updated abstract and discussion

@@
 
-\subsection*{Demography of domestication} %JRI
+%\subsection*{Demography of domestication} %JRI
 Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
@@
 
-\subsection*{Hard sweeps do not shape genome-wide diversity in maize} %JRI
+%\subsection*{Hard sweeps do not shape genome-wide diversity in maize} %JRI
 
-Our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of purifying selection  \cite{enard2014}.
+Irrespective of ancient or current population size, our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of purifying selection  \cite{enard2014}.
 Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability.
@@
 
-\subsection*{Demography influences the efficiency of purifying selection} %JRI 
+%\subsection*{Demography influences the efficiency of purifying selection} %JRI 
 One of our more striking findings is that the impact of purifying selection on maize and teosinte qualitatively changed over time. 
@@
 
-\section*{Rapid changes in linked selection} %JRI 
-Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
+%\section*{Rapid changes in linked selection} %JRI
+Most importantly, our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
 While a number of authors have tested for selection using methods that explicitly incorporate or are robust to demographic change \cite{eyre2009estimating, chen2010, zeng2010} and others have compared estimates of the efficiency of adaptive and purifying selection across species \cite{popadin2012} or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size \cite{corbett2015,leffler2012}.

### 2015-11-24 NatGen/manuscript.tex
figure update

@@
 \end{center}
-\caption{\textbf{A} and \textbf{B} Show pairwise diversity $\pi$, while \textbf{C} and \textbf{D} depict and Tajimas D in 1kb windows from genic and nongenic regions of maize and teosinte.
-Shown in \textbf{A} and \textbf{B} are means $\pm$ one standard deviation.
+\caption{\textbf{A} and \textbf{B} show mean pairwise diversity $\pi$, $\pm$ one standard deviation, while \textbf{C} and \textbf{D} depict and Tajimas D in 1kb windows from genic and nongenic regions of maize and teosinte.
 \label{fig:diversity} }

### 2015-11-25 NatGen/manuscript.tex
natgen
abstract

@@
 % fancy for Nature Genetics
-\fancyhead[RO]{\begin{picture}(600,1)(10,20)\put(440,32) {\textcolor{Dgreen}{\large{\sfbf{ARTICLES}}}}\end{picture}}
-\fancyhead[RE]{\begin{picture}(600,1)(10,20)\put(10,32)  {\textcolor{Dgreen}{\large{\sfbf{ARTICLES}}}}\end{picture}}
+\fancyhead[RO]{\begin{picture}(600,1)(10,20)\put(440,32) {\textcolor{Dgreen}{\large{\sfbf{ANALYSIS}}}}\end{picture}}
+\fancyhead[RE]{\begin{picture}(600,1)(10,20)\put(10,32)  {\textcolor{Dgreen}{\large{\sfbf{ANALYSIS}}}}\end{picture}}
 \fancyhead[C]{\begin{picture}(600,1)(10,20)\linethickness{50pt}\put(-43,52) {{\color{pTop}{\line(1,0){615}}}}\linethickness{0.5pt}\put(10,-678) {{\line(1,0){515}}}\end{picture}}
@@
   \noindent
-Across species, genetic diversity is shaped by the interaction of drift and selection.
-But the details of this interaction are not well understood.
+Genetic diversity is shaped by the interaction of drift and selection, but the details of this interaction are not well understood.
 Although a population's propensity to drift is typically summarized by its long-term effective population size ($N_e$ ), rapidly changing population demographics complicate this relationship.
 To better understand how changing demography impacts selection, we used whole-genome sequencing data to investigate patterns of linked selection in domesticated and wild maize (teosinte).
-We produce the first unconfounded estimate of $N_e$ at the onset of maize domestication, which shows that maize was reduced to approximately 5\% the population size of teosinte before it experienced rapid expansion post-domestication.
-As expected, a reduced population size during domestication decreased the efficiency selection, but rapid expansion after domestication has since increased the efficiency of selection to levels exceeding those seen in teosinte.
-This final observation is applicable to all species that have undergone recent demographic change, including humans, domesticates, model, and non-model species.
-It demonstrates that rapid demographic change can have immediate and wide-ranging impacts on diversity that conflict with would-be expectations based on $N_e$ alone.
+We produce the first whole-genome estimate of the demography of maize domestication, showing that maize was reduced to approximately 5\% the population size of teosinte before it experienced rapid expansion post-domestication to population sizes much larger than its ancestor.
+Evaluation of patterns of nucleotide diversity in and near genes shows little evidence of selection on beneficial amino acid substitutions, and that the domestication bottleneck lead to a decline in the efficiency of purifying selection in maize.
+Young alleles, however, show evidence of much stronger purifying selection in maize, reflecting the much larger effective size of present day populations.
+Our results demonstrate that recent demographic change --- a hallmark of many species including both humans and crops ----  can have immediate and wide-ranging impacts on diversity that conflict with would-be expectations based on $N_e$ alone.
 \end{abstract}
@@
 \bibliographystyle{naturemag}
+%\bibliographystyle{NatureSeriesT}
 \bibliography{Reference}

### 2015-11-25 NatGen/manuscript.tex
updates

@@
 Young alleles, however, show evidence of much stronger purifying selection in maize, reflecting the much larger effective size of present day populations.
-Our results demonstrate that recent demographic change --- a hallmark of many species including both humans and crops ----  can have immediate and wide-ranging impacts on diversity that conflict with would-be expectations based on $N_e$ alone.
+Our results demonstrate that recent demographic change --- a hallmark of many species including both humans and crops ---  can have immediate and wide-ranging impacts on diversity that conflict with would-be expectations based on $N_e$ alone.
 \end{abstract}
@@
 \section*{AUTHOR CONTRIBUTIONS}
-TMB and JRI devised this study. TB, LW, and KC analyzed the data. AD performed early-stage simulations. MBH provided advice. TB, JRI, and MBH wrote the manuscript.
+TMB and JRI devised this study. TMB, LW, and KC analyzed the data. AD performed early-stage simulations. MBH provided advice. TMB, JRI, and MBH wrote the manuscript.
 

### 2015-11-25 NatGen/manuscript.tex
updates

@@
 Beissinger\thanks{Dept.
-of Plant Sciences, University of California, Davis, CA, USA} $^,$\thanks{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA} $^,$\thanks{Division of Plant Sciences, University of Missouri, Columbia, MO, USA} \hspace{.5mm},
+of Plant Sciences, University of California, Davis, CA, USA} $^,$\thanks{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA} $^,$\thanks{Division of Plant Sciences, University of Missouri, Columbia, MO, USA} $^{,6}$\hspace{.5mm},
   Li Wang\thanks{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA} \hspace{.5mm},
@@
   Matthew B. Hufford$^{4,}$ \hspace{.5mm},
-  Jeffrey Ross-Ibarra$^{1,}$ \thanks{Genome Center and Center for Population Biology, University of California, Davis, CA, USA}
+  Jeffrey Ross-Ibarra$^{1,}$\thanks{Genome Center and Center for Population Biology, University of California, Davis, CA, USA} $^{,}$\thanks{Correspondence should be addressed to TMB (beissingert@missouri.edu) or JRI (rossibarra@ucdavis.edu).}
   }
@@
 
-\begin{figure*}[tb]
+\begin{figure*}[!tb]
 %\vspace*{.05in}
@@
 Future comparative analyses using a common statistical framework (e.g.
-\cite{corbett2015}) and considering additional ecological and life history factors (c.f.
+\cite{corbett2015} )and considering additional ecological and life history factors (c.f.
 \cite{leffler2012}) should allow explicit testing of this idea.

### 2015-12-06 PlosGen/manuscript (copy).tex
plos

% Template for PLoS
% Version 3.1 February 2015
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
% the final text of your manuscript.
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig." instead of "Figure".
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout or alignment
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment. For example, CO2 will be CO\textsubscript{2}.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% fixltx2e package for \textsubscript
\usepackage{fixltx2e}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
\usepackage{rotating}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.35in}

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
\begin{flushleft}
{\Large
\textbf\newline{Recent demography drives changes in linked selection across the maize genome}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Timothy M. Beissinger\textsuperscript{1,2,3,*},
Li Wang\textsuperscript{4},
Kate Crosby\textsuperscript{1},
Arun Durvasula\textsuperscript{1},
Matthew B. Hufford\textsuperscript{4},
Jeffrey Ross-Ibarra\textsuperscript{1,5,*},
\\
\bigskip
\bf{1} Department of Plant Sciences, University of California, Davis, Caliornia, USA\\
\bf{2} US Department of Agriculture, Agriculture Research Service, Columbia, Missouri, USA\\
\bf{3} Division of Plant Sciences, University of Missouri, Columbia, Missouri, USA\\
\bf{4} Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA\\
\bf{5} Genome Center and Center for Population Biology, University of California, Davis, California, USA
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
%\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
%\ddag These authors also contributed equally to this work.

% Current address notes
%\textcurrency a Insert current address of first author with an address update
% \textcurrency b Insert current address of second author with an address update
% \textcurrency c Insert current address of third author with an address update

% Deceased author note
%\dag Deceased

% Group/Consortium Author Note
%\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* rossibarra@ucdavis.edu or beissingert@missouri.edu

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author Summary}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\linenumbers

\section*{Introduction}
Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam Eq.~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\begin{equation}\label{eq:schemeP} 
D_{coll} = \frac{D_f+\frac{[S]^2}{K_D S_T} D_S} {1+\frac{[S]^2}{K_D S_T}}, 
D_{sm} = \frac{D_f+ \frac{[S]}{K_D} D_S}{1+\frac{[S]}{K_D}},
\end{equation}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
\subsection*{Etiam eget sapien nibh.}

% For figure citations, please use "Fig." instead of "Figure".
Nulla mi mi, Fig.~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\begin{figure}[h]
\caption{{\bf Figure Title first bold sentence Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}
Figure Caption Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
\label{fig1}
\end{figure}

\begin{enumerate}
\item{react}
\item{diffuse free particles}
\item{increment time by dt and go to 1}
\end{enumerate}

% Results and Discussion can be combined.
\section*{Results}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\caption{
{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
\begin{tabular}{|l|l|l|l|l|l|l|l|}
\hline
\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \hline
$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
\end{tabular}
\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
\end{flushleft}
\label{table1}
\end{adjustwidth}
\end{table}

\subsection*{\lorem\ and \ipsum\ Nunc blandit a tortor.}

Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit. Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat.

\subsection*{Sed ac quam id nisi malesuada congue.}

Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}. 
\subsection*{Subsection 1}
Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\subsection*{Subsection 2}
\paragraph{3rd Level Heading.} Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\section*{Discussion}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\subsection*{\lorem\ and \ipsum\ Nunc blandit a tortor.}

CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.

Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat. Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna. Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed. Nam condimentum sem eget mollis euismod. Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex. Nunc est dui, sodales sed mauris nec, auctor sagittis leo. Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor. For more information, see \nameref{S1_Text}.

\section*{Supporting Information}

% Include only the SI item label in the subsection heading. Use the \nameref{label} command to cite SI items in the text.
\subsection*{S1 Video}
\label{S1_Video}
{\bf Bold the first sentence.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\subsection*{S1 Text}
\label{S1_Text}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\subsection*{S1 Fig}
\label{S1_Fig}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\subsection*{S2 Fig}
\label{S2_Fig}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\subsection*{S1 Table}
\label{S1_Table}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\section*{Acknowledgments}
Cras egestas velit mauris, eu mollis turpis pellentesque sit amet. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada congue. Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.

\nolinenumbers

%\section*{References}
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% OR
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here.
% 
\begin{thebibliography}{10}
\bibitem{bib1}
Devaraju P, Gulati R, Antony PT, Mithun CB, Negi VS. Susceptibility to SLE in South Indian Tamils may be influenced by genetic selection pressure on TLR2 and TLR9 genes. Mol Immunol. 2014 Nov 22. pii: S0161-5890(14)00313-7. doi: 10.1016/j.molimm.2014.11.005

\bibitem{bib2}
Huynen MMTE, Martens P, Hilderlink HBM. The health impacts of globalisation: a conceptual framework. Global Health. 2005;1: 14. Available: http://www.globalizationandhealth.com/content/1/1/14.

\end{thebibliography}

\end{document}

### 2015-12-06 PlosGen/manuscript.tex
plos

--- 2015-11-25 NatGen/manuscript.tex
@@
-%\documentclass[twocolumn,twoside]{article}
-\documentclass[twoside, twocolumn, letterpaper]{article}
-\usepackage{graphics}
-\usepackage[hidelinks]{hyperref}
-\usepackage{color}
-\usepackage{NatGenArtT}
-%\usepackage{NatureLetT}
-\usepackage{times}
-\DeclareMathAlphabet{\msfsl}{OT1}{cmss}{m}{sl}
-\DeclareMathAlphabet{\msfrg}{OT1}{cmss}{n}{sl}
-\renewcommand{\baselinestretch}{1}
-\addtolength{\oddsidemargin}{-.2cm}
-\addtolength{\evensidemargin}{-1.2cm}
-\addtolength{\textwidth}{1.5cm}
-\addtolength{\topmargin}{-1.75cm}
-\addtolength{\textheight}{3.5cm}
-\renewcommand{\textfraction}{0.01}
-\renewcommand{\topfraction}{0.99}
-\renewcommand{\bottomfraction}{0.65}
-\renewcommand{\floatpagefraction}{0.90}
-\renewcommand{\dbltopfraction}{0.95}
-\renewcommand{\dblfloatpagefraction}{0.80}
-\renewcommand{\sfdefault}{phv}
+% Template for PLoS
+% Version 3.1 February 2015
+%
+% To compile to pdf, run:
+% latex plos.template
+% bibtex plos.template
+% latex plos.template
+% latex plos.template
+% dvipdf plos.template
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
+% the final text of your manuscript.
+%
+% There are no restrictions on package use within the LaTeX files except that 
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig." instead of "Figure".
+% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - tabs/spacing/line breaks within cells to alter layout or alignment
+% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
+% - colors, shading, or graphic objects
+% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
+%
+% Please be sure to include all portions of an equation in the math environment.
+%
+% Do not include text that is not math in the math environment. For example, CO2 will be CO\textsubscript{2}.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
 
-\usepackage[sort&compress]{natbib}
-\bibpunct{}{}{,}{s}{}{\textsuperscript{,}}
-\usepackage{amsmath}
-\usepackage{graphicx}
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
 
-\newcommand{\beginsupplement}{%
-        \setcounter{table}{0}
-        \renewcommand{\tablename}{Supplementary Table}%
-        \setcounter{figure}{0}
-        \renewcommand{\figurename}{Supplementary Figure}%
-}
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
 
+% Use Unicode characters when possible
+\usepackage[utf8]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% fixltx2e package for \textsubscript
+\usepackage{fixltx2e}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% rotating package for sideways tables
+\usepackage{rotating}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
 \makeatletter
-\renewcommand{\footnotesep}{-2pt}
-\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
 
-\usepackage{fancyhdr}
+% Leave date blank
+\date{}
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+\pagestyle{myheadings}
 \pagestyle{fancy}
 \fancyhf{}
+\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\sf PLOS}
 
-% fancy for Nature Genetics
-\fancyhead[RO]{\begin{picture}(600,1)(10,20)\put(440,32) {\textcolor{Dgreen}{\large{\sfbf{ANALYSIS}}}}\end{picture}}
-\fancyhead[RE]{\begin{picture}(600,1)(10,20)\put(10,32)  {\textcolor{Dgreen}{\large{\sfbf{ANALYSIS}}}}\end{picture}}
-\fancyhead[C]{\begin{picture}(600,1)(10,20)\linethickness{50pt}\put(-43,52) {{\color{pTop}{\line(1,0){615}}}}\linethickness{0.5pt}\put(10,-678) {{\line(1,0){515}}}\end{picture}}
+%% Include all macros below
 
-\fancyfoot[RO]{\thepage}
-\fancyfoot[LE]{\thepage}
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
 
-\renewcommand{\headrulewidth}{0pt}
-\fancypagestyle{plain}{
-    \fancyhf{}
-}
-\setcounter{footnote}{0}%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Recent demography drives changes in linked selection across the maize genome}
-
-\author{
-  Timothy M.
-Beissinger\thanks{Dept.
-of Plant Sciences, University of California, Davis, CA, USA} $^,$\thanks{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA} $^,$\thanks{Division of Plant Sciences, University of Missouri, Columbia, MO, USA} $^{,6}$\hspace{.5mm},
-  Li Wang\thanks{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA} \hspace{.5mm},
-  Kate Crosby$^{1,}$ \hspace{.5mm},
-  Arun Durvasula$^{1,}$ \hspace{.5mm},
-  Matthew B. Hufford$^{4,}$ \hspace{.5mm},
-  Jeffrey Ross-Ibarra$^{1,}$\thanks{Genome Center and Center for Population Biology, University of California, Davis, CA, USA} $^{,}$\thanks{Correspondence should be addressed to TMB (beissingert@missouri.edu) or JRI (rossibarra@ucdavis.edu).}
-  }
-
-\date{\small Manuscript intended for \emph{Nature Genetics}, \today}
+%% END MACROS SECTION
 
 \begin{document}
-\maketitle
+\vspace*{0.35in}
 
-\begin{abstract}
-  \noindent \bf
-  \noindent
+% Title must be 250 characters or less.
+% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
+\begin{flushleft}
+{\Large
+\textbf\newline{Recent demography drives changes in linked selection across the maize genome}
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Timothy M. Beissinger\textsuperscript{1,2,3,*},
+Li Wang\textsuperscript{4},
+Kate Crosby\textsuperscript{1},
+Arun Durvasula\textsuperscript{1},
+Matthew B. Hufford\textsuperscript{4},
+Jeffrey Ross-Ibarra\textsuperscript{1,5,*},
+\\
+\bigskip
+\bf{1} Department of Plant Sciences, University of California, Davis, Caliornia, USA\\
+\bf{2} US Department of Agriculture, Agriculture Research Service, Columbia, Missouri, USA\\
+\bf{3} Division of Plant Sciences, University of Missouri, Columbia, Missouri, USA\\
+\bf{4} Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA\\
+\bf{5} Genome Center and Center for Population Biology, University of California, Davis, California, USA
+\\
+\bigskip
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+%\Yinyang These authors contributed equally to this work.
+
+% Additional Equal Contribution Note
+% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+%\ddag These authors also contributed equally to this work.
+
+% Current address notes
+%\textcurrency a Insert current address of first author with an address update
+% \textcurrency b Insert current address of second author with an address update
+% \textcurrency c Insert current address of third author with an address update
+
+% Deceased author note
+%\dag Deceased
+
+% Group/Consortium Author Note
+%\textpilcrow Membership list can be found in the Acknowledgments section.
+
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* rossibarra@ucdavis.edu or beissingert@missouri.edu
+
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
 Genetic diversity is shaped by the interaction of drift and selection, but the details of this interaction are not well understood.
@@
 To better understand how changing demography impacts selection, we used whole-genome sequencing data to investigate patterns of linked selection in domesticated and wild maize (teosinte).
-We produce the first whole-genome estimate of the demography of maize domestication, showing that maize was reduced to approximately 5\% the population size of teosinte before it experienced rapid expansion post-domestication to population sizes much larger than its ancestor.
-Evaluation of patterns of nucleotide diversity in and near genes shows little evidence of selection on beneficial amino acid substitutions, and that the domestication bottleneck lead to a decline in the efficiency of purifying selection in maize.
-Young alleles, however, show evidence of much stronger purifying selection in maize, reflecting the much larger effective size of present day populations.
-Our results demonstrate that recent demographic change --- a hallmark of many species including both humans and crops ---  can have immediate and wide-ranging impacts on diversity that conflict with would-be expectations based on $N_e$ alone.
-\end{abstract}
+We show that maize went through a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid expansion post-domestication.
+We observe that hard sweeps on genic mutations are not the primary force driving maize evolution.
+As expected, a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize, but rapid expansion after domestication has since increased the efficiency of purifying selection to levels exceeding those seen in teosinte.
+This final observation demonstrates that rapid demographic change can have wide-ranging impacts on diversity that conflict with would-be expectations based on long-term $N_e$.
+
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
+\section*{Author Summary}
+Both selection and demographic change play important roles in shaping genetic diversity, but there are few empirical examples of the interplay between them.
+Although it is well known that both diversity and the efficacy of natural selection are reduced in smaller populations, the short-term consequences of demographic change on linked selection have not been well studied.
+We use wild and domesticated maize to demonstrate that rapid changes in population size impact the effect of selection on linked sites in ways not well described by long-term average population size.
+We also produce an unbiased estimate of maize demographic history and show that positive selection on new genic mutations plays a lesser role in shaping diversity than purifying selection against deleterious mutations.
+
+\linenumbers
+
+\section*{Introduction}
 
@@
 
+% Results and Discussion can be combined.
+
 \section*{RESULTS}
@@
 
+\begin{figure}[h]
+%  \begin{center}
+%  \includegraphics[width=.45\textwidth] {FigsAndFiles/Pi_and_Tajima.png}
+%\end{center}
+\caption{{\bf Diversity and Tajima's D in maize and teosinte.} \textbf{A} and \textbf{B} show mean pairwise diversity $\pi$, $\pm$ one standard deviation, while \textbf{C} and \textbf{D} depict and Tajimas D in 1kb windows from genic and nongenic regions of maize and teosinte.
+\label{fig:diversity} }
+\end{figure}
+
 \subsection*{Demography of maize domestication} %JRI
@@
 Maize and teosinte have continued to exchange migrants after the population split, with gene flow between the populations estimated at $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation from teosinte to maize and $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants from maize to teosinte.
+
+\begin{figure}[h]
+%\vspace*{.05in}
+%\centering
+%\includegraphics[width=.5\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
+\caption{{\bf Parameter estimates for a basic bottleneck model of maize domestication.} See methods for details. \label{fig:bottleneck} }
+\end{figure}
 
@@
 
-\begin{figure}[!tb]
-  \begin{center}
-  \includegraphics[width=.45\textwidth] {FigsAndFiles/Pi_and_Tajima.png}
-\end{center}
-\caption{\textbf{A} and \textbf{B} show mean pairwise diversity $\pi$, $\pm$ one standard deviation, while \textbf{C} and \textbf{D} depict and Tajimas D in 1kb windows from genic and nongenic regions of maize and teosinte.
-\label{fig:diversity} }
-\end{figure}
-
-\begin{figure}[!tb]
-%\vspace*{.05in}
-\centering
-\includegraphics[width=.5\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
-\caption{Parameter estimates for a basic bottleneck model of maize domestication. See methods for details. \label{fig:bottleneck} }
-\end{figure}
-
 \subsection*{Hard sweeps do not explain diversity differences} %JRI
@@
 
-%\mbh{I'm curious whether linked selection is affecting the diversity around a sizable proportion of synonymous substitutions.
-Would it be worthwhile to look at diversity around the subset of synonymous substitutions that are the least linked to missense substitutions just to make sure it is not substantially higher?
-Or perhaps generate a plot of diversity around synonymous substitutions vs.
-distance of these subs from missense subs?}
-
-\begin{figure*}[tb]
+\begin{figure*}[h]
 %\vspace*{.05in}
-\centering
-\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_Aug}
-\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_Aug}
-\caption{Pairwise diversity surrounding synonymous and missense  substitutions in {\bf A} maize and {\bf B} teosinte.
-Axes show absolute diversity values (right) and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution (left).
+%\centering
+%\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_Aug}
+%\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_Aug}
+\caption{ {\bf Pairwise diversity surrounding synonymous and missense  substitutions in {\bf A:} maize and {\bf B:} teosinte.} Axes show absolute diversity values (right) and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution (left).
 Lines depict a loess curve (span of 0.01) and shading represents bootstrap-based 95\% confidence intervals.
@@
 
+\begin{figure*}[h]
+%\vspace*{.05in}
+%\centering
+%\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_Downsampled_threeLines_manuscript.png}
+\caption{ {\bf Relative diversity versus distance to nearest gene in maize and teosinte.} 
+Shown are \textbf{A} pairwise nucleotide diversity and \textbf{B} singleton diversity.  
+Relative diversity is calculated compared to the mean diversity in windows $\geq 0.01 cM$ or $\geq 0.02 cM$ from the nearest
+gene for pairwise diversity and singletons, respectively. 
+  Lines depict cubic smoothing splines with smoothing parameters chosen via generalized cross validation and shading depicts bootstrap-based 95\% confidence intervals.
+  Inset plots depict a smaller range on the x-axis. \label{fig:purify}
+  }
+\end{figure*}
+
 Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple beneficial mutations or standing genetic variation, because in such cases diversity around the substitution may be reduced to a lesser degree \cite{innan2004,messer2013}.
@@
 We interpret these combined results as suggesting that purifying selection has predominantly shaped diversity near genes and left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in long-term effective population size.
-
-\begin{figure*}[!tb]
-%\vspace*{.05in}
-\centering
-\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_Downsampled_threeLines_manuscript.png}
-\caption{Relative diversity versus distance to nearest gene in maize and teosinte. 
-Shown are \textbf{A} pairwise nucleotide diversity and \textbf{B} singleton diversity.  
-Relative diversity is calculated compared to the mean diversity in windows $\geq 0.01 cM$ or $\geq 0.02 cM$ from the nearest
-gene for pairwise diversity and singletons, respectively. 
-  Lines depict cubic smoothing splines with smoothing parameters chosen via generalized cross validation and shading depicts bootstrap-based 95\% confidence intervals.
-  Inset plots depict a smaller range on the x-axis. \label{fig:purify}
-  }
-\end{figure*}
 
@@
 
-\section*{DISCUSSION}
+\section*{Discussion}
 
-%\subsection*{Demography of domestication} %JRI
+\subsection*{Demography of domestication} %JRI
 Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
@@
 
-%\subsection*{Hard sweeps do not shape genome-wide diversity in maize} %JRI
+\subsection*{Hard sweeps do not shape genome-wide diversity in maize} %JRI
 
@@
 
-%\subsection*{Demography influences the efficiency of purifying selection} %JRI 
+\subsection*{Demography influences the efficiency of purifying selection} %JRI 
 One of our more striking findings is that the impact of purifying selection on maize and teosinte qualitatively changed over time. 
@@
 
-%\section*{Rapid changes in linked selection} %JRI
-Most importantly, our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
+\section*{Rapid changes in linked selection} %JRI
+Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
 While a number of authors have tested for selection using methods that explicitly incorporate or are robust to demographic change \cite{eyre2009estimating, chen2010, zeng2010} and others have compared estimates of the efficiency of adaptive and purifying selection across species \cite{popadin2012} or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size \cite{corbett2015,leffler2012}.
@@
 
-\section*{METHODS}
-Methods and any associated references are available in a separate pdf file.
+% You may title this section "Methods" or "Models". 
+% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
+% authors may use "Analysis" 
+\section*{Materials and Methods}
 
-\section*{ACKNOWLEDGEMENTS}
-We are indebted to Graham Coop and Simon Aeschbacher for their constructive input during this study.
-We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.
-Funding was provided by NSF Plant Genome Research Project 1238014 and the USDA-Agricultural Research Service.
-
-\section*{AUTHOR CONTRIBUTIONS}
-TMB and JRI devised this study. TMB, LW, and KC analyzed the data. AD performed early-stage simulations. MBH provided advice. TMB, JRI, and MBH wrote the manuscript.
-
-\section*{COMPETING INTERESTS STATEMENT}
-The authors declare no competing financial interests.
-
-{\scriptsize \sf
-\renewcommand{\baselinestretch}{2.0}
-\bibliographystyle{naturemag}
-%\bibliographystyle{NatureSeriesT}
-\bibliography{Reference}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-\section*{ONLINE METHODS}
 \subsection*{BASH, R, and Python scripts}
@@
 
-\onecolumn
+\section*{Supporting Information}
+
+% Include only the SI item label in the subsection heading. Use the \nameref{label} command to cite SI items in the text.
 \input{supportingInfo}
 
+\section*{Acknowledgments}
+We are indebted to Graham Coop and Simon Aeschbacher for their constructive input during this study.
+We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.
+Funding was provided by NSF Plant Genome Research Project 1238014 and the USDA-Agricultural Research Service.
+
+\nolinenumbers
+
+%\section*{References}
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% OR
+%
+% Compile your BiBTeX database using our plos2015.bst
+% style file and paste the contents of your .bbl file
+% here.
+%
+
+\bibliography{Reference}
+
 \end{document}

### 2015-12-06 PlosGen/supportingInfo.tex
plos

\renewcommand\thefigure{S\arabic{figure}}
\renewcommand\thetable{S\arabic{table}}    
\setcounter{figure}{0}
\setcounter{table}{0}

\begin{figure}[h!]
%  \begin{center}
%    \includegraphics[width=\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_tga1Supp_June.png}
%    \end{center}
\caption{ {\bf Diversity surrounding the causitive substitution at the \emph{tga1} locus.}  \label{sFig:tga1}}
\end{figure}

\begin{figure}[h!]
%  \begin{center}
%    \includegraphics[width=.85\textwidth]{FigsAndFiles/TIL_BKN_bootstrapping_msmc.pdf}
%    \end{center}
\caption{{ \bf Effective population size estimated over time using MSMC.} Shown are estimates (solid lines) and boostrap resampling (dotted lines) for both maize (red) and teosinte (blue). Time is estimated assuming an annual generation time and a mutation rate of $\mu=3\times 10^{-8}$ \label{sFig:msmc}}
\end{figure}

\begin{figure}[h!]
%  \includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Conserved_Significance_June.png}
%  \includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Unconserved_Significance_June.png}
\caption{{ \bf Pairwise diversity surrounding synonymous and nonsynonymous substitutions in maize,} shown at \textbf{A:} highly conserved or \textbf{B:}  unconserved sites.  Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis. \label{sFig:consUncons}}
\end{figure}

\begin{figure}[h!]
%  \includegraphics[width=\textwidth]{FigsAndFiles/plotDiversity_TvM_Singletons.png}
\caption{{ \bf Singleton diversity surrounding synonymous and nonsynonymous substitutions in maize. \label{sFig:singleton}}}
\end{figure}

\begin{figure}[h!]
%  \includegraphics[width=\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_maizeAndTeoSingleVsPi.png}
\caption{ { \bf Relative diversity versus distance to nearest gene in maize and teosinte.} Relative diversity is calculated by comparing to the mean diversity in all windows $\geq 0.02 cM$ from the nearest gene. Lines depict cubic smoothing splines with smoothing parameters chosen via generalized cross validation and shading depicts bootstrap-based 95\% confidence intervals.  \label{sFig:singletonPi}}
\end{figure}

\begin{figure}[h!]
%  \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_Unselected_manuscript.png}
%    \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_unselected_Singletons_manuscript.png}
\caption{ {\bf Relative level of diversity versus distance to the nearest gene, in maize and teosinte, based on only sites that do not show evidence of hard or soft sweeps according to H12.} Two measures of diversity were investigated. {\bf A} displays pairwise diversity,
which is most influenced by intermediate frequency alleles and therefore depicts more ancient evolutionary patterns, and {\bf B} depicts singleton diversity, influenced by rare alleles and thus depicting evolutionary patterns in the recent past. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a smaller range on the x-axis. \label{sFig:H12}}
\end{figure}

\begin{figure}[h!]
%  \begin{center}
%  \includegraphics[width=.75\textwidth]{FigsAndFiles/tilPCA_aug.pdf}\\
%  \includegraphics[width=.75\textwidth]{FigsAndFiles/bknPCA_aug.pdf}\\
%  \end{center}
  \caption{ {\bf Principal component analysis of teosinte and maize individuals to ensure that no close relatives were inadvertantly included in our study.} Plots are based on a random sample of 10,000 SNPs. {\bf A} displays the percentage of total variance explained by each principal component for teosinte, while {\bf B} shows PC1 vs PC2 for all 13 teosinte individuals. Simlarly, {\bf C} depicts the percentage of total variance explained by each principal component for maize, and {\bf D} shows PC1 vs PC2 for all 23 maize individuals. \label{sFig:PCA}}
\end{figure}

\begin{table}[h!]  
%  \begin{center}
%  \begin{tabular}{c|c}
%    \bf Maize & \bf Teosinte \\ \hline \hline
%    BKN009 &  TIL01 \\
%    BKN010 & TIL02 \\
%    BKN011 & TIL03 \\
%    BKN014 & TIL04-TIP454 \\
%    BKN015 & TIL07 \\
%    BKN016 & TIL09 \\
%    BKN017 & TIL10 \\
%    BKN018 & TIL11 \\
%    BKN019 & TIL12 \\
%    BKN020 & TIL14-TIP498 \\
%    BKN022 & TIL15 \\
%    BKN023 & TIL16 \\
%    BKN025 & TIL17 \\
%    BKN026 & \\
%    BKN027 & \\
%    BKN029 & \\
%    BKN030 & \\
%    BKN031 & \\
%    BKN032 & \\
%    BKN033 & \\
%    BKN034 & \\
%    BKN035 & \\
%    BKN040 & \\ \hline
%  \end{tabular}
%  \end{center}
  \caption{ {\bf A list of maize and teosinte individuals included in this study.} Sequencing and details were previously described by \cite{chia2012}   } \label{sTab:list}
\end{table}

\begin{table}[h!]
%  \begin{center}
%  \def\arraystretch{2}
%  \begin{center}
%  \begin{tabular}{l|c|c|c}
%    \bf Parameter & \bf Initial value & \bf Upper bound & \bf Lower bound\\ \hline 
%    $\frac{N_b}{N_a}$ & 0.02 & $1\times10^{-7}$ & 2 \\ 
%    $\frac{N_{m}}{N_a}$ & 3 & $1\times10^{-7}$ & 200 \\
%    $\frac{T_b}{2N_a}$ & 0.04 & 0  & 1 \\ 
%    $\frac{M_{mt}}{N_a}$ & $1\times10^{-10}$ & $1\times10^{-7}$ & 0.001 \\
%    $\frac{M_{tm}}{N_a}$ & $1\times10^{-10}$ & $1\times10^{-7}$ & 0.001 \\
%  \end{tabular}
%  \def\arraystretch{1} % undo stretching
%  \end{center}
    \caption{{\bf Parameters, initial values, and boundaries used for model-fitting with $\delta\alpha\delta{i}$.} Parameters are shown in the units utilized by $\delta\alpha\delta{i}$, although in the text simplified units are reported.    }\label{sTab:dadi}
\end{table}

### 2016-01-04 cover.tex
cover start

\documentclass[]{article}

\begin{document}

Dear Editor, 
Please find enclosed a manuscript entitled, ``Recent demography drives changes in linked selection across the maize genome''. 
Many organisms, from humans to domesticated crops to models such as \texit{Arabidopsis} and \texti{Drosophila}, have undergone significant demographic change in their recent past.  
Such recent changes have often been ignored because their impact on total levels of genetic diversity and long-term effective population size ($N_e$) is usually minimal.

%still working
But a single number such as N e fails to encapsulate all of a population’s demographic history. 
We used resequencing data from domesticated maize and its wild ancestor teosinte to investigate the impacts of dynamic population size changes on patterns of linked selection. By comparing diversity in allelic classes of different ages, we show that the impact of purifying selection on linked diversity in maize and teosinte dramatically shifts as a result of the rapid expansion of maize post­domestication. This observation has broad implications across the tree of life: Our work shows that a detailed understanding of the role of linked selection and the fate of new mutations depends critically on these recent demographic shifts and is not captured well using only estimates of long­term N e. 
We hope that our observations regarding the impact of population size changes on patterns of selection and genome­wide diversity appeal to the broad readership of PLoS Genetics, including researchers who work on any species across the genetics and genomics spectrum. 
Thank you, Tim Beissinger & Jeffrey Ross­-Ibarra

\end{document}

### 2016-02-18 NatPlant2/manuscript.tex
nat_plant reviews

--- 2015-11-25 NatGen/manuscript.tex
@@
-%\documentclass[twocolumn,twoside]{article}
-\documentclass[twoside, twocolumn, letterpaper]{article}
-\usepackage{graphics}
-\usepackage[hidelinks]{hyperref}
+\documentclass[12pt,a4paper]{article}
+\usepackage{cite}
 \usepackage{color}
-\usepackage{NatGenArtT}
-%\usepackage{NatureLetT}
-\usepackage{times}
-\DeclareMathAlphabet{\msfsl}{OT1}{cmss}{m}{sl}
-\DeclareMathAlphabet{\msfrg}{OT1}{cmss}{n}{sl}
-\renewcommand{\baselinestretch}{1}
-\addtolength{\oddsidemargin}{-.2cm}
-\addtolength{\evensidemargin}{-1.2cm}
-\addtolength{\textwidth}{1.5cm}
-\addtolength{\topmargin}{-1.75cm}
-\addtolength{\textheight}{3.5cm}
-\renewcommand{\textfraction}{0.01}
-\renewcommand{\topfraction}{0.99}
-\renewcommand{\bottomfraction}{0.65}
-\renewcommand{\floatpagefraction}{0.90}
-\renewcommand{\dbltopfraction}{0.95}
-\renewcommand{\dblfloatpagefraction}{0.80}
-\renewcommand{\sfdefault}{phv}
-
-\usepackage[sort&compress]{natbib}
+\usepackage{pgf}
+\usepackage[margin=1in]{geometry}
+%\usepackage{hyperref}
+\usepackage[normalem]{ulem}
+\newcommand{\jri}[1]{\textcolor{red}{\scriptsize #1}}
+\newcommand{\mbh}[1]{\textcolor{blue}{\scriptsize #1}}
+\newcommand{\tmb}[1]{\textcolor{brown}{\scriptsize #1}}
+\newcommand{\citex}{\textcolor{red}{\bf CITE}}
+\newcommand{\X}{\textcolor{red}{\bf X}}
+\usepackage{natbib}
+\usepackage{authblk}
+\usepackage[colorlinks=true,
+            linkcolor=red,
+            urlcolor=blue,
+            citecolor=gray]{hyperref}
+\usepackage{natbib}
 \bibpunct{}{}{,}{s}{}{\textsuperscript{,}}
@@
 
-\newcommand{\beginsupplement}{%
-        \setcounter{table}{0}
-        \renewcommand{\tablename}{Supplementary Table}%
-        \setcounter{figure}{0}
-        \renewcommand{\figurename}{Supplementary Figure}%
-}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{More than one Author with different Affiliations}
+\author[*,1,2,3]{Timothy M. Beissinger}
+\author[4]{Li Wang}
+\author[1]{Kate Crosby}
+\author[1]{Arun Durvasula}
+\author[4]{Matthew B. Hufford}
+\author[*,1,5]{Jeffrey Ross-Ibarra}
+\affil[1]{Dept. of Plant Sciences, University of California, Davis, CA, USA}
+\affil[2]{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}
+\affil[3]{Division of Plant Sciences, University of Missouri, Columbia, MO, USA}
+\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA}
+\affil[5]{Genome Center and Center for Population Biology, University of California, Davis, CA, USA}
+\affil[*]{Correspondence may be addressed to rossibarra@ucdavis.edu or beissingert@missouri.edu}
 
-\makeatletter
-\renewcommand{\footnotesep}{-2pt}
-\makeatletter
+\renewcommand\Authands{ and }
 
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\fancyhf{}
+\begin{document}
 
-% fancy for Nature Genetics
-\fancyhead[RO]{\begin{picture}(600,1)(10,20)\put(440,32) {\textcolor{Dgreen}{\large{\sfbf{ANALYSIS}}}}\end{picture}}
-\fancyhead[RE]{\begin{picture}(600,1)(10,20)\put(10,32)  {\textcolor{Dgreen}{\large{\sfbf{ANALYSIS}}}}\end{picture}}
-\fancyhead[C]{\begin{picture}(600,1)(10,20)\linethickness{50pt}\put(-43,52) {{\color{pTop}{\line(1,0){615}}}}\linethickness{0.5pt}\put(10,-678) {{\line(1,0){515}}}\end{picture}}
-
-\fancyfoot[RO]{\thepage}
-\fancyfoot[LE]{\thepage}
-
-\renewcommand{\headrulewidth}{0pt}
-\fancypagestyle{plain}{
-    \fancyhf{}
-}
-\setcounter{footnote}{0}%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \title{Recent demography drives changes in linked selection across the maize genome}
 
-\author{
-  Timothy M.
-Beissinger\thanks{Dept.
-of Plant Sciences, University of California, Davis, CA, USA} $^,$\thanks{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA} $^,$\thanks{Division of Plant Sciences, University of Missouri, Columbia, MO, USA} $^{,6}$\hspace{.5mm},
-  Li Wang\thanks{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA} \hspace{.5mm},
-  Kate Crosby$^{1,}$ \hspace{.5mm},
-  Arun Durvasula$^{1,}$ \hspace{.5mm},
-  Matthew B. Hufford$^{4,}$ \hspace{.5mm},
-  Jeffrey Ross-Ibarra$^{1,}$\thanks{Genome Center and Center for Population Biology, University of California, Davis, CA, USA} $^{,}$\thanks{Correspondence should be addressed to TMB (beissingert@missouri.edu) or JRI (rossibarra@ucdavis.edu).}
-  }
-
-\date{\small Manuscript intended for \emph{Nature Genetics}, \today}
-
-\begin{document}
 \maketitle
@@
 \begin{abstract}
-  \noindent \bf
-  \noindent
 Genetic diversity is shaped by the interaction of drift and selection, but the details of this interaction are not well understood.
-Although a population's propensity to drift is typically summarized by its long-term effective population size ($N_e$ ), rapidly changing population demographics complicate this relationship.
+Although a population's propensity to drift is determined by its demographic history, typically summarized by its long-term effective population size ($N_e$ ), rapidly changing population demographics complicate this relationship.
 To better understand how changing demography impacts selection, we used whole-genome sequencing data to investigate patterns of linked selection in domesticated and wild maize (teosinte).
 We produce the first whole-genome estimate of the demography of maize domestication, showing that maize was reduced to approximately 5\% the population size of teosinte before it experienced rapid expansion post-domestication to population sizes much larger than its ancestor.
-Evaluation of patterns of nucleotide diversity in and near genes shows little evidence of selection on beneficial amino acid substitutions, and that the domestication bottleneck lead to a decline in the efficiency of purifying selection in maize.
+Evaluation of patterns of nucleotide diversity in and near genes shows little evidence of selection on beneficial amino acid substitutions, and that the domestication bottleneck led to a decline in the efficiency of purifying selection in maize.
 Young alleles, however, show evidence of much stronger purifying selection in maize, reflecting the much larger effective size of present day populations.
@@
 After establishing that genic sites are not evolving neutrally, we used sites $>5$kb from genes to estimate the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}).
+Here, we make the assumption that non-genic DNA is less selectively constrained than genic DNA.
+This assumption is consistent with observations purifying selection described in a later section, but we should highlight that there are known instances of non-genic sites under selection in maize\cite{studer2011}.
 The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ teosinte individuals.
@@
 
-In addition to our simple bottleneck model, we investigated two alternative approaches for demographic inference.
-First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using low frequency variants  informative of population expansion.
-This analysis yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
+In addition to our simple bottleneck model, and because our modest sample size of fully sequenced individuals has limited power to infer recent population expansion, we investigated two alternative approaches for demographic inference.
+First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using on a rare variant-based estimate of  $\theta$ \cite{fu1993} (see online methods).
+Because of the larger population size, this sample can be used to estimate the abundance of rare alleles and is therefore informative for recent population expansion.
+This yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
 Finally, we applied a model-free coalescent approach \cite{schiffels2014} using a subset of our samples. 
@@
 \end{center}
-\caption{\textbf{A} and \textbf{B} show mean pairwise diversity $\pi$, $\pm$ one standard deviation, while \textbf{C} and \textbf{D} depict and Tajimas D in 1kb windows from genic and nongenic regions of maize and teosinte.
+\caption{\textbf{A} and \textbf{B} show mean pairwise diversity $\pi$, $\pm$ one standard deviation, while \textbf{C} and \textbf{D} depict and Tajima's D in 1kb windows from genic and nongenic regions of maize and teosinte.
 \label{fig:diversity} }
@@
 If a substantial proportion of missense mutations have been fixed due to hard sweeps, diversity around these substitutions should be lower than around synonymous substitutions. 
-We observe this pattern around the causative amino acid substitution in the the maize domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005, wang2015}.
+We observe this pattern around the causative amino acid substitution in the maize domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005, wang2015}.
 Genome-wide, however, we observe no differences in diversity at sites near synonymous versus missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
@@
 \subsection*{Population expansion leads to stronger purifying selection in modern maize} %JRI
-Motivated by the rapid post-domestication expansion of maize evident in our demographic analyses, we reasoned that low-frequency --- and thus younger --- polymorphisms might show patterns distinct from pairwise diversity.
+Motivated by the rapid post-domestication expansion of maize evident in our demographic analyses, we reasoned that low-frequency --- and thus younger --- polymorphisms might show patterns distinct from pairwise diversity, which is determined primarily by intermediate frequency --- therefore comparably older --- alleles.
 Singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) appears nearly identical to results from pairwise diversity (Figure \ref{fig:hardSweeps}), providing little support for a substantial recent increase in the number or strength of hard sweeps occurring in maize.
@@
 
-%\subsection*{Demography of domestication} %JRI
+\subsection*{Demography of domestication} %JRI
 Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
@@
 
-One surprising result from our model is the estimated timing of domestication at $\approx 15,000$ years before present.
+One surprising result from our model is the estimated divergence time of maize and teosinte approximately $15,000$ generations before present.
 While this appears to conflict with archaeological estimates \cite{piperno2009}, we emphasize that this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
@@
 
-%\subsection*{Hard sweeps do not shape genome-wide diversity in maize} %JRI
-
-Irrespective of ancient or current population size, our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of purifying selection  \cite{enard2014}.
+\subsection*{Hard sweeps do not shape genome-wide diversity in maize} %JRI
+Our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of purifying selection  \cite{enard2014}.
 Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability.
-Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with ``soft sweeps''  from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
+Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with soft sweeps from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
 Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{van2012historical, beissinger2014} all support the importance of standing variation as primary sources of adaptive variation.
@@
 As many natural and domesticated populations have undergone considerable demographic change in their recent past, long-term comparisons of $N_e$ are likely not informative about current processes affecting allele frequency trajectories.
-
-\section*{METHODS}
-Methods and any associated references are available in a separate pdf file.
 
@@
 
-{\scriptsize \sf
-\renewcommand{\baselinestretch}{2.0}
-\bibliographystyle{naturemag}
-%\bibliographystyle{NatureSeriesT}
-\bibliography{Reference}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
 \section*{ONLINE METHODS}
@@
 \subsection*{Physical and genetic maps}
-Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{bukowski2015}. 
+Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (\url{ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/}) as described by \cite{bukowski2015}.
 All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following \cite{li2011statistical}.
@@
 We filtered these data to include only SNPs with data in $\geq 1,500$ individuals, and then projected the SFS down to a sample of 500 individuals by sampling each marker without replacement 1,000 times according to the observed allele frequencies.
-We then estimated $N_e$ from the data assuming $\mu = 3 \times 10^{-8}$ \cite{clark2005} and the relation  $4N_e\mu = \frac{S}{L}$ \cite{fu1993}, where where $S$ is the total number of singleton SNPs and $L$ is the total number of SNPs in the dataset.
+We then estimated $N_e$ from the data assuming $\mu = 3 \times 10^{-8}$ \cite{clark2005} and the relation  $4N_e\mu = \frac{S}{L}$ \cite{fu1993}, where $S$ is the total number of singleton SNPs and $L$ is the total number of SNPs in the dataset.
 
@@
 
+{\scriptsize \sf
+\renewcommand{\baselinestretch}{2.0}
+\bibliographystyle{naturemag}
+%\bibliographystyle{NatureSeriesT}
+\bibliography{Reference}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \onecolumn

### 2016-02-18 NatPlant2/manuscript.tex
new figure, text

@@
 
-In addition to our simple bottleneck model, and because our modest sample size of fully sequenced individuals has limited power to infer recent population expansion, we investigated two alternative approaches for demographic inference.
-First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size using on a rare variant-based estimate of  $\theta$ \cite{fu1993} (see online methods).
-Because of the larger population size, this sample can be used to estimate the abundance of rare alleles and is therefore informative for recent population expansion.
+Because our modest sample size of fully sequenced individuals has limited power to infer recent population expansion, we investigated two alternative approaches for demographic inference.  
+First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size.
+Because rare variants provide the best information about recent effective population sizes \jri{CITE}, we estimate $N_e$ using \citet{fu1993} singleton-based estimator of the population mutation rate $\theta=4N_e\mu$ and published values of the mutation rate \cite{clark2005} (see online methods for details).
 This yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
@@
 \section*{AUTHOR CONTRIBUTIONS}
-TMB and JRI devised this study. TMB, LW, and KC analyzed the data. AD performed early-stage simulations. MBH provided advice. TMB, JRI, and MBH wrote the manuscript.
+TMB and JRI devised this study. TMB, LW, JRI, and KC analyzed the data. AD performed early-stage simulations. TMB, JRI, and MBH wrote the manuscript.
 

### 2016-02-18 NatPlant2/manuscript.tex
citation

@@
 First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size.
-Because rare variants provide the best information about recent effective population sizes \jri{CITE}, we estimate $N_e$ using \citet{fu1993} singleton-based estimator of the population mutation rate $\theta=4N_e\mu$ and published values of the mutation rate \cite{clark2005} (see online methods for details).
+Because rare variants provide the best information about recent effective population sizes \cite{keinan2012}, we estimate $N_e$ using  a singleton-based estimator \cite{fu1993} of the population mutation rate $\theta=4N_e\mu$  and published values of the mutation rate \cite{clark2005} (see online methods for details).
 This yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.

### 2016-02-19 NatPlant2/manuscript.tex
more edits

--- 2016-02-18 NatPlant2/manuscript.tex
@@
 \subsection*{Demography of maize domestication} %JRI
-After establishing that genic sites are not evolving neutrally, we used sites $>5$kb from genes to estimate the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}).
-Here, we make the assumption that non-genic DNA is less selectively constrained than genic DNA.
-This assumption is consistent with observations purifying selection described in a later section, but we should highlight that there are known instances of non-genic sites under selection in maize\cite{studer2011}.
+We next estimated a demographic model of maize domestication (Figure \ref{fig:bottleneck}). 
+To minimize the impact of selection on our estimates \jri{cite http://onlinelibrary.wiley.com/doi/10.1111/mec.13390/full}, we only included sites \textgreater 5kb from genes.  
 The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ teosinte individuals.
@@
 This yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
-Finally, we applied a model-free coalescent approach \cite{schiffels2014} using a subset of our samples. 
+Finally, we employed a model-free coalescent approach \cite{schiffels2014} to estimate population size change using a subset of six genomes each of maize and teosinte.
 Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
@@
 We used the software $\delta$a$\delta$i \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS, using only sites $>5 kb$ from a gene to ameliorate the effects of linked selection.
-We modeled a teosinte population of constant effective size $N_a$, that at time $T_b$ generations in the past gave rise to a maize population of size $N_b$ which grew exponentially to size $N_m$ in the present (Figure \ref{fig:bottleneck}).
+To minimize the number of parameters estimated, we employed a simple demographic model which posits a teosinte population of constant effective size $N_a$.
+At time $T_b$ generations in the past, this population gave rise to a maize population of size $N_b$ which grew exponentially to size $N_m$ in the present (Figure \ref{fig:bottleneck}).
 The model includes migration of $M_{mt}$ individuals each generation from maize to teosinte and $M_{tm}$ individuals from teosinte to maize.

### 2016-02-19 NatPlant2/manuscript.tex
working late with coffee
a few updates

@@
 \usepackage[normalem]{ulem}
+%\usepackage{coffee4}
 \newcommand{\jri}[1]{\textcolor{red}{\scriptsize #1}}
@@
 \maketitle
+%\cofeAm{1}{1.0}{0}{5.5cm}{3cm}
 
@@
 After its split from teosinte, our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,000$.
-Maize and teosinte have continued to exchange migrants after the population split, with gene flow between the populations estimated at $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation from teosinte to maize and $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants from maize to teosinte.
+Although, our model provides only a rough approximation of migration rates, we included migration parameters during demographic inference because omitting these could bias our population size estimates.
+Still, we observe that maize and teosinte have continued to exchange migrants after the population split.
+We estimate that gene flow from teosinte to maize was $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation, and from maize to teosinte we estimate $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants per generation.
 

### 2016-02-19 NatPlant2/manuscript.tex
citation
updates

@@
 We next estimated a demographic model of maize domestication (Figure \ref{fig:bottleneck}). 
-To minimize the impact of selection on our estimates \jri{cite http://onlinelibrary.wiley.com/doi/10.1111/mec.13390/full}, we only included sites \textgreater 5kb from genes.  
+To minimize the impact of selection on our estimates \cite{ewing2016}, we only included sites \textgreater 5kb from genes.  
 The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ teosinte individuals.
@@
 After its split from teosinte, our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,000$.
-Although, our model provides only a rough approximation of migration rates, we included migration parameters during demographic inference because omitting these could bias our population size estimates.
+Although our model provides only a rough approximation of migration rates, we included migration parameters during demographic inference because omitting these could bias our population size estimates.
 Still, we observe that maize and teosinte have continued to exchange migrants after the population split.

### 2016-02-21 NatPlant2/manuscript.tex
sims added

@@
 \usepackage[normalem]{ulem}
-%\usepackage{coffee4}
 \newcommand{\jri}[1]{\textcolor{red}{\scriptsize #1}}
@@
 Genetic diversity is shaped by the interaction of drift and selection, but the details of this interaction are not well understood.
-Although a population's propensity to drift is determined by its demographic history, typically summarized by its long-term effective population size ($N_e$ ), rapidly changing population demographics complicate this relationship.
+The impact of genetic drift in a population is largely determined by its demographic history, typically summarized by its long-term effective population size ($N_e$ ).
+Rapidly changing population demographics complicate this relationship, however.
 To better understand how changing demography impacts selection, we used whole-genome sequencing data to investigate patterns of linked selection in domesticated and wild maize (teosinte).
@@
 Although our model provides only a rough approximation of migration rates, we included migration parameters during demographic inference because omitting these could bias our population size estimates.
-Still, we observe that maize and teosinte have continued to exchange migrants after the population split.
-We estimate that gene flow from teosinte to maize was $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation, and from maize to teosinte we estimate $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants per generation.
+We observe that maize and teosinte have continued to exchange migrants after the population split, with gene flow from teosinte to maize was $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation, and from maize to teosinte we estimate $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants per generation.
 
@@
 As before, this relationship remained after we removed the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12}).
-Finally, while direct comparison of pairwise and singleton diversity within taxa is consistent with   non-equilibrium dynamics in teosinte, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}).
+While direct comparison of pairwise and singleton diversity within taxa is consistent with non-equilibrium dynamics in teosinte, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}) and mirror results from simulations of purifying selection (Figure \ref{sFig:sims}).
 
@@
 Furthermore, by utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without making assumptions about its duration.
+This model paves the way for future work on the demography of domestication, evaluating for example the significance of differences in gene flow estimated here or removing assumptions about demographic history in teosinte.
 
@@
 Assuming the effective size is only $\approx 0.4\%$ of the census size (i.e. 1 ear for every 1000 male plants), this still implies a modern effective population size of more than four billion.
-While these genetic and census estimates are likely inaccurate, all of the evidence points to the fact that the effective size of modern maize is extremely large.
+While these genetic and census estimates are likely inaccurate, all of the evidence points to the fact that the modern effective size of maize is extremely large.
 
@@
 The efficiency of purifying selection is proportional to effective population size \cite{kimura1984}, and these results are thus consistent with our demographic analyses which show a domestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} followed by recent rapid expansion and a much larger modern $N_e$.
+Simple foward-in-time population genetic simulations qualitatively confirm these results, and further suggest that the observed patterns are likely cause by sites under relatively weak purifying selection \ref{sFig:sims}.
 
@@
 
+\subsection*{Simulations}
+We used the program \emph{bneck\_selection\_ind} included in version 0.4.4 of the forward-in-time population genetic simulation library \emph{fwdpp} \citep[\url{https://github.com/molpopgen/fwdpp}[]{thornton2014genetics}.
+All simulations used a population mutation rate of $\theta=20$, a population recombination rate of $\rho=20$, and simulated 150,000 burn-in generations at an ancestral population size of $N_1=15,000$ to establish equilibrium, after which the population instantly changed to size $N_2$ and then grew exponentially for 1,000 generations to size $N_3$.
+To simulate a constant size population emulating teosinte, we set $N_2=N_3=15,000$.
+For maize we simulated a bottleneck similar to that estimated in Figure \ref{fig:bottleneck} by setting $N_2=750$,   followed by exponential growth to a large modern population size of $N_3=150,000$. 
+For each taxon, we performed 1,000 simulations for each of five values of the strength of purifying selection: $s=\{0,10^{-6},10^{-5},10^{-4},10^{-3}\}$. 
+All mutations were assumed to be codominant.
+To mimic nonsynonymous changes at a coding locus, we assumed that $\frac{3}{4}$ of mutations were selected.
+We calculated summary statistic across all sites using version 0.3.4 of \emph{msstats} (\url{https://github.com/molpopgen/msstats/releases}).
+
 {\scriptsize \sf

### 2016-02-21 NatPlant2/supportingInfo.tex
nat_plant reviews
sims added

--- 2015-11-13 supportingInfo.tex
@@
     \end{center}
-\caption{Diversity surrounding the causitive substitution at the \emph{tga1} locus.  \label{sFig:tga1}}
+\caption{Diversity surrounding the causative substitution at the \emph{tga1} locus.  \label{sFig:tga1}}
 \end{figure}
@@
 \begin{figure}
+  \includegraphics[width=\textwidth]{FigsAndFiles/jri_sims.png}
+\caption{Simulations of diversity statistics in maize and teosinte with varying strengths of purifying selection.
+Points show the mean ($\pm$ standard error) of the population mutation rate $\theta$ estimated by singletons ($\eta_1$) and pairwise differences ($\pi$).
+\label{sFig:sims}}
+\end{figure}
+\clearpage
+
+\begin{figure}
   \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_Unselected_manuscript.png}
@@
 \end{figure}
-
-\begin{table}[h]
-\caption{A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by \cite{chia2012}   } \label{sTab:list}
-\end{table}
 \clearpage
@@
 \begin{table}[h]  
+  \begin{center}
   \begin{tabular}{c|c}
@@
   \end{tabular}
-\end{table}
-
-\begin{table}[h]
-\caption{ Parameters, initial values, and boundaries used for model-fitting with $\delta\alpha\delta{i}$.
-Parameters are shown in the units utilized by $\delta\alpha\delta{i}$, although in the text simplified units are reported.
-}\label{sTab:dadi}
+  \end{center}
+  \caption{A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by \cite{chia2012}   } \label{sTab:list}
 \end{table}
@@
 \begin{table}
+  \begin{center}
   \def\arraystretch{2}
@@
   \def\arraystretch{1} % undo stretching
+  \end{center}
+    \caption{ Parameters, initial values, and boundaries used for model-fitting with $\delta\alpha\delta{i}$.
+Parameters are shown in the units utilized by $\delta\alpha\delta{i}$, although in the text simplified units are reported.
+}\label{sTab:dadi}
 \end{table}

### 2016-02-22 NatPlant2/manuscript.tex
updates

@@
 \title{More than one Author with different Affiliations}
-\author[*,1,2,3]{Timothy M. Beissinger}
+\author[1,2,3,*]{Timothy M. Beissinger}
 \author[4]{Li Wang}
@@
 \author[4]{Matthew B. Hufford}
-\author[*,1,5]{Jeffrey Ross-Ibarra}
+\author[1,5,*]{Jeffrey Ross-Ibarra}
 \affil[1]{Dept. of Plant Sciences, University of California, Davis, CA, USA}
@@
 Drift and selection do not operate independently to determine genetic variability, however, in large part because linkage allows the effects of selection to be wide-ranging \cite{smith1974, li2012,slotte2014}.
-Linked selection can take the form of hitch-hiking, when the frequency of a neutral allele changes as a result of positive selection at a physically linked site \cite{smith1974}, or background selection, where diversity is reduced at loci linked to a site undergoing selection against deleterious alleles \cite{charlesworth1993}.
+Linked selection, which refers to selection in the presence of linkage \cite{slotte2014}, can take the form of hitch-hiking, when the frequency of a neutral allele changes as a result of positive selection at a physically linked site \cite{smith1974}, or background selection, where diversity is reduced at loci linked to a site undergoing selection against deleterious alleles \cite{charlesworth1993}.
 Recent work in \emph{Drosophila}, for example, has shown that virtually the entire genome is impacted by the combined effects of these processes \cite{sella2009,elyashiv2014,andolfatto2005}.

### 2016-02-22 NatPlant2/manuscript.tex
edit

@@
 Drift and selection do not operate independently to determine genetic variability, however, in large part because linkage allows the effects of selection to be wide-ranging \cite{smith1974, li2012,slotte2014}.
-Linked selection, which refers to selection in the presence of linkage \cite{slotte2014}, can take the form of hitch-hiking, when the frequency of a neutral allele changes as a result of positive selection at a physically linked site \cite{smith1974}, or background selection, where diversity is reduced at loci linked to a site undergoing selection against deleterious alleles \cite{charlesworth1993}.
+Linked selection, which refers to the effects of selection at one site on diversity at linked sites \cite{slotte2014}, can take the form of hitch-hiking, when the frequency of a neutral allele changes as a result of positive selection at a physically linked site \cite{smith1974}, or background selection, where diversity is reduced at loci linked to a site undergoing selection against deleterious alleles \cite{charlesworth1993}.
 Recent work in \emph{Drosophila}, for example, has shown that virtually the entire genome is impacted by the combined effects of these processes \cite{sella2009,elyashiv2014,andolfatto2005}.

### 2016-02-25 NatPlant2/manuscript.tex
updates

@@
 Finally, we employed a model-free coalescent approach \cite{schiffels2014} to estimate population size change using a subset of six genomes each of maize and teosinte.
-Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
+Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent with the other approaches, identifying population isolation beginning between 10,000 and 15,000 generations ago (Figure \ref{sFig:coalescense}), a clear domestication bottleneck, and ultimately rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
 Our assessment of the historical demography of maize and teosinte provides context for subsequent analyses of linked selection.

### 2016-02-25 NatPlant2/supportingInfo.tex
updates

@@
   \begin{center}
-    \includegraphics[width=.85\textwidth]{FigsAndFiles/TIL_BKN_bootstrapping_msmc.pdf}
+    \includegraphics[width=.85\textwidth]{FigsAndFiles/relativeCrossCoalescenceRate.pdf}
+    \end{center}
+\caption{Cross-coalescence rate between maize and teosinte estimated using MSMC.
+Time is estimated assuming an annual generation time and a mutation rate of $\mu=3\times 10^{-8}$ \label{sFig:coalescense}}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.85\textwidth]{FigsAndFiles/TIL_BKN_bootstrapping_msmc2.pdf}
     \end{center}

### 2016-02-25 NatPlant2/manuscript.tex
updates

--- 2016-02-22 NatPlant2/manuscript.tex
@@
 Finally, we employed a model-free coalescent approach \cite{schiffels2014} to estimate population size change using a subset of six genomes each of maize and teosinte.
-Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent, identifying a clear domestication bottleneck followed by rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
+Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent with the other approaches, identifying population isolation beginning between 10,000 and 15,000 generations ago, a clear domestication bottleneck, and ultimately rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
 Our assessment of the historical demography of maize and teosinte provides context for subsequent analyses of linked selection.
@@
 We then removed heterozygous genotypes and filtered sites with a mapping quality $<30$, a base quality $<20$, or a $|\textit{log}_2(\text{depth})|<1$.
-We ran MSMC with pattern parameters $20\times2+20\times4+10\times2$.
+We ran MSMC with pattern parameter $20\times2+20\times4+10\times2$ (Figure \ref{sFig:msmc}A) for population size inference.
+To estimate the rate of cross-coalescence we used four maize and four teosinte haplotypes with pattern parameter $20\times1 + 20\times2$ (Figure \ref{sFig:msmc}B).
 

### 2016-02-25 NatPlant2/supportingInfo.tex
updates

--- 2016-02-21 NatPlant2/supportingInfo.tex
@@
 \begin{figure}
-  \begin{center}
-    \includegraphics[width=.85\textwidth]{FigsAndFiles/TIL_BKN_bootstrapping_msmc.pdf}
-    \end{center}
-\caption{Effective population size estimated over time using MSMC.
-Shown are estimates (solid lines) and boostrap resampling (dotted lines) for both maize (red) and teosinte (blue).
-Time is estimated assuming an annual generation time and a mutation rate of $\mu=3\times 10^{-8}$ \label{sFig:msmc}}
+  \includegraphics[width=.5\textwidth]{FigsAndFiles/TIL_BKN_bootstrapping_msmc2_labeled.pdf}
+  \includegraphics[width=.5\textwidth]{FigsAndFiles/relativeCrossCoalescenceRate_labeled.pdf}
+\caption{MSMC Analyses.
+Shown in \textbf{A} are effective population size estimates over time.
+Estimates are depicted as solid lines and boostrap resampling is represented with dotted lines for both maize (red) and teosinte (blue).
+\textbf{B} depticts the cross-coalescence rate between maize and teosinte estimated using MSMC.
+In both panels, time is estimated assuming an annual generation time and a mutation rate of $\mu=3\times 10^{-8}$ \label{sFig:msmc}}
 \end{figure}

### 2016-02-29 NatPlant2/manuscript.tex
updates

@@
 Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability.
-Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with soft sweeps from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
+Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while published data for several loci are consistent with soft sweeps from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
 Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{van2012historical, beissinger2014} all support the importance of standing variation as primary sources of adaptive variation.

### 2016-04-08 NatPlant2/manuscript.tex
updates

@@
 
-\section*{ACKNOWLEDGEMENTS}
-We are indebted to Graham Coop and Simon Aeschbacher for their constructive input during this study.
-We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.
-Funding was provided by NSF Plant Genome Research Project 1238014 and the USDA-Agricultural Research Service.
-
-\section*{AUTHOR CONTRIBUTIONS}
-TMB and JRI devised this study. TMB, LW, JRI, and KC analyzed the data. AD performed early-stage simulations. TMB, JRI, and MBH wrote the manuscript.
-
-\section*{COMPETING INTERESTS STATEMENT}
-The authors declare no competing financial interests.
-
-\section*{ONLINE METHODS}
+\section*{METHODS}
 \subsection*{BASH, R, and Python scripts}
@@
 
+\section*{ACKNOWLEDGEMENTS}
+We are indebted to Graham Coop and Simon Aeschbacher for their constructive input during this study.
+We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.
+Funding was provided by NSF Plant Genome Research Project 1238014 and the USDA-Agricultural Research Service.
+
+\section*{AUTHOR CONTRIBUTIONS}
+TMB and JRI devised this study. TMB, LW, JRI, and KC analyzed the data. AD performed early-stage simulations. TMB, JRI, and MBH wrote the manuscript.
+
+\section*{COMPETING INTERESTS STATEMENT}
+The authors declare no competing financial interests.
+
 {\scriptsize \sf
 \renewcommand{\baselinestretch}{2.0}
-\bibliographystyle{naturemag}
-%\bibliographystyle{NatureSeriesT}
-\bibliography{Reference}
+%\bibliographystyle{naturemag}
+%\bibliography{Reference}
+
+\begin{thebibliography}{10}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{dobzhansky1957}
+\bibinfo{author}{Dobzhansky, T.} \& \bibinfo{author}{Pavlovsky, O.}
+\newblock \bibinfo{title}{An experimental study of interaction between genetic
+  drift and natural selection}.
+\newblock \emph{\bibinfo{journal}{Evolution}} \bibinfo{pages}{311--319}
+  (\bibinfo{year}{1957}).
+
+\bibitem{voight2005}
+\bibinfo{author}{Voight, B.~F.} \emph{et~al.}
+\newblock \bibinfo{title}{Interrogating multiple aspects of variation in a full
+  resequencing data set to infer human population size changes}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences of the United States of America}} \textbf{\bibinfo{volume}{102}},
+  \bibinfo{pages}{18508--18513} (\bibinfo{year}{2005}).
+
+\bibitem{luikart2003}
+\bibinfo{author}{Luikart, G.}, \bibinfo{author}{England, P.~R.},
+  \bibinfo{author}{Tallmon, D.}, \bibinfo{author}{Jordan, S.} \&
+  \bibinfo{author}{Taberlet, P.}
+\newblock \bibinfo{title}{The power and promise of population genomics: from
+  genotyping to genome typing}.
+\newblock \emph{\bibinfo{journal}{Nature Reviews Genetics}}
+  \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{981--994}
+  (\bibinfo{year}{2003}).
+
+\bibitem{gutenkunst2009}
+\bibinfo{author}{Gutenkunst, R.~N.}, \bibinfo{author}{Hernandez, R.~D.},
+  \bibinfo{author}{Williamson, S.~H.} \& \bibinfo{author}{Bustamante, C.~D.}
+\newblock \bibinfo{title}{Inferring the joint demographic history of multiple
+  populations from multidimensional snp frequency data}.
+\newblock \emph{\bibinfo{journal}{PLoS genetics}} \textbf{\bibinfo{volume}{5}},
+  \bibinfo{pages}{e1000695} (\bibinfo{year}{2009}).
+
+\bibitem{akey2009}
+\bibinfo{author}{Akey, J.~M.}
+\newblock \bibinfo{title}{Constructing genomic maps of positive selection in
+  humans: Where do we go from here?}
+\newblock \emph{\bibinfo{journal}{Genome research}}
+  \textbf{\bibinfo{volume}{19}}, \bibinfo{pages}{711--722}
+  (\bibinfo{year}{2009}).
+
+\bibitem{smith1974}
+\bibinfo{author}{Smith, J.~M.} \& \bibinfo{author}{Haigh, J.}
+\newblock \bibinfo{title}{The hitch-hiking effect of a favourable gene}.
+\newblock \emph{\bibinfo{journal}{Genetical research}}
+  \textbf{\bibinfo{volume}{23}}, \bibinfo{pages}{23--35}
+  (\bibinfo{year}{1974}).
+
+\bibitem{li2012}
+\bibinfo{author}{Li, J.} \emph{et~al.}
+\newblock \bibinfo{title}{Joint analysis of demography and selection in
+  population genetics: where do we stand and where could we go?}
+\newblock \emph{\bibinfo{journal}{Molecular Ecology}}
+  \textbf{\bibinfo{volume}{21}}, \bibinfo{pages}{28--44}
+  (\bibinfo{year}{2012}).
+
+\bibitem{slotte2014}
+\bibinfo{author}{Slotte, T.}
+\newblock \bibinfo{title}{The impact of linked selection on plant genomic
+  variation}.
+\newblock \emph{\bibinfo{journal}{Briefings in functional genomics}}
+  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{268--275}
+  (\bibinfo{year}{2014}).
+
+\bibitem{charlesworth1993}
+\bibinfo{author}{Charlesworth, B.}, \bibinfo{author}{Morgan, M.} \&
+  \bibinfo{author}{Charlesworth, D.}
+\newblock \bibinfo{title}{The effect of deleterious mutations on neutral
+  molecular variation.}
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{134}},
+  \bibinfo{pages}{1289--1303} (\bibinfo{year}{1993}).
+
+\bibitem{sella2009}
+\bibinfo{author}{Sella, G.}, \bibinfo{author}{Petrov, D.~A.},
+  \bibinfo{author}{Przeworski, M.} \& \bibinfo{author}{Andolfatto, P.}
+\newblock \bibinfo{title}{Pervasive natural selection in the drosophila
+  genome?}
+\newblock \emph{\bibinfo{journal}{PLoS genetics}} \textbf{\bibinfo{volume}{5}},
+  \bibinfo{pages}{e1000495} (\bibinfo{year}{2009}).
+
+\bibitem{elyashiv2014}
+\bibinfo{author}{Elyashiv, E.} \emph{et~al.}
+\newblock \bibinfo{title}{A genomic map of the effects of linked selection in
+  drosophila}.
+\newblock \emph{\bibinfo{journal}{arXiv preprint arXiv:1408.5461}}
+  (\bibinfo{year}{2014}).
+
+\bibitem{andolfatto2005}
+\bibinfo{author}{Andolfatto, P.}
+\newblock \bibinfo{title}{Adaptive evolution of non-coding dna in drosophila}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{437}},
+  \bibinfo{pages}{1149--1152} (\bibinfo{year}{2005}).
+
+\bibitem{cutter2013}
+\bibinfo{author}{Cutter, A.~D.} \& \bibinfo{author}{Payseur, B.~A.}
+\newblock \bibinfo{title}{Genomic signatures of selection at linked sites:
+  unifying the disparity among species}.
+\newblock \emph{\bibinfo{journal}{Nature Reviews Genetics}}
+  \textbf{\bibinfo{volume}{14}}, \bibinfo{pages}{262--274}
+  (\bibinfo{year}{2013}).
+
+\bibitem{corbett2015}
+\bibinfo{author}{Corbett-Detig, R.~B.}, \bibinfo{author}{Hartl, D.~L.} \&
+  \bibinfo{author}{Sackton, T.~B.}
+\newblock \bibinfo{title}{Natural selection constrains neutral diversity across
+  a wide range of species}.
+\newblock \emph{\bibinfo{journal}{PLoS Biol}} \textbf{\bibinfo{volume}{13}},
+  \bibinfo{pages}{e1002112} (\bibinfo{year}{2015}).
+\newblock \urlprefix\url{http://dx.doi.org/10.1371%2Fjournal.pbio.1002112}.
+
+\bibitem{leffler2012}
+\bibinfo{author}{Leffler, E.~M.} \emph{et~al.}
+\newblock \bibinfo{title}{Revisiting an old riddle: what determines genetic
+  diversity levels within species}.
+\newblock \emph{\bibinfo{journal}{PLoS Biol}} \textbf{\bibinfo{volume}{10}},
+  \bibinfo{pages}{e1001388} (\bibinfo{year}{2012}).
+
+\bibitem{duchen2013}
+\bibinfo{author}{Duchen, P.}, \bibinfo{author}{{\v{Z}}ivkovi{\'c}, D.},
+  \bibinfo{author}{Hutter, S.}, \bibinfo{author}{Stephan, W.} \&
+  \bibinfo{author}{Laurent, S.}
+\newblock \bibinfo{title}{Demographic inference reveals african and european
+  admixture in the north american drosophila melanogaster population}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{193}},
+  \bibinfo{pages}{291--301} (\bibinfo{year}{2013}).
+
+\bibitem{reich1998}
+\bibinfo{author}{Reich, D.~E.} \& \bibinfo{author}{Goldstein, D.~B.}
+\newblock \bibinfo{title}{Genetic evidence for a paleolithic human population
+  expansion in africa}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{95}}, \bibinfo{pages}{8119--8123}
+  (\bibinfo{year}{1998}).
+
+\bibitem{hyten2006}
+\bibinfo{author}{Hyten, D.~L.} \emph{et~al.}
+\newblock \bibinfo{title}{Impacts of genetic bottlenecks on soybean genome
+  diversity}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{103}}, \bibinfo{pages}{16666--16671}
+  (\bibinfo{year}{2006}).
+
+\bibitem{bovine2009}
+\bibinfo{author}{Consortium, B.~H.} \emph{et~al.}
+\newblock \bibinfo{title}{Genome-wide survey of snp variation uncovers the
+  genetic structure of cattle breeds}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{324}},
+  \bibinfo{pages}{528--532} (\bibinfo{year}{2009}).
+
+\bibitem{ellegren2014}
+\bibinfo{author}{Ellegren, H.}
+\newblock \bibinfo{title}{Genome sequencing and population genomics in
+  non-model organisms}.
+\newblock \emph{\bibinfo{journal}{Trends in ecology \& evolution}}
+  \textbf{\bibinfo{volume}{29}}, \bibinfo{pages}{51--63}
+  (\bibinfo{year}{2014}).
+
+\bibitem{smith1995}
+\bibinfo{author}{Smith, B.~D.}
+\newblock \emph{\bibinfo{title}{The emergence of agriculture}}
+  (\bibinfo{publisher}{Scientific American Library New York},
+  \bibinfo{year}{1995}).
+
+\bibitem{matsuoka2002}
+\bibinfo{author}{Matsuoka, Y.} \emph{et~al.}
+\newblock \bibinfo{title}{A single domestication for maize shown by multilocus
+  microsatellite genotyping}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{99}}, \bibinfo{pages}{6080--6084}
+  (\bibinfo{year}{2002}).
+
+\bibitem{wright2005}
+\bibinfo{author}{Wright, S.~I.} \emph{et~al.}
+\newblock \bibinfo{title}{The effects of artificial selection on the maize
+  genome}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{308}},
+  \bibinfo{pages}{1310--1314} (\bibinfo{year}{2005}).
+
+\bibitem{eyre1998}
+\bibinfo{author}{Eyre-Walker, A.}, \bibinfo{author}{Gaut, R.~L.},
+  \bibinfo{author}{Hilton, H.}, \bibinfo{author}{Feldman, D.~L.} \&
+  \bibinfo{author}{Gaut, B.~S.}
+\newblock \bibinfo{title}{Investigation of the bottleneck leading to the
+  domestication of maize}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{95}}, \bibinfo{pages}{4441--4446}
+  (\bibinfo{year}{1998}).
+
+\bibitem{tenaillon2004}
+\bibinfo{author}{Tenaillon, M.~I.}, \bibinfo{author}{U'Ren, J.},
+  \bibinfo{author}{Tenaillon, O.} \& \bibinfo{author}{Gaut, B.~S.}
+\newblock \bibinfo{title}{Selection versus demography: a multilocus
+  investigation of the domestication process in maize}.
+\newblock \emph{\bibinfo{journal}{Molecular Biology and Evolution}}
+  \textbf{\bibinfo{volume}{21}}, \bibinfo{pages}{1214--1225}
+  (\bibinfo{year}{2004}).
+
+\bibitem{chia2012}
+\bibinfo{author}{Chia, J.-M.} \emph{et~al.}
+\newblock \bibinfo{title}{Maize hapmap2 identifies extant variation from a
+  genome in flux}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}
+  \textbf{\bibinfo{volume}{44}}, \bibinfo{pages}{803--807}
+  (\bibinfo{year}{2012}).
+
+\bibitem{bukowski2015}
+\bibinfo{author}{Bukowski, R.} \emph{et~al.}
+\newblock \bibinfo{title}{Construction of the third generation zea mays
+  haplotype map}.
+\newblock \emph{\bibinfo{journal}{bioRxiv}} \bibinfo{pages}{026963}
+  (\bibinfo{year}{2015}).
+
+\bibitem{hufford2012}
+\bibinfo{author}{Hufford, M.~B.} \emph{et~al.}
+\newblock \bibinfo{title}{Comparative population genomics of maize
+  domestication and improvement}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}
+  \textbf{\bibinfo{volume}{44}}, \bibinfo{pages}{808--811}
+  (\bibinfo{year}{2012}).
+
+\bibitem{ewing2016}
+\bibinfo{author}{Ewing, G.~B.} \& \bibinfo{author}{Jensen, J.~D.}
+\newblock \bibinfo{title}{The consequences of not accounting for background
+  selection in demographic inference}.
+\newblock \emph{\bibinfo{journal}{Molecular ecology}}
+  \textbf{\bibinfo{volume}{25}}, \bibinfo{pages}{135--141}
+  (\bibinfo{year}{2016}).
+
+\bibitem{Hearne2015}
+\bibinfo{author}{Hearne, S.}, \bibinfo{author}{Chen, C.},
+  \bibinfo{author}{Buckler, E.} \& \bibinfo{author}{Mitchell, S.}
+\newblock \bibinfo{title}{Unimputed gbs derived snps for maize landrace
+  accessions represented in the seed-maize gwas panel}.
+\newblock \bibinfo{howpublished}{\url{http://hdl.handle.net/11529/10034}}
+  (\bibinfo{year}{2015}).
+\newblock \bibinfo{note}{Accessed: 2015-02-16}.
+
+\bibitem{keinan2012}
+\bibinfo{author}{Keinan, A.} \& \bibinfo{author}{Clark, A.~G.}
+\newblock \bibinfo{title}{Recent explosive human population growth has resulted
+  in an excess of rare genetic variants}.
+\newblock \emph{\bibinfo{journal}{science}} \textbf{\bibinfo{volume}{336}},
+  \bibinfo{pages}{740--743} (\bibinfo{year}{2012}).
+
+\bibitem{fu1993}
+\bibinfo{author}{Fu, Y.-X.} \& \bibinfo{author}{Li, W.-H.}
+\newblock \bibinfo{title}{Statistical tests of neutrality of mutations.}
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{133}},
+  \bibinfo{pages}{693--709} (\bibinfo{year}{1993}).
+
+\bibitem{clark2005}
+\bibinfo{author}{Clark, R.~M.}, \bibinfo{author}{Tavar{\'e}, S.} \&
+  \bibinfo{author}{Doebley, J.}
+\newblock \bibinfo{title}{Estimating a nucleotide substitution rate for maize
+  from polymorphism at a major domestication locus}.
+\newblock \emph{\bibinfo{journal}{Molecular biology and evolution}}
+  \textbf{\bibinfo{volume}{22}}, \bibinfo{pages}{2304--2312}
+  (\bibinfo{year}{2005}).
+
+\bibitem{schiffels2014}
+\bibinfo{author}{Schiffels, S.} \& \bibinfo{author}{Durbin, R.}
+\newblock \bibinfo{title}{Inferring human population size and separation
+  history from multiple genome sequences}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}  (\bibinfo{year}{2014}).
+
+\bibitem{wang2005}
+\bibinfo{author}{Wang, H.} \emph{et~al.}
+\newblock \bibinfo{title}{The origin of the naked grains of maize}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{436}},
+  \bibinfo{pages}{714--719} (\bibinfo{year}{2005}).
+
+\bibitem{wang2015}
+\bibinfo{author}{Wang, H.}, \bibinfo{author}{Studer, A.~J.},
+  \bibinfo{author}{Zhao, Q.}, \bibinfo{author}{Meeley, R.} \&
+  \bibinfo{author}{Doebley, J.~F.}
+\newblock \bibinfo{title}{Evidence that the origin of naked kernels during
+  maize domestication was caused by a single amino acid substitution in tga1}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \bibinfo{pages}{genetics--115}
+  (\bibinfo{year}{2015}).
+
+\bibitem{enard2014}
+\bibinfo{author}{Enard, D.}, \bibinfo{author}{Messer, P.~W.} \&
+  \bibinfo{author}{Petrov, D.~A.}
+\newblock \bibinfo{title}{Genome-wide signals of positive selection in human
+  evolution}.
+\newblock \emph{\bibinfo{journal}{Genome research}}
+  \textbf{\bibinfo{volume}{24}}, \bibinfo{pages}{885--895}
+  (\bibinfo{year}{2014}).
+
+\bibitem{rodgers2015}
+\bibinfo{author}{Rodgers-Melnick, E.} \emph{et~al.}
+\newblock \bibinfo{title}{Recombination in diverse maize is stable,
+  predictable, and associated with genetic load}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{112}}, \bibinfo{pages}{3823--3828}
+  (\bibinfo{year}{2015}).
+
+\bibitem{davydov2010}
+\bibinfo{author}{Davydov, E.~V.} \emph{et~al.}
+\newblock \bibinfo{title}{Identifying a high fraction of the human genome to be
+  under selective constraint using gerp++}.
+\newblock \emph{\bibinfo{journal}{PLoS Comput Biol}}
+  \textbf{\bibinfo{volume}{6}}, \bibinfo{pages}{e1001025}
+  (\bibinfo{year}{2010}).
+
+\bibitem{tenaillon2002}
+\bibinfo{author}{Tenaillon, M.~I.} \emph{et~al.}
+\newblock \bibinfo{title}{Patterns of diversity and recombination along
+  chromosome 1 of maize (zea mays ssp. mays l.)}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{162}},
+  \bibinfo{pages}{1401--1413} (\bibinfo{year}{2002}).
+
+\bibitem{innan2004}
+\bibinfo{author}{Innan, H.} \& \bibinfo{author}{Kim, Y.}
+\newblock \bibinfo{title}{Pattern of polymorphism after strong artificial
+  selection in a domestication event}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences of the United States of America}} \textbf{\bibinfo{volume}{101}},
+  \bibinfo{pages}{10667--10672} (\bibinfo{year}{2004}).
+
+\bibitem{messer2013}
+\bibinfo{author}{Messer, P.~W.} \& \bibinfo{author}{Petrov, D.~A.}
+\newblock \bibinfo{title}{Population genomics of rapid adaptation by soft
+  selective sweeps}.
+\newblock \emph{\bibinfo{journal}{Trends in ecology \& evolution}}
+  \textbf{\bibinfo{volume}{28}}, \bibinfo{pages}{659--669}
+  (\bibinfo{year}{2013}).
+
+\bibitem{garud2015}
+\bibinfo{author}{Garud, N.~R.}, \bibinfo{author}{Messer, P.~W.},
+  \bibinfo{author}{Buzbas, E.~O.} \& \bibinfo{author}{Petrov, D.~A.}
+\newblock \bibinfo{title}{Recent selective sweeps in north american drosophila
+  melanogaster show signatures of soft sweeps}.
+\newblock \emph{\bibinfo{journal}{PLoS genetics}}
+  \textbf{\bibinfo{volume}{11}}, \bibinfo{pages}{e1005004}
+  (\bibinfo{year}{2015}).
+
+\bibitem{piperno2009}
+\bibinfo{author}{Piperno, D.~R.}, \bibinfo{author}{Ranere, A.~J.},
+  \bibinfo{author}{Holst, I.}, \bibinfo{author}{Iriarte, J.} \&
+  \bibinfo{author}{Dickau, R.}
+\newblock \bibinfo{title}{Starch grain and phytolith evidence for early ninth
+  millennium bp maize from the central balsas river valley, mexico}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{106}}, \bibinfo{pages}{5019--5024}
+  (\bibinfo{year}{2009}).
+
+\bibitem{cimmyt1999}
+\bibinfo{author}{Program, T.~M.}
+\newblock \emph{\bibinfo{title}{Development, maintenance, and seed
+  multiplication of open-pollinated maize varieties}}
+  (\bibinfo{publisher}{CIMMYT}, \bibinfo{address}{Mexico, D.F.},
+  \bibinfo{year}{1999}), \bibinfo{edition}{2} edn.
+
+\bibitem{baden2001}
+\bibinfo{author}{Baden, W.~W.} \& \bibinfo{author}{Beekman, C.~S.}
+\newblock \bibinfo{title}{Culture and agriculture: A comment on sissel
+  schroeder, maize productivity in the eastern woodlands and great plains of
+  north america}.
+\newblock \emph{\bibinfo{journal}{American Antiquity}}
+  \bibinfo{pages}{505--515} (\bibinfo{year}{2001}).
+
+\bibitem{studer2011}
+\bibinfo{author}{Studer, A.}, \bibinfo{author}{Zhao, Q.},
+  \bibinfo{author}{Ross-Ibarra, J.} \& \bibinfo{author}{Doebley, J.}
+\newblock \bibinfo{title}{Identification of a functional transposon insertion
+  in the maize domestication gene tb1}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}
+  \textbf{\bibinfo{volume}{43}}, \bibinfo{pages}{1160--1163}
+  (\bibinfo{year}{2011}).
+
+\bibitem{gallavotti2004role}
+\bibinfo{author}{Gallavotti, A.} \emph{et~al.}
+\newblock \bibinfo{title}{The role of barren stalk1 in the architecture of
+  maize}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{432}},
+  \bibinfo{pages}{630--635} (\bibinfo{year}{2004}).
+
+\bibitem{wills2013}
+\bibinfo{author}{Wills, D.~M.} \emph{et~al.}
+\newblock \bibinfo{title}{From many, one: genetic control of prolificacy during
+  maize domestication}.
+\newblock \emph{\bibinfo{journal}{PLoS Genet}} \textbf{\bibinfo{volume}{9}},
+  \bibinfo{pages}{e1003604} (\bibinfo{year}{2013}).
+
+\bibitem{Takuno15062015}
+\bibinfo{author}{Takuno, S.} \emph{et~al.}
+\newblock \bibinfo{title}{Independent molecular basis of convergent highland
+  adaptation in maize}.
+\newblock \emph{\bibinfo{journal}{Genetics}}  (\bibinfo{year}{2015}).
+\newblock
+  \urlprefix\url{http://www.genetics.org/content/early/2015/06/15/genetics.115.178327.abstract}.
+\newblock
+  \eprint{http://www.genetics.org/content/early/2015/06/15/genetics.115.178327.full.pdf+html}.
+
+\bibitem{van2012historical}
+\bibinfo{author}{van Heerwaarden, J.}, \bibinfo{author}{Hufford, M.~B.} \&
+  \bibinfo{author}{Ross-Ibarra, J.}
+\newblock \bibinfo{title}{Historical genomics of north american maize}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{109}}, \bibinfo{pages}{12420--12425}
+  (\bibinfo{year}{2012}).
+
+\bibitem{beissinger2014}
+\bibinfo{author}{Beissinger, T.~M.} \emph{et~al.}
+\newblock \bibinfo{title}{A genome-wide scan for evidence of selection in a
+  maize population under long-term artificial selection for ear number}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{196}},
+  \bibinfo{pages}{829--840} (\bibinfo{year}{2014}).
+
+\bibitem{eyre2007}
+\bibinfo{author}{Eyre-Walker, A.} \& \bibinfo{author}{Keightley, P.~D.}
+\newblock \bibinfo{title}{The distribution of fitness effects of new
+  mutations}.
+\newblock \emph{\bibinfo{journal}{Nature Reviews Genetics}}
+  \textbf{\bibinfo{volume}{8}}, \bibinfo{pages}{610--618}
+  (\bibinfo{year}{2007}).
+
+\bibitem{wallace2014}
+\bibinfo{author}{Wallace, J.}, \bibinfo{author}{Larsson, S.} \&
+  \bibinfo{author}{Buckler, E.}
+\newblock \bibinfo{title}{Entering the second century of maize quantitative
+  genetics}.
+\newblock \emph{\bibinfo{journal}{Heredity}} \textbf{\bibinfo{volume}{112}},
+  \bibinfo{pages}{30--38} (\bibinfo{year}{2014}).
+
+\bibitem{weber2008}
+\bibinfo{author}{Weber, A.~L.} \emph{et~al.}
+\newblock \bibinfo{title}{The genetic architecture of complex traits in
+  teosinte (zea mays ssp. parviglumis): new evidence from association mapping}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{180}},
+  \bibinfo{pages}{1221--1232} (\bibinfo{year}{2008}).
+
+\bibitem{pyhajarvi2013}
+\bibinfo{author}{Pyh{\"a}j{\"a}rvi, T.}, \bibinfo{author}{Hufford, M.~B.},
+  \bibinfo{author}{Mezmouk, S.} \& \bibinfo{author}{Ross-Ibarra, J.}
+\newblock \bibinfo{title}{Complex patterns of local adaptation in teosinte}.
+\newblock \emph{\bibinfo{journal}{Genome biology and evolution}}
+  \textbf{\bibinfo{volume}{5}}, \bibinfo{pages}{1594--1609}
+  (\bibinfo{year}{2013}).
+
+\bibitem{sattath2011}
+\bibinfo{author}{Sattath, S.}, \bibinfo{author}{Elyashiv, E.},
+  \bibinfo{author}{Kolodny, O.}, \bibinfo{author}{Rinott, Y.} \&
+  \bibinfo{author}{Sella, G.}
+\newblock \bibinfo{title}{Pervasive adaptive protein evolution apparent in
+  diversity patterns around amino acid substitutions in drosophila simulans}.
+\newblock \emph{\bibinfo{journal}{PLoS genetics}} \textbf{\bibinfo{volume}{7}},
+  \bibinfo{pages}{e1001302} (\bibinfo{year}{2011}).
+
+\bibitem{williamson2014}
+\bibinfo{author}{Williamson, R.} \emph{et~al.}
+\newblock \bibinfo{title}{Evidence for widespread positive and negative
+  selection in coding and conserved noncoding regions of capsella grandiflora.}
+\newblock \emph{\bibinfo{journal}{PLoS genetics}}
+  \textbf{\bibinfo{volume}{10}}, \bibinfo{pages}{e1004622--e1004622}
+  (\bibinfo{year}{2014}).
+
+\bibitem{hernandez2011}
+\bibinfo{author}{Hernandez, R.~D.} \emph{et~al.}
+\newblock \bibinfo{title}{Classic selective sweeps were rare in recent human
+  evolution}.
+\newblock \emph{\bibinfo{journal}{science}} \textbf{\bibinfo{volume}{331}},
+  \bibinfo{pages}{920--924} (\bibinfo{year}{2011}).
+
+\bibitem{pritchard2010genetics}
+\bibinfo{author}{Pritchard, J.~K.}, \bibinfo{author}{Pickrell, J.~K.} \&
+  \bibinfo{author}{Coop, G.}
+\newblock \bibinfo{title}{The genetics of human adaptation: hard sweeps, soft
+  sweeps, and polygenic adaptation}.
+\newblock \emph{\bibinfo{journal}{Current Biology}}
+  \textbf{\bibinfo{volume}{20}}, \bibinfo{pages}{R208--R215}
+  (\bibinfo{year}{2010}).
+
+\bibitem{ross2009}
+\bibinfo{author}{Ross-Ibarra, J.}, \bibinfo{author}{Tenaillon, M.} \&
+  \bibinfo{author}{Gaut, B.~S.}
+\newblock \bibinfo{title}{Historical divergence and gene flow in the genus
+  zea}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{181}},
+  \bibinfo{pages}{1399--1413} (\bibinfo{year}{2009}).
+
+\bibitem{eyre2009estimating}
+\bibinfo{author}{Eyre-Walker, A.} \& \bibinfo{author}{Keightley, P.~D.}
+\newblock \bibinfo{title}{Estimating the rate of adaptive molecular evolution
+  in the presence of slightly deleterious mutations and population size
+  change}.
+\newblock \emph{\bibinfo{journal}{Molecular biology and evolution}}
+  \textbf{\bibinfo{volume}{26}}, \bibinfo{pages}{2097--2108}
+  (\bibinfo{year}{2009}).
+
+\bibitem{castelletti2014mite}
+\bibinfo{author}{Castelletti, S.}, \bibinfo{author}{Tuberosa, R.},
+  \bibinfo{author}{Pindo, M.} \& \bibinfo{author}{Salvi, S.}
+\newblock \bibinfo{title}{A mite transposon insertion is associated with
+  differential methylation at the maize flowering time qtl vgt1}.
+\newblock \emph{\bibinfo{journal}{G3: Genes| Genomes| Genetics}}
+  \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{805--812}
+  (\bibinfo{year}{2014}).
+
+\bibitem{mao2015}
+\bibinfo{author}{Mao, H.} \emph{et~al.}
+\newblock \bibinfo{title}{A transposable element in a nac gene is associated
+  with drought tolerance in maize seedlings}.
+\newblock \emph{\bibinfo{journal}{Nature Communications}}
+  \textbf{\bibinfo{volume}{6}} (\bibinfo{year}{2015}).
+
+\bibitem{yang2013}
+\bibinfo{author}{Yang, Q.} \emph{et~al.}
+\newblock \bibinfo{title}{{CACTA}-like transposable element in {ZmCCT}
+  attenuated photoperiod sensitivity and accelerated the postdomestication
+  spread of maize}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{110}}, \bibinfo{pages}{16969--16974}
+  (\bibinfo{year}{2013}).
+
+\bibitem{fraser2013gene}
+\bibinfo{author}{Fraser, H.~B.}
+\newblock \bibinfo{title}{Gene expression drives local adaptation in humans}.
+\newblock \emph{\bibinfo{journal}{Genome research}}
+  \textbf{\bibinfo{volume}{23}}, \bibinfo{pages}{1089--1096}
+  (\bibinfo{year}{2013}).
+
+\bibitem{Halligan:2013ir}
+\bibinfo{author}{Halligan, D.~L.} \emph{et~al.}
+\newblock \bibinfo{title}{{Contributions of Protein-Coding and Regulatory
+  Change to Adaptive Molecular Evolution in Murid Rodents}}.
+\newblock \emph{\bibinfo{journal}{PLoS Genetics}} \textbf{\bibinfo{volume}{9}},
+  \bibinfo{pages}{e1003995--14} (\bibinfo{year}{2013}).
+
+\bibitem{kimura1984}
+\bibinfo{author}{Kimura, M.}
+\newblock \emph{\bibinfo{title}{The neutral theory of molecular evolution}}
+  (\bibinfo{publisher}{Cambridge University Press}, \bibinfo{year}{1984}).
+
+\bibitem{Gunther:2010gt}
+\bibinfo{author}{G{\"u}nther, T.} \& \bibinfo{author}{Schmid, K.~J.}
+\newblock \bibinfo{title}{{Deleterious amino acid polymorphisms in Arabidopsis
+  thaliana and rice}}.
+\newblock \emph{\bibinfo{journal}{Theoretical and Applied Genetics}}
+  \textbf{\bibinfo{volume}{121}}, \bibinfo{pages}{157--168}
+  (\bibinfo{year}{2010}).
+
+\bibitem{renaut2015}
+\bibinfo{author}{Renaut, S.} \& \bibinfo{author}{Rieseberg, L.~H.}
+\newblock \bibinfo{title}{The accumulation of deleterious mutations as a
+  consequence of domestication and improvement in sunflowers and other
+  compositae crops}.
+\newblock \emph{\bibinfo{journal}{Molecular biology and evolution}}
+  \bibinfo{pages}{msv106} (\bibinfo{year}{2015}).
+
+\bibitem{coventry2010}
+\bibinfo{author}{Coventry, A.} \emph{et~al.}
+\newblock \bibinfo{title}{Deep resequencing reveals excess rare recent variants
+  consistent with explosive population growth}.
+\newblock \emph{\bibinfo{journal}{Nature communications}}
+  \textbf{\bibinfo{volume}{1}}, \bibinfo{pages}{131} (\bibinfo{year}{2010}).
+
+\bibitem{Mezmouk:2014es}
+\bibinfo{author}{Mezmouk, S.} \& \bibinfo{author}{Ross-Ibarra, J.}
+\newblock \bibinfo{title}{{The pattern and distribution of deleterious
+  mutations in maize.}}
+\newblock \emph{\bibinfo{journal}{G3 (Bethesda, Md.)}}
+  \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{163--171}
+  (\bibinfo{year}{2014}).
+
+\bibitem{eyre2010}
+\bibinfo{author}{Eyre-Walker, A.}
+\newblock \bibinfo{title}{Genetic architecture of a complex trait and its
+  implications for fitness and genome-wide association studies}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{107}}, \bibinfo{pages}{1752--1756}
+  (\bibinfo{year}{2010}).
+
+\bibitem{do2015}
+\bibinfo{author}{Do, R.} \emph{et~al.}
+\newblock \bibinfo{title}{No evidence that selection has been less effective at
+  removing deleterious mutations in europeans than in africans}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}
+  \textbf{\bibinfo{volume}{47}}, \bibinfo{pages}{126--131}
+  (\bibinfo{year}{2015}).
+
+\bibitem{simons2014}
+\bibinfo{author}{Simons, Y.~B.}, \bibinfo{author}{Turchin, M.~C.},
+  \bibinfo{author}{Pritchard, J.~K.} \& \bibinfo{author}{Sella, G.}
+\newblock \bibinfo{title}{The deleterious mutation load is insensitive to
+  recent population history}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}
+  \textbf{\bibinfo{volume}{46}}, \bibinfo{pages}{220--224}
+  (\bibinfo{year}{2014}).
+
+\bibitem{gazave2013}
+\bibinfo{author}{Gazave, E.}, \bibinfo{author}{Chang, D.},
+  \bibinfo{author}{Clark, A.~G.} \& \bibinfo{author}{Keinan, A.}
+\newblock \bibinfo{title}{Population growth inflates the per-individual number
+  of deleterious mutations and reduces their mean effect}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{195}},
+  \bibinfo{pages}{969--978} (\bibinfo{year}{2013}).
+
+\bibitem{lohmueller2014}
+\bibinfo{author}{Lohmueller, K.~E.}
+\newblock \bibinfo{title}{The impact of population demography and selection on
+  the genetic architecture of complex traits}.
+\newblock \emph{\bibinfo{journal}{PLoS Genetics}} \textbf{\bibinfo{volume}{10}}
+  (\bibinfo{year}{2014}).
+
+\bibitem{chen2010}
+\bibinfo{author}{Chen, H.}, \bibinfo{author}{Patterson, N.} \&
+  \bibinfo{author}{Reich, D.}
+\newblock \bibinfo{title}{Population differentiation as a test for selective
+  sweeps}.
+\newblock \emph{\bibinfo{journal}{Genome research}}
+  \textbf{\bibinfo{volume}{20}}, \bibinfo{pages}{393--402}
+  (\bibinfo{year}{2010}).
+
+\bibitem{zeng2010}
+\bibinfo{author}{Zeng, K.} \& \bibinfo{author}{Charlesworth, B.}
+\newblock \bibinfo{title}{The effects of demography and linkage on the
+  estimation of selection and mutation parameters}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{186}},
+  \bibinfo{pages}{1411--1424} (\bibinfo{year}{2010}).
+
+\bibitem{popadin2012}
+\bibinfo{author}{Popadin, K.~Y.}, \bibinfo{author}{Nikolaev, S.~I.},
+  \bibinfo{author}{Junier, T.}, \bibinfo{author}{Baranova, M.} \&
+  \bibinfo{author}{Antonarakis, S.~E.}
+\newblock \bibinfo{title}{Purifying selection in mammalian mitochondrial
+  protein-coding genes is highly effective and congruent with evolution of
+  nuclear genes}.
+\newblock \emph{\bibinfo{journal}{Molecular biology and evolution}}
+  \bibinfo{pages}{mss219} (\bibinfo{year}{2012}).
+
+\bibitem{Elyashiv:2010ga}
+\bibinfo{author}{Elyashiv, E.} \emph{et~al.}
+\newblock \bibinfo{title}{{Shifts in the intensity of purifying selection: An
+  analysis of genome-wide polymorphism data from two closely related yeast
+  species}}.
+\newblock \emph{\bibinfo{journal}{Genome Research}}
+  \textbf{\bibinfo{volume}{20}}, \bibinfo{pages}{1558--1573}
+  (\bibinfo{year}{2010}).
+
+\bibitem{lemmon2014}
+\bibinfo{author}{Lemmon, Z.~H.}, \bibinfo{author}{Bukowski, R.},
+  \bibinfo{author}{Sun, Q.} \& \bibinfo{author}{Doebley, J.~F.}
+\newblock \bibinfo{title}{The role of <italic>cis</italic> regulatory evolution
+  in maize domestication}.
+\newblock \emph{\bibinfo{journal}{PLoS Genet}} \textbf{\bibinfo{volume}{10}},
+  \bibinfo{pages}{e1004745} (\bibinfo{year}{2014}).
+\newblock \urlprefix\url{http://dx.doi.org/10.1371%2Fjournal.pgen.1004745}.
+
+\bibitem{jombart2011}
+\bibinfo{author}{Jombart, T.} \& \bibinfo{author}{Ahmed, I.}
+\newblock \bibinfo{title}{adegenet 1.3-1: new tools for the analysis of
+  genome-wide snp data}.
+\newblock \emph{\bibinfo{journal}{Bioinformatics}}
+  \textbf{\bibinfo{volume}{27}}, \bibinfo{pages}{3070--3071}
+  (\bibinfo{year}{2011}).
+
+\bibitem{schnable2009}
+\bibinfo{author}{Schnable, P.~S.} \emph{et~al.}
+\newblock \bibinfo{title}{The b73 maize genome: complexity, diversity, and
+  dynamics}.
+\newblock \emph{\bibinfo{journal}{science}} \textbf{\bibinfo{volume}{326}},
+  \bibinfo{pages}{1112--1115} (\bibinfo{year}{2009}).
+
+\bibitem{li2011statistical}
+\bibinfo{author}{Li, H.}
+\newblock \bibinfo{title}{A statistical framework for snp calling, mutation
+  discovery, association mapping and population genetical parameter estimation
+  from sequencing data}.
+\newblock \emph{\bibinfo{journal}{Bioinformatics}}
+  \textbf{\bibinfo{volume}{27}}, \bibinfo{pages}{2987--2993}
+  (\bibinfo{year}{2011}).
+
+\bibitem{glaubitz2014}
+\bibinfo{author}{Glaubitz, J.~C.} \emph{et~al.}
+\newblock \bibinfo{title}{Tassel-gbs: a high capacity genotyping by sequencing
+  analysis pipeline}.
+\newblock \emph{\bibinfo{journal}{PLoS One}} \textbf{\bibinfo{volume}{9}},
+  \bibinfo{pages}{E90346} (\bibinfo{year}{2014}).
+
+\bibitem{durinck2009}
+\bibinfo{author}{Durinck, S.}, \bibinfo{author}{Spellman, P.~T.},
+  \bibinfo{author}{Birney, E.} \& \bibinfo{author}{Huber, W.}
+\newblock \bibinfo{title}{Mapping identifiers for the integration of genomic
+  datasets with the r/bioconductor package biomart}.
+\newblock \emph{\bibinfo{journal}{Nature protocols}}
+  \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{1184--1191}
+  (\bibinfo{year}{2009}).
+
+\bibitem{durinck2005}
+\bibinfo{author}{Durinck, S.} \emph{et~al.}
+\newblock \bibinfo{title}{Biomart and bioconductor: a powerful link between
+  biological databases and microarray data analysis}.
+\newblock \emph{\bibinfo{journal}{Bioinformatics}}
+  \textbf{\bibinfo{volume}{21}}, \bibinfo{pages}{3439--3440}
+  (\bibinfo{year}{2005}).
+
+\bibitem{R2014}
+\bibinfo{author}{{R Core Team}}.
+\newblock \emph{\bibinfo{title}{R: A Language and Environment for Statistical
+  Computing}}.
+\newblock \bibinfo{organization}{R Foundation for Statistical Computing},
+  \bibinfo{address}{Vienna, Austria} (\bibinfo{year}{2014}).
+\newblock \urlprefix\url{http://www.R-project.org/}.
+
+\bibitem{korneliussen2014}
+\bibinfo{author}{Korneliussen, T.~S.}, \bibinfo{author}{Albrechtsen, A.} \&
+  \bibinfo{author}{Nielsen, R.}
+\newblock \bibinfo{title}{Angsd: analysis of next generation sequencing data}.
+\newblock \emph{\bibinfo{journal}{BMC bioinformatics}}
+  \textbf{\bibinfo{volume}{15}}, \bibinfo{pages}{356} (\bibinfo{year}{2014}).
+
+\bibitem{mclaren2010}
+\bibinfo{author}{McLaren, W.} \emph{et~al.}
+\newblock \bibinfo{title}{Deriving the consequences of genomic variants with
+  the ensembl api and snp effect predictor}.
+\newblock \emph{\bibinfo{journal}{Bioinformatics}}
+  \textbf{\bibinfo{volume}{26}}, \bibinfo{pages}{2069--2070}
+  (\bibinfo{year}{2010}).
+
+\bibitem{thornton2014genetics}
+\bibinfo{author}{Thornton, K.~R.}
+\newblock \bibinfo{title}{A c++ template library for efficient forward-time
+  population genetic simulation of large populations}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{198}},
+  \bibinfo{pages}{157--166} (\bibinfo{year}{2014}).
+
+\end{thebibliography}
+
 }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%% Supporting info below this line %%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \onecolumn
-\input{supportingInfo}
+\section*{Supporting Information}
+\renewcommand\thefigure{S\arabic{figure}}
+\renewcommand\thetable{S\arabic{table}}    
+\setcounter{figure}{0}
+\setcounter{table}{0}
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_tga1Supp_June.png}
+    \end{center}
+\caption{Diversity surrounding the causative substitution at the \emph{tga1} locus.  \label{sFig:tga1}}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \includegraphics[width=.5\textwidth]{FigsAndFiles/TIL_BKN_bootstrapping_msmc2_labeled.pdf}
+  \includegraphics[width=.5\textwidth]{FigsAndFiles/relativeCrossCoalescenceRate_labeled.pdf}
+\caption{MSMC Analyses.
+Shown in \textbf{A} are effective population size estimates over time.
+Estimates are depicted as solid lines and boostrap resampling is represented with dotted lines for both maize (red) and teosinte (blue).
+\textbf{B} depticts the relative cross-coalescence rate between maize and teosinte estimated using MSMC.
+In both panels, time is estimated assuming an annual generation time and a mutation rate of $\mu=3\times 10^{-8}$ \label{sFig:msmc}}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Conserved_Significance_June.png}
+  \includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Unconserved_Significance_June.png}
+\caption{ Pairwise diversity surrounding synonymous and nonsynonymous
+  substitutions in maize at \textbf{A} highly conserved or \textbf{B}  unconserved sites.
+Bootstrap-based 95\% confidence intervals are depicted via shading.
+Inset plots depict a larger range on the x-axis.
+\label{sFig:consUncons}}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{FigsAndFiles/plotDiversity_TvM_Singletons.png}
+\caption{ Singleton diversity surrounding synonymous and nonsynonymous
+  substitutions in maize. \label{sFig:singleton}}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_maizeAndTeoSingleVsPi.png}
+\caption{ Relative diversity versus distance to nearest gene in maize and teosinte.
+Relative diversity is calculated by comparing to the mean diversity in all windows $\geq 0.02 cM$ from the nearest gene.
+Lines depict cubic smoothing splines with smoothing parameters chosen via generalized cross validation and shading depicts bootstrap-based 95\% confidence intervals.
+\label{sFig:singletonPi}}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{FigsAndFiles/jri_sims.png}
+\caption{Simulations of diversity statistics in maize and teosinte with varying strengths of purifying selection.
+Points show the mean ($\pm$ standard error) of the population mutation rate $\theta$ estimated by singletons ($\eta_1$) and pairwise differences ($\pi$).
+\label{sFig:sims}}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_Unselected_manuscript.png}
+    \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_unselected_Singletons_manuscript.png}
+\caption{ Relative level of diversity versus distance to the nearest gene, in maize and teosinte, based on only sites that do not show evidence of hard or soft sweeps according to H12.
+Two measures of diversity were investigated.
+{\bf A} displays pairwise diversity,
+which is most influenced by intermediate frequency alleles and therefore depicts more ancient evolutionary patterns, and {\bf B} depicts singleton diversity, influenced by rare alleles and thus depicting evolutionary patterns in the recent past.
+Bootstrap-based 95\% confidence intervals are depicted via shading.
+Inset plots depict a smaller range on the x-axis.
+\label{sFig:H12}}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=.75\textwidth]{FigsAndFiles/tilPCA_aug.pdf}\\
+  \includegraphics[width=.75\textwidth]{FigsAndFiles/bknPCA_aug.pdf}\\
+  \end{center}
+  \caption{Principal component analysis of teosinte and maize individuals to ensure that no close relatives were inadvertantly included in our study.
+Plots are based on a random sample of 10,000 SNPs.
+{\bf A} displays the percentage of total variance explained by each principal component for teosinte, while {\bf B} shows PC1 vs PC2 for all 13 teosinte individuals.
+Simlarly, {\bf C} depicts the percentage of total variance explained by each principal component for maize, and {\bf D} shows PC1 vs PC2 for all 23 maize individuals.
+\label{sFig:PCA}}
+\end{figure}
+\clearpage
+
+\begin{table}[h]  
+  \begin{center}
+  \begin{tabular}{c|c}
+    \bf Maize & \bf Teosinte \\ \hline \hline
+    BKN009 &  TIL01 \\
+    BKN010 & TIL02 \\
+    BKN011 & TIL03 \\
+    BKN014 & TIL04-TIP454 \\
+    BKN015 & TIL07 \\
+    BKN016 & TIL09 \\
+    BKN017 & TIL10 \\
+    BKN018 & TIL11 \\
+    BKN019 & TIL12 \\
+    BKN020 & TIL14-TIP498 \\
+    BKN022 & TIL15 \\
+    BKN023 & TIL16 \\
+    BKN025 & TIL17 \\
+    BKN026 & \\
+    BKN027 & \\
+    BKN029 & \\
+    BKN030 & \\
+    BKN031 & \\
+    BKN032 & \\
+    BKN033 & \\
+    BKN034 & \\
+    BKN035 & \\
+    BKN040 & \\ \hline
+  \end{tabular}
+  \end{center}
+  \caption{A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by \cite{chia2012}   } \label{sTab:list}
+\end{table}
+\clearpage
+
+\begin{table}
+  \begin{center}
+  \def\arraystretch{2}
+%  \begin{center}
+  \begin{tabular}{l|c|c|c}
+    \bf Parameter & \bf Initial value & \bf Upper bound & \bf Lower bound\\ \hline 
+    $\frac{N_b}{N_a}$ & 0.02 & $1\times10^{-7}$ & 2 \\ 
+    $\frac{N_{m}}{N_a}$ & 3 & $1\times10^{-7}$ & 200 \\
+    $\frac{T_b}{2N_a}$ & 0.04 & 0  & 1 \\ 
+    $\frac{M_{mt}}{N_a}$ & $1\times10^{-10}$ & $1\times10^{-7}$ & 0.001 \\
+    $\frac{M_{tm}}{N_a}$ & $1\times10^{-10}$ & $1\times10^{-7}$ & 0.001 \\
+  \end{tabular}
+  \def\arraystretch{1} % undo stretching
+  \end{center}
+    \caption{ Parameters, initial values, and boundaries used for model-fitting with $\delta\alpha\delta{i}$.
+Parameters are shown in the units utilized by $\delta\alpha\delta{i}$, although in the text simplified units are reported.
+}\label{sTab:dadi}
+\end{table}
 

### 2016-04-12 NatPlant2/manuscript.tex
updates

--- 2016-02-25 NatPlant2/manuscript.tex
@@
 In contrast to the stochastic effects of drift, selection systematically alters allele frequencies by favoring particular alleles at the expense of others as a result of their effects on fitness.
-Researchers often study drift by excluding potentially selected sites \cite{voight2005, luikart2003, gutenkunst2009}, or selection by focusing on site-specific patterns under the assumption that genome-wide diversity reflects primarily the action of drift \cite{akey2009}.
+Researchers often study drift by excluding potentially selected sites \cite{voight2005, gutenkunst2009}, or selection by focusing on site-specific patterns under the assumption that genome-wide diversity reflects primarily the action of drift \cite{akey2009}.
 
-Drift and selection do not operate independently to determine genetic variability, however, in large part because linkage allows the effects of selection to be wide-ranging \cite{smith1974, li2012,slotte2014}.
+Drift and selection do not operate independently to determine genetic variability, however, in large part because linkage allows the effects of selection to be wide-ranging \cite{smith1974, slotte2014}.
 Linked selection, which refers to the effects of selection at one site on diversity at linked sites \cite{slotte2014}, can take the form of hitch-hiking, when the frequency of a neutral allele changes as a result of positive selection at a physically linked site \cite{smith1974}, or background selection, where diversity is reduced at loci linked to a site undergoing selection against deleterious alleles \cite{charlesworth1993}.
@@
 The impact of linked selection, in turn, is heavily influenced by the effective population size ($N_e$), as the  efficiency of natural selection is proportional to the product $N_es$, where $s$ is the strength of selection on a variant \cite{cutter2013, slotte2014, corbett2015,leffler2012}.
-The effective size of a population is not static, and nearly all species, including flies \cite{duchen2013}, humans \cite{reich1998}, domesticates \cite{hyten2006, bovine2009}, and non-model species \cite{ellegren2014} have experienced recent or ancient changes in $N_e$.
+The effective size of a population is not static, and nearly all species, including flies \cite{duchen2013}, humans \cite{reich1998}, domesticates \cite{bovine2009}, and non-model species \cite{ellegren2014} have experienced recent or ancient changes in $N_e$.
 Although much is known about how the long-term average $N_e$ affects linked selection \cite{cutter2013}, relatively little is understood about the immediate effects of more recent changes in $N_e$ on patterns of linked selection.
@@
 Because of its relatively simple demographic history and well-developed genomic resources, maize (\emph{Zea mays}) represents an excellent organism to study these effects. 
-Archaeological and genetic studies have established that maize domestication began in Central Mexico at least 9,000 years bp \cite{smith1995,matsuoka2002}, and involved a population bottleneck followed by recent expansion \cite{wright2005,eyre1998,tenaillon2004}.
+Archaeological and genetic studies have established that maize domestication began in Central Mexico at least 9,000 years bp \cite{matsuoka2002}, and involved a population bottleneck followed by recent expansion \cite{wright2005,eyre1998,tenaillon2004}.
 Because of this simple but dynamic demographic history, domesticated maize and its wild ancestor teosinte can be used to understand the effects of changing $N_e$ on linked selection. 
@@
 If a substantial proportion of missense mutations have been fixed due to hard sweeps, diversity around these substitutions should be lower than around synonymous substitutions. 
-We observe this pattern around the causative amino acid substitution in the maize domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2005, wang2015}.
+We observe this pattern around the causative amino acid substitution in the maize domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2015}.
 Genome-wide, however, we observe no differences in diversity at sites near synonymous versus missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
@@
 We investigated purifying selection in maize and teosinte by evaluating the reduction of diversity around genes.
-Pairwise diversity is strongly reduced within genes for both maize and teosinte (Figure \ref{fig:purify}A) but recovers quickly at sites outside of genes, consistent with the low levels of linkage disequilibrium generally observed in these subspecies \cite{tenaillon2002,chia2012}.
+Pairwise diversity is strongly reduced within genes for both maize and teosinte (Figure \ref{fig:purify}A) but recovers quickly at sites outside of genes, consistent with the low levels of linkage disequilibrium generally observed in these subspecies \cite{chia2012}.
 The reduction in relative diversity is more pronounced in teosinte, reaching lower levels in genes and occurring over a wider region.
 
-Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple beneficial mutations or standing genetic variation, because in such cases diversity around the substitution may be reduced to a lesser degree \cite{innan2004,messer2013}.
+Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple beneficial mutations or standing genetic variation, because in such cases diversity around the substitution may be reduced to a lesser degree \cite{messer2013}.
 Nonetheless, such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation to purifying selection for the observed reduction of diversity at linked sites in genes.
@@
 Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability.
-Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with soft sweeps from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}.
+Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while published data for several loci are consistent with soft sweeps from standing variation \cite{studer2011} or multiple mutations \cite{wills2013}.
 Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{van2012historical, beissinger2014} all support the importance of standing variation as primary sources of adaptive variation.
@@
 
-The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
+The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011}.
 Comparisons of the estimated percentages of nonsynonmyous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
@@
 One possibility, for example, is that, if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
-Support for this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the smaller \emph{Arabidopsis} genome.
+Support for this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the smaller \emph{Arabidopsis} genome.
 Our results, for example, are not dissimilar to findings in the comparably-sized mouse genome, where no differences are seen in diversity around nonsynonymous and synonymous substitutions in spite of a large $N_e$ and  as many as 80\% of adaptive substitutions occuring outside of genes\cite{Halligan:2013ir}.
@@
 Recent population bottlenecks and expansions have increased the relative abundance of rare and 
-deleterious variants in domesticated plants \cite{Gunther:2010gt,renaut2015} and human populations out of Africa \cite{keinan2012,coventry2010}, and such variants may play an important role in phenotypic variation \cite{Mezmouk:2014es, coventry2010, eyre2010}.
-Nonetheless, demographic history may have little impact on the overall genetic load of populations \cite{do2015,simons2014}, as decreases in $N_e$ that allow weakly deleterious variants to escape selection also help purge strongly deleterious ones, and the increase of new deleterious mutations in expanding populations is mitigated by their lower initial frequency and the increasing efficiency of purifying selection \cite{gazave2013, simons2014, lohmueller2014}.
+deleterious variants in domesticated plants \cite{renaut2015} and human populations out of Africa \cite{keinan2012,coventry2010}, and such variants may play an important role in phenotypic variation \cite{Mezmouk:2014es, coventry2010, eyre2010}.
+Nonetheless, demographic history may have little impact on the overall genetic load of populations \cite{do2015,simons2014}, as decreases in $N_e$ that allow weakly deleterious variants to escape selection also help purge strongly deleterious ones, and the increase of new deleterious mutations in expanding populations is mitigated by their lower initial frequency and the increasing efficiency of purifying selection \cite{simons2014, lohmueller2014}.
 
@@
 Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
-While a number of authors have tested for selection using methods that explicitly incorporate or are robust to demographic change \cite{eyre2009estimating, chen2010, zeng2010} and others have compared estimates of the efficiency of adaptive and purifying selection across species \cite{popadin2012} or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size \cite{corbett2015,leffler2012}.
+While a number of authors have tested for selection using methods that explicitly incorporate or are robust to demographic change \cite{eyre2009estimating, zeng2010} and others have compared estimates of the efficiency of adaptive and purifying selection across species \cite{popadin2012} or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size \cite{corbett2015,leffler2012}.
 Our results show that demographic change over short periods of time can quickly change the dynamics of linked selection: mutations arising in extant maize populations are much more strongly impacted by the effects of selection on linked sites than would be suggested by analyses using long-term effective population size.
@@
 
+\section*{ACKNOWLEDGEMENTS}
+We are indebted to Graham Coop and Simon Aeschbacher for their constructive input during this study.
+We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.
+Funding was provided by NSF Plant Genome Research Project 1238014 and the USDA-Agricultural Research Service.
+
+\section*{AUTHOR CONTRIBUTIONS}
+TMB and JRI devised this study. TMB, LW, JRI, and KC analyzed the data. AD performed early-stage simulations. TMB, JRI, and MBH wrote the manuscript.
+
+\section*{COMPETING INTERESTS STATEMENT}
+The authors declare no competing financial interests.
+
 {\scriptsize \sf
@@
 \bibliographystyle{naturemag}
-%\bibliographystyle{NatureSeriesT}
 \bibliography{Reference}

### 2016-04-12 NatPlant2/manuscript.tex
updates

@@
 Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while published data for several loci are consistent with soft sweeps from standing variation \cite{studer2011} or multiple mutations \cite{wills2013}.
-Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{van2012historical, beissinger2014} all support the importance of standing variation as primary sources of adaptive variation.
+Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{van2012historical, beissinger2014,hirsch2014} all support the importance of standing variation as primary sources of adaptive variation.
 Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013}.

### 2016-04-13 NatPlant2/manuscript.tex
minor changes

--- 2016-04-12 NatPlant2/manuscript.tex
@@
 One possibility, for example, is that, if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
-Support for this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the smaller \emph{Arabidopsis} genome.
-Our results, for example, are not dissimilar to findings in the comparably-sized mouse genome, where no differences are seen in diversity around nonsynonymous and synonymous substitutions in spite of a large $N_e$ and  as many as 80\% of adaptive substitutions occuring outside of genes\cite{Halligan:2013ir}.
+Support for this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the smaller \emph{Arabidopsis} genome \cite{hancock2011adaptation}.
+Our results, for example, are not dissimilar to findings in the comparably-sized mouse genome, where no differences are seen in diversity around nonsynonymous and synonymous substitutions in spite of a large $N_e$ and  as many as 80\% of adaptive substitutions occurring outside of genes\cite{Halligan:2013ir}.
 Future comparative analyses using a common statistical framework (e.g.
@@
 
-\subsection*{Demography influences the efficiency of purifying selection} %JRI 
+\subsection*{Demography influences the efficiency of purifying selection} 
 One of our more striking findings is that the impact of purifying selection on maize and teosinte qualitatively changed over time. 
@@
 Recent population bottlenecks and expansions have increased the relative abundance of rare and 
-deleterious variants in domesticated plants \cite{renaut2015} and human populations out of Africa \cite{keinan2012,coventry2010}, and such variants may play an important role in phenotypic variation \cite{Mezmouk:2014es, coventry2010, eyre2010}.
+deleterious variants in domesticated plants \cite{Gunther:2010gt,renaut2015} and human populations out of Africa \cite{keinan2012,coventry2010}, and such variants may play an important role in phenotypic variation \cite{Mezmouk:2014es, coventry2010, eyre2010}.
 Nonetheless, demographic history may have little impact on the overall genetic load of populations \cite{do2015,simons2014}, as decreases in $N_e$ that allow weakly deleterious variants to escape selection also help purge strongly deleterious ones, and the increase of new deleterious mutations in expanding populations is mitigated by their lower initial frequency and the increasing efficiency of purifying selection \cite{simons2014, lohmueller2014}.
@@
 We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.
-Funding was provided by NSF Plant Genome Research Project 1238014 and the USDA-Agricultural Research Service.
+Funding was provided by NSF Plant Genome Research Project 1238014, the USDA-Agricultural Research Service, and USDA Hatch project CA-D-PLS-2066-H.
 

### 2016-04-13 NatPlant2/manuscript.tex
last ref

@@
 The impact of linked selection, in turn, is heavily influenced by the effective population size ($N_e$), as the  efficiency of natural selection is proportional to the product $N_es$, where $s$ is the strength of selection on a variant \cite{cutter2013, slotte2014, corbett2015,leffler2012}.
-The effective size of a population is not static, and nearly all species, including flies \cite{duchen2013}, humans \cite{reich1998}, domesticates \cite{bovine2009}, and non-model species \cite{ellegren2014} have experienced recent or ancient changes in $N_e$.
+The effective size of a population is not static, and nearly all species, including flies \cite{duchen2013}, humans \cite{reich1998}, domesticates \cite{meyer2013}, and non-model species \cite{ellegren2014} have experienced recent or ancient changes in $N_e$.
 Although much is known about how the long-term average $N_e$ affects linked selection \cite{cutter2013}, relatively little is understood about the immediate effects of more recent changes in $N_e$ on patterns of linked selection.

### 2016-04-15 NatPlant2/SupportingInfoFinal.tex
final submission

--- 2016-04-12 NatPlant2/manuscript.tex
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{More than one Author with different Affiliations}
-\author[1,2,3,*]{Timothy M. Beissinger}
-\author[4]{Li Wang}
-\author[1]{Kate Crosby}
-\author[1]{Arun Durvasula}
-\author[4]{Matthew B. Hufford}
-\author[1,5,*]{Jeffrey Ross-Ibarra}
-\affil[1]{Dept. of Plant Sciences, University of California, Davis, CA, USA}
-\affil[2]{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}
-\affil[3]{Division of Plant Sciences, University of Missouri, Columbia, MO, USA}
-\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, USA}
-\affil[5]{Genome Center and Center for Population Biology, University of California, Davis, CA, USA}
-\affil[*]{Correspondence may be addressed to rossibarra@ucdavis.edu or beissingert@missouri.edu}
-
-\renewcommand\Authands{ and }
-
 \begin{document}
 
-\title{Recent demography drives changes in linked selection across the maize genome}
-
-\maketitle
-%\cofeAm{1}{1.0}{0}{5.5cm}{3cm}
-
-\begin{abstract}
-Genetic diversity is shaped by the interaction of drift and selection, but the details of this interaction are not well understood.
-The impact of genetic drift in a population is largely determined by its demographic history, typically summarized by its long-term effective population size ($N_e$ ).
-Rapidly changing population demographics complicate this relationship, however.
-To better understand how changing demography impacts selection, we used whole-genome sequencing data to investigate patterns of linked selection in domesticated and wild maize (teosinte).
-We produce the first whole-genome estimate of the demography of maize domestication, showing that maize was reduced to approximately 5\% the population size of teosinte before it experienced rapid expansion post-domestication to population sizes much larger than its ancestor.
-Evaluation of patterns of nucleotide diversity in and near genes shows little evidence of selection on beneficial amino acid substitutions, and that the domestication bottleneck led to a decline in the efficiency of purifying selection in maize.
-Young alleles, however, show evidence of much stronger purifying selection in maize, reflecting the much larger effective size of present day populations.
-Our results demonstrate that recent demographic change --- a hallmark of many species including both humans and crops ---  can have immediate and wide-ranging impacts on diversity that conflict with would-be expectations based on $N_e$ alone.
-\end{abstract}
-
-The genetic diversity of populations is determined by a constant interplay between genetic drift and natural selection. 
-Drift is a consequence of a finite population size and the random sampling of gametes each generation \cite{dobzhansky1957}. 
-In contrast to the stochastic effects of drift, selection systematically alters allele frequencies by favoring particular alleles at the expense of others as a result of their effects on fitness.
-Researchers often study drift by excluding potentially selected sites \cite{voight2005, gutenkunst2009}, or selection by focusing on site-specific patterns under the assumption that genome-wide diversity reflects primarily the action of drift \cite{akey2009}.
-
-Drift and selection do not operate independently to determine genetic variability, however, in large part because linkage allows the effects of selection to be wide-ranging \cite{smith1974, slotte2014}.
-Linked selection, which refers to the effects of selection at one site on diversity at linked sites \cite{slotte2014}, can take the form of hitch-hiking, when the frequency of a neutral allele changes as a result of positive selection at a physically linked site \cite{smith1974}, or background selection, where diversity is reduced at loci linked to a site undergoing selection against deleterious alleles \cite{charlesworth1993}.
-Recent work in \emph{Drosophila}, for example, has shown that virtually the entire genome is impacted by the combined effects of these processes \cite{sella2009,elyashiv2014,andolfatto2005}.
-
-The impact of linked selection, in turn, is heavily influenced by the effective population size ($N_e$), as the  efficiency of natural selection is proportional to the product $N_es$, where $s$ is the strength of selection on a variant \cite{cutter2013, slotte2014, corbett2015,leffler2012}.
-The effective size of a population is not static, and nearly all species, including flies \cite{duchen2013}, humans \cite{reich1998}, domesticates \cite{bovine2009}, and non-model species \cite{ellegren2014} have experienced recent or ancient changes in $N_e$.
-Although much is known about how the long-term average $N_e$ affects linked selection \cite{cutter2013}, relatively little is understood about the immediate effects of more recent changes in $N_e$ on patterns of linked selection.
-
-Because of its relatively simple demographic history and well-developed genomic resources, maize (\emph{Zea mays}) represents an excellent organism to study these effects. 
-Archaeological and genetic studies have established that maize domestication began in Central Mexico at least 9,000 years bp \cite{matsuoka2002}, and involved a population bottleneck followed by recent expansion \cite{wright2005,eyre1998,tenaillon2004}.
-Because of this simple but dynamic demographic history, domesticated maize and its wild ancestor teosinte can be used to understand the effects of changing $N_e$ on linked selection. 
-In this study, we leverage the maize-teosinte system to study these effects by first estimating the parameters of the maize domestication bottleneck using whole-genome resequencing data and then investigating the relative importance of different forms of linked selection on diversity in the ancient and more recent past.
-We show that, while patterns of overall nucleotide diversity reflect long-term differences in $Ne$, recent growth following domestication qualitatively changes these effects, thereby illustrating the importance of a comprehensive understanding of demography when considering the effects of selection genome-wide.
-
-\section*{RESULTS}
-
-\subsection*{Patterns of diversity differ between genic and  intergenic regions of the genome} %JRI
-To investigate how demography and linked selection have shaped patterns of diversity in maize and teosinte, we analyzed data from 23 maize and 13 teosinte genomes from the maize HapMap 2 and HapMap 3 projects \cite{chia2012, bukowski2015}.
-As a preliminary step, we evaluated levels of diversity inside and outside of genes across the genome.
-We find broad differences in genic and intergenic diversity consistent with earlier results  \cite{hufford2012}(Figure \ref{fig:diversity}).
-In maize, mean pairwise diversity ($\pi$) within genes was significantly lower than at sites at least 5 kb away from genes (0.00668 vs 0.00691, $p<2\times 10^{-44}$).
-Diversity differences in teosinte are even more pronounced (0.0088 vs. 0.0115, $p\approx 0$). 
-Differences were also apparent in the site frequency spectrum, with mean Tajima's D positive in genic regions in both maize (0.4) and teosinte (0.013) but negative outside of genes (-0.087 in maize and -0.25 in teosinte, $p\approx 0$ for both comparisons).
-These observations suggest that diversity in genes is not evolving neutrally, but instead is reduced by the impacts of selection on linked sites.
-
-\subsection*{Demography of maize domestication} %JRI
-We next estimated a demographic model of maize domestication (Figure \ref{fig:bottleneck}). 
-To minimize the impact of selection on our estimates \cite{ewing2016}, we only included sites \textgreater 5kb from genes.  
-The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ teosinte individuals.
-We estimate that maize split from teosinte $\approx 15,000$ generations in the past, with an initial size of only $\approx 5\% $ of the ancestral $N_a$. 
-After its split from teosinte, our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,000$.
-Although our model provides only a rough approximation of migration rates, we included migration parameters during demographic inference because omitting these could bias our population size estimates.
-We observe that maize and teosinte have continued to exchange migrants after the population split, with gene flow from teosinte to maize was $M_{tm} =  1.1 \times 10^{-5} \times N_a $  migrants per generation, and from maize to teosinte we estimate $M_{mt} =  1.4 \times 10^{-5} \times N_a$ migrants per generation.
-
-Because our modest sample size of fully sequenced individuals has limited power to infer recent population expansion, we investigated two alternative approaches for demographic inference.  
-First, we utilized genotyping data from more than 4,000 maize landraces \cite{Hearne2015} to estimate the modern maize effective population size.
-Because rare variants provide the best information about recent effective population sizes \cite{keinan2012}, we estimate $N_e$ using  a singleton-based estimator \cite{fu1993} of the population mutation rate $\theta=4N_e\mu$  and published values of the mutation rate \cite{clark2005} (see online methods for details).
-This yields a much higher estimate of the modern maize effective population size at  $N_m \approx 993,000$.
-Finally, we employed a model-free coalescent approach \cite{schiffels2014} to estimate population size change using a subset of six genomes each of maize and teosinte.
-Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent with the other approaches, identifying population isolation beginning between 10,000 and 15,000 generations ago, a clear domestication bottleneck, and ultimately rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
-Our assessment of the historical demography of maize and teosinte provides context for subsequent analyses of linked selection.
-
-\begin{figure}[!tb]
-  \begin{center}
-  \includegraphics[width=.45\textwidth] {FigsAndFiles/Pi_and_Tajima.png}
-\end{center}
-\caption{\textbf{A} and \textbf{B} show mean pairwise diversity $\pi$, $\pm$ one standard deviation, while \textbf{C} and \textbf{D} depict and Tajima's D in 1kb windows from genic and nongenic regions of maize and teosinte.
-\label{fig:diversity} }
-\end{figure}
-
-\begin{figure}[!tb]
-%\vspace*{.05in}
-\centering
-\includegraphics[width=.5\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
-\caption{Parameter estimates for a basic bottleneck model of maize domestication. See methods for details. \label{fig:bottleneck} }
-\end{figure}
-
-\subsection*{Hard sweeps do not explain diversity differences} %JRI
-When selection increases the frequency of a new beneficial mutation, a signature of reduced diversity is left at surrounding linked sites \cite{smith1974}.
-To evaluate whether patterns of such ``hard sweeps'' could explain observed differences in diversity between genic and intergenic regions of the genome, we compared diversity around missense and synonymous substitutions between \emph{Tripsacum} and either maize or teosinte.
-If a substantial proportion of missense mutations have been fixed due to hard sweeps, diversity around these substitutions should be lower than around synonymous substitutions. 
-We observe this pattern around the causative amino acid substitution in the maize domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2015}.
-Genome-wide, however, we observe no differences in diversity at sites near synonymous versus missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
-
-%\mbh{I'm curious whether linked selection is affecting the diversity around a sizable proportion of synonymous substitutions.
-Would it be worthwhile to look at diversity around the subset of synonymous substitutions that are the least linked to missense substitutions just to make sure it is not substantially higher?
-Or perhaps generate a plot of diversity around synonymous substitutions vs.
-distance of these subs from missense subs?}
-
-\begin{figure*}[tb]
-%\vspace*{.05in}
-\centering
-\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_Aug}
-\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_Aug}
-\caption{Pairwise diversity surrounding synonymous and missense  substitutions in {\bf A} maize and {\bf B} teosinte.
-Axes show absolute diversity values (right) and values relative to mean nucleotide diversity in windows $\geq 0.01 cM$ from a substitution (left).
-Lines depict a loess curve (span of 0.01) and shading represents bootstrap-based 95\% confidence intervals.
-Inset plots depict a larger range on the x-axis.
-\label{fig:hardSweeps}}
-\end{figure*}
-
-Previous analyses have suggested that this approach may have limited power because a relatively high proportion of missense substitutions will be found in genes that, due to weak purifying selection, have higher genetic diversity \cite{enard2014}.
-To address this concern, we took advantage of genome-wide estimates of evolutionary constraint \cite{rodgers2015} calculated using genomic evolutionary rate profile (GERP) scores \cite{davydov2010}. 
-We then evaluated substitutions only in subsets of genes in the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection. 
-As expected, we see higher diversity around substitutions in genes under weak purifying selection, but we still find no difference in diversity near synonymous and missense substitutions in either subset of the data (Figure \ref{sFig:consUncons}).
-Taken together, these data suggest hard sweeps do not play a major role in patterning genic diversity in either maize or teosinte.
-
-\subsection*{Diversity is strongly influenced by purifying selection} %JRI
-In the case of purifying or background selection, diversity is reduced in functional regions of the genome via removal of deleterious mutations \cite{charlesworth1993}.
-We investigated purifying selection in maize and teosinte by evaluating the reduction of diversity around genes.
-Pairwise diversity is strongly reduced within genes for both maize and teosinte (Figure \ref{fig:purify}A) but recovers quickly at sites outside of genes, consistent with the low levels of linkage disequilibrium generally observed in these subspecies \cite{chia2012}.
-The reduction in relative diversity is more pronounced in teosinte, reaching lower levels in genes and occurring over a wider region.
-
-Our previous comparison of synonymous and missense substitutions has low power to detect the effects of selection acting on multiple beneficial mutations or standing genetic variation, because in such cases diversity around the substitution may be reduced to a lesser degree \cite{messer2013}.
-Nonetheless, such ``soft sweeps'' are still expected to occur more frequently in functional regions of the genome and could provide an alternative explanation to purifying selection for the observed reduction of diversity at linked sites in genes.
-To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
-Qualitative differences between maize and teosinte in patterns of diversity within and outside of genes remained unchanged even after removing genes in the top 20\% quantile of H12 (Figure \ref{sFig:H12}A).
-%\mbh{This makes me wonder about the top 20\%. Is the pattern of diversity reduction different in this fraction of the genome?  If so, that seems like something that should be mentioned.}
-%TB Says: Sites in the top 20% of H12 are unlikely to be under background selection, so maize vs teo background selection pattern at these sites seems strange to investigate.
-We interpret these combined results as suggesting that purifying selection has predominantly shaped diversity near genes and left a more pronounced signature in the teosinte genome due to the increased efficacy of selection resulting from differences in long-term effective population size.
-
-\begin{figure*}[!tb]
-%\vspace*{.05in}
-\centering
-\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_Downsampled_threeLines_manuscript.png}
-\caption{Relative diversity versus distance to nearest gene in maize and teosinte. 
-Shown are \textbf{A} pairwise nucleotide diversity and \textbf{B} singleton diversity.  
-Relative diversity is calculated compared to the mean diversity in windows $\geq 0.01 cM$ or $\geq 0.02 cM$ from the nearest
-gene for pairwise diversity and singletons, respectively. 
-  Lines depict cubic smoothing splines with smoothing parameters chosen via generalized cross validation and shading depicts bootstrap-based 95\% confidence intervals.
-  Inset plots depict a smaller range on the x-axis. \label{fig:purify}
-  }
-\end{figure*}
-
-\subsection*{Population expansion leads to stronger purifying selection in modern maize} %JRI
-Motivated by the rapid post-domestication expansion of maize evident in our demographic analyses, we reasoned that low-frequency --- and thus younger --- polymorphisms might show patterns distinct from pairwise diversity, which is determined primarily by intermediate frequency --- therefore comparably older --- alleles.
-Singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) appears nearly identical to results from pairwise diversity (Figure \ref{fig:hardSweeps}), providing little support for a substantial recent increase in the number or strength of hard sweeps occurring in maize.
-
-In contrast, we observe a significant shift in the effects of purifying selection: singleton polymorphisms are more strongly reduced in and near genes in maize than in teosinte, even after downsampling our maize data to account for differences in sample size (Figure \ref{fig:purify}B).
-This result is the opposite of the pattern observed for $\pi$, where teosinte demonstrated a stronger reduction of diversity in and around genes than did maize. 
-As before, this relationship remained after we removed the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12}).
-While direct comparison of pairwise and singleton diversity within taxa is consistent with non-equilibrium dynamics in teosinte, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}) and mirror results from simulations of purifying selection (Figure \ref{sFig:sims}).
-
-\section*{DISCUSSION}
-
-\subsection*{Demography of domestication} %JRI
-Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.
-We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites.
-This implies that genic polymorphism data are unable to tell the complete or accurate demographic history of maize, but the rapid recovery of diversity outside of genes demonstrates that sites far from genes can be reasonably used for demographic inference.
-Furthermore, by utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without making assumptions about its duration.
-This model paves the way for future work on the demography of domestication, evaluating for example the significance of differences in gene flow estimated here or removing assumptions about demographic history in teosinte.
-
-One surprising result from our model is the estimated divergence time of maize and teosinte approximately $15,000$ generations before present.
-While this appears to conflict with archaeological estimates \cite{piperno2009}, we emphasize that this estimate reflects the fact that the genetic split between populations likely preceded anatomical changes that can be identified in the archaeological record.
-We also note that our result may be inflated due to population structure, as our geographically diverse sample of teosinte may include populations diverged from those that gave rise to maize.
-
-The estimated bottleneck of $\approx 5\%$ of the ancestral teosinte population seems low given that maize landraces exhibit $\approx 80\%$ of the diversity of teosinte \cite{hufford2012}, but our model suggests that the effects of the bottleneck on diversity are likely ameliorated by both gene flow and rapid population growth (Figure \ref{fig:bottleneck}).
-Although we estimate that the modern effective size of maize is larger than teosinte, the small size of our sample reduces our power to identify the low frequency alleles most sensitive to rapid population growth \cite{keinan2012}, and our model is unable to incorporate growth faster than exponential.
-Both alternative approaches we employ estimate a much larger modern effective size of maize in the range of $\approx 10^6 - 10^9$, an order of magnitude or more than the current size of teosinte.
-Census data suggest these estimates are plausible: there are 47.9 million ha of open-pollinated maize in production \cite{cimmyt1999}, likely planted at a density of $\approx 25,000$ individuals per hectare \cite{baden2001}.
-Assuming the effective size is only $\approx 0.4\%$ of the census size (i.e. 1 ear for every 1000 male plants), this still implies a modern effective population size of more than four billion.
-While these genetic and census estimates are likely inaccurate, all of the evidence points to the fact that the modern effective size of maize is extremely large.
-
-\subsection*{Hard sweeps do not shape genome-wide diversity in maize} %JRI
-Our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of purifying selection  \cite{enard2014}.
-Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability.
-Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while published data for several loci are consistent with soft sweeps from standing variation \cite{studer2011} or multiple mutations \cite{wills2013}.
-Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{van2012historical, beissinger2014,hirsch2014} all support the importance of standing variation as primary sources of adaptive variation.
-Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013}.
-Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.]{studer2011}.
-Indeed, many adaptive traits in both maize \cite{wallace2014} and teosinte \cite{weber2008} are highly quantitative,  and adaptation in both maize \cite{hufford2012} and teosinte \cite{pyhajarvi2013} has involved selection on regulatory variation.
-
-The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011}.
-Comparisons of the estimated percentages of nonsynonmyous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
-While differences in long-term $N_e$ likely explains some of the observed variation across species, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}), perhaps suggesting other factors may contribute to these differences as well.
-One possibility, for example, is that, if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
-Support for this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the smaller \emph{Arabidopsis} genome.
-Our results, for example, are not dissimilar to findings in the comparably-sized mouse genome, where no differences are seen in diversity around nonsynonymous and synonymous substitutions in spite of a large $N_e$ and  as many as 80\% of adaptive substitutions occuring outside of genes\cite{Halligan:2013ir}.
-Future comparative analyses using a common statistical framework (e.g.
-\cite{corbett2015} )and considering additional ecological and life history factors (c.f.
-\cite{leffler2012}) should allow explicit testing of this idea.
-
-\subsection*{Demography influences the efficiency of purifying selection} %JRI 
-One of our more striking findings is that the impact of purifying selection on maize and teosinte qualitatively changed over time. 
-We observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), but the opposite trend when we evaluate diversity using singleton polymorphisms (Figure \ref{fig:purify}B).
-The efficiency of purifying selection is proportional to effective population size \cite{kimura1984}, and these results are thus consistent with our demographic analyses which show a domestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} followed by recent rapid expansion and a much larger modern $N_e$.
-Simple foward-in-time population genetic simulations qualitatively confirm these results, and further suggest that the observed patterns are likely cause by sites under relatively weak purifying selection \ref{sFig:sims}.
-
-Although demographic change affects the efficiency of purifying selection, it may have limited implications for genetic load. 
-Recent population bottlenecks and expansions have increased the relative abundance of rare and 
-deleterious variants in domesticated plants \cite{renaut2015} and human populations out of Africa \cite{keinan2012,coventry2010}, and such variants may play an important role in phenotypic variation \cite{Mezmouk:2014es, coventry2010, eyre2010}.
-Nonetheless, demographic history may have little impact on the overall genetic load of populations \cite{do2015,simons2014}, as decreases in $N_e$ that allow weakly deleterious variants to escape selection also help purge strongly deleterious ones, and the increase of new deleterious mutations in expanding populations is mitigated by their lower initial frequency and the increasing efficiency of purifying selection \cite{simons2014, lohmueller2014}.
-
-\section*{Rapid changes in linked selection} %JRI
-Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
-While a number of authors have tested for selection using methods that explicitly incorporate or are robust to demographic change \cite{eyre2009estimating, zeng2010} and others have compared estimates of the efficiency of adaptive and purifying selection across species \cite{popadin2012} or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size \cite{corbett2015,leffler2012}.
-Our results show that demographic change over short periods of time can quickly change the dynamics of linked selection: mutations arising in extant maize populations are much more strongly impacted by the effects of selection on linked sites than would be suggested by analyses using long-term effective population size.
-As many natural and domesticated populations have undergone considerable demographic change in their recent past, long-term comparisons of $N_e$ are likely not informative about current processes affecting allele frequency trajectories.
-
-\section*{METHODS}
-\subsection*{BASH, R, and Python scripts}
-All scripts used for analysis are available in an online repository at \url{https://github.com/timbeissinger/Maize-Teo-Scripts}.
-
-\subsection*{Plant materials}
-We made use of published sequences from inbred accessions of teosinte (Z.
-{\it mays} ssp.
-{\it parviglumis}) and maize landraces from the Maize HapMap3 panel as part of the Panzea project  \cite{chia2012, lemmon2014, bukowski2015}.
-From these data, we removed 4 teosinte individuals that were not ssp.
-\textit{parviglumis} or appeared as outliers in an initial principal component analysis conducted with the package adegenet \cite{jombart2011} (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table \ref{sTab:list}).
-We also utilized a single individual of (\textit{Tripsacum dactyloides}) as an outgroup.
-All bam files are available at \url{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}.
-
-\subsection*{Physical and genetic maps}
-Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (\url{ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/}) as described by \cite{bukowski2015}.
-All analyses made use of uniquely mapping reads with mapping quality score $\geq  30$ and bases with base quality score $\geq 20$; quality scores around indels were adjusted following \cite{li2011statistical}.
-We converted physical coordinates to genetic coordinates via linear interpolation of the previously published 1cM resolution NAM genetic map \cite{glaubitz2014}.
-
-\subsection*{Estimating the site frequency spectrum}
-We estimated both the genome-wide site frequency spectrum (SFS) as well as a separate SFS for genic (within annotated transcript) and intergenic ($\geq 5kb$ from a transcript) regions. 
-We used the biomaRt package \cite{durinck2009,durinck2005} of R \cite{R2014} to parse annotations from genebuild version 5b of AGPv3. 
-We estimated single population and joint SFS with the software ANGSD \cite{korneliussen2014}, including all positions with at least one aligned read in $\geq 80\%$ of samples in one or both populations.
-We assumed individuals were fully inbred and treated each line as a single haplotype.
-Because ANGSD cannot calculate a folded joint SFS, we first polarized SNPs using the maize reference genome and then folded spectra using $\delta$a$\delta$i \cite{gutenkunst2009}.
-
-\subsection*{Demographic inference}
-We used the software $\delta$a$\delta$i \cite{gutenkunst2009} to estimate parameters of a domestication bottleneck from the joint maize-teosinte SFS, using only sites $>5 kb$ from a gene to ameliorate the effects of linked selection.
-To minimize the number of parameters estimated, we employed a simple demographic model which posits a teosinte population of constant effective size $N_a$.
-At time $T_b$ generations in the past, this population gave rise to a maize population of size $N_b$ which grew exponentially to size $N_m$ in the present (Figure \ref{fig:bottleneck}).
-The model includes migration of $M_{mt}$ individuals each generation from maize to teosinte and $M_{tm}$ individuals from teosinte to maize.
-We estimated $N_a$ using $\delta$a$\delta$i's estimation of $\theta=4N_a\mu$ from the data and a mutation rate of $\mu = 3 \times 10^{-8}$ \cite{clark2005}.
-We estimated all other parameters using 1,000 $\delta$a$\delta$i optimizations and allowing initial values between runs to be randomly perturbed by a factor of 2.  
-Optimized parameters along with their initial values and upper and lower bounds can be found in table \ref{sTab:dadi}.
-We report parameter estimates from the optimization run with the highest log-likelihood.
-
-We further made use of a large genotyping data set of more than 4,000 partially imputed maize landraces \cite{Hearne2015} to estimate the modern maize $N_e$ from singleton counts. 
-We filtered these data to include only SNPs with data in $\geq 1,500$ individuals, and then projected the SFS down to a sample of 500 individuals by sampling each marker without replacement 1,000 times according to the observed allele frequencies.
-We then estimated $N_e$ from the data assuming $\mu = 3 \times 10^{-8}$ \cite{clark2005} and the relation  $4N_e\mu = \frac{S}{L}$ \cite{fu1993}, where $S$ is the total number of singleton SNPs and $L$ is the total number of SNPs in the dataset.
-
-As a final estimate of demography, we employed MSMC \cite{schiffels2014}  to complement our model-based demographic inference. 
-We used six each of maize and teosinte (BKN022, BKN025, BKN029, BKN030, BKN031, BKN033, TIL01, TIL03, TIL09, TIL10, TIL11 and TIL14), treating each inbred genome as a single haplotype.
-We called SNPs in ANGSD \cite{korneliussen2014} using a SNP p-value of $1e-6$ against a reference genome masked using SNPable (\url{http://lh3lh3.users.sourceforge.net/snpable.shtml}). 
-We then removed heterozygous genotypes and filtered sites with a mapping quality $<30$, a base quality $<20$, or a $|\textit{log}_2(\text{depth})|<1$.
-We ran MSMC with pattern parameter $20\times2+20\times4+10\times2$ (Figure \ref{sFig:msmc}A) for population size inference.
-To estimate the rate of cross-coalescence we used four maize and four teosinte haplotypes with pattern parameter $20\times1 + 20\times2$ (Figure \ref{sFig:msmc}B).
-
-\subsection*{Diversity}
-We made use of the software ANGSD \cite{korneliussen2014} for diversity calculations and genotype calling. 
-We calculated diversity statistics in maize and teosinte in 1 kb non-overlapping windows using filters as described above for the SFS. 
-We used allele counts to estimate the number of singleton polymorphisms in each window, and used binomial sampling to create a second maize data set down-sampled to have the same number of samples as teosinte.
-We called genotypes in maize, teosinte, and \textit{ Tripsacum} at sites with a SNP p-value $<10^{-6}$ and when the genotype posterior probability $>0.95$. 
-We identified substitutions in maize and teosinte as all sites with a fixed difference with \textit{Tripsacum} and $\leq 20\%$ missing data. 
-Substitutions were classified as synonymous, or missense using the ensembl variant effects predictor \cite{mclaren2010}.
-For each window with $\geq 100bp$ of data we computed the genetic distance between the window center and the nearest synonymous and missense substitution as well as the genetic distance to the center of the nearest gene transcript.
-
-\subsection*{Selection scan}
-We scanned the genome to identify sites that have experienced recent positive selection using the H12 statistic \cite{garud2015} in sliding windows of 200 SNPs with a step of 25 SNPs.
-
-\subsection*{Simulations}
-We used the program \emph{bneck\_selection\_ind} included in version 0.4.4 of the forward-in-time population genetic simulation library \emph{fwdpp} \citep[\url{https://github.com/molpopgen/fwdpp}[]{thornton2014genetics}.
-All simulations used a population mutation rate of $\theta=20$, a population recombination rate of $\rho=20$, and simulated 150,000 burn-in generations at an ancestral population size of $N_1=15,000$ to establish equilibrium, after which the population instantly changed to size $N_2$ and then grew exponentially for 1,000 generations to size $N_3$.
-To simulate a constant size population emulating teosinte, we set $N_2=N_3=15,000$.
-For maize we simulated a bottleneck similar to that estimated in Figure \ref{fig:bottleneck} by setting $N_2=750$,   followed by exponential growth to a large modern population size of $N_3=150,000$. 
-For each taxon, we performed 1,000 simulations for each of five values of the strength of purifying selection: $s=\{0,10^{-6},10^{-5},10^{-4},10^{-3}\}$. 
-All mutations were assumed to be codominant.
-To mimic nonsynonymous changes at a coding locus, we assumed that $\frac{3}{4}$ of mutations were selected.
-We calculated summary statistic across all sites using version 0.3.4 of \emph{msstats} (\url{https://github.com/molpopgen/msstats/releases}).
-
-\section*{ACKNOWLEDGEMENTS}
-We are indebted to Graham Coop and Simon Aeschbacher for their constructive input during this study.
-We thank Robert Bukowski and Qi Sun for providing early-access data from maize HapMap3.
-Funding was provided by NSF Plant Genome Research Project 1238014 and the USDA-Agricultural Research Service.
-
-\section*{AUTHOR CONTRIBUTIONS}
-TMB and JRI devised this study. TMB, LW, JRI, and KC analyzed the data. AD performed early-stage simulations. TMB, JRI, and MBH wrote the manuscript.
-
-\section*{COMPETING INTERESTS STATEMENT}
-The authors declare no competing financial interests.
-
-{\scriptsize \sf
-\renewcommand{\baselinestretch}{2.0}
-\bibliographystyle{naturemag}
-\bibliography{Reference}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%% Supporting info below this line %%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\onecolumn
 \section*{Supporting Information}
@@
 \begin{figure}[h]
-  \begin{center}
+  \begin{center}`
     \includegraphics[width=\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_tga1Supp_June.png}

### 2016-04-15 NatPlant2/manuscript.tex
final submission

@@
 
+%\renewcommand{\abstractname}{\vspace{-\baselineskip}}
+\renewcommand{\abstractname}{Abstract (opening paragraph)}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \renewcommand\Authands{ and }
+
+\newcounter{firstbib} % bibliography counter
 
@@
 Finally, we employed a model-free coalescent approach \cite{schiffels2014} to estimate population size change using a subset of six genomes each of maize and teosinte.
-Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent with the other approaches, identifying population isolation beginning between 10,000 and 15,000 generations ago, a clear domestication bottleneck, and ultimately rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure \ref{sFig:msmc}).
+Though this analysis suggests non-equilibrium dynamics for teosinte not included in our initial model, it is nonetheless broadly consistent with the other approaches, identifying population isolation beginning between 10,000 and 15,000 generations ago, a clear domestication bottleneck, and ultimately rapid population expansion in maize to an extremely large extant size of $\approx 10^9$ (Figure S2).
 Our assessment of the historical demography of maize and teosinte provides context for subsequent analyses of linked selection.
@@
 If a substantial proportion of missense mutations have been fixed due to hard sweeps, diversity around these substitutions should be lower than around synonymous substitutions. 
-We observe this pattern around the causative amino acid substitution in the maize domestication locus \emph{tga1} (Figure \ref{sFig:tga1}), likely the result of a hard sweep during domestication \cite{wang2015}.
+We observe this pattern around the causative amino acid substitution in the maize domestication locus \emph{tga1} (Figure S1), likely the result of a hard sweep during domestication \cite{wang2015}.
 Genome-wide, however, we observe no differences in diversity at sites near synonymous versus missense substitutions in either maize or teosinte (Figure \ref{fig:hardSweeps}).
@@
 We then evaluated substitutions only in subsets of genes in the highest and lowest 10\% quantile of mean GERP score, putatively representing genes under the strongest and weakest purifying selection. 
-As expected, we see higher diversity around substitutions in genes under weak purifying selection, but we still find no difference in diversity near synonymous and missense substitutions in either subset of the data (Figure \ref{sFig:consUncons}).
+As expected, we see higher diversity around substitutions in genes under weak purifying selection, but we still find no difference in diversity near synonymous and missense substitutions in either subset of the data (Figure S3).
 Taken together, these data suggest hard sweeps do not play a major role in patterning genic diversity in either maize or teosinte.
@@
 To test this possibility, we performed a genome-wide scan for selection using the H12 statistic, a method expected to be sensitive to both hard and soft sweeps \cite{garud2015}. 
-Qualitative differences between maize and teosinte in patterns of diversity within and outside of genes remained unchanged even after removing genes in the top 20\% quantile of H12 (Figure \ref{sFig:H12}A).
+Qualitative differences between maize and teosinte in patterns of diversity within and outside of genes remained unchanged even after removing genes in the top 20\% quantile of H12 (Figure S7A).
 %\mbh{This makes me wonder about the top 20\%. Is the pattern of diversity reduction different in this fraction of the genome?  If so, that seems like something that should be mentioned.}
@@
 Motivated by the rapid post-domestication expansion of maize evident in our demographic analyses, we reasoned that low-frequency --- and thus younger --- polymorphisms might show patterns distinct from pairwise diversity, which is determined primarily by intermediate frequency --- therefore comparably older --- alleles.
-Singleton diversity around missense and synonymous substitutions (Figure \ref{sFig:singleton}) appears nearly identical to results from pairwise diversity (Figure \ref{fig:hardSweeps}), providing little support for a substantial recent increase in the number or strength of hard sweeps occurring in maize.
+Singleton diversity around missense and synonymous substitutions (Figure S4) appears nearly identical to results from pairwise diversity (Figure \ref{fig:hardSweeps}), providing little support for a substantial recent increase in the number or strength of hard sweeps occurring in maize.
 
@@
 This result is the opposite of the pattern observed for $\pi$, where teosinte demonstrated a stronger reduction of diversity in and around genes than did maize. 
-As before, this relationship remained after we removed the 20\% of genes with the highest H12 values (Figure \ref{sFig:H12}).
-While direct comparison of pairwise and singleton diversity within taxa is consistent with non-equilibrium dynamics in teosinte, these too reveal much stronger differences in maize (Figure \ref{sFig:singletonPi}) and mirror results from simulations of purifying selection (Figure \ref{sFig:sims}).
+As before, this relationship remained after we removed the 20\% of genes with the highest H12 values (Figure S7).
+While direct comparison of pairwise and singleton diversity within taxa is consistent with non-equilibrium dynamics in teosinte, these too reveal much stronger differences in maize (Figure S5) and mirror results from simulations of purifying selection (Figure S6).
 
@@
 Comparisons of the estimated percentages of nonsynonmyous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
-While differences in long-term $N_e$ likely explains some of the observed variation across species, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}), perhaps suggesting other factors may contribute to these differences as well.
+While differences in long-term $N_e$ likely explains some of the observed variation across species, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure S4), perhaps suggesting other factors may contribute to these differences as well.
 One possibility, for example, is that, if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role.
@@
 The efficiency of purifying selection is proportional to effective population size \cite{kimura1984}, and these results are thus consistent with our demographic analyses which show a domestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} followed by recent rapid expansion and a much larger modern $N_e$.
-Simple foward-in-time population genetic simulations qualitatively confirm these results, and further suggest that the observed patterns are likely cause by sites under relatively weak purifying selection \ref{sFig:sims}.
+Simple foward-in-time population genetic simulations qualitatively confirm these results, and further suggest that the observed patterns are likely cause by sites under relatively weak purifying selection (Figure S6).
 
@@
 From these data, we removed 4 teosinte individuals that were not ssp.
-\textit{parviglumis} or appeared as outliers in an initial principal component analysis conducted with the package adegenet \cite{jombart2011} (Figure \ref{sFig:PCA}), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table \ref{sTab:list}).
+\textit{parviglumis} or appeared as outliers in an initial principal component analysis conducted with the package adegenet \cite{jombart2011} (Figure S8), leaving 13 teosinte and 23 maize that were used for all subsequent analyses (Table S1).
 We also utilized a single individual of (\textit{Tripsacum dactyloides}) as an outgroup.
-All bam files are available at \url{/iplant/home/shared/panzea/hapmap3/bam\_internal/v3\_bams\_bwamem}.
+All bam files are available from Cyverse at \url{/iplant/home/shared/data_commons/curated/Beissinger_MaizeTeo_2016}.
 
@@
 We estimated all other parameters using 1,000 $\delta$a$\delta$i optimizations and allowing initial values between runs to be randomly perturbed by a factor of 2.  
-Optimized parameters along with their initial values and upper and lower bounds can be found in table \ref{sTab:dadi}.
-We report parameter estimates from the optimization run with the highest log-likelihood.
+Optimized parameters along with their initial values and upper and lower bounds can be found in table S2. We report parameter estimates from the optimization run with the highest log-likelihood.
 
@@
 We then removed heterozygous genotypes and filtered sites with a mapping quality $<30$, a base quality $<20$, or a $|\textit{log}_2(\text{depth})|<1$.
-We ran MSMC with pattern parameter $20\times2+20\times4+10\times2$ (Figure \ref{sFig:msmc}A) for population size inference.
-To estimate the rate of cross-coalescence we used four maize and four teosinte haplotypes with pattern parameter $20\times1 + 20\times2$ (Figure \ref{sFig:msmc}B).
+We ran MSMC with pattern parameter $20\times2+20\times4+10\times2$ (Figure S2A) for population size inference.
+To estimate the rate of cross-coalescence we used four maize and four teosinte haplotypes with pattern parameter $20\times1 + 20\times2$ (Figure \S2B).
 
@@
 
+\section*{Correspondence}
+Correspondence may be addressed to Tim Beissinger (beissingert@missouri.edu) or Jeffrey Ross-Ibarra (rossibarra@ucdavis.edu)
+
 \section*{ACKNOWLEDGEMENTS}
@@
 
-{\scriptsize \sf
-\renewcommand{\baselinestretch}{2.0}
-\bibliographystyle{naturemag}
-\bibliography{Reference}
-}
+%{\scriptsize \sf
+%\renewcommand{\baselinestretch}{2.0}
+%\bibliographystyle{naturemag}
+%\bibliography{Reference}
+%}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%% Supporting info below this line %%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\renewcommand\refname{References (Main Text)}
 
-\onecolumn
-\section*{Supporting Information}
-\renewcommand\thefigure{S\arabic{figure}}
-\renewcommand\thetable{S\arabic{table}}    
-\setcounter{figure}{0}
-\setcounter{table}{0}
+\begin{thebibliography}{10}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
 
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[width=\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_tga1Supp_June.png}
-    \end{center}
-\caption{Diversity surrounding the causative substitution at the \emph{tga1} locus.  \label{sFig:tga1}}
-\end{figure}
-\clearpage
+\bibitem{dobzhansky1957}
+\bibinfo{author}{Dobzhansky, T.} \& \bibinfo{author}{Pavlovsky, O.}
+\newblock \bibinfo{title}{An experimental study of interaction between genetic
+  drift and natural selection}.
+\newblock \emph{\bibinfo{journal}{Evolution}} \bibinfo{pages}{311--319}
+  (\bibinfo{year}{1957}).
 
-\begin{figure}
-  \includegraphics[width=.5\textwidth]{FigsAndFiles/TIL_BKN_bootstrapping_msmc2_labeled.pdf}
-  \includegraphics[width=.5\textwidth]{FigsAndFiles/relativeCrossCoalescenceRate_labeled.pdf}
-\caption{MSMC Analyses.
-Shown in \textbf{A} are effective population size estimates over time.
-Estimates are depicted as solid lines and boostrap resampling is represented with dotted lines for both maize (red) and teosinte (blue).
-\textbf{B} depticts the relative cross-coalescence rate between maize and teosinte estimated using MSMC.
-In both panels, time is estimated assuming an annual generation time and a mutation rate of $\mu=3\times 10^{-8}$ \label{sFig:msmc}}
-\end{figure}
-\clearpage
+\bibitem{voight2005}
+\bibinfo{author}{Voight, B.~F.} \emph{et~al.}
+\newblock \bibinfo{title}{Interrogating multiple aspects of variation in a full
+  resequencing data set to infer human population size changes}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences of the United States of America}} \textbf{\bibinfo{volume}{102}},
+  \bibinfo{pages}{18508--18513} (\bibinfo{year}{2005}).
 
-\begin{figure}
-  \includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Conserved_Significance_June.png}
-  \includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Unconserved_Significance_June.png}
-\caption{ Pairwise diversity surrounding synonymous and nonsynonymous
-  substitutions in maize at \textbf{A} highly conserved or \textbf{B}  unconserved sites.
-Bootstrap-based 95\% confidence intervals are depicted via shading.
-Inset plots depict a larger range on the x-axis.
-\label{sFig:consUncons}}
-\end{figure}
-\clearpage
+\bibitem{gutenkunst2009}
+\bibinfo{author}{Gutenkunst, R.~N.}, \bibinfo{author}{Hernandez, R.~D.},
+  \bibinfo{author}{Williamson, S.~H.} \& \bibinfo{author}{Bustamante, C.~D.}
+\newblock \bibinfo{title}{Inferring the joint demographic history of multiple
+  populations from multidimensional snp frequency data}.
+\newblock \emph{\bibinfo{journal}{PLoS genetics}} \textbf{\bibinfo{volume}{5}},
+  \bibinfo{pages}{e1000695} (\bibinfo{year}{2009}).
 
-\begin{figure}
-  \includegraphics[width=\textwidth]{FigsAndFiles/plotDiversity_TvM_Singletons.png}
-\caption{ Singleton diversity surrounding synonymous and nonsynonymous
-  substitutions in maize. \label{sFig:singleton}}
-\end{figure}
-\clearpage
+\bibitem{akey2009}
+\bibinfo{author}{Akey, J.~M.}
+\newblock \bibinfo{title}{Constructing genomic maps of positive selection in
+  humans: Where do we go from here?}
+\newblock \emph{\bibinfo{journal}{Genome research}}
+  \textbf{\bibinfo{volume}{19}}, \bibinfo{pages}{711--722}
+  (\bibinfo{year}{2009}).
 
-\begin{figure}
-  \includegraphics[width=\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_maizeAndTeoSingleVsPi.png}
-\caption{ Relative diversity versus distance to nearest gene in maize and teosinte.
-Relative diversity is calculated by comparing to the mean diversity in all windows $\geq 0.02 cM$ from the nearest gene.
-Lines depict cubic smoothing splines with smoothing parameters chosen via generalized cross validation and shading depicts bootstrap-based 95\% confidence intervals.
-\label{sFig:singletonPi}}
-\end{figure}
-\clearpage
+\bibitem{smith1974}
+\bibinfo{author}{Smith, J.~M.} \& \bibinfo{author}{Haigh, J.}
+\newblock \bibinfo{title}{The hitch-hiking effect of a favourable gene}.
+\newblock \emph{\bibinfo{journal}{Genetical research}}
+  \textbf{\bibinfo{volume}{23}}, \bibinfo{pages}{23--35}
+  (\bibinfo{year}{1974}).
 
-\begin{figure}
-  \includegraphics[width=\textwidth]{FigsAndFiles/jri_sims.png}
-\caption{Simulations of diversity statistics in maize and teosinte with varying strengths of purifying selection.
-Points show the mean ($\pm$ standard error) of the population mutation rate $\theta$ estimated by singletons ($\eta_1$) and pairwise differences ($\pi$).
-\label{sFig:sims}}
-\end{figure}
-\clearpage
+\bibitem{slotte2014}
+\bibinfo{author}{Slotte, T.}
+\newblock \bibinfo{title}{The impact of linked selection on plant genomic
+  variation}.
+\newblock \emph{\bibinfo{journal}{Briefings in functional genomics}}
+  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{268--275}
+  (\bibinfo{year}{2014}).
 
-\begin{figure}
-  \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_Unselected_manuscript.png}
-    \includegraphics[width=.5\textwidth]{FigsAndFiles/distanceToGene_unselected_Singletons_manuscript.png}
-\caption{ Relative level of diversity versus distance to the nearest gene, in maize and teosinte, based on only sites that do not show evidence of hard or soft sweeps according to H12.
-Two measures of diversity were investigated.
-{\bf A} displays pairwise diversity,
-which is most influenced by intermediate frequency alleles and therefore depicts more ancient evolutionary patterns, and {\bf B} depicts singleton diversity, influenced by rare alleles and thus depicting evolutionary patterns in the recent past.
-Bootstrap-based 95\% confidence intervals are depicted via shading.
-Inset plots depict a smaller range on the x-axis.
-\label{sFig:H12}}
-\end{figure}
-\clearpage
+\bibitem{charlesworth1993}
+\bibinfo{author}{Charlesworth, B.}, \bibinfo{author}{Morgan, M.} \&
+  \bibinfo{author}{Charlesworth, D.}
+\newblock \bibinfo{title}{The effect of deleterious mutations on neutral
+  molecular variation.}
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{134}},
+  \bibinfo{pages}{1289--1303} (\bibinfo{year}{1993}).
 
-\begin{figure}
-  \begin{center}
-  \includegraphics[width=.75\textwidth]{FigsAndFiles/tilPCA_aug.pdf}\\
-  \includegraphics[width=.75\textwidth]{FigsAndFiles/bknPCA_aug.pdf}\\
-  \end{center}
-  \caption{Principal component analysis of teosinte and maize individuals to ensure that no close relatives were inadvertantly included in our study.
-Plots are based on a random sample of 10,000 SNPs.
-{\bf A} displays the percentage of total variance explained by each principal component for teosinte, while {\bf B} shows PC1 vs PC2 for all 13 teosinte individuals.
-Simlarly, {\bf C} depicts the percentage of total variance explained by each principal component for maize, and {\bf D} shows PC1 vs PC2 for all 23 maize individuals.
-\label{sFig:PCA}}
-\end{figure}
-\clearpage
+\bibitem{sella2009}
+\bibinfo{author}{Sella, G.}, \bibinfo{author}{Petrov, D.~A.},
+  \bibinfo{author}{Przeworski, M.} \& \bibinfo{author}{Andolfatto, P.}
+\newblock \bibinfo{title}{Pervasive natural selection in the drosophila
+  genome?}
+\newblock \emph{\bibinfo{journal}{PLoS genetics}} \textbf{\bibinfo{volume}{5}},
+  \bibinfo{pages}{e1000495} (\bibinfo{year}{2009}).
 
-\begin{table}[h]  
-  \begin{center}
-  \begin{tabular}{c|c}
-    \bf Maize & \bf Teosinte \\ \hline \hline
-    BKN009 &  TIL01 \\
-    BKN010 & TIL02 \\
-    BKN011 & TIL03 \\
-    BKN014 & TIL04-TIP454 \\
-    BKN015 & TIL07 \\
-    BKN016 & TIL09 \\
-    BKN017 & TIL10 \\
-    BKN018 & TIL11 \\
-    BKN019 & TIL12 \\
-    BKN020 & TIL14-TIP498 \\
-    BKN022 & TIL15 \\
-    BKN023 & TIL16 \\
-    BKN025 & TIL17 \\
-    BKN026 & \\
-    BKN027 & \\
-    BKN029 & \\
-    BKN030 & \\
-    BKN031 & \\
-    BKN032 & \\
-    BKN033 & \\
-    BKN034 & \\
-    BKN035 & \\
-    BKN040 & \\ \hline
-  \end{tabular}
-  \end{center}
-  \caption{A list of maize and teosinte individuals included in this study. Sequencing and details were previously described by \cite{chia2012}   } \label{sTab:list}
-\end{table}
-\clearpage
+\bibitem{elyashiv2014}
+\bibinfo{author}{Elyashiv, E.} \emph{et~al.}
+\newblock \bibinfo{title}{A genomic map of the effects of linked selection in
+  drosophila}.
+\newblock \emph{\bibinfo{journal}{arXiv preprint arXiv:1408.5461}}
+  (\bibinfo{year}{2014}).
 
-\begin{table}
-  \begin{center}
-  \def\arraystretch{2}
-%  \begin{center}
-  \begin{tabular}{l|c|c|c}
-    \bf Parameter & \bf Initial value & \bf Upper bound & \bf Lower bound\\ \hline 
-    $\frac{N_b}{N_a}$ & 0.02 & $1\times10^{-7}$ & 2 \\ 
-    $\frac{N_{m}}{N_a}$ & 3 & $1\times10^{-7}$ & 200 \\
-    $\frac{T_b}{2N_a}$ & 0.04 & 0  & 1 \\ 
-    $\frac{M_{mt}}{N_a}$ & $1\times10^{-10}$ & $1\times10^{-7}$ & 0.001 \\
-    $\frac{M_{tm}}{N_a}$ & $1\times10^{-10}$ & $1\times10^{-7}$ & 0.001 \\
-  \end{tabular}
-  \def\arraystretch{1} % undo stretching
-  \end{center}
-    \caption{ Parameters, initial values, and boundaries used for model-fitting with $\delta\alpha\delta{i}$.
-Parameters are shown in the units utilized by $\delta\alpha\delta{i}$, although in the text simplified units are reported.
-}\label{sTab:dadi}
-\end{table}
+\bibitem{andolfatto2005}
+\bibinfo{author}{Andolfatto, P.}
+\newblock \bibinfo{title}{Adaptive evolution of non-coding dna in drosophila}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{437}},
+  \bibinfo{pages}{1149--1152} (\bibinfo{year}{2005}).
+
+\bibitem{cutter2013}
+\bibinfo{author}{Cutter, A.~D.} \& \bibinfo{author}{Payseur, B.~A.}
+\newblock \bibinfo{title}{Genomic signatures of selection at linked sites:
+  unifying the disparity among species}.
+\newblock \emph{\bibinfo{journal}{Nature Reviews Genetics}}
+  \textbf{\bibinfo{volume}{14}}, \bibinfo{pages}{262--274}
+  (\bibinfo{year}{2013}).
+
+\bibitem{corbett2015}
+\bibinfo{author}{Corbett-Detig, R.~B.}, \bibinfo{author}{Hartl, D.~L.} \&
+  \bibinfo{author}{Sackton, T.~B.}
+\newblock \bibinfo{title}{Natural selection constrains neutral diversity across
+  a wide range of species}.
+\newblock \emph{\bibinfo{journal}{PLoS Biol}} \textbf{\bibinfo{volume}{13}},
+  \bibinfo{pages}{e1002112} (\bibinfo{year}{2015}).
+\newblock \urlprefix\url{http://dx.doi.org/10.1371%2Fjournal.pbio.1002112}.
+
+\bibitem{leffler2012}
+\bibinfo{author}{Leffler, E.~M.} \emph{et~al.}
+\newblock \bibinfo{title}{Revisiting an old riddle: what determines genetic
+  diversity levels within species}.
+\newblock \emph{\bibinfo{journal}{PLoS Biol}} \textbf{\bibinfo{volume}{10}},
+  \bibinfo{pages}{e1001388} (\bibinfo{year}{2012}).
+
+\bibitem{duchen2013}
+\bibinfo{author}{Duchen, P.}, \bibinfo{author}{{\v{Z}}ivkovi{\'c}, D.},
+  \bibinfo{author}{Hutter, S.}, \bibinfo{author}{Stephan, W.} \&
+  \bibinfo{author}{Laurent, S.}
+\newblock \bibinfo{title}{Demographic inference reveals african and european
+  admixture in the north american drosophila melanogaster population}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{193}},
+  \bibinfo{pages}{291--301} (\bibinfo{year}{2013}).
+
+\bibitem{reich1998}
+\bibinfo{author}{Reich, D.~E.} \& \bibinfo{author}{Goldstein, D.~B.}
+\newblock \bibinfo{title}{Genetic evidence for a paleolithic human population
+  expansion in africa}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{95}}, \bibinfo{pages}{8119--8123}
+  (\bibinfo{year}{1998}).
+
+\bibitem{meyer2013}
+\bibinfo{author}{Meyer, R.~S.} \& \bibinfo{author}{Purugganan, M.~D.}
+\newblock \bibinfo{title}{Evolution of crop species: genetics of domestication
+  and diversification}.
+\newblock \emph{\bibinfo{journal}{Nature Reviews Genetics}}
+  \textbf{\bibinfo{volume}{14}}, \bibinfo{pages}{840--852}
+  (\bibinfo{year}{2013}).
+
+\bibitem{ellegren2014}
+\bibinfo{author}{Ellegren, H.}
+\newblock \bibinfo{title}{Genome sequencing and population genomics in
+  non-model organisms}.
+\newblock \emph{\bibinfo{journal}{Trends in ecology \& evolution}}
+  \textbf{\bibinfo{volume}{29}}, \bibinfo{pages}{51--63}
+  (\bibinfo{year}{2014}).
+
+\bibitem{matsuoka2002}
+\bibinfo{author}{Matsuoka, Y.} \emph{et~al.}
+\newblock \bibinfo{title}{A single domestication for maize shown by multilocus
+  microsatellite genotyping}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{99}}, \bibinfo{pages}{6080--6084}
+  (\bibinfo{year}{2002}).
+
+\bibitem{wright2005}
+\bibinfo{author}{Wright, S.~I.} \emph{et~al.}
+\newblock \bibinfo{title}{The effects of artificial selection on the maize
+  genome}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{308}},
+  \bibinfo{pages}{1310--1314} (\bibinfo{year}{2005}).
+
+\bibitem{eyre1998}
+\bibinfo{author}{Eyre-Walker, A.}, \bibinfo{author}{Gaut, R.~L.},
+  \bibinfo{author}{Hilton, H.}, \bibinfo{author}{Feldman, D.~L.} \&
+  \bibinfo{author}{Gaut, B.~S.}
+\newblock \bibinfo{title}{Investigation of the bottleneck leading to the
+  domestication of maize}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{95}}, \bibinfo{pages}{4441--4446}
+  (\bibinfo{year}{1998}).
+
+\bibitem{tenaillon2004}
+\bibinfo{author}{Tenaillon, M.~I.}, \bibinfo{author}{U'Ren, J.},
+  \bibinfo{author}{Tenaillon, O.} \& \bibinfo{author}{Gaut, B.~S.}
+\newblock \bibinfo{title}{Selection versus demography: a multilocus
+  investigation of the domestication process in maize}.
+\newblock \emph{\bibinfo{journal}{Molecular Biology and Evolution}}
+  \textbf{\bibinfo{volume}{21}}, \bibinfo{pages}{1214--1225}
+  (\bibinfo{year}{2004}).
+
+\bibitem{chia2012}
+\bibinfo{author}{Chia, J.-M.} \emph{et~al.}
+\newblock \bibinfo{title}{Maize hapmap2 identifies extant variation from a
+  genome in flux}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}
+  \textbf{\bibinfo{volume}{44}}, \bibinfo{pages}{803--807}
+  (\bibinfo{year}{2012}).
+
+\bibitem{bukowski2015}
+\bibinfo{author}{Bukowski, R.} \emph{et~al.}
+\newblock \bibinfo{title}{Construction of the third generation zea mays
+  haplotype map}.
+\newblock \emph{\bibinfo{journal}{bioRxiv}} \bibinfo{pages}{026963}
+  (\bibinfo{year}{2015}).
+
+\bibitem{hufford2012}
+\bibinfo{author}{Hufford, M.~B.} \emph{et~al.}
+\newblock \bibinfo{title}{Comparative population genomics of maize
+  domestication and improvement}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}
+  \textbf{\bibinfo{volume}{44}}, \bibinfo{pages}{808--811}
+  (\bibinfo{year}{2012}).
+
+\bibitem{ewing2016}
+\bibinfo{author}{Ewing, G.~B.} \& \bibinfo{author}{Jensen, J.~D.}
+\newblock \bibinfo{title}{The consequences of not accounting for background
+  selection in demographic inference}.
+\newblock \emph{\bibinfo{journal}{Molecular ecology}}
+  \textbf{\bibinfo{volume}{25}}, \bibinfo{pages}{135--141}
+  (\bibinfo{year}{2016}).
+
+\bibitem{Hearne2015}
+\bibinfo{author}{Hearne, S.}, \bibinfo{author}{Chen, C.},
+  \bibinfo{author}{Buckler, E.} \& \bibinfo{author}{Mitchell, S.}
+\newblock \bibinfo{title}{Unimputed gbs derived snps for maize landrace
+  accessions represented in the seed-maize gwas panel}.
+\newblock \bibinfo{howpublished}{\url{http://hdl.handle.net/11529/10034}}
+  (\bibinfo{year}{2015}).
+\newblock \bibinfo{note}{Accessed: 2015-02-16}.
+
+\bibitem{keinan2012}
+\bibinfo{author}{Keinan, A.} \& \bibinfo{author}{Clark, A.~G.}
+\newblock \bibinfo{title}{Recent explosive human population growth has resulted
+  in an excess of rare genetic variants}.
+\newblock \emph{\bibinfo{journal}{science}} \textbf{\bibinfo{volume}{336}},
+  \bibinfo{pages}{740--743} (\bibinfo{year}{2012}).
+
+\bibitem{fu1993}
+\bibinfo{author}{Fu, Y.-X.} \& \bibinfo{author}{Li, W.-H.}
+\newblock \bibinfo{title}{Statistical tests of neutrality of mutations.}
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{133}},
+  \bibinfo{pages}{693--709} (\bibinfo{year}{1993}).
+
+\bibitem{clark2005}
+\bibinfo{author}{Clark, R.~M.}, \bibinfo{author}{Tavar{\'e}, S.} \&
+  \bibinfo{author}{Doebley, J.}
+\newblock \bibinfo{title}{Estimating a nucleotide substitution rate for maize
+  from polymorphism at a major domestication locus}.
+\newblock \emph{\bibinfo{journal}{Molecular biology and evolution}}
+  \textbf{\bibinfo{volume}{22}}, \bibinfo{pages}{2304--2312}
+  (\bibinfo{year}{2005}).
+
+\bibitem{schiffels2014}
+\bibinfo{author}{Schiffels, S.} \& \bibinfo{author}{Durbin, R.}
+\newblock \bibinfo{title}{Inferring human population size and separation
+  history from multiple genome sequences}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}  (\bibinfo{year}{2014}).
+
+\bibitem{wang2015}
+\bibinfo{author}{Wang, H.}, \bibinfo{author}{Studer, A.~J.},
+  \bibinfo{author}{Zhao, Q.}, \bibinfo{author}{Meeley, R.} \&
+  \bibinfo{author}{Doebley, J.~F.}
+\newblock \bibinfo{title}{Evidence that the origin of naked kernels during
+  maize domestication was caused by a single amino acid substitution in tga1}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \bibinfo{pages}{genetics--115}
+  (\bibinfo{year}{2015}).
+
+\bibitem{enard2014}
+\bibinfo{author}{Enard, D.}, \bibinfo{author}{Messer, P.~W.} \&
+  \bibinfo{author}{Petrov, D.~A.}
+\newblock \bibinfo{title}{Genome-wide signals of positive selection in human
+  evolution}.
+\newblock \emph{\bibinfo{journal}{Genome research}}
+  \textbf{\bibinfo{volume}{24}}, \bibinfo{pages}{885--895}
+  (\bibinfo{year}{2014}).
+
+\bibitem{rodgers2015}
+\bibinfo{author}{Rodgers-Melnick, E.} \emph{et~al.}
+\newblock \bibinfo{title}{Recombination in diverse maize is stable,
+  predictable, and associated with genetic load}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{112}}, \bibinfo{pages}{3823--3828}
+  (\bibinfo{year}{2015}).
+
+\bibitem{davydov2010}
+\bibinfo{author}{Davydov, E.~V.} \emph{et~al.}
+\newblock \bibinfo{title}{Identifying a high fraction of the human genome to be
+  under selective constraint using gerp++}.
+\newblock \emph{\bibinfo{journal}{PLoS Comput Biol}}
+  \textbf{\bibinfo{volume}{6}}, \bibinfo{pages}{e1001025}
+  (\bibinfo{year}{2010}).
+
+\bibitem{messer2013}
+\bibinfo{author}{Messer, P.~W.} \& \bibinfo{author}{Petrov, D.~A.}
+\newblock \bibinfo{title}{Population genomics of rapid adaptation by soft
+  selective sweeps}.
+\newblock \emph{\bibinfo{journal}{Trends in ecology \& evolution}}
+  \textbf{\bibinfo{volume}{28}}, \bibinfo{pages}{659--669}
+  (\bibinfo{year}{2013}).
+
+\bibitem{garud2015}
+\bibinfo{author}{Garud, N.~R.}, \bibinfo{author}{Messer, P.~W.},
+  \bibinfo{author}{Buzbas, E.~O.} \& \bibinfo{author}{Petrov, D.~A.}
+\newblock \bibinfo{title}{Recent selective sweeps in north american drosophila
+  melanogaster show signatures of soft sweeps}.
+\newblock \emph{\bibinfo{journal}{PLoS genetics}}
+  \textbf{\bibinfo{volume}{11}}, \bibinfo{pages}{e1005004}
+  (\bibinfo{year}{2015}).
+
+\bibitem{piperno2009}
+\bibinfo{author}{Piperno, D.~R.}, \bibinfo{author}{Ranere, A.~J.},
+  \bibinfo{author}{Holst, I.}, \bibinfo{author}{Iriarte, J.} \&
+  \bibinfo{author}{Dickau, R.}
+\newblock \bibinfo{title}{Starch grain and phytolith evidence for early ninth
+  millennium bp maize from the central balsas river valley, mexico}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{106}}, \bibinfo{pages}{5019--5024}
+  (\bibinfo{year}{2009}).
+
+\bibitem{cimmyt1999}
+\bibinfo{author}{Program, T.~M.}
+\newblock \emph{\bibinfo{title}{Development, maintenance, and seed
+  multiplication of open-pollinated maize varieties}}
+  (\bibinfo{publisher}{CIMMYT}, \bibinfo{address}{Mexico, D.F.},
+  \bibinfo{year}{1999}), \bibinfo{edition}{2} edn.
+
+\bibitem{baden2001}
+\bibinfo{author}{Baden, W.~W.} \& \bibinfo{author}{Beekman, C.~S.}
+\newblock \bibinfo{title}{Culture and agriculture: A comment on sissel
+  schroeder, maize productivity in the eastern woodlands and great plains of
+  north america}.
+\newblock \emph{\bibinfo{journal}{American Antiquity}}
+  \bibinfo{pages}{505--515} (\bibinfo{year}{2001}).
+
+\bibitem{studer2011}
+\bibinfo{author}{Studer, A.}, \bibinfo{author}{Zhao, Q.},
+  \bibinfo{author}{Ross-Ibarra, J.} \& \bibinfo{author}{Doebley, J.}
+\newblock \bibinfo{title}{Identification of a functional transposon insertion
+  in the maize domestication gene tb1}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}
+  \textbf{\bibinfo{volume}{43}}, \bibinfo{pages}{1160--1163}
+  (\bibinfo{year}{2011}).
+
+\bibitem{wills2013}
+\bibinfo{author}{Wills, D.~M.} \emph{et~al.}
+\newblock \bibinfo{title}{From many, one: genetic control of prolificacy during
+  maize domestication}.
+\newblock \emph{\bibinfo{journal}{PLoS Genet}} \textbf{\bibinfo{volume}{9}},
+  \bibinfo{pages}{e1003604} (\bibinfo{year}{2013}).
+
+\bibitem{Takuno15062015}
+\bibinfo{author}{Takuno, S.} \emph{et~al.}
+\newblock \bibinfo{title}{Independent molecular basis of convergent highland
+  adaptation in maize}.
+\newblock \emph{\bibinfo{journal}{Genetics}}  (\bibinfo{year}{2015}).
+\newblock
+  \urlprefix\url{http://www.genetics.org/content/early/2015/06/15/genetics.115.178327.abstract}.
+\newblock
+  \eprint{http://www.genetics.org/content/early/2015/06/15/genetics.115.178327.full.pdf+html}.
+
+\bibitem{van2012historical}
+\bibinfo{author}{van Heerwaarden, J.}, \bibinfo{author}{Hufford, M.~B.} \&
+  \bibinfo{author}{Ross-Ibarra, J.}
+\newblock \bibinfo{title}{Historical genomics of north american maize}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{109}}, \bibinfo{pages}{12420--12425}
+  (\bibinfo{year}{2012}).
+
+\bibitem{beissinger2014}
+\bibinfo{author}{Beissinger, T.~M.} \emph{et~al.}
+\newblock \bibinfo{title}{A genome-wide scan for evidence of selection in a
+  maize population under long-term artificial selection for ear number}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{196}},
+  \bibinfo{pages}{829--840} (\bibinfo{year}{2014}).
+
+\bibitem{eyre2007}
+\bibinfo{author}{Eyre-Walker, A.} \& \bibinfo{author}{Keightley, P.~D.}
+\newblock \bibinfo{title}{The distribution of fitness effects of new
+  mutations}.
+\newblock \emph{\bibinfo{journal}{Nature Reviews Genetics}}
+  \textbf{\bibinfo{volume}{8}}, \bibinfo{pages}{610--618}
+  (\bibinfo{year}{2007}).
+
+\bibitem{wallace2014}
+\bibinfo{author}{Wallace, J.}, \bibinfo{author}{Larsson, S.} \&
+  \bibinfo{author}{Buckler, E.}
+\newblock \bibinfo{title}{Entering the second century of maize quantitative
+  genetics}.
+\newblock \emph{\bibinfo{journal}{Heredity}} \textbf{\bibinfo{volume}{112}},
+  \bibinfo{pages}{30--38} (\bibinfo{year}{2014}).
+
+\bibitem{weber2008}
+\bibinfo{author}{Weber, A.~L.} \emph{et~al.}
+\newblock \bibinfo{title}{The genetic architecture of complex traits in
+  teosinte (zea mays ssp. parviglumis): new evidence from association mapping}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{180}},
+  \bibinfo{pages}{1221--1232} (\bibinfo{year}{2008}).
+
+\bibitem{pyhajarvi2013}
+\bibinfo{author}{Pyh{\"a}j{\"a}rvi, T.}, \bibinfo{author}{Hufford, M.~B.},
+  \bibinfo{author}{Mezmouk, S.} \& \bibinfo{author}{Ross-Ibarra, J.}
+\newblock \bibinfo{title}{Complex patterns of local adaptation in teosinte}.
+\newblock \emph{\bibinfo{journal}{Genome biology and evolution}}
+  \textbf{\bibinfo{volume}{5}}, \bibinfo{pages}{1594--1609}
+  (\bibinfo{year}{2013}).
+
+\bibitem{sattath2011}
+\bibinfo{author}{Sattath, S.}, \bibinfo{author}{Elyashiv, E.},
+  \bibinfo{author}{Kolodny, O.}, \bibinfo{author}{Rinott, Y.} \&
+  \bibinfo{author}{Sella, G.}
+\newblock \bibinfo{title}{Pervasive adaptive protein evolution apparent in
+  diversity patterns around amino acid substitutions in drosophila simulans}.
+\newblock \emph{\bibinfo{journal}{PLoS genetics}} \textbf{\bibinfo{volume}{7}},
+  \bibinfo{pages}{e1001302} (\bibinfo{year}{2011}).
+
+\bibitem{williamson2014}
+\bibinfo{author}{Williamson, R.} \emph{et~al.}
+\newblock \bibinfo{title}{Evidence for widespread positive and negative
+  selection in coding and conserved noncoding regions of capsella grandiflora.}
+\newblock \emph{\bibinfo{journal}{PLoS genetics}}
+  \textbf{\bibinfo{volume}{10}}, \bibinfo{pages}{e1004622--e1004622}
+  (\bibinfo{year}{2014}).
+
+\bibitem{hernandez2011}
+\bibinfo{author}{Hernandez, R.~D.} \emph{et~al.}
+\newblock \bibinfo{title}{Classic selective sweeps were rare in recent human
+  evolution}.
+\newblock \emph{\bibinfo{journal}{science}} \textbf{\bibinfo{volume}{331}},
+  \bibinfo{pages}{920--924} (\bibinfo{year}{2011}).
+
+\bibitem{ross2009}
+\bibinfo{author}{Ross-Ibarra, J.}, \bibinfo{author}{Tenaillon, M.} \&
+  \bibinfo{author}{Gaut, B.~S.}
+\newblock \bibinfo{title}{Historical divergence and gene flow in the genus
+  zea}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{181}},
+  \bibinfo{pages}{1399--1413} (\bibinfo{year}{2009}).
+
+\bibitem{eyre2009estimating}
+\bibinfo{author}{Eyre-Walker, A.} \& \bibinfo{author}{Keightley, P.~D.}
+\newblock \bibinfo{title}{Estimating the rate of adaptive molecular evolution
+  in the presence of slightly deleterious mutations and population size
+  change}.
+\newblock \emph{\bibinfo{journal}{Molecular biology and evolution}}
+  \textbf{\bibinfo{volume}{26}}, \bibinfo{pages}{2097--2108}
+  (\bibinfo{year}{2009}).
+
+\bibitem{mao2015}
+\bibinfo{author}{Mao, H.} \emph{et~al.}
+\newblock \bibinfo{title}{A transposable element in a nac gene is associated
+  with drought tolerance in maize seedlings}.
+\newblock \emph{\bibinfo{journal}{Nature Communications}}
+  \textbf{\bibinfo{volume}{6}} (\bibinfo{year}{2015}).
+
+\bibitem{yang2013}
+\bibinfo{author}{Yang, Q.} \emph{et~al.}
+\newblock \bibinfo{title}{{CACTA}-like transposable element in {ZmCCT}
+  attenuated photoperiod sensitivity and accelerated the postdomestication
+  spread of maize}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{110}}, \bibinfo{pages}{16969--16974}
+  (\bibinfo{year}{2013}).
+
+\bibitem{fraser2013gene}
+\bibinfo{author}{Fraser, H.~B.}
+\newblock \bibinfo{title}{Gene expression drives local adaptation in humans}.
+\newblock \emph{\bibinfo{journal}{Genome research}}
+  \textbf{\bibinfo{volume}{23}}, \bibinfo{pages}{1089--1096}
+  (\bibinfo{year}{2013}).
+
+\bibitem{hancock2011adaptation}
+\bibinfo{author}{Hancock, A.~M.} \emph{et~al.}
+\newblock \bibinfo{title}{Adaptation to climate across the arabidopsis thaliana
+  genome}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{334}},
+  \bibinfo{pages}{83--86} (\bibinfo{year}{2011}).
+
+\bibitem{Halligan:2013ir}
+\bibinfo{author}{Halligan, D.~L.} \emph{et~al.}
+\newblock \bibinfo{title}{{Contributions of Protein-Coding and Regulatory
+  Change to Adaptive Molecular Evolution in Murid Rodents}}.
+\newblock \emph{\bibinfo{journal}{PLoS Genetics}} \textbf{\bibinfo{volume}{9}},
+  \bibinfo{pages}{e1003995--14} (\bibinfo{year}{2013}).
+
+\bibitem{kimura1984}
+\bibinfo{author}{Kimura, M.}
+\newblock \emph{\bibinfo{title}{The neutral theory of molecular evolution}}
+  (\bibinfo{publisher}{Cambridge University Press}, \bibinfo{year}{1984}).
+
+\bibitem{Gunther:2010gt}
+\bibinfo{author}{G{\"u}nther, T.} \& \bibinfo{author}{Schmid, K.~J.}
+\newblock \bibinfo{title}{{Deleterious amino acid polymorphisms in Arabidopsis
+  thaliana and rice}}.
+\newblock \emph{\bibinfo{journal}{Theoretical and Applied Genetics}}
+  \textbf{\bibinfo{volume}{121}}, \bibinfo{pages}{157--168}
+  (\bibinfo{year}{2010}).
+
+\bibitem{renaut2015}
+\bibinfo{author}{Renaut, S.} \& \bibinfo{author}{Rieseberg, L.~H.}
+\newblock \bibinfo{title}{The accumulation of deleterious mutations as a
+  consequence of domestication and improvement in sunflowers and other
+  compositae crops}.
+\newblock \emph{\bibinfo{journal}{Molecular biology and evolution}}
+  \bibinfo{pages}{msv106} (\bibinfo{year}{2015}).
+
+\bibitem{coventry2010}
+\bibinfo{author}{Coventry, A.} \emph{et~al.}
+\newblock \bibinfo{title}{Deep resequencing reveals excess rare recent variants
+  consistent with explosive population growth}.
+\newblock \emph{\bibinfo{journal}{Nature communications}}
+  \textbf{\bibinfo{volume}{1}}, \bibinfo{pages}{131} (\bibinfo{year}{2010}).
+
+\bibitem{Mezmouk:2014es}
+\bibinfo{author}{Mezmouk, S.} \& \bibinfo{author}{Ross-Ibarra, J.}
+\newblock \bibinfo{title}{{The pattern and distribution of deleterious
+  mutations in maize.}}
+\newblock \emph{\bibinfo{journal}{G3 (Bethesda, Md.)}}
+  \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{163--171}
+  (\bibinfo{year}{2014}).
+
+\bibitem{eyre2010}
+\bibinfo{author}{Eyre-Walker, A.}
+\newblock \bibinfo{title}{Genetic architecture of a complex trait and its
+  implications for fitness and genome-wide association studies}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{107}}, \bibinfo{pages}{1752--1756}
+  (\bibinfo{year}{2010}).
+
+\bibitem{do2015}
+\bibinfo{author}{Do, R.} \emph{et~al.}
+\newblock \bibinfo{title}{No evidence that selection has been less effective at
+  removing deleterious mutations in europeans than in africans}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}
+  \textbf{\bibinfo{volume}{47}}, \bibinfo{pages}{126--131}
+  (\bibinfo{year}{2015}).
+
+\bibitem{simons2014}
+\bibinfo{author}{Simons, Y.~B.}, \bibinfo{author}{Turchin, M.~C.},
+  \bibinfo{author}{Pritchard, J.~K.} \& \bibinfo{author}{Sella, G.}
+\newblock \bibinfo{title}{The deleterious mutation load is insensitive to
+  recent population history}.
+\newblock \emph{\bibinfo{journal}{Nature genetics}}
+  \textbf{\bibinfo{volume}{46}}, \bibinfo{pages}{220--224}
+  (\bibinfo{year}{2014}).
+
+\bibitem{lohmueller2014}
+\bibinfo{author}{Lohmueller, K.~E.}
+\newblock \bibinfo{title}{The impact of population demography and selection on
+  the genetic architecture of complex traits}.
+\newblock \emph{\bibinfo{journal}{PLoS Genetics}} \textbf{\bibinfo{volume}{10}}
+  (\bibinfo{year}{2014}).
+
+\bibitem{zeng2010}
+\bibinfo{author}{Zeng, K.} \& \bibinfo{author}{Charlesworth, B.}
+\newblock \bibinfo{title}{The effects of demography and linkage on the
+  estimation of selection and mutation parameters}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{186}},
+  \bibinfo{pages}{1411--1424} (\bibinfo{year}{2010}).
+
+\bibitem{popadin2012}
+\bibinfo{author}{Popadin, K.~Y.}, \bibinfo{author}{Nikolaev, S.~I.},
+  \bibinfo{author}{Junier, T.}, \bibinfo{author}{Baranova, M.} \&
+  \bibinfo{author}{Antonarakis, S.~E.}
+\newblock \bibinfo{title}{Purifying selection in mammalian mitochondrial
+  protein-coding genes is highly effective and congruent with evolution of
+  nuclear genes}.
+\newblock \emph{\bibinfo{journal}{Molecular biology and evolution}}
+  \bibinfo{pages}{mss219} (\bibinfo{year}{2012}).
+
+\bibitem{Elyashiv:2010ga}
+\bibinfo{author}{Elyashiv, E.} \emph{et~al.}
+\newblock \bibinfo{title}{{Shifts in the intensity of purifying selection: An
+  analysis of genome-wide polymorphism data from two closely related yeast
+  species}}.
+\newblock \emph{\bibinfo{journal}{Genome Research}}
+  \textbf{\bibinfo{volume}{20}}, \bibinfo{pages}{1558--1573}
+  (\bibinfo{year}{2010}).
+
+  \setcounter{firstbib}{\value{NAT@ctr}}
+
+\end{thebibliography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\renewcommand\refname{References (Methods)}
+
+  \begin{thebibliography}{10}
+\setcounter{NAT@ctr}{\value{firstbib}}
+    \expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{lemmon2014}
+\bibinfo{author}{Lemmon, Z.~H.}, \bibinfo{author}{Bukowski, R.},
+  \bibinfo{author}{Sun, Q.} \& \bibinfo{author}{Doebley, J.~F.}
+\newblock \bibinfo{title}{The role of cis regulatory evolution
+  in maize domestication}.
+\newblock \emph{\bibinfo{journal}{PLoS Genet}} \textbf{\bibinfo{volume}{10}},
+  \bibinfo{pages}{e1004745} (\bibinfo{year}{2014}).
+\newblock \urlprefix\url{http://dx.doi.org/10.1371%2Fjournal.pgen.1004745}.
+
+\bibitem{jombart2011}
+\bibinfo{author}{Jombart, T.} \& \bibinfo{author}{Ahmed, I.}
+\newblock \bibinfo{title}{adegenet 1.3-1: new tools for the analysis of
+  genome-wide snp data}.
+\newblock \emph{\bibinfo{journal}{Bioinformatics}}
+  \textbf{\bibinfo{volume}{27}}, \bibinfo{pages}{3070--3071}
+  (\bibinfo{year}{2011}).
+
+\bibitem{schnable2009}
+\bibinfo{author}{Schnable, P.~S.} \emph{et~al.}
+\newblock \bibinfo{title}{The b73 maize genome: complexity, diversity, and
+  dynamics}.
+\newblock \emph{\bibinfo{journal}{science}} \textbf{\bibinfo{volume}{326}},
+  \bibinfo{pages}{1112--1115} (\bibinfo{year}{2009}).
+
+\bibitem{li2011statistical}
+\bibinfo{author}{Li, H.}
+\newblock \bibinfo{title}{A statistical framework for snp calling, mutation
+  discovery, association mapping and population genetical parameter estimation
+  from sequencing data}.
+\newblock \emph{\bibinfo{journal}{Bioinformatics}}
+  \textbf{\bibinfo{volume}{27}}, \bibinfo{pages}{2987--2993}
+  (\bibinfo{year}{2011}).
+
+\bibitem{glaubitz2014}
+\bibinfo{author}{Glaubitz, J.~C.} \emph{et~al.}
+\newblock \bibinfo{title}{Tassel-gbs: a high capacity genotyping by sequencing
+  analysis pipeline}.
+\newblock \emph{\bibinfo{journal}{PLoS One}} \textbf{\bibinfo{volume}{9}},
+  \bibinfo{pages}{E90346} (\bibinfo{year}{2014}).
+
+\bibitem{durinck2009}
+\bibinfo{author}{Durinck, S.}, \bibinfo{author}{Spellman, P.~T.},
+  \bibinfo{author}{Birney, E.} \& \bibinfo{author}{Huber, W.}
+\newblock \bibinfo{title}{Mapping identifiers for the integration of genomic
+  datasets with the r/bioconductor package biomart}.
+\newblock \emph{\bibinfo{journal}{Nature protocols}}
+  \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{1184--1191}
+  (\bibinfo{year}{2009}).
+
+\bibitem{durinck2005}
+\bibinfo{author}{Durinck, S.} \emph{et~al.}
+\newblock \bibinfo{title}{Biomart and bioconductor: a powerful link between
+  biological databases and microarray data analysis}.
+\newblock \emph{\bibinfo{journal}{Bioinformatics}}
+  \textbf{\bibinfo{volume}{21}}, \bibinfo{pages}{3439--3440}
+  (\bibinfo{year}{2005}).
+
+\bibitem{R2014}
+\bibinfo{author}{{R Core Team}}.
+\newblock \emph{\bibinfo{title}{R: A Language and Environment for Statistical
+  Computing}}.
+\newblock \bibinfo{organization}{R Foundation for Statistical Computing},
+  \bibinfo{address}{Vienna, Austria} (\bibinfo{year}{2014}).
+\newblock \urlprefix\url{http://www.R-project.org/}.
+
+\bibitem{korneliussen2014}
+\bibinfo{author}{Korneliussen, T.~S.}, \bibinfo{author}{Albrechtsen, A.} \&
+  \bibinfo{author}{Nielsen, R.}
+\newblock \bibinfo{title}{Angsd: analysis of next generation sequencing data}.
+\newblock \emph{\bibinfo{journal}{BMC bioinformatics}}
+  \textbf{\bibinfo{volume}{15}}, \bibinfo{pages}{356} (\bibinfo{year}{2014}).
+
+\bibitem{mclaren2010}
+\bibinfo{author}{McLaren, W.} \emph{et~al.}
+\newblock \bibinfo{title}{Deriving the consequences of genomic variants with
+  the ensembl api and snp effect predictor}.
+\newblock \emph{\bibinfo{journal}{Bioinformatics}}
+  \textbf{\bibinfo{volume}{26}}, \bibinfo{pages}{2069--2070}
+  (\bibinfo{year}{2010}).
+
+\bibitem{thornton2014genetics}
+\bibinfo{author}{Thornton, K.~R.}
+\newblock \bibinfo{title}{A c++ template library for efficient forward-time
+  population genetic simulation of large populations}.
+\newblock \emph{\bibinfo{journal}{Genetics}} \textbf{\bibinfo{volume}{198}},
+  \bibinfo{pages}{157--166} (\bibinfo{year}{2014}).
+
+\end{thebibliography}
 
